---
title: "Player Playstyle Data Preprocessing"
author: "Thomas Hakman"
format: 
  html:
    toc: true
    toc-depth: 3
    code-fold: true
    theme: flatly
    fig-width: 8
    fig-height: 6
execute:
  echo: true
  message: false
  warning: false
  cache: true
editor: visual
---

```{r setup}
#| label: setup
#| include: false

# Set global options
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  cache = TRUE,
  fig.path = "figures/"
)
```

# Introduction

This document processes player telemetry and survey data from Plants vs. Zombies: Battle for Neighborville. The preprocessing steps prepare the data for analysis of the relationship between player playstyles, game performance metrics, and well-being outcomes.

# Data Preparation

## Load Packages

```{r packages}
#| label: load-packages
# Load required packages
library(tidyverse)
library(here)
library(lubridate)
library(scales)
library(janitor)
library(magrittr)
library(psych)
library(readxl)
library(knitr)
```

## Load Helper Functions

```{r helper-functions}
#| label: helper-functions

# Helper functions for PvZ data analysis
# These functions support the main preprocessing steps

# Function to read both waves of telemetry data into a single table
read_both_waves <- function(stub = "OII_PVZ_Authentications", data_dir = "data/raw/ea", ...) {
  # Filenames of wave 1 and 2 authentications files
  list.files(
    here::here(data_dir), pattern = stub, full.names = TRUE
  ) %>% 
    # Put wave 1 first (they are in reverse order)
    rev() %>% 
    # Read into a list of data frames
    purrr::map(readr::read_csv, ...) %>% 
    # Bind into one data frame with wave ID
    dplyr::bind_rows(.id = "wave")
}

# Function to create performance metrics from player data
calculate_performance_metrics <- function(df, prefix = "total") {
  # Handle cases where denominators might be zero
  df <- df %>%
    dplyr::mutate(
      # Set death count to 1 if 0 to avoid division by zero
      dplyr::across(
        dplyr::contains("death_count"), 
        ~ ifelse(.x == 0, 1, .x)
      )
    )
  
  # Create column names based on prefix
  kill_count_col <- paste0(prefix, ifelse(prefix == "", "", "_"), "kill_count")
  death_count_col <- paste0(prefix, ifelse(prefix == "", "", "_"), "death_count")
  shots_hit_col <- paste0(prefix, ifelse(prefix == "", "", "_"), "shots_hit")
  shots_fired_col <- paste0(prefix, ifelse(prefix == "", "", "_"), "shots_fired")
  crit_hit_col <- paste0(prefix, ifelse(prefix == "", "", "_"), "critical_hit_count")
  damage_col <- paste0(prefix, ifelse(prefix == "", "", "_"), "damage_dealt")
  sessions_col <- paste0(prefix, ifelse(prefix == "", "", "_"), "sessions")
  seconds_col <- paste0(prefix, ifelse(prefix == "", "", "_"), "seconds")
  score_col <- paste0(prefix, ifelse(prefix == "", "", "_"), "score")
  
  # Calculate and return metrics
  df %>%
    dplyr::mutate(
      # KD ratio
      !!paste0(prefix, ifelse(prefix == "", "", "_"), "kd") := 
        .data[[kill_count_col]] / .data[[death_count_col]],
      
      # Hit accuracy
      !!paste0(prefix, ifelse(prefix == "", "", "_"), "hit_accuracy") := 
        .data[[shots_hit_col]] / .data[[shots_fired_col]],
      
      # Critical hit ratio
      !!paste0(prefix, ifelse(prefix == "", "", "_"), "crit_ratio") := 
        .data[[crit_hit_col]] / .data[[shots_hit_col]],
      
      # Damage per second
      !!paste0(prefix, ifelse(prefix == "", "", "_"), "dps") := 
        .data[[damage_col]] / .data[[seconds_col]],
      
      # Damage per game
      !!paste0(prefix, ifelse(prefix == "", "", "_"), "average_damage") := 
        .data[[damage_col]] / .data[[sessions_col]],
      
      # Score per second
      !!paste0(prefix, ifelse(prefix == "", "", "_"), "sps") := 
        .data[[score_col]] / .data[[seconds_col]],
      
      # Score per game
      !!paste0(prefix, ifelse(prefix == "", "", "_"), "average_score") := 
        .data[[score_col]] / .data[[sessions_col]]
    ) %>%
    # Replace NaN with 0
    dplyr::mutate(
      dplyr::across(
        tidyselect::everything(), 
        ~ ifelse(is.nan(.x), 0, .x)
      )
    )
}

# Function to summarize mode-specific playing time
summarize_mode_hours <- function(df, mode_col, start_col = "game_start_time", end_col = "game_end_time") {
  df %>%
    dplyr::filter(.data[[mode_col]] == 1) %>%
    dplyr::mutate(duration = .data[[end_col]] - .data[[start_col]]) %>%
    dplyr::mutate(seconds = as.numeric(duration)) %>%
    dplyr::group_by(player_id) %>%
    dplyr::summarise(
      seconds = sum(seconds, na.rm = TRUE),
      .groups = "drop"
    ) %>%
    dplyr::mutate(hours = seconds / 60 / 60) %>%
    dplyr::rename_with(
      ~ paste0(gsub("_session", "", mode_col), "_", .),
      c(seconds, hours)
    )
}

# Function to calculate SPANE scales
calculate_spane <- function(df, prefix = "spane") {
  positive_items <- c(1, 3, 5, 7, 10, 12)
  negative_items <- c(2, 4, 6, 8, 9, 11)
  
  df %>%
    dplyr::mutate(
      # Positive affect
      !!paste0(prefix, "_positive") := rowMeans(
        dplyr::select(
          .,
          paste0(prefix, "_", positive_items)
        ),
        na.rm = TRUE
      ),
      # Negative affect
      !!paste0(prefix, "_negative") := rowMeans(
        dplyr::select(
          .,
          paste0(prefix, "_", negative_items)
        ),
        na.rm = TRUE
      ),
      # Affect balance
      !!paste0(prefix, "_balance") := .data[[paste0(prefix, "_positive")]] - 
        .data[[paste0(prefix, "_negative")]]
    )
}

# Function to create motivation scales
calculate_motivation_scales <- function(df) {
  df %>%
    dplyr::mutate(
      autonomy = rowMeans(
        dplyr::select(., dplyr::starts_with("autonomy_")), 
        na.rm = TRUE
      ),
      competence = rowMeans(
        dplyr::select(., dplyr::starts_with("competence_")), 
        na.rm = TRUE
      ),
      relatedness = rowMeans(
        dplyr::select(., dplyr::starts_with("relatedness_")), 
        na.rm = TRUE
      ),
      enjoyment = rowMeans(
        dplyr::select(., dplyr::starts_with("enjoyment_")), 
        na.rm = TRUE
      ),
      extrinsic = rowMeans(
        dplyr::select(., dplyr::starts_with("extrinsic_")), 
        na.rm = TRUE
      )
    )
}
```

## Load Raw Data

The data comes from a study conducted by EA and the Oxford Internet Institute. Players completed a survey on well-being and motivations, and EA provided telemetry data on their gameplay.

```{r raw-data-paths}
#| label: raw-data-paths

# Store all file paths in a central location for easy updates
data_paths <- list(
  survey = here("data/raw/ea/Oxford PvZ - Pilot Wave + Wave 1 Download.xlsx"),
  telemetry_files = list.files(
    here("data/raw/ea"), 
    pattern = "OII_PVZ_", 
    full.names = TRUE
  )
)

# The read_both_waves function is defined in the helper functions section above
```

## Load Survey Data

```{r survey-load}
#| label: survey-load

survey <- read_excel(data_paths$survey)

survey <- survey %>% 
  mutate(
    # Ensure date data type
    date = mdy_hm(date),
    # Additional date variable for filtering telemetry
    start_date = date
  ) %>% 
  # Rename ID column to match telemetry files
  rename(player_id = UID)
```

## Clean Survey Data

```{r survey-clean}
#| label: survey-clean

# Rename variables with meaningful names
survey <- 
  survey %>% 
  select(
    # meta-info
    player_id,
    start_date, date,
    survey_duration = qtime,
    of_age = Q2,
    consent = Q3,
    
    # demographics
    played = Q1,
    country = hCountry,
    survey_language = PRELANG,
    age = D1,
    gender = D2,
    gender_other = D2r3oe,
    experience = D3,
    
    # well-being
    spane_ = starts_with("A1"), # general well-being in the two weeks
    straightliner_spane = hA1Straight,
    
    # motivations
    played_with_others = NSat1, # filter question for relatedness items
    autonomy_ = NSat2r1:NSat2r3,
    competence_ = NSat2r4:NSat2r6,
    relatedness_ = NSat2r7:NSat2r9,
    enjoyment_ = NSat2r10:NSat2r13,
    extrinsic_ = NSat2r14:NSat2r17,
    straightliner_motivations = hNSat2Straight,
    
    # self-reported play
    active_play_hours = GT1c1,
    active_play_minutes = GT1c2,
    within_estimate = GT2,
    between_estimate = GT3,
    spane_game_ = starts_with("GT4")
  )

# Assign proper variable types and factor levels
survey <- survey %>%   
  mutate(
    across(
      c(
        player_id,
        of_age:survey_language,
        gender,
        played_with_others
      ),
      as.factor
    ),
    
    across(
      c(of_age:played, played_with_others),
      ~ fct_recode(
        .x,
        "Yes" = "1",
        "No" = "2"
      )
    ),
    
    gender = fct_recode(
      gender,
      "Male" = "1",
      "Female" = "2",
      "Other" = "3",
      "Prefer not to say" = "4"
    ),
    
    country = fct_recode(
      country,
      "US" = "1",
      "UK" = "2",
      "Canada" = "5"
    ),
    
    survey_language = fct_recode(
      survey_language,
      "English" = "1",
      "French" = "2"
    )
  )

# Apply reverse scoring for appropriate items
survey <- survey %>% 
  mutate(
    across(
      c(
        starts_with("spane"), # 1 was never
        between_estimate, # those estimates also had 1: much more than typically
        within_estimate,
        relatedness_3,
        enjoyment_2,
        enjoyment_3
      ),
      ~ 8 - .x
    )
  )

# Create indicators for straightliners
survey <- survey %>% 
  mutate(across(contains("straightliner"), ~.x==1))
```

## Create Survey Scales

```{r survey-scales}
#| label: survey-scales

# General SPANE scale (well-being)
survey <- survey %>% 
  mutate(
    # Use purrr to create positive and negative scale means
    spane_positive = map_dbl(1:nrow(.), ~mean(c(
      survey$spane_1[.x], 
      survey$spane_3[.x], 
      survey$spane_5[.x], 
      survey$spane_7[.x], 
      survey$spane_10[.x], 
      survey$spane_12[.x]
    ), na.rm = TRUE)),
    
    spane_negative = map_dbl(1:nrow(.), ~mean(c(
      survey$spane_2[.x], 
      survey$spane_4[.x], 
      survey$spane_6[.x], 
      survey$spane_8[.x], 
      survey$spane_9[.x], 
      survey$spane_11[.x]
    ), na.rm = TRUE)),
    
    spane_balance = spane_positive - spane_negative
  )

# Motivation scales
survey <- survey %>% 
  mutate(
    # Use purrr's map functions to compute means for different motivation scales
    autonomy = pmap_dbl(select(., starts_with("autonomy")), 
                       ~mean(c(...), na.rm = TRUE)),
    competence = pmap_dbl(select(., starts_with("competence")), 
                         ~mean(c(...), na.rm = TRUE)),
    relatedness = pmap_dbl(select(., starts_with("relatedness")), 
                          ~mean(c(...), na.rm = TRUE)),
    enjoyment = pmap_dbl(select(., starts_with("enjoyment")), 
                        ~mean(c(...), na.rm = TRUE)),
    extrinsic = pmap_dbl(select(., starts_with("extrinsic")), 
                        ~mean(c(...), na.rm = TRUE))
  )

# SPANE because of playing game
survey <- survey %>% 
  mutate(
    # Use purrr for game-related SPANE scales
    spane_game_positive = map_dbl(1:nrow(.), ~mean(c(
      survey$spane_game_1[.x], 
      survey$spane_game_3[.x], 
      survey$spane_game_5[.x], 
      survey$spane_game_7[.x], 
      survey$spane_game_10[.x], 
      survey$spane_game_12[.x]
    ), na.rm = TRUE)),
    
    spane_game_negative = map_dbl(1:nrow(.), ~mean(c(
      survey$spane_game_2[.x], 
      survey$spane_game_4[.x], 
      survey$spane_game_6[.x], 
      survey$spane_game_8[.x], 
      survey$spane_game_9[.x], 
      survey$spane_game_11[.x]
    ), na.rm = TRUE)),
    
    spane_game_balance = spane_game_positive - spane_game_negative
  )

# Hours of estimated play
survey <- survey %>%
  mutate(active_play_minutes = active_play_minutes / 60) %>% 
  mutate(active_play = rowSums(select(., starts_with("active_play")), na.rm = TRUE))

# Save processed survey data
write_csv(survey, here("data/processed/survey_processed.csv"))
```

# Telemetry Data Processing

## Load Game Time Data

```{r game-time-load}
#| label: game-time-load

game_time_raw <- read_both_waves("OII_PVZ_Game_Time") %>%
  # Only include survey participants 
  filter(player_id %in% survey$player_id)
```

## Process Game Time Data

```{r game-time-process}
#| label: game-time-process

# Handle missing start times
game_time <- game_time_raw %>% 
  mutate(
    # If start time is missing, set to end time (will result in 0 duration)
    game_start_time = coalesce(game_start_time, game_end_time)
  ) %>%
  # Collapse rows by session
  group_by(player_id, game_session) %>% 
  mutate(
    game_start_time = min(game_start_time),
    game_end_time = max(game_end_time),
  ) %>% 
  # Take the first of each (identical) person-session combo
  slice(1) %>% 
  ungroup()

# Filter to sessions within 14 days before survey
game_time_2week <- game_time %>% 
  left_join(select(survey, player_id, start_date)) %>% 
  filter(game_start_time >= (start_date - days(14)))

# Save processed game time data
write_csv(game_time_2week, here("data/processed/game_time_2week.csv"))
```

## Load Authentication Data

```{r auth-load}
#| label: auth-load

authentications_raw <- read_both_waves("OII_PVZ_Authentications") %>%
  # Only include survey participants
  filter(player_id %in% survey$player_id)

# Filter to 14 days before survey
authentications_2week <- authentications_raw %>% 
  left_join(select(survey, player_id, start_date)) %>% 
  filter(date >= (as_date(start_date) - days(14)))

# Count authentications per player
authentications_players <- authentications_2week %>% 
  count(player_id, name = "authentications")

# Save processed authentication data
write_csv(authentications_players, here("data/processed/authentications_players.csv"))
```

## Load Character Performance Data

```{r character-load}
#| label: character-load

characters_raw <- read_both_waves("OII_PVZ_Character") %>%
  # Only include survey participants
  filter(player_id %in% survey$player_id)

# Add session timing info and filter to 14 days before survey
characters_2week <- characters_raw %>%
  left_join(
    game_time_2week %>% 
      select(player_id, game_session, game_start_time, game_end_time),
    by = c("player_id", "game_session")
  ) %>%
  left_join(select(survey, player_id, start_date)) %>% 
  filter(game_start_time >= (start_date - days(14)))

# Save for further processing
write_csv(characters_2week, here("data/processed/characters_2week.csv"))
```

# Game Mode Classification

```{r game-mode-classification}
#| label: game-mode-classification

# Create game mode classification variables
characters_2week <- characters_2week %>%
  mutate(
    # Social hub sessions (no combat)
    social_hub = case_when(
      game_level_key == "dsub_social_terrain_summer" & is.na(score) ~ 1,
      TRUE ~ 0
    ),
    
    # PvE sessions
    pve_session = case_when(
      game_level_key %in% c(
        "dsub_region_cheesemines", 
        "dsub_region_dreadwood", 
        "dsub_region_towncenter"
      ) ~ 1,
      TRUE ~ 0
    ),
    
    # PvP sessions
    pvp_session = case_when(
      game_level_key %in% c(
        "dsub_rush_dreadwood", "dsub_rush_towncenter", 
        "dsub_tvm_cheesemines", "dsub_tvm_zrefinery", 
        "dsub_tvm_suburbs", "dsub_tvm_dreadwood", 
        "dsub_rush_farm", "dsub_rush_cheesemines", 
        "dsub_rush_shoreline", "dsub_rush_resort", 
        "dsub_4v4_battlearena_v2", "dsub_4v4_superarena"
      ) ~ 1,
      # Giddy Park with combat data is also PvP
      game_level_key == "dsub_social_terrain_summer" & !is.na(score) ~ 1,
      TRUE ~ 0
    )
  )

# Remove sessions without performance data (mainly social hub)
characters_2week_nona <- characters_2week %>% 
  drop_na(total_kill_count:shots_hit)

# Save processed data
write_csv(characters_2week, here("data/processed/characters_with_modes.csv"))
write_csv(characters_2week_nona, here("data/processed/characters_with_modes_nona.csv"))
```

# Game Time Calculations

```{r game-time-calculations}
#| label: game-time-calculations

# Define a function to calculate time per player for a specific game mode
calculate_game_time <- function(data, mode_filter, time_prefix) {
  data %>%
    filter({{ mode_filter }} == 1) %>%
    mutate(duration = game_end_time - game_start_time) %>%
    mutate(seconds = as.numeric(duration)) %>%
    group_by(player_id) %>%
    summarise(seconds = sum(seconds)) %>%
    mutate(hours = seconds / 60 / 60) %>%
    rename_with(~ paste0(time_prefix, "_", .), .cols = c("seconds", "hours")) %>%
    ungroup()
}

# Calculate total game time per player
game_time_players_total <- characters_2week %>% 
  mutate(duration = game_end_time - game_start_time) %>% 
  mutate(seconds = as.numeric(duration)) %>% 
  group_by(player_id) %>% 
  summarise(seconds = sum(seconds)) %>%
  mutate(hours = seconds / 60 / 60) %>% 
  ungroup()

# Calculate game time for different modes using our function
game_time_players_pvp <- calculate_game_time(characters_2week, pvp_session, "pvp")
game_time_players_pve <- calculate_game_time(characters_2week, pve_session, "pve")
game_time_players_social <- calculate_game_time(characters_2week, social_hub, "social")

# Combine game time data with purrr's reduce
game_time_players <- list(
  game_time_players_total, 
  game_time_players_social, 
  game_time_players_pve, 
  game_time_players_pvp
) %>%
  reduce(left_join, by = "player_id")

# Replace NA values with 0 (no time spent in that mode)
game_time_players[is.na(game_time_players)] <- 0

# Save processed game time data
write_csv(game_time_players, here("data/processed/game_time_players.csv"))
```

# Performance Metrics Calculation

```{r performance-setup}
#| label: performance-setup
#| cache: true
#| dependson: c("game-mode-classification", "game-time-calculations")

# Create an empty list to store intermediate results
metrics_list <- list()
```

```{r overall-performance}
#| label: overall-performance
#| cache: true
#| dependson: c("performance-setup")

# Calculate overall performance metrics per player
characters_players <- characters_2week %>%
  group_by(player_id) %>% 
  summarise(across(total_kill_count:shots_hit, sum, na.rm = TRUE)) %>% 
  ungroup()

# Store in our list
metrics_list$characters_players <- characters_players

# Force garbage collection
gc()
```

```{r game-sessions}
#| label: game-sessions
#| cache: true
#| dependson: c("performance-setup")

# Calculate game sessions per player
game_sessions_total <- characters_2week %>% 
  count(player_id) %>%
  rename(n_total_game_sessions = n)

# Store in our list
metrics_list$game_sessions_total <- game_sessions_total

# Force garbage collection
gc()
```

```{r sessions-by-mode}
#| label: sessions-by-mode
#| cache: true
#| dependson: c("performance-setup")

# Calculate sessions by game mode
game_sessions <- characters_2week %>% 
  group_by(player_id) %>% 
  summarise(across(social_hub:pvp_session, sum, na.rm = TRUE)) %>% 
  ungroup()

# Store in our list
metrics_list$game_sessions <- game_sessions

# Force garbage collection
gc()
```

```{r pvp-performance}
#| label: pvp-performance
#| cache: true
#| dependson: c("performance-setup")

# PvP-specific performance metrics
pvp_telemetry <- characters_2week %>%
  filter(pvp_session == 1) %>%
  group_by(player_id) %>% 
  summarise(across(total_kill_count:shots_hit, sum, na.rm = TRUE)) %>%
  rename_with(~ paste0("pvp_", .), -player_id)

# Store in our list
metrics_list$pvp_telemetry <- pvp_telemetry

# Force garbage collection
gc()
```

```{r pve-performance}
#| label: pve-performance
#| cache: true
#| dependson: c("performance-setup")

# PvE-specific performance metrics
pve_telemetry <- characters_2week %>%
  filter(pve_session == 1) %>%
  group_by(player_id) %>% 
  summarise(across(total_kill_count:shots_hit, sum, na.rm = TRUE)) %>%
  rename_with(~ paste0("pve_", .), -player_id)

# Store in our list
metrics_list$pve_telemetry <- pve_telemetry

# Force garbage collection
gc()
```

```{r combine-metrics}
#| label: combine-metrics
#| cache: true
#| dependson: c("overall-performance", "game-sessions", "sessions-by-mode", "pvp-performance", "pve-performance")

# Load additional data we need from previous processing
metrics_list$game_time_players <- game_time_players

# Combine telemetry data using purrr's reduce function
telemetry_players <- list(
  metrics_list$characters_players,
  metrics_list$game_time_players,
  metrics_list$game_sessions,
  metrics_list$game_sessions_total,
  metrics_list$pvp_telemetry,
  metrics_list$pve_telemetry
) %>%
  reduce(left_join, by = "player_id")

# Replace NA values with 0
telemetry_players <- telemetry_players %>%
  mutate(across(everything(), ~replace_na(.x, 0)))

# Clear intermediate objects and list to free memory
rm(metrics_list)
gc()

# Save processed telemetry data
write_csv(telemetry_players, here("data/processed/telemetry_players.csv"))
```

# Derived Performance Metrics

```{r derived-metrics}
#| label: derived-metrics

# Define a function to calculate performance ratios
calculate_ratio <- function(numerator, denominator) {
  numerator / denominator
}

# Calculate derived performance metrics
telemetry_players <- telemetry_players %>%
  # Handle divisions by zero for K/D ratio
  mutate(
    across(
      c(total_death_count, pvp_total_death_count, pve_total_death_count),
      ~ ifelse(.x == 0, 1, .x)
    )
  ) %>%
  # Calculate all ratio metrics with purrr
  mutate(
    # K/D ratios
    total_kd = calculate_ratio(total_kill_count, total_death_count),
    pve_kd = calculate_ratio(pve_total_kill_count, pve_total_death_count),
    pvp_kd = calculate_ratio(pvp_total_kill_count, pvp_total_death_count),
    
    # Hit accuracy ratios
    total_hit_accuracy = calculate_ratio(shots_hit, shots_fired),
    pve_hit_accuracy = calculate_ratio(pve_shots_hit, pve_shots_fired),
    pvp_hit_accuracy = calculate_ratio(pvp_shots_hit, pvp_shots_fired),
    
    # Critical hit ratios
    total_crit_ratio = calculate_ratio(critical_hit_count, shots_hit),
    pve_crit_ratio = calculate_ratio(pve_critical_hit_count, pve_shots_hit),
    pvp_crit_ratio = calculate_ratio(pvp_critical_hit_count, pvp_shots_hit),
    
    # Damage per game
    total_average_damage = calculate_ratio(damage_dealt, n_total_game_sessions),
    pve_average_damage = calculate_ratio(pve_damage_dealt, pve_session),
    pvp_average_damage = calculate_ratio(pvp_damage_dealt, pvp_session),
    
    # Damage per second
    total_dps = calculate_ratio(damage_dealt, seconds),
    pve_dps = calculate_ratio(pve_damage_dealt, pve_seconds),
    pvp_dps = calculate_ratio(pvp_damage_dealt, pvp_seconds),
    
    # Score per game
    total_average_score = calculate_ratio(score, n_total_game_sessions),
    pve_average_score = calculate_ratio(pve_score, pve_session),
    pvp_average_score = calculate_ratio(pvp_score, pvp_session),
    
    # Score per second
    total_sps = calculate_ratio(score, seconds),
    pve_sps = calculate_ratio(pve_score, pve_seconds),
    pvp_sps = calculate_ratio(pvp_score, pvp_seconds)
  )

# Replace NaN values with 0
telemetry_players <- telemetry_players %>%
  mutate(across(everything(), ~ifelse(is.nan(.x), 0, .x)))

# Save processed metrics
write_csv(telemetry_players, here("data/processed/telemetry_with_metrics.csv"))
```

# Participants and Procedure

# Combine Survey and Telemetry Data

```{r merge-data}
#| label: merge-data

# Merge survey with telemetry data
pvz_df <- survey %>% 
  right_join(telemetry_players, by = "player_id")

# Check for participants without valid performance metrics
# These would be participants with zeros or NAs for all key performance indicators
pvz_df <- pvz_df %>%
  # Create a flag for participants with no meaningful telemetry data
  mutate(
    has_telemetry = !(
      # Check if key performance metrics are all zeros or NAs
      (total_kill_count == 0 | is.na(total_kill_count)) &
      (total_death_count == 0 | is.na(total_death_count)) &
      (damage_dealt == 0 | is.na(damage_dealt)) &
      (shots_fired == 0 | is.na(shots_fired)) &
      (n_total_game_sessions == 0 | is.na(n_total_game_sessions))
    )
  )

# Print summary of participants with and without telemetry
cat("Participants without valid telemetry:", sum(!pvz_df$has_telemetry, na.rm = TRUE), "\n")
cat("Participants with valid telemetry:", sum(pvz_df$has_telemetry, na.rm = TRUE), "\n")

# Remove participants without valid telemetry data
pvz_df_clean <- pvz_df %>%
  filter(has_telemetry == TRUE)

# Calculate ratio metrics
pvz_df_clean <- pvz_df_clean %>%
  mutate(
    ratio_social_hours = social_hours / hours,
    ratio_pve_hours = pve_hours / hours,
    ratio_pvp_hours = pvp_hours / hours
  )

# Remove participants with very low playtime (< 2 minutes)
pvz_df_clean <- pvz_df_clean %>%
  filter(hours > 0.03)

# Print final dataset dimensions
cat("Final dataset dimensions:", dim(pvz_df_clean)[1], "participants x", dim(pvz_df_clean)[2], "variables\n")

# Save final dataset
write_csv(pvz_df_clean, here("data/processed/pvz_complete.csv"))
```

```{r survey-demographics-calculation}
#| label: survey-demographics-calculation
#| include: false

# Calculate overall survey demographics
survey_demographics <- survey %>%
  summarize(
    total_participants = n(),
    avg_age = round(mean(age, na.rm = TRUE), 1),
    sd_age = round(sd(age, na.rm = TRUE), 1)
  )

# Calculate gender distribution in original survey
gender_survey <- survey %>%
  count(gender) %>%
  mutate(percentage = n / sum(n) * 100)

# Count participants with each gender
male_count <- gender_survey %>% filter(gender == "Male") %>% pull(n)
female_count <- gender_survey %>% filter(gender == "Female") %>% pull(n)
other_count <- gender_survey %>% filter(gender == "Other") %>% pull(n)
undisclosed_count <- gender_survey %>% filter(gender == "Prefer not to say") %>% pull(n)
```

As detailed in Johannes et al. (2020), participant recruitment happened in two waves, first in early August 2020 and second in late September 2020. The response rate was \~ 0.21%. In total `r survey_demographics$total_participants` PvZ players participated in the well-being survey (`r male_count` male, `r female_count` female, `r other_count` other, and `r undisclosed_count` undisclosed) with an average age of `r survey_demographics$avg_age` (SD = `r survey_demographics$sd_age`). However, since we are interested in modelling playing styles from in-game telemetry only the participants with valid telemetry data were included in this study. After preprocessing, our final sample consists of `r nrow(pvz_df_clean)` participants with valid performance metrics.

```{r final-sample-demographics}
#| label: final-sample-demographics

# Display demographics of final sample
cat("Demographics of final sample:\n")
cat("- Sample size:", nrow(pvz_df_clean), "\n")
cat("- Average age:", round(mean(pvz_df_clean$age, na.rm = TRUE), 1), "(SD =", round(sd(pvz_df_clean$age, na.rm = TRUE), 1), ")\n")

# Gender distribution
gender_dist <- pvz_df_clean %>%
  count(gender) %>%
  mutate(percentage = n / sum(n) * 100)

kable(gender_dist, caption = "Gender distribution in final sample", digits = 2)

# Print summary of final dataset
glimpse(pvz_df_clean)
```

# Next Steps

The preprocessing steps are now complete. The processed data has been saved and is ready for analysis in the `2_analysis.qmd` file. The main dataset for analysis is `pvz_df_clean`.