---
title: "Player Playstyle Analysis"
author: "Thomas Hakman"
format: 
  html:
    toc: true
    toc-depth: 3
    code-fold: true
    theme: flatly
    fig-width: 8
    fig-height: 6
execute:
  echo: true
  message: false
  warning: false
  cache: true
editor: visual
---

```{r setup}
#| label: setup
#| include: false

# Set global options
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  cache = FALSE,
  fig.path = "figures/"
)
```

# Introduction

This document contains the analysis of player telemetry and survey data from Plants vs. Zombies: Battle for Neighborville. It focuses on analyzing player playstyles, game performance metrics, and their relationship with well-being outcomes.

test test

# Load Required Packages and Data

```{r install-packages, include=FALSE}
#| label: install-packages

# First ensure pacman is installed if not already
if (!requireNamespace("pacman", quietly = TRUE)) {
  install.packages("pacman")
}

# Use pacman to load all required packages (will install any missing packages)
pacman::p_load(
  tidyverse,  # For data manipulation and visualization
  here,       # For project-relative paths
  lubridate,  # For date/time manipulation
  scales,     # For scale transformations
  janitor,    # For data cleaning
  magrittr,   # For additional pipe operators
  psych,      # For psychological statistics
  corrplot,   # For correlation plots
  ggpubr,     # For publication-ready plots
  factoextra, # For cluster analysis visualization
  cluster,    # For clustering algorithms
  fmsb,       # For radar charts
  knitr,      # For dynamic report generation
  tibble,     # For improved data frames
  kableExtra  # For enhanced table formatting
)
```

```{r helper-functions}
#| label: helper-functions

# Function to scale values to 0-1 range for radar charts
scale_to_01 <- function(x, min_val = min(x, na.rm = TRUE), max_val = max(x, na.rm = TRUE)) {
  if (max_val == min_val) return(rep(0.5, length(x)))
  (x - min_val) / (max_val - min_val)
}

```

```{r statistical-functions}
#| label: statistical-functions

# Function to run ANOVA and extract results
run_anova <- function(dv) {
  formula <- as.formula(paste(dv, "~ player_type"))
  model <- aov(formula, data = pvz_data)
  result <- broom::tidy(model)
  result$eta_squared <- sjstats::eta_sq(model)[["eta.sq"]]
  result$variable <- dv
  return(result)
}

# Function to compute Hopkins statistic for cluster tendency
compute_hopkins <- function(data, n = nrow(data)-1) {
  # Standardize data for clustering
  data_scaled <- scale(data)
  
  # Compute Hopkins statistic
  hopkins_result <- clustertend::hopkins(data_scaled, n = n)
  
  return(hopkins_result)
}
```

```{r load-processed-data}
#| label: load-processed-data

pvz_data <- read_csv(here("data/processed/pvz_complete.csv"))

# Create a copy of pvz_data for analysis to maintain consistency
pvz_df_analysis <- pvz_data

glimpse(pvz_df_analysis)
```

```{r select-columns}
#| label: select-columns

pvz_df_analysis <- pvz_df_analysis %>%
  select(
    # Demographic variables
    player_id, country, age, gender, played_with_others,
    
    # Well-being variables
    spane_balance, spane_positive, spane_negative,
    autonomy, competence, relatedness, enjoyment, extrinsic,
    spane_game_balance, spane_game_positive, spane_game_negative,
    
    # Total playtime
    hours,
    
    # PVP metrics
    ratio_pvp_hours,
    pvp_kd,
    pvp_hit_accuracy,
    pvp_dps,
    pvp_average_damage,
    pvp_average_score,
    
    # PVE metrics
    ratio_pve_hours,
    pve_kd,
    pve_hit_accuracy,
    pve_dps,
    pve_average_damage,
    pve_average_score
  )
```

# Exploratory Data Analysis

## plot session timeline for a player

```{r session-timeline-helper}
#| label: session-timeline-helper
#| include: false

# Function to plot session timeline for a player
plot_player_sessions <- function(player_data, player_id = NULL, player_label = "Player") {
  # If a player_id is provided, filter data for that player
  if (!is.null(player_id)) {
    player_data <- player_data %>%
      filter(player_id == !!player_id)
  }
  
  # Convert timestamps to POSIXct if they're not already
  if (!inherits(player_data$game_start_time, "POSIXct")) {
    player_data$game_start_time <- as.POSIXct(player_data$game_start_time, format = "%Y-%m-%d %H:%M:%S")
    player_data$game_end_time <- as.POSIXct(player_data$game_end_time, format = "%Y-%m-%d %H:%M:%S")
  }
  
  # Extract day of week, date, and time component
  player_data$day_of_week <- weekdays(player_data$game_start_time)
  player_data$date <- as.Date(player_data$game_start_time)
  player_data$start_time_only <- as.POSIXct(format(player_data$game_start_time, format = "%H:%M:%S"), format = "%H:%M:%S")
  player_data$end_time_only <- as.POSIXct(format(player_data$game_end_time, format = "%H:%M:%S"), format = "%H:%M:%S")
  
  # Identify sessions that cross midnight (where end_time is earlier than start_time)
  player_data$crosses_midnight <- player_data$end_time_only < player_data$start_time_only
  
  # Create a new dataframe that will include split sessions for those crossing midnight
  player_data_expanded <- player_data %>%
    filter(!crosses_midnight) # Keep all sessions that don't cross midnight
  
  # Handle sessions that cross midnight by splitting them
  midnight_sessions <- player_data %>%
    filter(crosses_midnight)
  
  if (nrow(midnight_sessions) > 0) {
    # For each session crossing midnight, create two rows:
    # 1. From start time to 23:59:59
    # 2. From 00:00:00 to end time
    midnight <- as.POSIXct(strptime("23:59:59", format = "%H:%M:%S"), format = "%H:%M:%S")
    midnight_start <- as.POSIXct(strptime("00:00:00", format = "%H:%M:%S"), format = "%H:%M:%S")
    
    # First part: from start time to midnight
    midnight_part1 <- midnight_sessions
    midnight_part1$end_time_only <- midnight
    
    # Second part: from midnight start to end time
    midnight_part2 <- midnight_sessions
    midnight_part2$start_time_only <- midnight_start
    # The second part is on the next day of the week and the next date
    midnight_part2$day_of_week <- factor(
      weekdays(midnight_part2$game_start_time + 86400), # Add one day (86400 seconds)
      levels = c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday")
    )
    midnight_part2$date <- midnight_part2$date + 1
    
    # Add both parts to the expanded dataset
    player_data_expanded <- bind_rows(player_data_expanded, midnight_part1, midnight_part2)
  }
  
  # Recode match_type based on game_level_key 
  player_data_expanded$match_type <- "0"  # Initialize with 0 as in the original code
  
  # Social matches
  player_data_expanded$match_type[player_data_expanded$game_level_key == "dsub_social_terrain_beach"] <- "social match"
  player_data_expanded$match_type[player_data_expanded$game_level_key == "dsub_social_terrain_summer"] <- "social match"
  player_data_expanded$match_type[player_data_expanded$game_level_key == "dsub_socialspace"] <- "social match"
  
  # PVE matches
  player_data_expanded$match_type[player_data_expanded$game_level_key == "dsub_region_cheesemines"] <- "pve match"
  player_data_expanded$match_type[player_data_expanded$game_level_key == "dsub_region_dreadwood"] <- "pve match"
  player_data_expanded$match_type[player_data_expanded$game_level_key == "dsub_region_towncenter"] <- "pve match" 
  
  # PVP matches (Rush)
  player_data_expanded$match_type[player_data_expanded$game_level_key == "dsub_rush_dreadwood"] <- "pvp match (rush)"
  player_data_expanded$match_type[player_data_expanded$game_level_key == "dsub_rush_towncenter"] <- "pvp match (rush)"
  player_data_expanded$match_type[player_data_expanded$game_level_key == "dsub_rush_farm"] <- "pvp match (rush)"
  player_data_expanded$match_type[player_data_expanded$game_level_key == "dsub_rush_cheesemines"] <- "pvp match (rush)"
  player_data_expanded$match_type[player_data_expanded$game_level_key == "dsub_rush_shoreline"] <- "pvp match (rush)"
  player_data_expanded$match_type[player_data_expanded$game_level_key == "dsub_rush_resort"] <- "pvp match (rush)"
  
  # PVP matches (Vanquish)
  player_data_expanded$match_type[player_data_expanded$game_level_key == "dsub_tvm_cheesemines"] <- "pvp match (vanquish)"
  player_data_expanded$match_type[player_data_expanded$game_level_key == "dsub_tvm_zrefinery"] <- "pvp match (vanquish)"
  player_data_expanded$match_type[player_data_expanded$game_level_key == "dsub_tvm_suburbs"] <- "pvp match (vanquish)"
  player_data_expanded$match_type[player_data_expanded$game_level_key == "dsub_tvm_dreadwood"] <- "pvp match (vanquish)"
  
  # PVP matches (Arena)
  player_data_expanded$match_type[player_data_expanded$game_level_key == "dsub_4v4_battlearena_v2"] <- "pvp match (arena)"
  player_data_expanded$match_type[player_data_expanded$game_level_key == "dsub_4v4_superarena"] <- "pvp match (arena)"
  
  # Podium celebration
  player_data_expanded$match_type[player_data_expanded$game_level_key == "dsub_podiumcelebration"] <- "podium celebration"
  
  # Ensure day of week is a factor with ordered levels
  player_data_expanded$day_of_week <- factor(
    player_data_expanded$day_of_week,
    levels = c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday")
  )
  
  # Create a position adjustment for stacking sessions by different dates for the same day of week
  player_data_expanded <- player_data_expanded %>%
    group_by(day_of_week) %>%
    mutate(
      # Get unique dates for this day of week and assign a position index to each date
      unique_dates = list(unique(date)),
      date_index = match(date, unique(date)),
      # Scale the vertical offset to be small but noticeable
      y_offset = (date_index - 1) * 0.2,  # Creates a small vertical offset for different dates
      # Create adjusted y-position that includes the offset
      y_position = as.numeric(day_of_week) + y_offset
    ) %>%
    ungroup()
  
  # Sort by day of week, date, and start time
  player_data_expanded <- player_data_expanded %>%
    arrange(day_of_week, date, game_start_time, game_end_time)
  
  # Add a session ID within each day to distinguish multiple sessions on same day
  player_data_expanded <- player_data_expanded %>%
    group_by(day_of_week, date) %>%
    mutate(session_id_within_day = row_number()) %>%
    ungroup()
  
  # Convert match_type to factor with proper levels for consistent ordering
  player_data_expanded$match_type <- factor(
    player_data_expanded$match_type,
    levels = c("social match", "pve match", "pvp match (rush)", "pvp match (vanquish)", "pvp match (arena)", "podium celebration", "0"),
    labels = c("Social Match", "PvE Match", "PvP (Rush)", "PvP (Vanquish)", "PvP (Arena)", "Podium Celebration", "Other")
  )
  
  # Create the plot with enhanced styling
  plot <- player_data_expanded %>%
    ggplot2::ggplot(aes(start_time_only, y_position, col = match_type, 
                        group = interaction(day_of_week, date, session_id_within_day))) +
    # Use improved color palette with better contrast
    ggplot2::scale_color_manual(
      name = "Match Type", 
      values = c(
        "Social Match" = "#00BFC4",       # Teal
        "PvE Match" = "#7CAE00",          # Green
        "PvP (Rush)" = "#F8766D",         # Red
        "PvP (Vanquish)" = "#FF9E1C",     # Orange
        "PvP (Arena)" = "#FFD700",        # Gold
        "Podium Celebration" = "#C77CFF", # Purple
        "Other" = "#A0A0A0"               # Gray
      )
    ) +
    # Thicker segments with rounded ends for better visibility
    ggplot2::geom_segment(
      aes(xend = end_time_only, yend = y_position),
      linewidth = 2,
      lineend = "round"
    ) +
    # Add points at start/end for better visibility of short sessions
    ggplot2::geom_point(size = 1.5, alpha = 0.6) +
    ggplot2::geom_point(aes(x = end_time_only), size = 1.5, alpha = 0.6) +
    # Better time axis formatting - using fixed limits for consistency
    ggplot2::scale_x_datetime(
      labels = scales::date_format("%H:%M"),
      breaks = scales::date_breaks("2 hours"),
      limits = c(
        as.POSIXct(strptime("00:00", format = "%H:%M")),
        as.POSIXct(strptime("23:59", format = "%H:%M"))
      ),
      expand = c(0.01, 0.01),
      date_minor_breaks = "1 hour"
    ) +
    # Use the original day of week as the y-axis labels
    ggplot2::scale_y_continuous(
      breaks = 1:7,
      labels = levels(player_data_expanded$day_of_week),
      limits = c(0.5, 7.5)
    ) +
    # Empty labs - axis labels will be added to the combined figure instead
    ggplot2::labs(
      x = NULL, 
      y = NULL
    ) +
    # Improved theme with better typography and layout
    ggplot2::theme_minimal(base_size = 11, base_family = "sans") +
    ggplot2::theme(
      legend.position = "none", # Move legend to the combined figure
      plot.margin = margin(30, 10, 30, 10), # Increased margins for better separation
      
      axis.text.y = element_text(face = "bold"),
      axis.title = element_text(face = "bold"),
      
      panel.grid.major.x = element_line(color = "grey90"),
      panel.grid.minor.x = element_line(color = "grey95"),
      panel.grid.major.y = element_line(color = "grey90"),
      panel.grid.minor.y = element_blank(),
      
      panel.background = element_rect(fill = "white", color = NA),
      plot.background = element_rect(fill = "white", color = NA)
    )
  
  return(plot)
}
```

```{r session-timeline-data-prep}
#| label: session-timeline-data-prep
#| include: false
#| cache: true

# Load game session data if not already included
game_sessions <- read_csv(here("data/processed/game_time_2week.csv"))

# Make sure timestamps are in the correct format
game_sessions$game_start_time <- as.POSIXct(game_sessions$game_start_time, format = "%Y-%m-%d %H:%M:%S")
game_sessions$game_end_time <- as.POSIXct(game_sessions$game_end_time, format = "%Y-%m-%d %H:%M:%S")

# First, get a list of player IDs in the data
player_ids <- unique(game_sessions$player_id)

# Choose specific indices
# If those indices don't exist in our data, we'll use the first three players
player_indices <- c(
  min(401, length(player_ids)),  # Equivalent to slice(401) 
  min(6, length(player_ids)),    # Equivalent to slice(6)
  min(3, length(player_ids))     # Equivalent to slice(3)
)

selected_player_ids <- player_ids[player_indices]
```

```{r create-individual-plots}
#| label: create-individual-plots
#| include: false
#| cache: true
#| dependson: session-timeline-data-prep

# Create individual plots for each player with standardized x-axis
plot1 <- plot_player_sessions(
  game_sessions %>% filter(player_id == selected_player_ids[1])
)

plot2 <- plot_player_sessions(
  game_sessions %>% filter(player_id == selected_player_ids[2])
)

# For the third plot, use the same time range for consistency
plot3 <- plot_player_sessions(
  game_sessions %>% filter(player_id == selected_player_ids[3])
)
```

```{r combine-session-plots}
#| label: combine-session-plots
#| fig.width: 12
#| fig.height: 16
#| dependson: create-individual-plots

# Combine plots into a single figure with enhanced styling and proper spacing
figure <- ggarrange(
  plot2, plot3, plot1,
  labels = c("Player A", "Player B", "Player C"),
  font.label = list(size = 14, face = "bold", color = "black"),
  ncol = 1, 
  nrow = 3, 
  common.legend = TRUE, 
  legend = "bottom",
  heights = c(1, 1, 1),
  align = "v",
  hjust = -0.1
)

# Add shared x and y labels to the combined figure
fig_sessiontime <- annotate_figure(figure,
                        bottom = text_grob("Time of Day (hours)", 
                                          face = "bold", size = 12),
                        left = text_grob("Day of Week", 
                                         face = "bold", size = 12, rot = 90))

# Display the combined figure
fig_sessiontime
```

# Descriptive Statistics

```{r apa-descriptive-stats}
#| label: tbl-descriptive-stats
#| tbl-cap: "Descriptive Statistics for Key Variables"
#| fig.width: 10
#| fig.height: 6
#| tbl-cap-location: top
#| echo: false
#| warning: false
#| message: false

# Create a simple APA-style descriptive statistics table
# First, check variable types to handle them appropriately
variable_types <- sapply(pvz_df_analysis, class)

# Handle numeric variables
numeric_vars <- pvz_df_analysis %>%
  select(where(is.numeric)) %>%
  select(
    # Demographics
    age,
    
    # Playtime 
    hours, ratio_pvp_hours, ratio_pve_hours,
    
    # Performance metrics
    pvp_kd, pve_kd, 
    pvp_hit_accuracy, pve_hit_accuracy,
    pvp_dps, pve_dps,
    pvp_average_damage, pve_average_damage,
    pvp_average_score, pve_average_score,
    
    # Well-being and motivation
    spane_balance, spane_game_balance, 
    autonomy, competence, relatedness,
    enjoyment, extrinsic
  )

# Calculate descriptive statistics for numeric variables
numeric_stats <- numeric_vars %>%
  summarise(across(everything(), 
                  list(
                    n = ~sum(!is.na(.)),
                    M = ~mean(., na.rm = TRUE),
                    SD = ~sd(., na.rm = TRUE),
                    Min = ~min(., na.rm = TRUE),
                    Max = ~max(., na.rm = TRUE)
                  ))) %>%
  pivot_longer(everything(), 
               names_to = c("variable", "stat"), 
               names_pattern = "(.*)_(.*)",
               values_to = "value") %>%
  pivot_wider(names_from = stat, values_from = value)

# For categorical variables like gender, just count frequencies for the footnote
if ("gender" %in% names(pvz_df_analysis)) {
  gender_counts <- pvz_df_analysis %>%
    count(gender) %>%
    mutate(pct = n / sum(n) * 100) %>%
    mutate(pct = round(pct, 1)) %>%
    mutate(frequency = paste0(n, " (", pct, "%)"))
  
  # Note: We'll add a footnote to explain gender distribution separately
  gender_note <- paste("Gender distribution: ", 
                     paste(paste0(gender_counts$gender, ": ", gender_counts$frequency), 
                           collapse = "; "))
} else {
  gender_note <- NULL
}

# Format the data for the table
descriptive_stats <- numeric_stats %>%
  # Format variable names for better readability
  mutate(variable = case_when(
    variable == "age" ~ "Age",
    variable == "hours" ~ "Total Hours Played",
    variable == "ratio_pvp_hours" ~ "Proportion of PvP Playtime",
    variable == "ratio_pve_hours" ~ "Proportion of PvE Playtime",
    variable == "pvp_kd" ~ "PvP Kill/Death Ratio",
    variable == "pve_kd" ~ "PvE Kill/Death Ratio",
    variable == "pvp_hit_accuracy" ~ "PvP Hit Accuracy",
    variable == "pve_hit_accuracy" ~ "PvE Hit Accuracy",
    variable == "pvp_dps" ~ "PvP Damage per Second",
    variable == "pve_dps" ~ "PvE Damage per Second",
    variable == "pvp_average_damage" ~ "PvP Average Damage per Game",
    variable == "pve_average_damage" ~ "PvE Average Damage per Game",
    variable == "pvp_average_score" ~ "PvP Average Score",
    variable == "pve_average_score" ~ "PvE Average Score",
    variable == "spane_balance" ~ "Overall Well-being (SPANE)",
    variable == "spane_game_balance" ~ "Game-related Well-being",
    variable == "autonomy" ~ "Autonomy",
    variable == "competence" ~ "Competence",
    variable == "relatedness" ~ "Relatedness",
    variable == "enjoyment" ~ "Intrinsic Enjoyment",
    variable == "extrinsic" ~ "Extrinsic Motivation",
    TRUE ~ variable
  )) %>%
  # Create category variable for grouping
  mutate(category = case_when(
    variable == "Age" ~ "Demographics",
    variable %in% c("Total Hours Played", "Proportion of PvP Playtime", "Proportion of PvE Playtime") ~ "Playtime",
    variable %in% c("PvP Kill/Death Ratio", "PvE Kill/Death Ratio", 
                   "PvP Hit Accuracy", "PvE Hit Accuracy",
                   "PvP Damage per Second", "PvE Damage per Second",
                   "PvP Average Damage per Game", "PvE Average Damage per Game",
                   "PvP Average Score", "PvE Average Score") ~ "Performance",
    variable %in% c("Overall Well-being (SPANE)", "Game-related Well-being") ~ "Well-being",
    variable %in% c("Autonomy", "Competence", "Relatedness", "Intrinsic Enjoyment", "Extrinsic Motivation") ~ "Motivation",
    TRUE ~ "Other"
  )) %>%
  # Round numeric values
  mutate(
    M = round(M, 2),
    SD = round(SD, 2),
    Min = round(Min, 2),
    Max = round(Max, 2)
)

# Create custom ordering for variables to ensure autonomy, competence, relatedness are listed in order
variable_order <- c(
  # Demographics
  "Age",
  
  # Playtime
  "Total Hours Played", "Proportion of PvP Playtime", "Proportion of PvE Playtime",
  
  # Performance
  "PvP Kill/Death Ratio", "PvE Kill/Death Ratio", 
  "PvP Hit Accuracy", "PvE Hit Accuracy",
  "PvP Damage per Second", "PvE Damage per Second",
  "PvP Average Damage per Game", "PvE Average Damage per Game",
  "PvP Average Score", "PvE Average Score",
  
  # Well-being
  "Overall Well-being (SPANE)", "Game-related Well-being",
  
  # Motivation (in specified order)
  "Autonomy", "Competence", "Relatedness", "Intrinsic Enjoyment", "Extrinsic Motivation"
)

# Apply the custom ordering
descriptive_stats <- descriptive_stats %>%
  mutate(variable = factor(variable, levels = variable_order)) %>%
  arrange(variable)

# Create the APA-styled table with kable
apa_table <- descriptive_stats %>%
  select(variable, n, M, SD, Min, Max) %>%
  kable(
    # Caption is handled by chunk options now
    col.names = c("Variable", "*N*", "*M*", "*SD*", "Min", "Max"),
    align = c("l", "c", "c", "c", "c", "c"),
    booktabs = TRUE,
    format = "html",
    digits = 2
  ) %>%
  kable_styling(
    bootstrap_options = c("striped", "hover", "condensed"),
    full_width = FALSE,
    position = "left",
    font_size = 12
  ) %>%
  row_spec(0, bold = TRUE) %>%  # Bold header row per APA
  pack_rows(index = table(factor(descriptive_stats$category, 
                              levels = unique(descriptive_stats$category))), 
            hline_after = TRUE)  # Add lines after each group

# Add footnote for gender if it exists
if (!is.null(gender_note)) {
  apa_table <- apa_table %>%
    footnote(general = gender_note, general_title = "Note: ")
}

# Display the table
apa_table
```

# Player Clustering Analysis with PAM

This section identifies distinct player types through clustering analysis based on gameplay metrics. We use PAM (Partitioning Around Medoids) because it's more robust to outliers than K-means.

## Feature Selection and Preparation

```{r cluster_feature_selection}
#| label: cluster_feature_selection
#| echo: true
#| warning: false
#| message: false

# Select relevant features for clustering
player_cluster_features <- pvz_df_analysis %>% 
  select(
    # Playstyle variables (game mode preferences)
    ratio_pvp_hours, ratio_pve_hours,
    
    # PvP performance metrics
    pvp_kd, pvp_hit_accuracy, pvp_dps, pvp_average_damage, pvp_average_score,
    
    # PvE performance metrics
    pve_kd, pve_hit_accuracy, pve_dps, pve_average_damage, pve_average_score
  )

# Remove any NAs in the dataset to ensure clean clustering
player_cluster_features <- na.omit(player_cluster_features)

# Standardize the data (important for clustering)
player_data_scaled <- scale(player_cluster_features)
rownames(player_data_scaled) <- pvz_df_analysis$player_id[rownames(player_data_scaled) %in% rownames(pvz_df_analysis)]
```

## Assessing Cluster Tendency

Before performing clustering, we assess whether natural clusters exist in the data.

```{r cluster_tendency_assessment}
#| label: cluster_tendency_assessment
#| fig.width: 12
#| fig.height: 12
#| echo: true
#| warning: false
#| message: false

# Load required packages for cluster tendency assessment
if (!requireNamespace("clustertend", quietly = TRUE)) {
  install.packages("clustertend")
}
if (!requireNamespace("factoextra", quietly = TRUE)) {
  install.packages("factoextra")
}
if (!requireNamespace("gridExtra", quietly = TRUE)) {
  install.packages("gridExtra")
}
library(clustertend)
library(factoextra)
library(gridExtra)

# Create a random dataset with the same dimensions for comparison
set.seed(123) # For reproducibility
random_cluster_data <- apply(player_data_scaled, 2,
                     function(x){runif(length(x), min(x), max(x))})
random_cluster_data <- as.data.frame(random_cluster_data)

# Visual assessment using PCA visualization
pca_real_data <- prcomp(player_data_scaled)
pca_random_data <- prcomp(random_cluster_data)

# Create PCA plots for comparison
pca_plot_real <- fviz_pca_ind(pca_real_data, 
                  title = "PCA - Player Performance Data",
                  geom = "point", 
                  ggtheme = theme_minimal(),
                  legend = "none",
                  alpha.ind = 0.7) +
      theme(plot.title = element_text(face = "bold", size = 14))

pca_plot_random <- fviz_pca_ind(pca_random_data, 
                  title = "PCA - Random Data",
                  geom = "point", 
                  ggtheme = theme_minimal(),
                  legend = "none",
                  alpha.ind = 0.7) +
      theme(plot.title = element_text(face = "bold", size = 14))

# Visual assessment using distance matrices
dist_real_data <- get_dist(player_data_scaled, method = "euclidean")
dist_random_data <- get_dist(random_cluster_data, method = "euclidean")

# Create distance matrix plots
dist_plot_real <- fviz_dist(dist_real_data, 
                show_labels = FALSE,
                gradient = list(low = "#00AFBB", mid = "white", high = "#FC4E07")) +
       ggtitle("Distance Matrix - Player Data") +
       theme(plot.title = element_text(face = "bold", size = 14))

dist_plot_random <- fviz_dist(dist_random_data, 
                show_labels = FALSE,
                gradient = list(low = "#00AFBB", mid = "white", high = "#FC4E07")) +
       ggtitle("Distance Matrix - Random Data") +
       theme(plot.title = element_text(face = "bold", size = 14))

# Arrange plots in a 2x2 grid for better comparison
grid.arrange(
  pca_plot_real, pca_plot_random,
  dist_plot_real, dist_plot_random,
  ncol = 2
)

# Interpretation text
cat("## Interpretation of Clustering Tendency\n\n",
    "1. **PCA Visualization**: If the real data shows more visible grouping than random data, natural clusters likely exist.\n\n",
    "2. **Distance Matrix**: In data with cluster structure, the distance matrix shows dark blocks along the diagonal,\n",
    "   indicating groups of observations that are similar to each other.\n\n")
```

## Determining Optimal Number of Clusters

We use multiple methods to determine the appropriate number of clusters.

```{r optimal_cluster_determination}
#| label: optimal_cluster_determination
#| fig.width: 10
#| fig.height: 8
#| cache: true

# Load required packages
pacman::p_load(cluster, fpc, factoextra, gridExtra)

# PAM optimal cluster determination
# -------------------------------------
# Silhouette method for PAM
pam_sil <- function(data, max_k = 10) {
  sil_values <- numeric(max_k - 1)
  for (i in 2:max_k) {
    pam_result <- pam(data, k = i, metric = "euclidean")
    sil_values[i-1] <- pam_result$silinfo$avg.width
  }
  k_values <- 2:max_k
  return(data.frame(k = k_values, sil = sil_values))
}

pam_sil_results <- pam_sil(player_data_scaled)
p_pam_silhouette <- ggplot(pam_sil_results, aes(x = k, y = sil)) +
  geom_line() +
  geom_point() +
  scale_x_continuous(breaks = 2:10) +
  labs(title = "PAM: Silhouette Method",
       x = "Number of clusters k",
       y = "Average silhouette width") +
  theme_minimal()

# Calinski-Harabasz Index (CH index) for PAM
ch_index_pam <- function(data, max_k = 10) {
  ch_values <- numeric(max_k - 1)
  for (i in 2:max_k) {
    pam_result <- pam(data, k = i)
    ch_values[i-1] <- fpc::calinhara(data, pam_result$clustering)
  }
  k_values <- 2:max_k
  return(data.frame(k = k_values, ch_index = ch_values))
}

ch_pam_results <- ch_index_pam(player_data_scaled)
p_pam_ch <- ggplot(ch_pam_results, aes(x = k, y = ch_index)) +
  geom_line() +
  geom_point() +
  scale_x_continuous(breaks = 2:10) +
  labs(title = "PAM: Calinski-Harabasz Index",
       x = "Number of clusters k",
       y = "CH Index (higher is better)") +
  theme_minimal()

# Arrange PAM validation plots in a grid
grid.arrange(p_pam_silhouette, p_pam_ch,
             ncol = 2, nrow = 1,
             top = grid::textGrob("Determining Optimal Number of Clusters with PAM", 
                                 gp = grid::gpar(fontsize = 16, font = 2)))

# Print summary of results for easier interpretation
cat("\n\nSummary of Optimal k Recommendations for PAM:\n")
cat("- Silhouette Method: Highest peak suggests optimal k =", 
    pam_sil_results$k[which.max(pam_sil_results$sil)], "\n")
cat("- Calinski-Harabasz Index: Highest value suggests optimal k =", 
    ch_pam_results$k[which.max(ch_pam_results$ch_index)], "\n")

# Try k=3, k=4, and k=5 with PAM to visually inspect clusters
set.seed(123)
pam3_clusters <- pam(player_data_scaled, k = 3)
pam4_clusters <- pam(player_data_scaled, k = 4)
pam5_clusters <- pam(player_data_scaled, k = 5)

p_pam_k3 <- fviz_cluster(list(data = player_data_scaled, cluster = pam3_clusters$clustering),
                   palette = "Set2", ggtheme = theme_minimal(),
                   main = "PAM: k = 3")
p_pam_k4 <- fviz_cluster(list(data = player_data_scaled, cluster = pam4_clusters$clustering),
                   palette = "Set2", ggtheme = theme_minimal(),
                   main = "PAM: k = 4")
p_pam_k5 <- fviz_cluster(list(data = player_data_scaled, cluster = pam5_clusters$clustering),
                   palette = "Set2", ggtheme = theme_minimal(),
                   main = "PAM: k = 5")

grid.arrange(p_pam_k3, p_pam_k4, p_pam_k5, ncol = 3,
            top = grid::textGrob("Visual Comparison of Different k Values Using PAM", 
                                gp = grid::gpar(fontsize = 16, font = 2)))

# Calculate key statistics for each PAM cluster option
pam_cluster_stats <- function(data, k) {
  set.seed(123)
  pam_result <- pam(data, k = k)
  
  # Calculate silhouette coefficient for each cluster
  sil <- pam_result$silinfo$widths
  sil_per_cluster <- tapply(sil[, 3], sil[, 1], mean)
  
  # Return results
  return(list(
    k = k,
    avg_silhouette = pam_result$silinfo$avg.width,
    silhouette_per_cluster = sil_per_cluster,
    cluster_sizes = table(pam_result$clustering),
    medoid_indices = pam_result$id.med
  ))
}

# Get statistics for k=3, k=4, and k=5 using PAM
stats_pam_k3 <- pam_cluster_stats(player_data_scaled, 3)
stats_pam_k4 <- pam_cluster_stats(player_data_scaled, 4)
stats_pam_k5 <- pam_cluster_stats(player_data_scaled, 5)

# Print summary of PAM cluster statistics
cat("\nPAM Cluster Statistics Comparison:\n")

cat("\nPAM k=3 Clusters:\n")
cat("Avg Silhouette:", stats_pam_k3$avg_silhouette, "\n")
cat("Cluster Sizes:", stats_pam_k3$cluster_sizes, "\n")
cat("Silhouette per Cluster:", stats_pam_k3$silhouette_per_cluster, "\n")
cat("Medoid indices:", stats_pam_k3$medoid_indices, "\n")

cat("\nPAM k=4 Clusters:\n")
cat("Avg Silhouette:", stats_pam_k4$avg_silhouette, "\n")
cat("Cluster Sizes:", stats_pam_k4$cluster_sizes, "\n")
cat("Silhouette per Cluster:", stats_pam_k4$silhouette_per_cluster, "\n")
cat("Medoid indices:", stats_pam_k4$medoid_indices, "\n")

cat("\nPAM k=5 Clusters:\n")
cat("Avg Silhouette:", stats_pam_k5$avg_silhouette, "\n")
cat("Cluster Sizes:", stats_pam_k5$cluster_sizes, "\n")
cat("Silhouette per Cluster:", stats_pam_k5$silhouette_per_cluster, "\n")
cat("Medoid indices:", stats_pam_k5$medoid_indices, "\n")

# Key advantages of PAM
cat("\nAdvantages of PAM for Clustering with Outliers Present:\n")
cat("1. PAM uses actual data points (medoids) as cluster centers, making it less influenced by outliers\n")
cat("2. PAM minimizes the sum of dissimilarities rather than sum of squared Euclidean distances\n")
cat("3. PAM is more robust to noise and outliers as it's not affected by extreme values\n")
cat("4. PAM works with any distance metric, not just Euclidean distance\n")
cat("5. Medoids are actual observations that represent cluster characteristics well\n")
```

## Performing PAM Clustering

```{r pam_clustering}
#| label: pam_clustering
#| cache: true

# Ensure cluster package is loaded for PAM clustering
pacman::p_load(cluster)

# Perform PAM clustering with k=4 
# PAM is more robust to outliers than k-means as it uses medoids (actual data points) as centers
set.seed(42) # For reproducibility
k_clusters <- 4
pam_result <- pam(player_data_scaled, k = k_clusters, metric = "euclidean", stand = FALSE)

# Create a function to generate descriptive cluster names
generate_player_type_names <- function(cluster_data) {
  # Calculate cluster medoids for key variables
  centers <- cluster_data %>%
    group_by(cluster) %>%
    summarise(
      pvp_ratio = mean(ratio_pvp_hours),
      pve_ratio = mean(ratio_pve_hours),
      pvp_kd = mean(pvp_kd),
      pve_kd = mean(pve_kd),
      pvp_accuracy = mean(pvp_hit_accuracy),
      pve_accuracy = mean(pve_hit_accuracy),
      pvp_score = mean(pvp_average_score),
      pve_score = mean(pve_average_score)
    )
  
  # Assign our more meaningful player type names
  cluster_names <- character(nrow(centers))
  
  # Create map from cluster metric profile to our interpretations
  for (i in 1:nrow(centers)) {
    # Determine primary characteristics
    is_pvp_focused <- centers$pvp_ratio[i] > centers$pve_ratio[i]
    is_high_pvp_skill <- centers$pvp_kd[i] > median(cluster_data$pvp_kd, na.rm = TRUE) && 
                          centers$pvp_accuracy[i] > median(cluster_data$pvp_hit_accuracy, na.rm = TRUE)
    is_high_pve_skill <- centers$pve_kd[i] > median(cluster_data$pve_kd, na.rm = TRUE) && 
                          centers$pve_accuracy[i] > median(cluster_data$pve_hit_accuracy, na.rm = TRUE)
    is_balanced <- abs(centers$pvp_ratio[i] - centers$pve_ratio[i]) < 0.2 # Threshold for balanced play
    
    # Determine player type based on characteristics
    if (is_pvp_focused && is_high_pvp_skill) {
      cluster_names[i] <- "PvP Focused High-Skill Players"
    } else if (!is_pvp_focused && is_high_pve_skill) {
      cluster_names[i] <- "PvE Specialists"
    } else if (is_pvp_focused && !is_high_pvp_skill) {
      cluster_names[i] <- "Casual PvP Players"
    } else if (is_balanced) {
      cluster_names[i] <- "Balanced Mixed-Mode Players"
    } else {
      # Fallback for any cluster that doesn't match our specific interpretations
      cluster_names[i] <- "Mixed Playstyle Players"
    }
  }
  
  return(data.frame(cluster = 1:nrow(centers), player_type = cluster_names))
}

# Add cluster assignments to the original data
pvz_df_clustered <- pvz_df_analysis %>%
  mutate(cluster = factor(pam_result$clustering))

# Generate descriptive names for clusters
player_type_labels <- generate_player_type_names(pvz_df_clustered)

# Add descriptive names to the clustered data
pvz_df_clustered <- pvz_df_clustered %>%
  mutate(cluster_num = as.numeric(as.character(cluster))) %>%
  left_join(player_type_labels, by = c("cluster_num" = "cluster")) %>%
  select(-cluster_num) %>%
  mutate(player_type = factor(player_type))

# IMPORTANT: Ensure pvz_df_clustered is the ONLY dataframe used for all subsequent analyses
# Save the clustered data for further analysis
write_csv(pvz_df_clustered, here("data/processed/pvz_with_clusters_pam_nohours.csv"))

# Output a summary of the PAM clustering results
cat("PAM Clustering Results Summary:\n")
cat("- Number of clusters:", k_clusters, "\n")
cat("- Silhouette coefficient:", pam_result$silinfo$avg.width, "\n")
cat("- Cluster sizes:\n")
print(table(pam_result$clustering))
cat("\nCluster to Player Type Mapping:\n")
print(player_type_labels)

# Extract and display the medoids
medoid_indices <- pam_result$medoids
cat("\nCluster Medoids (Representative Players):\n")
print(pam_result$id.med)

# This clear statement of player types helps ensure consistency
cat("\nThe 4 player types identified by PAM clustering are:\n")
for (type in unique(player_type_labels$player_type)) {
  cat("- ", type, "\n")
}
```

## Visualizing Clusters

```{r visualize_clusters_pca}
#| label: visualize_clusters_pca
#| fig.width: 12
#| fig.height: 8

# Perform PCA for visualization
cluster_pca_result <- prcomp(player_data_scaled)
cluster_pca_data <- as.data.frame(cluster_pca_result$x[, 1:2])
cluster_pca_data$player_type <- pvz_df_clustered$player_type

# Prepare data for plotting with cluster medoids highlighted
cluster_pca_data$is_medoid <- FALSE
# Mark medoids in the PCA data
medoid_id <- pam_result$id.med
cluster_pca_data$is_medoid[medoid_id] <- TRUE

# Create rotation data frame for variable arrows
rotation_df <- as.data.frame(cluster_pca_result$rotation[, 1:2])
rotation_df$var_name <- rownames(rotation_df)
# Scale arrows for better visualization
arrow_scale <- 3
rotation_df$PC1 <- rotation_df$PC1 * arrow_scale
rotation_df$PC2 <- rotation_df$PC2 * arrow_scale

# Create enhanced PCA plot
ggplot(cluster_pca_data, aes(x = PC1, y = PC2, color = player_type)) +
  # Plot points with different shapes for medoids
  geom_point(aes(shape = is_medoid), size = ifelse(cluster_pca_data$is_medoid, 5, 3), alpha = 0.7) +
  scale_shape_manual(values = c("FALSE" = 16, "TRUE" = 18), 
                     labels = c("Regular Observation", "Cluster Medoid"),
                     name = "Observation Type") +
  scale_color_brewer(palette = "Set2", name = "Player Type") +
  # Add biplot arrows for variable contributions - using rotation_df
  geom_segment(data = rotation_df,
               aes(x = 0, y = 0, xend = PC1, yend = PC2),
               arrow = arrow(length = unit(0.2, "cm")), 
               color = "darkgrey") +
  # Add variable labels
  geom_text(data = rotation_df,
            aes(x = PC1 * 1.1, y = PC2 * 1.1, label = var_name),
            color = "black", size = 3) +
  # Add ellipses around clusters
  stat_ellipse(aes(fill = player_type), geom = "polygon", alpha = 0.2, level = 0.95) +
  labs(
    title = "PCA Visualization of Player Clusters with PAM",
    subtitle = "Cluster medoids (representative players) are highlighted with diamonds",
    x = paste0("PC1 (", round(summary(cluster_pca_result)$importance[2, 1] * 100, 1), "% of variance)"),
    y = paste0("PC2 (", round(summary(cluster_pca_result)$importance[2, 2] * 100, 1), "% of variance)")
  ) +
  theme_minimal() +
  theme(
    legend.position = "right",
    plot.title = element_text(face = "bold", size = 14),
    plot.subtitle = element_text(size = 10),
    axis.title = element_text(face = "bold")
  )
```

## Cluster Characteristics

```{r cluster_profile_summary}
#| label: cluster_profile_summary
#| fig.width: 12
#| fig.height: 8

# Calculate mean values of key variables by cluster
player_type_profiles <- pvz_df_clustered %>%
  group_by(player_type) %>%
  summarise(
    n = n(),
    pct = n() / nrow(pvz_df_clustered) * 100,
    hours = mean(hours),
    ratio_pvp = mean(ratio_pvp_hours) * 100,
    ratio_pve = mean(ratio_pve_hours) * 100,
    pvp_kd = mean(pvp_kd),
    pve_kd = mean(pve_kd),
    pvp_accuracy = mean(pvp_hit_accuracy) * 100,
    pve_accuracy = mean(pve_hit_accuracy) * 100,
    pvp_damage = mean(pvp_average_damage),
    pve_damage = mean(pve_average_damage),
    pvp_score = mean(pvp_average_score),
    pve_score = mean(pve_average_score)
  )

# Display cluster profiles
kable(player_type_profiles, digits = 2,
      caption = "Profile Summary of Player Types") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
                full_width = FALSE)
```

## Visualizing Player Types with Radar Charts

```{r radar_chart_functions}
#| label: radar_chart_functions

# Function to scale values to 0-1 range for radar charts
scale_to_01 <- function(x, min_val = min(x, na.rm = TRUE), max_val = max(x, na.rm = TRUE)) {
  if (max_val == min_val) return(rep(0.5, length(x)))
  (x - min_val) / (max_val - min_val)
}

# Add a z-score standardization function
z_score_scale <- function(x) {
  # Check for zero standard deviation to avoid division by zero
  if (sd(x, na.rm = TRUE) == 0) return(rep(0, length(x)))
  return((x - mean(x, na.rm = TRUE)) / sd(x, na.rm = TRUE))
}

# Helper function to convert z-scores to a 0-1 scale for radar chart display
z_to_radar_scale <- function(z, min_z = -3, max_z = 3) {
  # Clip extreme values to min_z and max_z
  z_clipped <- pmax(pmin(z, max_z), min_z)
  # Scale to 0-1 range
  scaled <- (z_clipped - min_z) / (max_z - min_z)
  return(scaled)
}

# Function to create an enhanced radar chart
create_player_type_radar <- function(cluster_data, cluster_color, cluster_title, actual_values) {
  # Get column names from the data
  data_cols <- colnames(cluster_data)
  
  # Create max and min rows with exactly matching column names
  max_min <- data.frame(matrix(c(rep(1, length(data_cols)), rep(0, length(data_cols))), 
                             nrow = 2, 
                             byrow = TRUE))
  # Set column names to match exactly
  colnames(max_min) <- data_cols
  
  # Combine max, min, and cluster data
  plot_data <- rbind(max_min, cluster_data)
  
  # Create the enhanced radar chart
  radarchart(
    plot_data, 
    pcol = cluster_color,
    pfcol = adjustcolor(cluster_color, alpha.f = 0.5),  # Increased opacity
    plwd = 3,  # Thicker lines
    vlcex = 0.9,  # Larger text
    title = cluster_title,
    cglcol = "grey80",
    cglty = 1,
    axislabcol = "grey30",
    caxislabels = seq(0, 1, 0.25),
    cglwd = 0.8,
    centerzero = FALSE,
    seg = 4,  # Fewer segments for cleaner look
    pty = 32  # More smooth polygon edges
  )
  
  # Add values as text at the bottom
  legend_text <- sprintf("PvP Ratio: %.2f, PvP K/D: %.2f, PvE Ratio: %.2f, PvE K/D: %.2f", 
                         actual_values$`PvP Ratio`, 
                         actual_values$`PvP K/D`,
                         actual_values$`PvE Ratio`,
                         actual_values$`PvE K/D`)
  
  mtext(legend_text, side = 1, line = 2, cex = 0.8, col = cluster_color)
}
```

```{r prepare_radar_data}
#| label: prepare_radar_data

# Define radar chart metrics
radar_metrics <- c(
  # Place PvE metrics first (left side)
  "PvE Ratio", "PvE K/D", "PvE Accuracy", "PvE Score", "PvE DPS", 
  # Place PvP metrics second (right side)
  "PvP Ratio", "PvP K/D", "PvP Accuracy", "PvP Score", "PvP DPS"
)

# Extract player type names to use as chart titles - ENSURE these are the PAM-generated types
# We're using the actual player_type factor levels from the clustered data
player_type_names <- levels(pvz_df_clustered$player_type)
cat("Radar charts will be created for these player types:\n")
print(player_type_names)

# Calculate z-scores for each metric across all players
pvz_metrics_z <- pvz_df_clustered %>%
  summarise(
    ratio_pve_hours_z = list(z_score_scale(ratio_pve_hours)),
    pve_kd_z = list(z_score_scale(pve_kd)),
    pve_hit_accuracy_z = list(z_score_scale(pve_hit_accuracy)),
    pve_average_score_z = list(z_score_scale(pve_average_score)),
    pve_dps_z = list(z_score_scale(pve_dps)),
    ratio_pvp_hours_z = list(z_score_scale(ratio_pvp_hours)),
    pvp_kd_z = list(z_score_scale(pvp_kd)),
    pvp_hit_accuracy_z = list(z_score_scale(pvp_hit_accuracy)),
    pvp_average_score_z = list(z_score_scale(pvp_average_score)),
    pvp_dps_z = list(z_score_scale(pvp_dps))
  )

# Calculate radar chart data for each player type using z-scores
radar_data_by_type <- pvz_df_clustered %>%
  group_by(player_type) %>%
  summarise(
    # Calculate mean values for each metric within this player type
    ratio_pve_hours_mean = mean(ratio_pve_hours, na.rm = TRUE),
    pve_kd_mean = mean(pve_kd, na.rm = TRUE),
    pve_hit_accuracy_mean = mean(pve_hit_accuracy, na.rm = TRUE),
    pve_average_score_mean = mean(pve_average_score, na.rm = TRUE),
    pve_dps_mean = mean(pve_dps, na.rm = TRUE),
    ratio_pvp_hours_mean = mean(ratio_pvp_hours, na.rm = TRUE),
    pvp_kd_mean = mean(pvp_kd, na.rm = TRUE),
    pvp_hit_accuracy_mean = mean(pvp_hit_accuracy, na.rm = TRUE),
    pvp_average_score_mean = mean(pvp_average_score, na.rm = TRUE),
    pvp_dps_mean = mean(pvp_dps, na.rm = TRUE)
  ) %>%
  ungroup() %>%
  # Convert type means to z-scores
  mutate(
    `PvE Ratio` = z_to_radar_scale(z_score_scale(ratio_pve_hours_mean)),
    `PvE K/D` = z_to_radar_scale(z_score_scale(pve_kd_mean)),
    `PvE Accuracy` = z_to_radar_scale(z_score_scale(pve_hit_accuracy_mean)),
    `PvE Score` = z_to_radar_scale(z_score_scale(pve_average_score_mean)),
    `PvE DPS` = z_to_radar_scale(z_score_scale(pve_dps_mean)),
    `PvP Ratio` = z_to_radar_scale(z_score_scale(ratio_pvp_hours_mean)),
    `PvP K/D` = z_to_radar_scale(z_score_scale(pvp_kd_mean)),
    `PvP Accuracy` = z_to_radar_scale(z_score_scale(pvp_hit_accuracy_mean)),
    `PvP Score` = z_to_radar_scale(z_score_scale(pvp_average_score_mean)),
    `PvP DPS` = z_to_radar_scale(z_score_scale(pvp_dps_mean))
  ) %>%
  # Drop the intermediate columns
  select(
    player_type,
    `PvE Ratio`, `PvE K/D`, `PvE Accuracy`, `PvE Score`, `PvE DPS`,
    `PvP Ratio`, `PvP K/D`, `PvP Accuracy`, `PvP Score`, `PvP DPS`
  )
```

```{r plot_radar_charts}
#| label: plot_radar_charts
#| fig.height: 16
#| fig.width: 10

# Define colors for each player type (ensuring they're unique)
# We're using the same number of colors as player types to ensure a 1:1 mapping
player_type_colors <- c("#E41A1C", "#377EB8", "#4DAF4A", "#984EA3")[1:length(player_type_names)]

# Set up the plotting layout - adjust based on number of player types
if(length(player_type_names) <= 4) {
  par(mfrow = c(2, 2))
} else {
  par(mfrow = c(3, 2))
}

# Loop through player types to create radar charts
for (i in 1:length(player_type_names)) {
  # Skip if we've run out of player types
  if (i > length(player_type_names)) break
  
  # Extract data for this player type
  type_data <- radar_data_by_type %>%
    filter(player_type == player_type_names[i]) %>%
    select(-player_type) %>%
    as.data.frame()
  
  # Get actual values for reference
  actual_values <- pvz_df_clustered %>%
    filter(player_type == player_type_names[i]) %>%
    summarise(
      `PvP Ratio` = mean(ratio_pvp_hours),
      `PvP K/D` = mean(pvp_kd),
      `PvE Ratio` = mean(ratio_pve_hours),
      `PvE K/D` = mean(pve_kd)
    )
  
  # Choose a color for this player type
  type_color <- player_type_colors[i]
  type_title <- paste("Player Type:", player_type_names[i])
  
  # Create the enhanced radar chart
  create_player_type_radar(type_data, type_color, type_title, actual_values)
}

# Add a global note about scaling
mtext("Note: Values shown as standardized scores (z-scores converted to 0-1 scale)", 
      side = 1, line = -1, outer = TRUE, cex = 0.8)

# Reset the layout
par(mfrow = c(1, 1))
```

## Comparing Player Types with Bar Charts

```{r player_type_comparison_bars}
#| label: player_type_comparison_bars
#| fig.width: 12
#| fig.height: 14

# Create a long format dataset for easier visualization
player_type_metrics <- pvz_df_clustered %>%
  group_by(player_type) %>%
  summarise(
    `Total Hours` = mean(hours),
    `PvP Ratio (%)` = mean(ratio_pvp_hours) * 100,
    `PvE Ratio (%)` = mean(ratio_pve_hours) * 100,
    `PvP K/D Ratio` = mean(pvp_kd),
    `PvE K/D Ratio` = mean(pve_kd),
    `PvP Accuracy (%)` = mean(pvp_hit_accuracy) * 100,
    `PvE Accuracy (%)` = mean(pve_hit_accuracy) * 100,
    `PvP Avg Score` = mean(pvp_average_score),
    `PvE Avg Score` = mean(pve_average_score)
  ) %>%
  pivot_longer(
    cols = -player_type,
    names_to = "Metric",
    values_to = "Value"
  )

# Group metrics for more organized plotting
player_type_metrics <- player_type_metrics %>%
  mutate(
    MetricGroup = case_when(
      Metric %in% c("Total Hours") ~ "Playtime",
      Metric %in% c("PvP Ratio (%)", "PvE Ratio (%)") ~ "Game Mode Preference",
      Metric %in% c("PvP K/D Ratio", "PvE K/D Ratio") ~ "Combat Effectiveness",
      Metric %in% c("PvP Accuracy (%)", "PvE Accuracy (%)") ~ "Accuracy",
      Metric %in% c("PvP Avg Score", "PvE Avg Score") ~ "Performance",
      TRUE ~ "Other"
    )
  )

# Create grouped bar charts for each metric category
ggplot(player_type_metrics, aes(x = Metric, y = Value, fill = player_type)) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.9)) +
  facet_wrap(~ MetricGroup, scales = "free", ncol = 2) +
  scale_fill_brewer(palette = "Set2") +
  labs(
    title = "Comparison of Player Type Characteristics",
    x = NULL,
    y = NULL,
    fill = "Player Type"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    strip.background = element_rect(fill = "grey90"),
    strip.text = element_text(face = "bold")
  )
```

# Relationship Between Playstyle and Well-Being

## Overall Well-Being by Player Type

```{r wellbeing-by-cluster}
#| label: wellbeing-by-cluster
#| fig.width: 10
#| fig.height: 6

# Verify we're using the correct player types
cat("Analyzing well-being metrics for these player types:\n")
print(levels(pvz_df_clustered$player_type))

# Compare well-being metrics across clusters
wellbeing_by_cluster <- pvz_df_clustered %>%
  group_by(player_type) %>%
  summarise(
    spane_balance = mean(spane_balance, na.rm = TRUE),
    spane_positive = mean(spane_positive, na.rm = TRUE),
    spane_negative = mean(spane_negative, na.rm = TRUE),
    spane_game_balance = mean(spane_game_balance, na.rm = TRUE),
    autonomy = mean(autonomy, na.rm = TRUE),
    competence = mean(competence, na.rm = TRUE),
    relatedness = mean(relatedness, na.rm = TRUE),
    enjoyment = mean(enjoyment, na.rm = TRUE),
    extrinsic = mean(extrinsic, na.rm = TRUE),
    n = n()
  )

# Display the summary table
kable(wellbeing_by_cluster, digits = 2,
      caption = "Well-being metrics by player type") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
                full_width = FALSE)

# Create boxplots for SPANE balance by player type
ggplot(pvz_df_clustered, aes(x = player_type, y = spane_balance, fill = player_type)) +
  geom_boxplot(alpha = 0.7) +
  scale_fill_brewer(palette = "Set2") +
  labs(
    title = "Overall Well-Being (SPANE Balance) by Player Type",
    x = "Player Type",
    y = "SPANE Balance Score"
  ) +
  theme_minimal() +
  theme(
    legend.position = "none",
    axis.text.x = element_text(angle = 45, hjust = 1)
  )
```

## Playtime and Well-Being Relationships

```{r playtime-wellbeing}
#| label: playtime-wellbeing
#| fig.width: 12
#| fig.height: 8

# Verify we're using the correct player types
cat("Examining playtime-wellbeing relationships for these player types:\n")
print(levels(pvz_df_clustered$player_type))

# Create scatter plots for total playtime vs. well-being
p1 <- ggplot(pvz_df_clustered, aes(x = hours, y = spane_balance, color = player_type)) +
  geom_point(alpha = 0.6) +
  geom_smooth(method = "lm", se = TRUE, alpha = 0.2) +
  scale_color_brewer(palette = "Set2") +
  labs(
    title = "Total Playtime vs. Overall Well-Being",
    x = "Total Hours Played",
    y = "SPANE Balance Score",
    color = "Player Type"
  ) +
  theme_minimal()

p2 <- ggplot(pvz_df_clustered, aes(x = hours, y = spane_game_balance, color = player_type)) +
  geom_point(alpha = 0.6) +
  geom_smooth(method = "lm", se = TRUE, alpha = 0.2) +
  scale_color_brewer(palette = "Set2") +
  labs(
    title = "Total Playtime vs. Game-Related Well-being",
    x = "Total Hours Played",
    y = "Game SPANE Balance Score",
    color = "Player Type"
  ) +
  theme_minimal()

# Combine the plots
ggarrange(p1, p2, ncol = 2, common.legend = TRUE, legend = "right")
```

```{r gamemode-wellbeing}
#| label: gamemode-wellbeing
#| fig.width: 12
#| fig.height: 10

# Create scatter plots for game mode playtime vs. well-being
# Since we don't have direct hour measurements for each game mode,
# we'll use the ratio variables multiplied by total hours

# Calculate estimated hours for each mode
# Note: We're careful to NOT overwrite the player_type information here
pvz_df_with_est_hours <- pvz_df_clustered %>%
  mutate(
    est_pvp_hours = hours * ratio_pvp_hours,
    est_pve_hours = hours * ratio_pve_hours
  )

p1 <- ggplot(pvz_df_with_est_hours, aes(x = est_pvp_hours, y = spane_balance, color = player_type)) +
  geom_point(alpha = 0.6) +
  geom_smooth(method = "lm", se = TRUE, alpha = 0.2) +
  scale_color_brewer(palette = "Set2") +
  labs(
    title = "PvP Playtime vs. Overall Well-Being",
    x = "Estimated PvP Hours Played",
    y = "SPANE Balance Score",
    color = "Player Type"
  ) +
  theme_minimal()

p2 <- ggplot(pvz_df_with_est_hours, aes(x = est_pve_hours, y = spane_balance, color = player_type)) +
  geom_point(alpha = 0.6) +
  geom_smooth(method = "lm", se = TRUE, alpha = 0.2) +
  scale_color_brewer(palette = "Set2") +
  labs(
    title = "PvE Playtime vs. Overall Well-Being",
    x = "Estimated PvE Hours Played",
    y = "SPANE Balance Score",
    color = "Player Type"
  ) +
  theme_minimal()

# Combine the plots - using just two plots since we don't have social hours data
ggarrange(p1, p2, ncol = 2, nrow = 1, common.legend = TRUE, legend = "right")
```

## Motivation and Well-Being by Player Type

```{r motivation-by-cluster}
#| label: motivation-by-cluster
#| fig.width: 12
#| fig.height: 8

# Print out the player types again to ensure consistency
cat("Analyzing motivation factors for these player types:\n")
print(levels(pvz_df_clustered$player_type))

# Prepare data for motivation visualization
motivation_data <- pvz_df_clustered %>%
  select(player_id, player_type, autonomy, competence, relatedness, enjoyment, extrinsic) %>%
  pivot_longer(
    cols = c(autonomy, competence, relatedness, enjoyment, extrinsic),
    names_to = "motivation_type",
    values_to = "score"
  ) %>%
  mutate(
    motivation_type = factor(
      motivation_type, 
      levels = c("autonomy", "competence", "relatedness", "enjoyment", "extrinsic"),
      labels = c("Autonomy", "Competence", "Relatedness", "Enjoyment", "Extrinsic")
    )
  )

# Create bar chart for motivation by player type
ggplot(motivation_data, aes(x = motivation_type, y = score, fill = player_type)) +
  geom_bar(stat = "summary", fun = "mean", position = "dodge") +
  scale_fill_brewer(palette = "Set2") +
  labs(
    title = "Motivation Factors by Player Type",
    x = "Motivation Type",
    y = "Mean Score",
    fill = "Player Type"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

# Discussion and Conclusion

This analysis has identified four distinct player types in Plants vs. Zombies: Battle for Neighborville based on gameplay metrics and preferences. Using PAM clustering (which is robust to outliers), we identified four clearly differentiated player segments with unique characteristics and motivations:

Key findings include:

1. **PvP Focused High-Skill Players**: These players demonstrate exceptional competence in competitive play, with high kill/death ratios and accuracy in PvP modes. They likely value mastery, competition, and strategic depth.

2. **PvE Specialists**: These players focus primarily on cooperative or solo gameplay against AI opponents. They excel in PvE content with above-average performance metrics, suggesting they value progression, predictability, and potentially storytelling elements.

3. **Casual PvP Players**: While preferring competitive modes, these players have more moderate performance metrics than their high-skill counterparts. They likely play for social enjoyment and variety rather than competitive mastery.

4. **Balanced Mixed-Mode Players**: These versatile players engage with both PvP and PvE content without strongly specializing in either. Their balanced approach suggests they value game variety and may adjust their playstyle based on social context or mood.

The identified player types show different relationships between playstyle and well-being, suggesting that how players engage with the game matters more than simply how much they play. Motivation factors such as autonomy, competence, and enjoyment vary meaningfully across player types, indicating different psychological needs being met through gameplay.

Game-related well-being tends to be more positive than general well-being across all player types, suggesting that the game provides a positive experience regardless of playstyle. However, the patterns and strength of these relationships differ based on player type.

These findings have important implications for game design:

1. **Segmented Content Development**: Developers should ensure robust content for both hardcore competitive players and those who prefer cooperative experiences.

2. **Skill-Based Considerations**: The clear separation between high-skill and casual players suggests a need for effective matchmaking and progression systems that accommodate different skill levels.

3. **Cross-Mode Incentives**: For balanced players who enjoy variety, designing rewards that encourage engagement across different game modes could enhance satisfaction.

4. **Tailored Social Features**: Different player types likely have different social needs—from team coordination for competitive players to casual social spaces for others.

Future research could explore how these patterns evolve over longer periods and how specific game design elements might impact the well-being of different player types. Longitudinal studies could also examine if players migrate between these types over time or remain consistent in their preferences.

```{r save-results}
#| label: save-results
#| include: false

# Final cleanup and save session info for reproducibility
sessionInfo()
```