---
title: "Player Playstyle Analysis"
author: "Thomas Hakman"
format: 
  html:
    toc: true
    toc-depth: 3
    code-fold: true
    theme: flatly
    fig-width: 8
    fig-height: 6
execute:
  echo: true
  message: false
  warning: false
  cache: true
editor: visual
---

```{r setup}
#| label: setup
#| include: false

# Set global options
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  cache = FALSE,
  fig.path = "figures/"
)
```

# Introduction

This document contains the analysis of player telemetry and survey data from Plants vs. Zombies: Battle for Neighborville. It focuses on analyzing player playstyles, game performance metrics, and their relationship with well-being outcomes.

test test

# Load Required Packages and Data

```{r install-packages, include=FALSE}
#| label: install-packages

# First ensure pacman is installed if not already
if (!requireNamespace("pacman", quietly = TRUE)) {
  install.packages("pacman")
}

# Use pacman to load all required packages (will install any missing packages)
pacman::p_load(
  tidyverse,  # For data manipulation and visualization
  here,       # For project-relative paths
  lubridate,  # For date/time manipulation
  scales,     # For scale transformations
  janitor,    # For data cleaning
  magrittr,   # For additional pipe operators
  psych,      # For psychological statistics
  corrplot,   # For correlation plots
  ggpubr,     # For publication-ready plots
  factoextra, # For cluster analysis visualization
  cluster,    # For clustering algorithms
  fmsb,       # For radar charts
  knitr,      # For dynamic report generation
  tibble,     # For improved data frames
  kableExtra  # For enhanced table formatting
)
```

```{r helper-functions}
#| label: helper-functions

# Function to scale values to 0-1 range for radar charts
scale_to_01 <- function(x, min_val = min(x, na.rm = TRUE), max_val = max(x, na.rm = TRUE)) {
  if (max_val == min_val) return(rep(0.5, length(x)))
  (x - min_val) / (max_val - min_val)
}

```

```{r statistical-functions}
#| label: statistical-functions

# Function to run ANOVA and extract results
run_anova <- function(dv) {
  formula <- as.formula(paste(dv, "~ player_type"))
  model <- aov(formula, data = pvz_data)
  result <- broom::tidy(model)
  result$eta_squared <- sjstats::eta_sq(model)[["eta.sq"]]
  result$variable <- dv
  return(result)
}

# Function to compute Hopkins statistic for cluster tendency
compute_hopkins <- function(data, n = nrow(data)-1) {
  # Standardize data for clustering
  data_scaled <- scale(data)
  
  # Compute Hopkins statistic
  hopkins_result <- clustertend::hopkins(data_scaled, n = n)
  
  return(hopkins_result)
}
```

```{r load-processed-data}
#| label: load-processed-data

pvz_data <- read_csv(here("data/processed/pvz_complete.csv"))

# Create a copy of pvz_data for analysis to maintain consistency
pvz_df_analysis <- pvz_data

glimpse(pvz_df_analysis)
```

```{r select-columns}
#| label: select-columns

pvz_df_analysis <- pvz_df_analysis %>%
  select(
    # Demographic variables
    player_id, country, age, gender, played_with_others,
    
    # Well-being variables
    spane_balance, spane_positive, spane_negative,
    autonomy, competence, relatedness, enjoyment, extrinsic,
    spane_game_balance, spane_game_positive, spane_game_negative,
    
    # Total playtime
    hours,
    
    # PVP metrics
    ratio_pvp_hours,
    pvp_kd,
    pvp_hit_accuracy,
    pvp_dps,
    pvp_average_damage,
    pvp_average_score,
    
    # PVE metrics
    ratio_pve_hours,
    pve_kd,
    pve_hit_accuracy,
    pve_dps,
    pve_average_damage,
    pve_average_score
  )
```

# Exploratory Data Analysis

## plot session timeline for a player

```{r plot-session-timeline}
#| label: plot-session-timeline
#| fig.width: 12
#| fig.height: 16

# Function to plot session timeline for a player
plot_player_sessions <- function(player_data, player_id = NULL, player_label = "Player") {
  # If a player_id is provided, filter data for that player
  if (!is.null(player_id)) {
    player_data <- player_data %>%
      filter(player_id == !!player_id)
  }
  
  # Convert timestamps to POSIXct if they're not already
  if (!inherits(player_data$game_start_time, "POSIXct")) {
    player_data$game_start_time <- as.POSIXct(player_data$game_start_time, format = "%Y-%m-%d %H:%M:%S")
    player_data$game_end_time <- as.POSIXct(player_data$game_end_time, format = "%Y-%m-%d %H:%M:%S")
  }
  
  # Extract day of week, date, and time component
  player_data$day_of_week <- weekdays(player_data$game_start_time)
  player_data$date <- as.Date(player_data$game_start_time)
  player_data$start_time_only <- as.POSIXct(format(player_data$game_start_time, format = "%H:%M:%S"), format = "%H:%M:%S")
  player_data$end_time_only <- as.POSIXct(format(player_data$game_end_time, format = "%H:%M:%S"), format = "%H:%M:%S")
  
  # Identify sessions that cross midnight (where end_time is earlier than start_time)
  player_data$crosses_midnight <- player_data$end_time_only < player_data$start_time_only
  
  # Create a new dataframe that will include split sessions for those crossing midnight
  player_data_expanded <- player_data %>%
    filter(!crosses_midnight) # Keep all sessions that don't cross midnight
  
  # Handle sessions that cross midnight by splitting them
  midnight_sessions <- player_data %>%
    filter(crosses_midnight)
  
  if (nrow(midnight_sessions) > 0) {
    # For each session crossing midnight, create two rows:
    # 1. From start time to 23:59:59
    # 2. From 00:00:00 to end time
    midnight <- as.POSIXct(strptime("23:59:59", format = "%H:%M:%S"), format = "%H:%M:%S")
    midnight_start <- as.POSIXct(strptime("00:00:00", format = "%H:%M:%S"), format = "%H:%M:%S")
    
    # First part: from start time to midnight
    midnight_part1 <- midnight_sessions
    midnight_part1$end_time_only <- midnight
    
    # Second part: from midnight start to end time
    midnight_part2 <- midnight_sessions
    midnight_part2$start_time_only <- midnight_start
    # The second part is on the next day of the week and the next date
    midnight_part2$day_of_week <- factor(
      weekdays(midnight_part2$game_start_time + 86400), # Add one day (86400 seconds)
      levels = c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday")
    )
    midnight_part2$date <- midnight_part2$date + 1
    
    # Add both parts to the expanded dataset
    player_data_expanded <- bind_rows(player_data_expanded, midnight_part1, midnight_part2)
  }
  
  # Recode match_type based on game_level_key 
  player_data_expanded$match_type <- "0"  # Initialize with 0 as in the original code
  
  # Social matches
  player_data_expanded$match_type[player_data_expanded$game_level_key == "dsub_social_terrain_beach"] <- "social match"
  player_data_expanded$match_type[player_data_expanded$game_level_key == "dsub_social_terrain_summer"] <- "social match"
  player_data_expanded$match_type[player_data_expanded$game_level_key == "dsub_socialspace"] <- "social match"
  
  # PVE matches
  player_data_expanded$match_type[player_data_expanded$game_level_key == "dsub_region_cheesemines"] <- "pve match"
  player_data_expanded$match_type[player_data_expanded$game_level_key == "dsub_region_dreadwood"] <- "pve match"
  player_data_expanded$match_type[player_data_expanded$game_level_key == "dsub_region_towncenter"] <- "pve match" 
  
  # PVP matches (Rush)
  player_data_expanded$match_type[player_data_expanded$game_level_key == "dsub_rush_dreadwood"] <- "pvp match (rush)"
  player_data_expanded$match_type[player_data_expanded$game_level_key == "dsub_rush_towncenter"] <- "pvp match (rush)"
  player_data_expanded$match_type[player_data_expanded$game_level_key == "dsub_rush_farm"] <- "pvp match (rush)"
  player_data_expanded$match_type[player_data_expanded$game_level_key == "dsub_rush_cheesemines"] <- "pvp match (rush)"
  player_data_expanded$match_type[player_data_expanded$game_level_key == "dsub_rush_shoreline"] <- "pvp match (rush)"
  player_data_expanded$match_type[player_data_expanded$game_level_key == "dsub_rush_resort"] <- "pvp match (rush)"
  
  # PVP matches (Vanquish)
  player_data_expanded$match_type[player_data_expanded$game_level_key == "dsub_tvm_cheesemines"] <- "pvp match (vanquish)"
  player_data_expanded$match_type[player_data_expanded$game_level_key == "dsub_tvm_zrefinery"] <- "pvp match (vanquish)"
  player_data_expanded$match_type[player_data_expanded$game_level_key == "dsub_tvm_suburbs"] <- "pvp match (vanquish)"
  player_data_expanded$match_type[player_data_expanded$game_level_key == "dsub_tvm_dreadwood"] <- "pvp match (vanquish)"
  
  # PVP matches (Arena)
  player_data_expanded$match_type[player_data_expanded$game_level_key == "dsub_4v4_battlearena_v2"] <- "pvp match (arena)"
  player_data_expanded$match_type[player_data_expanded$game_level_key == "dsub_4v4_superarena"] <- "pvp match (arena)"
  
  # Podium celebration
  player_data_expanded$match_type[player_data_expanded$game_level_key == "dsub_podiumcelebration"] <- "podium celebration"
  
  # Ensure day of week is a factor with ordered levels
  player_data_expanded$day_of_week <- factor(
    player_data_expanded$day_of_week,
    levels = c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday")
  )
  
  # Create a position adjustment for stacking sessions by different dates for the same day of week
  player_data_expanded <- player_data_expanded %>%
    group_by(day_of_week) %>%
    mutate(
      # Get unique dates for this day of week and assign a position index to each date
      unique_dates = list(unique(date)),
      date_index = match(date, unique(date)),
      # Scale the vertical offset to be small but noticeable
      y_offset = (date_index - 1) * 0.2,  # Creates a small vertical offset for different dates
      # Create adjusted y-position that includes the offset
      y_position = as.numeric(day_of_week) + y_offset
    ) %>%
    ungroup()
  
  # Sort by day of week, date, and start time
  player_data_expanded <- player_data_expanded %>%
    arrange(day_of_week, date, game_start_time, game_end_time)
  
  # Add a session ID within each day to distinguish multiple sessions on same day
  player_data_expanded <- player_data_expanded %>%
    group_by(day_of_week, date) %>%
    mutate(session_id_within_day = row_number()) %>%
    ungroup()
  
  # Convert match_type to factor with proper levels for consistent ordering
  player_data_expanded$match_type <- factor(
    player_data_expanded$match_type,
    levels = c("social match", "pve match", "pvp match (rush)", "pvp match (vanquish)", "pvp match (arena)", "podium celebration", "0"),
    labels = c("Social Match", "PvE Match", "PvP (Rush)", "PvP (Vanquish)", "PvP (Arena)", "Podium Celebration", "Other")
  )
  
  # Create the plot with enhanced styling
  plot <- player_data_expanded %>%
    ggplot2::ggplot(aes(start_time_only, y_position, col = match_type, 
                        group = interaction(day_of_week, date, session_id_within_day))) +
    # Use improved color palette with better contrast
    ggplot2::scale_color_manual(
      name = "Match Type", 
      values = c(
        "Social Match" = "#00BFC4",       # Teal
        "PvE Match" = "#7CAE00",          # Green
        "PvP (Rush)" = "#F8766D",         # Red
        "PvP (Vanquish)" = "#FF9E1C",     # Orange
        "PvP (Arena)" = "#FFD700",        # Gold
        "Podium Celebration" = "#C77CFF", # Purple
        "Other" = "#A0A0A0"               # Gray
      )
    ) +
    # Thicker segments with rounded ends for better visibility
    ggplot2::geom_segment(
      aes(xend = end_time_only, yend = y_position),
      linewidth = 2,
      lineend = "round"
    ) +
    # Add points at start/end for better visibility of short sessions
    ggplot2::geom_point(size = 1.5, alpha = 0.6) +
    ggplot2::geom_point(aes(x = end_time_only), size = 1.5, alpha = 0.6) +
    # Better time axis formatting - using fixed limits for consistency
    ggplot2::scale_x_datetime(
      labels = scales::date_format("%H:%M"),
      breaks = scales::date_breaks("2 hours"),
      limits = c(
        as.POSIXct(strptime("00:00", format = "%H:%M")),
        as.POSIXct(strptime("23:59", format = "%H:%M"))
      ),
      expand = c(0.01, 0.01),
      date_minor_breaks = "1 hour"
    ) +
    # Use the original day of week as the y-axis labels
    ggplot2::scale_y_continuous(
      breaks = 1:7,
      labels = levels(player_data_expanded$day_of_week),
      limits = c(0.5, 7.5)
    ) +
    # Empty labs - axis labels will be added to the combined figure instead
    ggplot2::labs(
      x = NULL, 
      y = NULL
    ) +
    # Improved theme with better typography and layout
    ggplot2::theme_minimal(base_size = 11, base_family = "sans") +
    ggplot2::theme(
      legend.position = "none", # Move legend to the combined figure
      plot.margin = margin(30, 10, 30, 10), # Increased margins for better separation
      
      axis.text.y = element_text(face = "bold"),
      axis.title = element_text(face = "bold"),
      
      panel.grid.major.x = element_line(color = "grey90"),
      panel.grid.minor.x = element_line(color = "grey95"),
      panel.grid.major.y = element_line(color = "grey90"),
      panel.grid.minor.y = element_blank(),
      
      panel.background = element_rect(fill = "white", color = NA),
      plot.background = element_rect(fill = "white", color = NA)
    )
  
  return(plot)
}

# Load game session data if not already included
game_sessions <- read_csv(here("data/processed/game_time_2week.csv"))

# Make sure timestamps are in the correct format
game_sessions$game_start_time <- as.POSIXct(game_sessions$game_start_time, format = "%Y-%m-%d %H:%M:%S")
game_sessions$game_end_time <- as.POSIXct(game_sessions$game_end_time, format = "%Y-%m-%d %H:%M:%S")


# First, get a list of player IDs in the data
player_ids <- unique(game_sessions$player_id)

# Choose specific indices
# If those indices don't exist in our data, we'll use the first three players
player_indices <- c(
  min(401, length(player_ids)),  # Equivalent to slice(401) 
  min(6, length(player_ids)),    # Equivalent to slice(6)
  min(3, length(player_ids))     # Equivalent to slice(3)
)

selected_player_ids <- player_ids[player_indices]

# Create individual plots for each player with standardized x-axis
plot1 <- plot_player_sessions(
  game_sessions %>% filter(player_id == selected_player_ids[1])
)

plot2 <- plot_player_sessions(
  game_sessions %>% filter(player_id == selected_player_ids[2])
)

# For the third plot, use the same time range for consistency
plot3 <- plot_player_sessions(
  game_sessions %>% filter(player_id == selected_player_ids[3])
)

# Combine plots into a single figure with enhanced styling and proper spacing
figure <- ggarrange(
  plot2, plot3, plot1,
  labels = c("Player A", "Player B", "Player C"),
  font.label = list(size = 14, face = "bold", color = "black"),
  ncol = 1, 
  nrow = 3, 
  common.legend = TRUE, 
  legend = "bottom",
  heights = c(1, 1, 1),
  align = "v",
  hjust = -0.1
)

# Add shared x and y labels to the combined figure
figure <- annotate_figure(figure,
                        bottom = text_grob("Time of Day (hours)", 
                                          face = "bold", size = 12),
                        left = text_grob("Day of Week", 
                                         face = "bold", size = 12, rot = 90))

# Display the combined figure
figure
```

# Descriptive Statistics

```{r apa-descriptive-stats}
#| label: tbl-descriptive-stats
#| tbl-cap: "Descriptive Statistics for Key Variables"
#| fig.width: 10
#| fig.height: 6
#| tbl-cap-location: top
#| echo: false
#| warning: false
#| message: false

# Create a simple APA-style descriptive statistics table
# First, check variable types to handle them appropriately
variable_types <- sapply(pvz_df_analysis, class)

# Handle numeric variables
numeric_vars <- pvz_df_analysis %>%
  select(where(is.numeric)) %>%
  select(
    # Demographics
    age,
    
    # Playtime 
    hours, ratio_pvp_hours, ratio_pve_hours,
    
    # Performance metrics
    pvp_kd, pve_kd, 
    pvp_hit_accuracy, pve_hit_accuracy,
    pvp_dps, pve_dps,
    pvp_average_damage, pve_average_damage,
    pvp_average_score, pve_average_score,
    
    # Well-being and motivation
    spane_balance, spane_game_balance, 
    autonomy, competence, relatedness,
    enjoyment, extrinsic
  )

# Calculate descriptive statistics for numeric variables
numeric_stats <- numeric_vars %>%
  summarise(across(everything(), 
                  list(
                    n = ~sum(!is.na(.)),
                    M = ~mean(., na.rm = TRUE),
                    SD = ~sd(., na.rm = TRUE),
                    Min = ~min(., na.rm = TRUE),
                    Max = ~max(., na.rm = TRUE)
                  ))) %>%
  pivot_longer(everything(), 
               names_to = c("variable", "stat"), 
               names_pattern = "(.*)_(.*)",
               values_to = "value") %>%
  pivot_wider(names_from = stat, values_from = value)

# For categorical variables like gender, just count frequencies for the footnote
if ("gender" %in% names(pvz_df_analysis)) {
  gender_counts <- pvz_df_analysis %>%
    count(gender) %>%
    mutate(pct = n / sum(n) * 100) %>%
    mutate(pct = round(pct, 1)) %>%
    mutate(frequency = paste0(n, " (", pct, "%)"))
  
  # Note: We'll add a footnote to explain gender distribution separately
  gender_note <- paste("Gender distribution: ", 
                     paste(paste0(gender_counts$gender, ": ", gender_counts$frequency), 
                           collapse = "; "))
} else {
  gender_note <- NULL
}

# Format the data for the table
descriptive_stats <- numeric_stats %>%
  # Format variable names for better readability
  mutate(variable = case_when(
    variable == "age" ~ "Age",
    variable == "hours" ~ "Total Hours Played",
    variable == "ratio_pvp_hours" ~ "Proportion of PvP Playtime",
    variable == "ratio_pve_hours" ~ "Proportion of PvE Playtime",
    variable == "pvp_kd" ~ "PvP Kill/Death Ratio",
    variable == "pve_kd" ~ "PvE Kill/Death Ratio",
    variable == "pvp_hit_accuracy" ~ "PvP Hit Accuracy",
    variable == "pve_hit_accuracy" ~ "PvE Hit Accuracy",
    variable == "pvp_dps" ~ "PvP Damage per Second",
    variable == "pve_dps" ~ "PvE Damage per Second",
    variable == "pvp_average_damage" ~ "PvP Average Damage per Game",
    variable == "pve_average_damage" ~ "PvE Average Damage per Game",
    variable == "pvp_average_score" ~ "PvP Average Score",
    variable == "pve_average_score" ~ "PvE Average Score",
    variable == "spane_balance" ~ "Overall Well-being (SPANE)",
    variable == "spane_game_balance" ~ "Game-related Well-being",
    variable == "autonomy" ~ "Autonomy",
    variable == "competence" ~ "Competence",
    variable == "relatedness" ~ "Relatedness",
    variable == "enjoyment" ~ "Intrinsic Enjoyment",
    variable == "extrinsic" ~ "Extrinsic Motivation",
    TRUE ~ variable
  )) %>%
  # Create category variable for grouping
  mutate(category = case_when(
    variable == "Age" ~ "Demographics",
    variable %in% c("Total Hours Played", "Proportion of PvP Playtime", "Proportion of PvE Playtime") ~ "Playtime",
    variable %in% c("PvP Kill/Death Ratio", "PvE Kill/Death Ratio", 
                   "PvP Hit Accuracy", "PvE Hit Accuracy",
                   "PvP Damage per Second", "PvE Damage per Second",
                   "PvP Average Damage per Game", "PvE Average Damage per Game",
                   "PvP Average Score", "PvE Average Score") ~ "Performance",
    variable %in% c("Overall Well-being (SPANE)", "Game-related Well-being") ~ "Well-being",
    variable %in% c("Autonomy", "Competence", "Relatedness", "Intrinsic Enjoyment", "Extrinsic Motivation") ~ "Motivation",
    TRUE ~ "Other"
  )) %>%
  # Round numeric values
  mutate(
    M = round(M, 2),
    SD = round(SD, 2),
    Min = round(Min, 2),
    Max = round(Max, 2)
  )

# Create custom ordering for variables to ensure autonomy, competence, relatedness are listed in order
variable_order <- c(
  # Demographics
  "Age",
  
  # Playtime
  "Total Hours Played", "Proportion of PvP Playtime", "Proportion of PvE Playtime",
  
  # Performance
  "PvP Kill/Death Ratio", "PvE Kill/Death Ratio", 
  "PvP Hit Accuracy", "PvE Hit Accuracy",
  "PvP Damage per Second", "PvE Damage per Second",
  "PvP Average Damage per Game", "PvE Average Damage per Game",
  "PvP Average Score", "PvE Average Score",
  
  # Well-being
  "Overall Well-being (SPANE)", "Game-related Well-being",
  
  # Motivation (in specified order)
  "Autonomy", "Competence", "Relatedness", "Intrinsic Enjoyment", "Extrinsic Motivation"
)

# Apply the custom ordering
descriptive_stats <- descriptive_stats %>%
  mutate(variable = factor(variable, levels = variable_order)) %>%
  arrange(variable)

# Create the APA-styled table with kable
apa_table <- descriptive_stats %>%
  select(variable, n, M, SD, Min, Max) %>%
  kable(
    # Caption is handled by chunk options now
    col.names = c("Variable", "*N*", "*M*", "*SD*", "Min", "Max"),
    align = c("l", "c", "c", "c", "c", "c"),
    booktabs = TRUE,
    format = "html",
    digits = 2
  ) %>%
  kable_styling(
    bootstrap_options = c("striped", "hover", "condensed"),
    full_width = FALSE,
    position = "left",
    font_size = 12
  ) %>%
  row_spec(0, bold = TRUE) %>%  # Bold header row per APA
  pack_rows(index = table(factor(descriptive_stats$category, 
                              levels = unique(descriptive_stats$category))), 
            hline_after = TRUE)  # Add lines after each group

# Add footnote for gender if it exists
if (!is.null(gender_note)) {
  apa_table <- apa_table %>%
    footnote(general = gender_note, general_title = "Note: ")
}

# Display the table
apa_table
```

# Player Clustering Analysis

## Assessing Cluster Tendency

```{r assess-cluster-tendency}
#| label: assess-cluster-tendency
#| fig.width: 12
#| fig.height: 6  
#| echo: true
#| warning: false
#| message: false

# Load required packages for cluster tendency assessment
if (!requireNamespace("clustertend", quietly = TRUE)) {
  install.packages("clustertend")
}
library(clustertend)
library(factoextra)

# 1. Prepare datasets for visual assessment of clustering tendency

# Select relevant features for clustering
cluster_features <- pvz_df_analysis %>% 
  select(
    # Playtime variables
    hours, ratio_pvp_hours, ratio_pve_hours,
    
    # PvP performance metrics
    pvp_kd, pvp_hit_accuracy, pvp_dps, pvp_average_damage, pvp_average_score,
    
    # PvE performance metrics
    pve_kd, pve_hit_accuracy, pve_dps, pve_average_damage, pve_average_score
  )

# Remove any NAs in the dataset to ensure clean clustering
cluster_features <- na.omit(cluster_features)

# Standardize the data (important for clustering)
cluster_data_scaled <- scale(cluster_features)

# Create a random dataset with the same dimensions for comparison
set.seed(123) # For reproducibility
random_data <- apply(cluster_data_scaled, 2,
                     function(x){runif(length(x), min(x), max(x))})
random_data <- as.data.frame(random_data)

# 2. Visual assessment using PCA visualization
pca_real <- prcomp(cluster_data_scaled)
pca_random <- prcomp(random_data)

# Create PCA plot for real data
p1 <- fviz_pca_ind(pca_real, 
                  title = "PCA - Player Performance Data",
                  geom = "point", 
                  ggtheme = theme_minimal(),
                  legend = "none",
                  alpha.ind = 0.7) +
      theme(plot.title = element_text(face = "bold", size = 14))

# Create PCA plot for random data
p2 <- fviz_pca_ind(pca_random, 
                  title = "PCA - Random Data",
                  geom = "point", 
                  ggtheme = theme_minimal(),
                  legend = "none",
                  alpha.ind = 0.7) +
      theme(plot.title = element_text(face = "bold", size = 14))

# Arrange plots side by side for comparison
gridExtra::grid.arrange(p1, p2, ncol = 2)

# Skip the Hopkins statistic calculation for now since it's causing issues
cat("Visual assessment of clustering tendency:\n",
    "Compare the two PCA plots above. If the real data shows more visible clustering ",
    "than the random data, this suggests natural clusters exist in the dataset.\n\n",
    "The distance matrix visualization below provides another perspective on clustering tendency.")

# 3. Visual assessment using distance matrix
# Create and visualize distance matrices
dist_real <- get_dist(cluster_data_scaled, method = "euclidean")
dist_random <- get_dist(random_data, method = "euclidean")

# Add titles to distance matrix plots using ggplot methods
p3 <- fviz_dist(dist_real, 
               show_labels = FALSE,
               gradient = list(low = "#00AFBB", mid = "white", high = "#FC4E07")) +
      ggtitle("Distance Matrix - Player Data") +
      theme(plot.title = element_text(face = "bold", size = 14))

p4 <- fviz_dist(dist_random, 
               show_labels = FALSE,
               gradient = list(low = "#00AFBB", mid = "white", high = "#FC4E07")) +
      ggtitle("Distance Matrix - Random Data") +
      theme(plot.title = element_text(face = "bold", size = 14))

# Arrange distance matrix plots side by side
gridExtra::grid.arrange(p3, p4, ncol = 2)

cat("Distance matrix visualization interpretation:\n",
    "In a dataset with clear cluster structure, the distance matrix will show dark blocks along the diagonal,\n",
    "indicating groups of observations that are similar to each other (small distances).\n",
    "Random data typically shows a more uniform pattern with no clear blocks.")
```

## Feature Selection for Clustering

```{r prepare-clustering}
#| label: prepare-clustering

# Select features for clustering
clustering_features <- pvz_df_analysis %>%
  select(
    # Playtime variables
    hours, 
    ratio_pvp_hours, ratio_pve_hours,
    
    # Performance metrics
    pvp_kd, pve_kd,
    pvp_hit_accuracy, pve_hit_accuracy,
    pvp_dps, pve_dps,
    pvp_average_damage, pve_average_damage,
    pvp_average_score, pve_average_score
  )

# Scale the data for clustering (important for k-means)
clustering_data_scaled <- scale(clustering_features)
rownames(clustering_data_scaled) <- pvz_df_analysis$player_id
```

## Determine Optimal Number of Clusters

```{r optimal-clusters}
#| label: optimal-clusters
#| fig.width: 10
#| fig.height: 6
#| cache: true

# Elbow method
fviz_nbclust(clustering_data_scaled, kmeans, method = "wss") +
  labs(title = "Elbow Method for Optimal k") +
  theme_minimal()

# Silhouette method
fviz_nbclust(clustering_data_scaled, kmeans, method = "silhouette") +
  labs(title = "Silhouette Method for Optimal k") +
  theme_minimal()

# Gap statistic method (this can be compute-intensive)
# fviz_nbclust(clustering_data_scaled, kmeans, method = "gap_stat", nboot = 50) +
#   labs(title = "Gap Statistic Method for Optimal k") +
#   theme_minimal()
```

## K-means Clustering

```{r kmeans-clustering}
#| label: kmeans-clustering
#| fig.width: 10
#| fig.height: 6
#| cache: true

# Perform k-means clustering with k=3
set.seed(42) # For reproducibility
k <- 3
kmeans_result <- kmeans(clustering_data_scaled, centers = k, nstart = 25)

# Add cluster assignments to the original data
pvz_df_clustered <- pvz_df_analysis %>%
  mutate(cluster = factor(kmeans_result$cluster))

# Create a function to generate descriptive cluster names
generate_cluster_names <- function(cluster_data) {
  # Calculate cluster centers for key variables
  centers <- cluster_data %>%
    group_by(cluster) %>%
    summarise(
      pvp_ratio = mean(ratio_pvp_hours),
      pve_ratio = mean(ratio_pve_hours),
      pvp_kd = mean(pvp_kd),
      pve_kd = mean(pve_kd),
      pvp_accuracy = mean(pvp_hit_accuracy),
      pve_accuracy = mean(pve_hit_accuracy),
      hours = mean(hours)
    )
  
  # Assign descriptive names based on the characteristics
  cluster_names <- character(nrow(centers))
  
  for (i in 1:nrow(centers)) {
    # Determine the primary play mode
    mode_ratios <- c(centers$pvp_ratio[i], centers$pve_ratio[i])
    mode_names <- c("PvP", "PvE")
    primary_mode <- mode_names[which.max(mode_ratios)]
    
    # Determine skill level based on KD and accuracy
    skill <- if ((centers$pvp_kd[i] > median(cluster_data$pvp_kd) && 
                 centers$pvp_accuracy[i] > median(cluster_data$pvp_hit_accuracy)) ||
                 (centers$pve_kd[i] > median(cluster_data$pve_kd) && 
                 centers$pve_accuracy[i] > median(cluster_data$pve_hit_accuracy))) {
      "High-skill"
    } else if ((centers$pvp_kd[i] < median(cluster_data$pvp_kd) && 
               centers$pvp_accuracy[i] < median(cluster_data$pvp_hit_accuracy)) ||
               (centers$pve_kd[i] < median(cluster_data$pve_kd) && 
               centers$pve_accuracy[i] < median(cluster_data$pve_hit_accuracy))) {
      "Casual"
    } else {
      "Mixed-skill"
    }
    
    # Determine playtime intensity
    playtime <- if (centers$hours[i] > median(cluster_data$hours)) {
      "Dedicated"
    } else {
      "Moderate"
    }
    
    # Combine characteristics into a descriptive name
    cluster_names[i] <- paste(playtime, primary_mode, skill, "Players")
  }
  
  return(data.frame(cluster = 1:nrow(centers), player_type = cluster_names))
}

# Generate descriptive names for clusters
cluster_names <- generate_cluster_names(pvz_df_clustered)

# Add descriptive names to the clustered data
pvz_df_clustered <- pvz_df_clustered %>%
  mutate(cluster_num = as.numeric(as.character(cluster))) %>%
  left_join(cluster_names, by = c("cluster_num" = "cluster")) %>%
  select(-cluster_num) %>%
  mutate(cluster = factor(cluster, labels = cluster_names$player_type))

# Save the clustered data for further analysis
write_csv(pvz_df_clustered, here("data/processed/pvz_with_clusters.csv"))
```

## Visualize Clusters

```{r visualize-clusters}
#| label: visualize-clusters
#| fig.width: 12
#| fig.height: 8

# Perform PCA for visualization
pca_result <- prcomp(clustering_data_scaled)
pca_data <- as.data.frame(pca_result$x[, 1:2])
pca_data$cluster <- pvz_df_clustered$cluster

# Create PCA biplot
pca_plot <- fviz_pca_biplot(pca_result, 
                  label = "var",
                  habillage = pvz_df_clustered$cluster,
                  addEllipses = TRUE,
                  ellipse.level = 0.95,
                  palette = "Set2",
                  title = "PCA of Player Clusters",
                  ggtheme = theme_minimal()) +
  labs(color = "Player Type")

# Display the plot
print(pca_plot)
```

## Cluster Characteristics

```{r cluster-characteristics}
#| label: cluster-characteristics
#| fig.width: 12
#| fig.height: 8

# Calculate mean values of key variables by cluster
cluster_profiles <- pvz_df_clustered %>%
  group_by(cluster) %>%
  summarise(
    n = n(),
    pct = n() / nrow(pvz_df_clustered) * 100,
    hours = mean(hours),
    ratio_pvp = mean(ratio_pvp_hours) * 100,
    ratio_pve = mean(ratio_pve_hours) * 100,
    pvp_kd = mean(pvp_kd),
    pve_kd = mean(pve_kd),
    pvp_accuracy = mean(pvp_hit_accuracy) * 100,
    pve_accuracy = mean(pve_hit_accuracy) * 100,
    pvp_damage = mean(pvp_average_damage),
    pve_damage = mean(pve_average_damage),
    pvp_score = mean(pvp_average_score),
    pve_score = mean(pve_average_score)
  )

# Display cluster profiles
kable(cluster_profiles, digits = 2,
      caption = "Profile Summary of Player Clusters")
```

```{r radar-charts}
# Prepare data for radar charts
radar_metrics <- c(
  "Total Hours", 
  # Place PvE metrics first (left side)
  "PvE Ratio", "PvE K/D", "PvE Accuracy", "PvE Score",
  # Place PvP metrics second (right side)
  "PvP Ratio", "PvP K/D", "PvP Accuracy", "PvP Score"
)

# Extract cluster names to use as chart titles
cluster_names <- levels(pvz_df_clustered$cluster)

# Add a z-score standardization function
# This scales values based on how many standard deviations they are from the mean
z_score_scale <- function(x) {
  # Check for zero standard deviation to avoid division by zero
  if (sd(x, na.rm = TRUE) == 0) return(rep(0, length(x)))
  return((x - mean(x, na.rm = TRUE)) / sd(x, na.rm = TRUE))
}

# Helper function to convert z-scores to a 0-1 scale for radar chart display
# Typically -3 to +3 z-scores cover 99.7% of data, so we'll use this as limits
z_to_radar_scale <- function(z, min_z = -3, max_z = 3) {
  # Clip extreme values to min_z and max_z
  z_clipped <- pmax(pmin(z, max_z), min_z)
  # Scale to 0-1 range
  scaled <- (z_clipped - min_z) / (max_z - min_z)
  return(scaled)
}

# Calculate z-scores for each metric across all players
pvz_metrics_z <- pvz_df_clustered %>%
  summarise(
    hours_z = list(z_score_scale(hours)),
    ratio_pve_hours_z = list(z_score_scale(ratio_pve_hours)),
    pve_kd_z = list(z_score_scale(pve_kd)),
    pve_hit_accuracy_z = list(z_score_scale(pve_hit_accuracy)),
    pve_average_score_z = list(z_score_scale(pve_average_score)),
    ratio_pvp_hours_z = list(z_score_scale(ratio_pvp_hours)),
    pvp_kd_z = list(z_score_scale(pvp_kd)),
    pvp_hit_accuracy_z = list(z_score_scale(pvp_hit_accuracy)),
    pvp_average_score_z = list(z_score_scale(pvp_average_score))
  )

# Calculate radar chart data for each cluster using the z-scores
radar_data <- pvz_df_clustered %>%
  group_by(cluster) %>%
  summarise(
    # Calculate mean z-score for each metric within this cluster
    hours_cluster_mean = mean(hours, na.rm = TRUE),
    ratio_pve_hours_cluster_mean = mean(ratio_pve_hours, na.rm = TRUE),
    pve_kd_cluster_mean = mean(pve_kd, na.rm = TRUE),
    pve_hit_accuracy_cluster_mean = mean(pve_hit_accuracy, na.rm = TRUE),
    pve_average_score_cluster_mean = mean(pve_average_score, na.rm = TRUE),
    ratio_pvp_hours_cluster_mean = mean(ratio_pvp_hours, na.rm = TRUE),
    pvp_kd_cluster_mean = mean(pvp_kd, na.rm = TRUE),
    pvp_hit_accuracy_cluster_mean = mean(pvp_hit_accuracy, na.rm = TRUE),
    pvp_average_score_cluster_mean = mean(pvp_average_score, na.rm = TRUE)
  ) %>%
  ungroup() %>%
  # Convert cluster means to z-scores
  mutate(
    `Total Hours` = z_to_radar_scale(z_score_scale(hours_cluster_mean)),
    `PvE Ratio` = z_to_radar_scale(z_score_scale(ratio_pve_hours_cluster_mean)),
    `PvE K/D` = z_to_radar_scale(z_score_scale(pve_kd_cluster_mean)),
    `PvE Accuracy` = z_to_radar_scale(z_score_scale(pve_hit_accuracy_cluster_mean)),
    `PvE Score` = z_to_radar_scale(z_score_scale(pve_average_score_cluster_mean)),
    `PvP Ratio` = z_to_radar_scale(z_score_scale(ratio_pvp_hours_cluster_mean)),
    `PvP K/D` = z_to_radar_scale(z_score_scale(pvp_kd_cluster_mean)),
    `PvP Accuracy` = z_to_radar_scale(z_score_scale(pvp_hit_accuracy_cluster_mean)),
    `PvP Score` = z_to_radar_scale(z_score_scale(pvp_average_score_cluster_mean))
  ) %>%
  # Drop the intermediate columns
  select(
    cluster,
    `Total Hours`,
    `PvE Ratio`, `PvE K/D`, `PvE Accuracy`, `PvE Score`,
    `PvP Ratio`, `PvP K/D`, `PvP Accuracy`, `PvP Score`
  )

# Add a helper function for enhanced radar charts
#| label: radar-chart-enhancements

# Function to create an enhanced radar chart
create_enhanced_radar <- function(cluster_data, cluster_color, cluster_title, actual_values) {
  # Get column names from the data
  data_cols <- colnames(cluster_data)
  
  # Create max and min rows with exactly matching column names
  max_min <- data.frame(matrix(c(rep(1, length(data_cols)), rep(0, length(data_cols))), 
                             nrow = 2, 
                             byrow = TRUE))
  # Set column names to match exactly
  colnames(max_min) <- data_cols
  
  # Combine max, min, and cluster data
  plot_data <- rbind(max_min, cluster_data)
  
  # Create the enhanced radar chart
  radarchart(
    plot_data, 
    pcol = cluster_color,
    pfcol = adjustcolor(cluster_color, alpha.f = 0.5),  # Increased opacity
    plwd = 3,  # Thicker lines
    vlcex = 0.9,  # Larger text
    title = cluster_title,
    cglcol = "grey80",
    cglty = 1,
    axislabcol = "grey30",
    caxislabels = seq(0, 1, 0.25),
    cglwd = 0.8,
    centerzero = FALSE,
    seg = 4,  # Fewer segments for cleaner look
    pty = 32  # More smooth polygon edges
  )
  
  # Add values as text at the bottom
  legend_text <- sprintf("Hours: %.1f, PvP K/D: %.2f, PvE K/D: %.2f", 
                         actual_values$`Total Hours`, 
                         actual_values$`PvP K/D`, 
                         actual_values$`PvE K/D`)
  
  mtext(legend_text, side = 1, line = 2, cex = 0.8, col = cluster_color)
}
```

```{r radar-chart-cluster1}
#| label: radar-chart-cluster1
#| fig.width: 8
#| fig.height: 7

# Create radar chart for the first cluster
if (length(cluster_names) >= 1) {
  # Extract data for this cluster
  cluster_data <- radar_data %>%
    filter(cluster == cluster_names[1]) %>%
    select(-cluster) %>%
    # Ensure consistent column order (PvE left, PvP right)
    select(
      `Total Hours`,
      # PvE metrics (left)
      `PvE Ratio`, `PvE K/D`, `PvE Accuracy`, `PvE Score`,
      # PvP metrics (right)
      `PvP Ratio`, `PvP K/D`, `PvP Accuracy`, `PvP Score`
    ) %>%
    as.data.frame()
  
  # Get actual values for reference
  actual_values <- pvz_df_clustered %>%
    filter(cluster == cluster_names[1]) %>%
    summarise(
      `Total Hours` = mean(hours),
      `PvP K/D` = mean(pvp_kd),
      `PvE K/D` = mean(pve_kd)
    )
  
  # Choose a color for this cluster
  cluster_color <- "#E41A1C"  # Red
  cluster_title <- paste("Cluster 1:", cluster_names[1])
  
  # Create the enhanced radar chart
  create_enhanced_radar(cluster_data, cluster_color, cluster_title, actual_values)
  
  # Display a note about scaling
  mtext("Note: Values shown as standardized scores (z-scores converted to 0-1 scale)", 
        side = 1, line = 4, cex = 0.8)
}
```

```{r radar-chart-cluster2}
#| label: radar-chart-cluster2
#| fig.width: 8
#| fig.height: 7

# Create radar chart for the second cluster
if (length(cluster_names) >= 2) {
  # Extract data for this cluster
  cluster_data <- radar_data %>%
    filter(cluster == cluster_names[2]) %>%
    select(-cluster) %>%
    # Ensure consistent column order (PvE left, PvP right)
    select(
      `Total Hours`,
      # PvE metrics (left)
      `PvE Ratio`, `PvE K/D`, `PvE Accuracy`, `PvE Score`,
      # PvP metrics (right)
      `PvP Ratio`, `PvP K/D`, `PvP Accuracy`, `PvP Score`
    ) %>%
    as.data.frame()
  
  # Get actual values for reference
  actual_values <- pvz_df_clustered %>%
    filter(cluster == cluster_names[2]) %>%
    summarise(
      `Total Hours` = mean(hours),
      `PvP K/D` = mean(pvp_kd),
      `PvE K/D` = mean(pve_kd)
    )
  
  # Choose a color for this cluster
  cluster_color <- "#377EB8"  # Blue
  cluster_title <- paste("Cluster 2:", cluster_names[2])
  
  # Create the enhanced radar chart
  create_enhanced_radar(cluster_data, cluster_color, cluster_title, actual_values)
  
  # Display a note about scaling
  mtext("Note: Values shown as standardized scores (z-scores converted to 0-1 scale)", 
        side = 1, line = 4, cex = 0.8)
}
```

```{r radar-chart-cluster3}
#| label: radar-chart-cluster3
#| fig.width: 8
#| fig.height: 7

# Create radar chart for the third cluster
if (length(cluster_names) >= 3) {
  # Extract data for this cluster
  cluster_data <- radar_data %>%
    filter(cluster == cluster_names[3]) %>%
    select(-cluster) %>%
    # Ensure consistent column order (PvE left, PvP right)
    select(
      `Total Hours`,
      # PvE metrics (left)
      `PvE Ratio`, `PvE K/D`, `PvE Accuracy`, `PvE Score`,
      # PvP metrics (right)
      `PvP Ratio`, `PvP K/D`, `PvP Accuracy`, `PvP Score`
    ) %>%
    as.data.frame()
  
  # Get actual values for reference
  actual_values <- pvz_df_clustered %>%
    filter(cluster == cluster_names[3]) %>%
    summarise(
      `Total Hours` = mean(hours),
      `PvP K/D` = mean(pvp_kd),
      `PvE K/D` = mean(pve_kd)
    )
  
  # Choose a color for this cluster
  cluster_color <- "#4DAF4A"  # Green
  cluster_title <- paste("Cluster 3:", cluster_names[3])
  
  # Create the enhanced radar chart
  create_enhanced_radar(cluster_data, cluster_color, cluster_title, actual_values)
  
  # Display a note about scaling
  mtext("Note: Values shown as standardized scores (z-scores converted to 0-1 scale)", 
        side = 1, line = 4, cex = 0.8)
}
```


### Comparison of Cluster Characteristics with Barplots

```{r cluster-bar-comparisons}
#| label: cluster-bar-comparisons
#| fig.width: 12
#| fig.height: 14

# Create a long format dataset for easier visualization
cluster_long <- pvz_df_clustered %>%
  group_by(cluster) %>%
  summarise(
    `Total Hours` = mean(hours),
    `PvP Ratio` = mean(ratio_pvp_hours) * 100,
    `PvE Ratio` = mean(ratio_pve_hours) * 100,
    `PvP K/D Ratio` = mean(pvp_kd),
    `PvE K/D Ratio` = mean(pve_kd),
    `PvP Accuracy` = mean(pvp_hit_accuracy) * 100,
    `PvE Accuracy` = mean(pve_hit_accuracy) * 100,
    `PvP Avg Score` = mean(pvp_average_score),
    `PvE Avg Score` = mean(pve_average_score)
  ) %>%
  pivot_longer(
    cols = -cluster,
    names_to = "Metric",
    values_to = "Value"
  )

# Group metrics for more organized plotting
cluster_long <- cluster_long %>%
  mutate(
    MetricGroup = case_when(
      Metric %in% c("Total Hours") ~ "Playtime",
      Metric %in% c("PvP Ratio", "PvE Ratio") ~ "Game Mode Preference",
      Metric %in% c("PvP K/D Ratio", "PvE K/D Ratio") ~ "Combat Effectiveness",
      Metric %in% c("PvP Accuracy", "PvE Accuracy") ~ "Accuracy",
      Metric %in% c("PvP Avg Score", "PvE Avg Score") ~ "Performance",
      TRUE ~ "Other"
    )
  )

# Create grouped bar charts for each metric category
ggplot(cluster_long, aes(x = Metric, y = Value, fill = cluster)) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.9)) +
  facet_wrap(~ MetricGroup, scales = "free", ncol = 2) +
  scale_fill_brewer(palette = "Set2") +
  labs(
    title = "Comparison of Cluster Characteristics",
    x = NULL,
    y = NULL,
    fill = "Player Type"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    strip.background = element_rect(fill = "grey90"),
    strip.text = element_text(face = "bold")
  )
```

# Relationship Between Playtime and Well-Being

## Overall Well-Being by Player Type

```{r wellbeing-by-cluster}
#| label: wellbeing-by-cluster
#| fig.width: 10
#| fig.height: 6

# Compare well-being metrics across clusters
wellbeing_by_cluster <- pvz_df_clustered %>%
  group_by(cluster) %>%
  summarise(
    spane_balance = mean(spane_balance),
    spane_positive = mean(spane_positive),
    spane_negative = mean(spane_negative),
    spane_game_balance = mean(spane_game_balance),
    autonomy = mean(autonomy),
    competence = mean(competence),
    relatedness = mean(relatedness),
    enjoyment = mean(enjoyment),
    extrinsic = mean(extrinsic),
    n = n()
  )

# Display the summary table
kable(wellbeing_by_cluster, digits = 2,
      caption = "Well-being metrics by player cluster")

# Create boxplots for SPANE balance by cluster
ggplot(pvz_df_clustered, aes(x = cluster, y = spane_balance, fill = cluster)) +
  geom_boxplot(alpha = 0.7) +
  scale_fill_brewer(palette = "Set2") +
  labs(
    title = "Overall Well-Being (SPANE Balance) by Player Type",
    x = "Player Type",
    y = "SPANE Balance Score"
  ) +
  theme_minimal() +
  theme(
    legend.position = "none",
    axis.text.x = element_text(angle = 45, hjust = 1)
  )

# Create boxplots for game-related well-being by cluster
ggplot(pvz_df_clustered, aes(x = cluster, y = spane_game_balance, fill = cluster)) +
  geom_boxplot(alpha = 0.7) +
  scale_fill_brewer(palette = "Set2") +
  labs(
    title = "Game-Related Well-Being by Player Type",
    x = "Player Type",
    y = "Game SPANE Balance Score"
  ) +
  theme_minimal() +
  theme(
    legend.position = "none",
    axis.text.x = element_text(angle = 45, hjust = 1)
  )
```

## Playtime and Well-Being Relationships

```{r playtime-wellbeing}
#| label: playtime-wellbeing
#| fig.width: 12
#| fig.height: 8

# Create scatter plots for total playtime vs. well-being
p1 <- ggplot(pvz_df_clustered, aes(x = hours, y = spane_balance, color = cluster)) +
  geom_point(alpha = 0.6) +
  geom_smooth(method = "lm", se = TRUE, alpha = 0.2) +
  scale_color_brewer(palette = "Set2") +
  labs(
    title = "Total Playtime vs. Overall Well-Being",
    x = "Total Hours Played",
    y = "SPANE Balance Score",
    color = "Player Type"
  ) +
  theme_minimal()

p2 <- ggplot(pvz_df_clustered, aes(x = hours, y = spane_game_balance, color = cluster)) +
  geom_point(alpha = 0.6) +
  geom_smooth(method = "lm", se = TRUE, alpha = 0.2) +
  scale_color_brewer(palette = "Set2") +
  labs(
    title = "Total Playtime vs. Game-Related Well-Being",
    x = "Total Hours Played",
    y = "Game SPANE Balance Score",
    color = "Player Type"
  ) +
  theme_minimal()

# Combine the plots
ggarrange(p1, p2, ncol = 2, common.legend = TRUE, legend = "right")
```

```{r gamemode-wellbeing}
#| label: gamemode-wellbeing
#| fig.width: 12
#| fig.height: 10

# Create scatter plots for game mode playtime vs. well-being
# Since we don't have direct hour measurements for each game mode,
# we'll use the ratio variables multiplied by total hours

# Calculate estimated hours for each mode
pvz_df_clustered <- pvz_df_clustered %>%
  mutate(
    est_pvp_hours = hours * ratio_pvp_hours,
    est_pve_hours = hours * ratio_pve_hours
  )

p1 <- ggplot(pvz_df_clustered, aes(x = est_pvp_hours, y = spane_balance, color = cluster)) +
  geom_point(alpha = 0.6) +
  geom_smooth(method = "lm", se = TRUE, alpha = 0.2) +
  scale_color_brewer(palette = "Set2") +
  labs(
    title = "PvP Playtime vs. Overall Well-Being",
    x = "Estimated PvP Hours Played",
    y = "SPANE Balance Score",
    color = "Player Type"
  ) +
  theme_minimal()

p2 <- ggplot(pvz_df_clustered, aes(x = est_pve_hours, y = spane_balance, color = cluster)) +
  geom_point(alpha = 0.6) +
  geom_smooth(method = "lm", se = TRUE, alpha = 0.2) +
  scale_color_brewer(palette = "Set2") +
  labs(
    title = "PvE Playtime vs. Overall Well-Being",
    x = "Estimated PvE Hours Played",
    y = "SPANE Balance Score",
    color = "Player Type"
  ) +
  theme_minimal()

# Combine the plots - using just two plots since we don't have social hours data
ggarrange(p1, p2, ncol = 2, nrow = 1, common.legend = TRUE, legend = "right")
```

## Motivation and Well-Being by Player Type

```{r motivation-by-cluster}
#| label: motivation-by-cluster
#| fig.width: 12
#| fig.height: 8

# Prepare data for motivation visualization
motivation_data <- pvz_df_clustered %>%
  select(player_id, cluster, autonomy, competence, relatedness, enjoyment, extrinsic) %>%
  pivot_longer(
    cols = c(autonomy, competence, relatedness, enjoyment, extrinsic),
    names_to = "motivation_type",
    values_to = "score"
  ) %>%
  mutate(
    motivation_type = factor(
      motivation_type, 
      levels = c("autonomy", "competence", "relatedness", "enjoyment", "extrinsic"),
      labels = c("Autonomy", "Competence", "Relatedness", "Enjoyment", "Extrinsic")
    )
  )

# Create bar chart for motivation by player type
ggplot(motivation_data, aes(x = motivation_type, y = score, fill = cluster)) +
  geom_bar(stat = "summary", fun = "mean", position = "dodge") +
  scale_fill_brewer(palette = "Set2") +
  labs(
    title = "Motivation Factors by Player Type",
    x = "Motivation Type",
    y = "Mean Score",
    fill = "Player Type"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

# Discussion and Conclusion

This analysis has identified distinct player types in Plants vs. Zombies: Battle for Neighborville based on playstyle metrics, including game mode preference, playtime intensity, and performance measures. The three player clusters exhibit different patterns of well-being and motivation, suggesting that the relationship between gameplay and psychological outcomes varies by play style.

Key findings include:

1.  Players show distinct preferences for different game modes (PvP, PvE, and Social), with varying levels of skill and time investment.

2.  The identified player types show different relationships between playtime and well-being, suggesting that how players engage with the game matters more than simply how much they play.

3.  Motivation factors such as autonomy, competence, and enjoyment vary across player types, indicating different psychological needs being met through gameplay.

4.  Game-related well-being tends to be more positive than general well-being across all player types, suggesting that the game provides a positive experience regardless of playstyle.

These findings have implications for game design, suggesting that catering to different playstyles and motivations can enhance player well-being. Future research could explore how these patterns evolve over longer periods and how game design changes might impact different player types.

```{r save-results}
#| label: save-results
#| include: false

# Final cleanup and save session info for reproducibility
sessionInfo()
```