---
title: "Player Playstyle Analysis (3 Clusters)"
author: "Thomas Hakman"
format: 
  html:
    toc: true
    toc-depth: 3
    code-fold: true
    theme: flatly
    fig-width: 8
    fig-height: 6
    embed-resources: true    # Bundle all resources in the HTML file
    self-contained: true     # Create a standalone HTML file
execute:
  echo: false
  message: false
  warning: false
  cache: false
  eval: true                 # Ensure code is evaluated
  output: true               # Ensure output is displayed
  fig-show: "asis"           # Ensure figures are displayed inline
editor: visual
bibliography: bibliography.bib
---

```{r setup}
#| label: setup
#| include: false

# Set global options
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  cache = FALSE,
  fig.show = "asis"  # Ensures figures are displayed inline in the viewer
)

# Basic httpgd setup for VSCode (if available)
if (interactive() && requireNamespace("httpgd", quietly = TRUE)) {
  library(httpgd)
  if (!httpgd::hgd_is_active()) {
    httpgd::hgd_open()
  }
  options(vsc.plot = FALSE)  # Configure VSCode to use httpgd
  options(device = function() httpgd::hgd())
}

# Define a capitalize function for facet labels
capitalize <- function(x) {
  x <- as.character(x)
  substr(x, 1, 1) <- toupper(substr(x, 1, 1))
  x
}

# Create a list of wellbeing variables (focus on our two primary outcomes)
wellbeing_vars <- c("spane_game_balance", "enjoyment")
```

# Abstract

This document presents a comprehensive analysis of playstyle patterns and their relationship with well-being among players of Plants vs. Zombies: Battle for Neighborville. Using cluster analysis, we identified three distinct player types based on their gameplay behavior and preferences. Through robust linear mixed-effects modeling, we examined how these playstyles moderate the relationship between gameplay time and well-being and enjoyment. Our findings reveal that the relationship between playtime and well-being and enjoyment differ significantly across player types, suggesting that how players engage with the game is potentially more important than simply how much they play.

# Introduction

The literature on whether video games harm or improve well-being remains divided. Several studies report negative associations between playtime and measures of mental health or well-being [@fergusonLackConsensusScholars2019; @liuDigitalCommunicationMedia2019; @mathurFindingCommonGround2019]. Other studies find negative associations only among highly engaged players [@coldercarrasVideoGamingHyperconnected2017; @przybylski2017a]. Yet others, find null or negligible associations [@johannesVideoGamePlay2021; @vuorreTimeSpentPlaying2022]. The positive aspects are similarly mixed. Some studies find positive associations between playtime and well-being [@halbrookWhenHowVideo2019; @johannesVideoGamePlay2021; @villaniVideogamesEmotionRegulation2018], whereas others report them just for moderate usage [@przybylski2017a] or specific game genres such as MMOs [@raithMassivelyMultiplayerOnline2021]. Overall, existing evidence suggests that the relations between video games and well-being are complex, small, and potentially bidirectional. However, the inconsistency of these findings has prevented any firm conclusions from being drawn. This lack of consensus has led scholars to question the value of time spent playing games as a factor meriting study in its own right [@orbenOnlyHolisticIterative2020]

Most research on the well-being effects of video games applies what could be called a *playtime paradigm*, measuring how long a player engages with the video game, whilst overlooking, *how* actual play behaviour that is enacted by participants. Whilst the concept of playtime is helpful to ascertain more clearly how extreme cases of over- and under-usage may impact differentially on well-being, it gives researchers no insight into the myriad actions, behaviours, and feelings that players perform and experience. Hence, it has become increasingly clear that playtime as a unidimensional or homogenous construct has less validity than once thought [@kayeConceptualMethodologicalMayhem2020] and might not be informative on its own when it comes to understanding genuine player behaviour and well-being outcomes [@granicScreenTimeIdentity2020]. This playtime paradigm is problematic, as it oversimplifies a complex phenomenon by failing to recognize that video games are multimodal virtual worlds that offer a wide variety of different experiences to each player [@kayeConceptualMethodologicalMayhem2020; @orben2019a]. Subsuming all these activities and experiences under one simple count of frequency and duration fails to acknowledge that they serve different functions and possibly show different effects [@dienlinImpactDigitalTechnology2020]. Since, this limitation calls the previous studies' marginal findings into question, there is a strong need to start considering player-specific differences in in-game play behaviour and performance such as behavioural profiles or play styles.

Playing styles can be best thought of as collections of in-game events and features that may be interpreted to provide concise descriptions of the general way in which players behave [\[AP3\]](#_msocom_3) in a game to achieve the game's objectives [@ravariInvestigatingRelationPlaying2022]. Playing styles have also been referred to as behavioural profiles [@drachenGunsSwordsData2012; @drachen2014a], and have been studied in a variety of contexts. Several studies have investigated the relationship between playing style and personality [@beanVideoGamersPersonality2016; @vanlankveldGamesPersonalityProfiling2011], playing style and motivation [@melhartYourGameplaySays2019; @yeeGamerMotivationProfile2016], and playing style and national culture [@ravariInvestigatingRelationPlaying2022]. However, the effects of playing style on well-being have been studied very little. Especially, the extent to which differences in players' playing styles influence the association between playtime and subjective well-being remains unknown. To provide insight into these player-specific effects, this study will link playing styles based on naturalistic in-game play behaviour, with objective measures of playtime, and self-reported measures of well-being to: (1) demonstrate how playing styles can be modelled from in-game behavioural data (i.e., telemetry data), (2) show how playing styles relate to well-being, and (3) determine the extent to which playing styles moderate the relationship between playtime and subjective well-being.

**Clustering playing styles from behavioural telemetry data**

Assessing in-game play behaviour objectively can be achieved with the analysis of behavioural telemetry data (also known as digital trace data), which is the second-by-second logged behaviour of players in a game [@el-nasrGameDataScience2021]. Yet, the general inaccessibility of this data has proven to be a major hurdle that has stalled most research efforts. Hence, a committee report on immersive and addictive technologies presented to the UK parliament by the Secretary of State for Digital, Culture, Media and Sport recommended that the government "require games companies to share aggregated player data with researchers" [@britain2020a, p. 2]. Nevertheless, recent collaborations between academics and industry partners are starting to overcome this limitation by increasingly making behavioural telemetry data available for independent scientific analysis [@johannesVideoGamePlay2021; @vuorreTimeSpentPlaying2022]. The availability of high-dimensional telemetry data in combination with unsupervised clustering techniques allows us to identify playing styles based on in-game player behaviour and performance. Establishing playing styles from telemetry data can be achieved through telemetry-driven profiling. Telemetry-driven profiling focuses on condensing varied, high-dimensional, and high-volume data into descriptions that highlight underlying patterns in the behaviour of the player base [@drachen2014a; @santhoshTelemetryAnalyticsBest2013]. These playing styles provide researchers with a "quantifiable way to understand who the players are and how they play the game or games being investigated" [@sifaProfilingGamesUnderstanding2018, p. 337].

In this study we address the inconsistency of published research linking video game play time to psychological outcomes introducing playing styles as an explanatory factor which might reconcile earlier observations of positive, negative, and null associations linking time spent playing games to player well-being. To this end, we introduce a data-driven player modelling approach to explore the unknown association and potential moderation between how much players play, how players play—as manifested through their behavioural gameplay data— and their subjective well-being. In line with this we investigate three research questions: First, can distinct playing styles be identified from behavioural telemetry data? Second, do groups of players with distinct playing styles report different levels of subjective well-being? And finally, do playing styles moderate the relationship between playtime and subjective well-being?

[\[AP3\]](#_msoanchor_3)I would lead here with Richard Bartle.

<https://mud.co.uk/richard/hcds.htm>

<https://mud.co.uk/richard/DesigningVirtualWorlds.pdf>


```{r install-packages, include=FALSE}
#| label: install-packages

# First ensure pacman is installed if not already
if (!requireNamespace("pacman", quietly = TRUE)) {
  install.packages("pacman")
}

# Use pacman to load all required packages (will install any missing packages)
pacman::p_load(
  tidyverse,  # For data manipulation and visualization
  here,       # For project-relative paths
  lubridate,  # For date/time manipulation
  scales,     # For scale transformations
  janitor,    # For data cleaning
  magrittr,   # For additional pipe operators
  psych,      # For psychological statistics
  corrplot,   # For correlation plots
  ggpubr,     # For publication-ready plots
  factoextra, # For cluster analysis visualization
  cluster,    # For clustering algorithms
  fmsb,       # For radar charts
  knitr,      # For dynamic report generation
  tibble,     # For improved data frames
  kableExtra, # For enhanced table formatting
  fpc,        # For cluster validation
  gridExtra,  # For arranging multiple plots
  hopkins,    # For cluster tendency assessment
  effectsize, # For effect size calculations
  robustlmm,  # For robust linear mixed-effects models
  broom,      # For tidy model outputs
  broom.mixed, # For tidying mixed-effects model output
  patchwork,  # For combining plots
  ggeffects,  # For visualizing regression model predictions
  grid        # For grid graphics (needed for tableGrob)
)
```

```{r helper-functions}
#| label: helper-functions

# Function to scale values to 0-1 range for radar charts
scale_to_01 <- function(x, min_val = min(x, na.rm = TRUE), max_val = max(x, na.rm = TRUE)) {
  if (max_val == min_val) return(rep(0.5, length(x)))
  (x - min_val) / (max_val - min_val)
}

```

```{r statistical-functions}
#| label: statistical-functions

# Function to run ANOVA and extract results
run_anova <- function(dv) {
  formula <- as.formula(paste(dv, "~ player_type"))
  model <- aov(formula, data = pvz_data)
  result <- broom::tidy(model)
  result$eta_squared <- effectsize::eta_squared(model)[["Eta2"]]
  result$variable <- dv
  return(result)
}

# Function to compute Hopkins statistic for cluster tendency
compute_hopkins <- function(data, m = nrow(data)-1) {
  # Standardize data for clustering
  data_scaled <- scale(data)
  
  # Compute Hopkins statistic
  hopkins_result <- hopkins::hopkins(data_scaled, m = m)
  
  return(hopkins_result)
}
```

```{r load-processed-data}
#| label: load-processed-data

# Load the dataset with outliers removed
pvz_data <- read_csv(here("data/processed/pvz_clean_no_outliers.csv"))

# Create a copy of pvz_data for analysis to maintain consistency
pvz_df_analysis <- pvz_data

```

```{r select-columns}
#| label: select-columns

pvz_df_analysis <- pvz_df_analysis %>%
  select(
    # Demographic variables
    player_id, country, age, gender, played_with_others,
    
    # Well-being variables - focus only on our two key outcomes
    spane_game_balance, enjoyment,
    
    # Total playtime
    hours,
    
    # PVP metrics
    ratio_pvp_hours,
    pvp_kd,
    pvp_hit_accuracy,
    pvp_dps,
    pvp_average_damage,
    pvp_average_score,
    
    # PVE metrics
    ratio_pve_hours,
    pve_kd,
    pve_hit_accuracy,
    pve_dps,
    pve_average_damage,
    pve_average_score
  )
```

```{r telemetry-data-filter}
#| label: telemetry-data-filter
#| include: false

# Filter for players with valid performance metrics (> 0)
pvz_df_analysis <- pvz_df_analysis %>%
  filter(
    ratio_pvp_hours > 0 | ratio_pve_hours > 0 | pvp_kd > 0 | pve_kd > 0 | pvp_hit_accuracy > 0 | pve_hit_accuracy > 0 | pvp_dps > 0 | pve_dps > 0 | pvp_average_damage > 0 | pve_average_damage > 0 | pvp_average_score > 0 | pve_average_score > 0
  )

# Remove players with less than 0.25 hours (15 minutes) of total playtime
pvz_df_analysis <- pvz_df_analysis %>%
  filter(hours >= 0.25)

# Remove any rows with NA values in key analysis variables
pvz_df_analysis <- pvz_df_analysis %>%
  filter(!is.na(ratio_pvp_hours) & !is.na(ratio_pve_hours) & 
         !is.na(pvp_kd) & !is.na(pve_kd) & 
         !is.na(pvp_hit_accuracy) & !is.na(pve_hit_accuracy) &
         !is.na(pvp_dps) & !is.na(pve_dps) &
         !is.na(pvp_average_damage) & !is.na(pve_average_damage) &
         !is.na(pvp_average_score) & !is.na(pve_average_score))

# Show the number of players left
print(paste("Number of players left after filtering:", nrow(pvz_df_analysis)))
```

# Methods

In this study we conducted a secondary analysis of the *Plants versus Zombies: Battle for Neighbourville* (PvZ) data gathered originally reported in @johannesVideoGamePlay2021. As detailed in their paper, they collaborated with Electronic Arts to collect objective game telemetry data and player survey responses. Here we will discuss the subset of the data used for this study. For details on the entire data which includes *Animal Crossing: New Horizons*, see Johannes and colleagues (2020).

## Participants and procedure

```{r survey-demographics-calculation}
#| label: survey-demographics-calculation
#| include: false

# Calculate demographics for the filtered dataset
demographics <- pvz_df_analysis %>%
  summarize(
    total_participants = n(),
    avg_age = round(mean(age, na.rm = TRUE), 1),
    sd_age = round(sd(age, na.rm = TRUE), 1)
  )

# Calculate gender distribution in filtered dataset
gender_distribution <- pvz_df_analysis %>%
  count(gender) %>%
  mutate(percentage = n / sum(n) * 100)

# Count participants with each gender
male_count <- gender_distribution %>% filter(gender == "Male") %>% pull(n)
female_count <- gender_distribution %>% filter(gender == "Female") %>% pull(n)
other_count <- gender_distribution %>% filter(gender == "Other") %>% pull(n)
undisclosed_count <- gender_distribution %>% filter(gender == "Prefer not to say") %>% pull(n)
```

```{r final-sample-demographics}
#| label: final-sample-demographics

# Display demographics of final sample
cat("Demographics of final sample:
")
cat("- Sample size:", nrow(pvz_df_analysis), "
")
cat("- Average age:", round(mean(pvz_df_analysis$age, na.rm = TRUE), 1), "(SD =", round(sd(pvz_df_analysis$age, na.rm = TRUE), 1), ")
")

# Gender distribution
gender_dist <- pvz_df_analysis %>%
  count(gender) %>%
  mutate(percentage = n / sum(n) * 100)

kable(gender_dist, caption = "Gender distribution in final sample", digits = 2)

# Exploratory Data Analysis
```

As detailed in @johannesVideoGamePlay2021, participant recruitment happened in two waves, first in early August 2020 and second in late September 2020. The response rate was \~ 0.21%. In total 518 PvZ players participated in the well-being survey (404 male, 94 female, 2 other, and 17 undisclosed) with an average age of 35 (SD = 12). However, after filtering for players with valid performance metrics and those with at least 15 minutes of total playtime, our final sample consists of `r nrow(pvz_df_analysis)` participants. The demographic characteristics of this filtered sample include an average age of `r demographics$avg_age` (SD = `r demographics$sd_age`). The gender distribution includes `r male_count` male, `r female_count` female, `r other_count` other, and `r undisclosed_count` participants who preferred not to disclose their gender.

## Measures

### Playtime

```{r session-timeline-helper}
#| label: session-timeline-helper
#| include: false

# Function to plot session timeline for a player
plot_player_sessions <- function(player_data, player_id = NULL, player_label = "Player") {
  # If a player_id is provided, filter data for that player
  if (!is.null(player_id)) {
    player_data <- player_data %>%
      filter(player_id == !!player_id)
  }
  
  # Convert timestamps to POSIXct if they're not already
  if (!inherits(player_data$game_start_time, "POSIXct")) {
    player_data$game_start_time <- as.POSIXct(player_data$game_start_time, format = "%Y-%m-%d %H:%M:%S")
    player_data$game_end_time <- as.POSIXct(player_data$game_end_time, format = "%Y-%m-%d %H:%M:%S")
  }
  
  # Extract day of week, date, and time component
  player_data$day_of_week <- weekdays(player_data$game_start_time)
  player_data$date <- as.Date(player_data$game_start_time)
  player_data$start_time_only <- as.POSIXct(format(player_data$game_start_time, format = "%H:%M:%S"), format = "%H:%M:%S")
  player_data$end_time_only <- as.POSIXct(format(player_data$game_end_time, format = "%H:%M:%S"), format = "%H:%M:%S")
  
  # Identify sessions that cross midnight (where end_time is earlier than start_time)
  player_data$crosses_midnight <- player_data$end_time_only < player_data$start_time_only
  
  # Create a new dataframe that will include split sessions for those crossing midnight
  player_data_expanded <- player_data %>%
    filter(!crosses_midnight) # Keep all sessions that don't cross midnight
  
  # Handle sessions that cross midnight by splitting them
  midnight_sessions <- player_data %>%
    filter(crosses_midnight)
  
  if (nrow(midnight_sessions) > 0) {
    # For each session crossing midnight, create two rows:
    # 1. From start time to 23:59:59
    # 2. From 00:00:00 to end time
    midnight <- as.POSIXct(strptime("23:59:59", format = "%H:%M:%S"), format = "%H:%M:%S")
    midnight_start <- as.POSIXct(strptime("00:00:00", format = "%H:%M:%S"), format = "%H:%M:%S")
    
    # First part: from start time to midnight
    midnight_part1 <- midnight_sessions
    midnight_part1$end_time_only <- midnight
    
    # Second part: from midnight start to end time
    midnight_part2 <- midnight_sessions
    midnight_part2$start_time_only <- midnight_start
    # The second part is on the next day of the week and the next date
    midnight_part2$day_of_week <- factor(
      weekdays(midnight_part2$game_start_time + 86400), # Add one day (86400 seconds)
      levels = c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday")
    )
    midnight_part2$date <- midnight_part2$date + 1
    
    # Add both parts to the expanded dataset
    player_data_expanded <- bind_rows(player_data_expanded, midnight_part1, midnight_part2)
  }
  
  # Recode match_type based on game_level_key 
  player_data_expanded$match_type <- "0"  # Initialize with 0 as in the original code
  
  # Social matches
  player_data_expanded$match_type[player_data_expanded$game_level_key == "dsub_social_terrain_beach"] <- "social match"
  player_data_expanded$match_type[player_data_expanded$game_level_key == "dsub_social_terrain_summer"] <- "social match"
  player_data_expanded$match_type[player_data_expanded$game_level_key == "dsub_socialspace"] <- "social match"
  
  # PVE matches
  player_data_expanded$match_type[player_data_expanded$game_level_key == "dsub_region_cheesemines"] <- "pve match"
  player_data_expanded$match_type[player_data_expanded$game_level_key == "dsub_region_dreadwood"] <- "pve match"
  player_data_expanded$match_type[player_data_expanded$game_level_key == "dsub_region_towncenter"] <- "pve match" 
  
  # PVP matches (Rush)
  player_data_expanded$match_type[player_data_expanded$game_level_key == "dsub_rush_dreadwood"] <- "pvp match (rush)"
  player_data_expanded$match_type[player_data_expanded$game_level_key == "dsub_rush_towncenter"] <- "pvp match (rush)"
  player_data_expanded$match_type[player_data_expanded$game_level_key == "dsub_rush_farm"] <- "pvp match (rush)"
  player_data_expanded$match_type[player_data_expanded$game_level_key == "dsub_rush_cheesemines"] <- "pvp match (rush)"
  player_data_expanded$match_type[player_data_expanded$game_level_key == "dsub_rush_shoreline"] <- "pvp match (rush)"
  player_data_expanded$match_type[player_data_expanded$game_level_key == "dsub_rush_resort"] <- "pvp match (rush)"
  
  # PVP matches (Vanquish)
  player_data_expanded$match_type[player_data_expanded$game_level_key == "dsub_tvm_cheesemines"] <- "pvp match (vanquish)"
  player_data_expanded$match_type[player_data_expanded$game_level_key == "dsub_tvm_zrefinery"] <- "pvp match (vanquish)"
  player_data_expanded$match_type[player_data_expanded$game_level_key == "dsub_tvm_suburbs"] <- "pvp match (vanquish)"
  player_data_expanded$match_type[player_data_expanded$game_level_key == "dsub_tvm_dreadwood"] <- "pvp match (vanquish)"
  
  # PVP matches (Arena)
  player_data_expanded$match_type[player_data_expanded$game_level_key == "dsub_4v4_battlearena_v2"] <- "pvp match (arena)"
  player_data_expanded$match_type[player_data_expanded$game_level_key == "dsub_4v4_superarena"] <- "pvp match (arena)"
  
  # Podium celebration
  player_data_expanded$match_type[player_data_expanded$game_level_key == "dsub_podiumcelebration"] <- "podium celebration"
  
  # Ensure day of week is a factor with ordered levels
  player_data_expanded$day_of_week <- factor(
    player_data_expanded$day_of_week,
    levels = c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday")
  )
  
  # Create a position adjustment for stacking sessions by different dates for the same day of week
  player_data_expanded <- player_data_expanded %>%
    group_by(day_of_week) %>%
    mutate(
      # Get unique dates for this day of week and assign a position index to each date
      unique_dates = list(unique(date)),
      date_index = match(date, unique(date)),
      # Scale the vertical offset to be small but noticeable
      y_offset = (date_index - 1) * 0.2,  # Creates a small vertical offset for different dates
      # Create adjusted y-position that includes the offset
      y_position = as.numeric(day_of_week) + y_offset
    ) %>%
    ungroup()
  
  # Sort by day of week, date, and start time
  player_data_expanded <- player_data_expanded %>%
    arrange(day_of_week, date, game_start_time, game_end_time)
  
  # Add a session ID within each day to distinguish multiple sessions on same day
  player_data_expanded <- player_data_expanded %>%
    group_by(day_of_week, date) %>%
    mutate(session_id_within_day = row_number()) %>%
    ungroup()
  
  # Convert match_type to factor with proper levels for consistent ordering
  player_data_expanded$match_type <- factor(
    player_data_expanded$match_type,
    levels = c("social match", "pve match", "pvp match (rush)", "pvp match (vanquish)", "pvp match (arena)", "podium celebration", "0"),
    labels = c("Social Lobby", "PvE Match", "PvP (Rush)", "PvP (Vanquish)", "PvP (Arena)", "Podium Celebration", "Other")
  )
  
  # Create the plot with enhanced styling
  plot <- player_data_expanded %>%
    ggplot2::ggplot(aes(start_time_only, y_position, col = match_type, 
                        group = interaction(day_of_week, date, session_id_within_day))) +
    # Use improved color palette with better contrast
    ggplot2::scale_color_manual(
      name = "Match Type", 
      values = c(
        "Social Lobby" = "#00BFC4",       # Teal
        "PvE Match" = "#7CAE00",          # Green
        "PvP (Rush)" = "#F8766D",         # Red
        "PvP (Vanquish)" = "#FF9E1C",     # Orange
        "PvP (Arena)" = "#FFD700",        # Gold
        "Podium Celebration" = "#C77CFF", # Purple
        "Other" = "#A0A0A0"               # Gray
      )
    ) +
    # Thicker segments with rounded ends for better visibility
    ggplot2::geom_segment(
      aes(xend = end_time_only, yend = y_position),
      linewidth = 2,
      lineend = "round"
    ) +
    # Add points at start/end for better visibility of short sessions
    ggplot2::geom_point(size = 1.5, alpha = 0.6) +
    ggplot2::geom_point(aes(x = end_time_only), size = 1.5, alpha = 0.6) +
    # Better time axis formatting - using fixed limits for consistency
    ggplot2::scale_x_datetime(
      labels = scales::date_format("%H:%M"),
      breaks = scales::date_breaks("2 hours"),
      limits = c(
        as.POSIXct(strptime("00:00", format = "%H:%M")),
        as.POSIXct(strptime("23:59", format = "%H:%M"))
      ),
      expand = c(0.01, 0.01),
      date_minor_breaks = "1 hour"
    ) +
    # Use the original day of week as the y-axis labels
    ggplot2::scale_y_continuous(
      breaks = 1:7,
      labels = levels(player_data_expanded$day_of_week),
      limits = c(0.5, 7.5)
    ) +
    # Empty labs - axis labels will be added to the combined figure instead
    ggplot2::labs(
      x = NULL, 
      y = NULL
    ) +
    # Improved theme with better typography and layout
    ggplot2::theme_minimal(base_size = 11, base_family = "sans") +
    ggplot2::theme(
      legend.position = "none", # Move legend to the combined figure
      plot.margin = margin(30, 10, 30, 10), # Increased margins for better separation
      
      axis.text.y = element_text(face = "bold"),
      axis.title = element_text(face = "bold"),
      
      panel.grid.major.x = element_line(color = "grey90"),
      panel.grid.minor.x = element_line(color = "grey95"),
      panel.grid.major.y = element_line(color = "grey90"),
      panel.grid.minor.y = element_blank(),
      
      panel.background = element_rect(fill = "white", color = NA),
      plot.background = element_rect(fill = "white", color = NA)
    )
  
  return(plot)
}
```

```{r session-timeline-data-prep}
#| label: session-timeline-data-prep
#| include: false
#| cache: false

# Load game session data if not already included
game_sessions <- read_csv(here("data/processed/game_time_2week.csv"))

# Make sure timestamps are in the correct format
game_sessions$game_start_time <- as.POSIXct(game_sessions$game_start_time, format = "%Y-%m-%d %H:%M:%S")
game_sessions$game_end_time <- as.POSIXct(game_sessions$game_end_time, format = "%Y-%m-%d %H:%M:%S")

# First, get a list of player IDs in the data
player_ids <- unique(game_sessions$player_id)

# Choose specific indices
# If those indices don't exist in our data, we'll use the first three players
player_indices <- c(
  min(401, length(player_ids)),  # Equivalent to slice(401) 
  min(6, length(player_ids)),    # Equivalent to slice(6)
  min(3, length(player_ids))     # Equivalent to slice(3)
)

selected_player_ids <- player_ids[player_indices]
```

```{r create-individual-plots}
#| label: create-individual-plots
#| include: false
#| cache: false
#| dependson: session-timeline-data-prep

# Create individual plots for each player with standardized x-axis
plot1 <- plot_player_sessions(
  game_sessions %>% filter(player_id == selected_player_ids[1])
)

plot2 <- plot_player_sessions(
  game_sessions %>% filter(player_id == selected_player_ids[2])
)

# For the third plot, use the same time range for consistency
plot3 <- plot_player_sessions(
  game_sessions %>% filter(player_id == selected_player_ids[3])
)
```

```{r combine-session-plots}
#| label: combine-session-plots
#| fig.width: 12
#| fig.height: 16
#| dependson: create-individual-plots

# Combine plots into a single figure with enhanced styling and proper spacing
figure <- ggarrange(
  plot2, plot3, plot1,
  labels = c("Player A", "Player B", "Player C"),
  font.label = list(size = 14, face = "bold", color = "black"),
  ncol = 1, 
  nrow = 3, 
  common.legend = TRUE, 
  legend = "bottom",
  heights = c(1, 1, 1),
  align = "v",
  hjust = -0.1
)

# Add shared x and y labels to the combined figure
fig_sessiontime <- annotate_figure(figure,
                        bottom = text_grob("Time of Day (hours)", 
                                          face = "bold", size = 12),
                        left = text_grob("Day of Week", 
                                         face = "bold", size = 12, rot = 90))

# Display the combined figure
fig_sessiontime
```

Calculating playtime in video games is complex. Simply measuring the total time an account has been online does not distinguish between active and idle time. Idle time is when the player's account is online but not actively engaging with the objectives or mechanics of the game. Idle time can happen when a player's computer is logged onto their account, but they are away from the keyboard. Other forms of idle time are spending time in home screens, customization menus, lobbies, and social spaces.

In this study, active playtime was calculated by aggregating session data. Sessions represent the different levels, game modes, or worlds a player can be present in. Every player is always in a session and could only be in one session at a time, however, multiple players could be in the same session. For example, when a player launches PvZ, they enter the hub world (where they can select what type of level or mode to play), entering the hub world starts the session timer, which ends as they choose a level or leave to join a match, which then starts a new session timer for that level. In total, 36,621 game sessions were provided, of which 25,937 happened in the two weeks prior to a player taking the survey. These sessions took place across 19 different maps that each facilitated different forms of gameplay, these maps could be grouped into 13,594 (51.27%) Player-versus-Player (PVP), 4,854 (18.31%) Player-versus-Environment (PVE), and 7,489 (28.24%) Social Hub maps. In PVP maps players battle against each other in various competitive modes such as objective-based matches, team deathmatches, and 2v2 arena matches. In PVE maps players explore open worlds that are designed for adventure with story-driven quests, side missions, and hidden collectables. The Social Hub served as an idle space for navigation and customization between matches and was excluded. This left 17,505 total sessions, 12,445 PVP sessions (71.09%) and 4,509 PVE sessions (25.76%). Players had clear preferences for which game mode they chose to play (See figure 1 for an Illustration), which already indicated clear differences in playing styles. 

### Behavioural telemetry

```{r apa-descriptive-stats}
#| label: tbl-descriptive-stats
#| tbl-cap: "Descriptive Statistics for Key Variables"
#| fig.width: 10
#| fig.height: 6
#| tbl-cap-location: top
#| echo: false
#| warning: false
#| message: false

# Create a simple APA-style descriptive statistics table
# First, check variable types to handle them appropriately
variable_types <- sapply(pvz_df_analysis, class)

# Handle numeric variables
numeric_vars <- pvz_df_analysis %>%
  select(where(is.numeric)) %>%
  select(
    # Demographics
    age,
    
    # Playtime 
    hours, ratio_pvp_hours, ratio_pve_hours,
    
    # Performance metrics
    pvp_kd, pve_kd, 
    pvp_hit_accuracy, pve_hit_accuracy,
    pvp_dps, pve_dps,
    pvp_average_damage, pve_average_damage,
    pvp_average_score, pve_average_score,
    
    # Well-being and enjoyment
    spane_game_balance, enjoyment
  )

# Calculate descriptive statistics for numeric variables
numeric_stats <- numeric_vars %>%
  summarise(across(everything(), 
                  list(
                    n = ~sum(!is.na(.)),
                    M = ~mean(., na.rm = TRUE),
                    SD = ~sd(., na.rm = TRUE),
                    Min = ~min(., na.rm = TRUE),
                    Max = ~max(., na.rm = TRUE)
                  ))) %>%
  pivot_longer(everything(), 
               names_to = c("variable", "stat"), 
               names_pattern = "(.*)_(.*)",
               values_to = "value") %>%
  pivot_wider(names_from = stat, values_from = value)

# For categorical variables like gender, just count frequencies for the footnote
if ("gender" %in% names(pvz_df_analysis)) {
  gender_counts <- pvz_df_analysis %>%
    count(gender) %>%
    mutate(pct = n / sum(n) * 100) %>%
    mutate(pct = round(pct, 1)) %>%
    mutate(frequency = paste0(n, " (", pct, "%)"))
  
  # Note: We'll add a footnote to explain gender distribution separately
  gender_note <- paste("Gender distribution: ", 
                     paste(paste0(gender_counts$gender, ": ", gender_counts$frequency), 
                           collapse = "; "))
} else {
  gender_note <- NULL
}

# Format the data for the table
descriptive_stats <- numeric_stats %>%
  # Format variable names for better readability
  mutate(variable = case_when(
    variable == "age" ~ "Age",
    variable == "hours" ~ "Total Hours Played",
    variable == "ratio_pvp_hours" ~ "Proportion of PvP Playtime",
    variable == "ratio_pve_hours" ~ "Proportion of PvE Playtime",
    variable == "pvp_kd" ~ "PvP Kill/Death Ratio",
    variable == "pve_kd" ~ "PvE Kill/Death Ratio",
    variable == "pvp_hit_accuracy" ~ "PvP Hit Accuracy",
    variable == "pve_hit_accuracy" ~ "PvE Hit Accuracy",
    variable == "pvp_dps" ~ "PvP Damage per Second",
    variable == "pve_dps" ~ "PvE Damage per Second",
    variable == "pvp_average_damage" ~ "PvP Average Damage per Game",
    variable == "pve_average_damage" ~ "PvE Average Damage per Game",
    variable == "pvp_average_score" ~ "PvP Average Score",
    variable == "pve_average_score" ~ "PvE Average Score",
    variable == "spane_game_balance" ~ "Game-Specific Well-being (SPANE)",
    variable == "enjoyment" ~ "Enjoyment",
    TRUE ~ variable
  )) %>%
  # Create category variable for grouping
  mutate(category = case_when(
    variable == "Age" ~ "Demographics",
    variable %in% c("Total Hours Played", "Proportion of PvP Playtime", "Proportion of PvE Playtime") ~ "Playtime",
    variable %in% c("PvP Kill/Death Ratio", "PvE Kill/Death Ratio", 
                   "PvP Hit Accuracy", "PvE Hit Accuracy",
                   "PvP Damage per Second", "PvE Damage per Second",
                   "PvP Average Damage per Game", "PvE Average Damage per Game",
                   "PvP Average Score", "PvE Average Score") ~ "Performance",
    variable %in% c("Game-Specific Well-being (SPANE)", "Enjoyment") ~ "Well-being",
    TRUE ~ "Other"
  )) %>%
  # Round numeric values
  mutate(
    M = round(M, 2),
    SD = round(SD, 2),
    Min = round(Min, 2),
    Max = round(Max, 2)
)

# Create custom ordering for variables to ensure autonomy, competence, relatedness are listed in order
variable_order <- c(
  # Demographics
  "Age",
  
  # Playtime
  "Total Hours Played", "Proportion of PvP Playtime", "Proportion of PvE Playtime",
  
  # Performance
  "PvP Kill/Death Ratio", "PvE Kill/Death Ratio", 
  "PvP Hit Accuracy", "PvE Hit Accuracy",
  "PvP Damage per Second", "PvE Damage per Second",
  "PvP Average Damage per Game", "PvE Average Damage per Game",
  "PvP Average Score", "PvE Average Score",
  
  # Well-being
  "Game-Specific Well-being (SPANE)", "Enjoyment"
)

# Apply the custom ordering
descriptive_stats <- descriptive_stats %>%
  mutate(variable = factor(variable, levels = variable_order)) %>%
  arrange(variable)

# Create the APA-styled table with kable
apa_table <- descriptive_stats %>%
  select(variable, n, M, SD, Min, Max) %>%
  kable(
    # Caption is handled by chunk options now
    col.names = c("Variable", "*N*", "*M*", "*SD*", "Min", "Max"),
    align = c("l", "c", "c", "c", "c", "c"),
    booktabs = TRUE,
    format = "html",
    digits = 2
  ) %>%
  kable_styling(
    bootstrap_options = c("striped", "hover", "condensed"),
    full_width = FALSE,
    position = "left",
    font_size = 12
  ) %>%
  row_spec(0, bold = TRUE) %>%  # Bold header row per APA
  pack_rows(index = table(factor(descriptive_stats$category, 
                              levels = unique(descriptive_stats$category))), 
            hline_after = TRUE)  # Add lines after each group

# Add footnote for gender if it exists
if (!is.null(gender_note)) {
  apa_table <- apa_table %>%
    footnote(general = gender_note, general_title = "Note: ")
}

# Display the table
apa_table
```

The telemetry featured many fine-grained records of game events per player per session. Metrics were selected and calculated only if they had been previously validated as an informative measure of player behaviour or performance in previous studies [@drachenGunsSwordsData2012; @lugrinImmersiveFPSGames2013; @melhartYourGameplaySays2019]. To effectively measure in-game performance all metrics were calculated to be independent of playtime, hence common game metrics such as 'player level' or 'total kills' are purposefully excluded (e.g., a player's level might simply be higher because they played more and not because they are better). Furthermore, every game metric is calculated twice, once for PVP and once for PVE, because performance indicators should not be computed or compared across these different game modes (e.g., it might be easier to score a point or make a kill on a computer-controlled enemy in PVE, than it is versus matchmade players of equal rank in PVP). The final metrics excluding playtimes and map types were: PVP ratio, PVP Kill/Death Ratio, PVP Hit Accuracy, PVP Damage per second, PVP Average damage per game, PVP Average score per game, PVE playtime, PVE Ratio, PVE Kill/Death Ratio, PVE Hit Accuracy, PVE Damage per second, PVE Average damage per game, PVE Average score per game. See table I for descriptive statistics.

### Well-being

Well-being was assessed through the Scale of Positive and Negative Experiences [@dienerNewWellbeingMeasures2010], which measures the players' affective dimension of well-being [@dienerAdvancesOpenQuestions2018]. The respondents were asked to think about how they had been feeling in the past two weeks in relation to playing PvZ and reported how often they experienced each of six negative and six positive feelings on a scale from 1 (Very rarely or never) to 7 (Very often or always). An overall measure of well-being was calculated by subtracting the mean negative from the mean positive score. This measure was checked for straight lining (e.g. when respondents give identical answers to all items) to identify records with low data quality commonly found in internet surveys [@leinerTooFastToo2019], however, no unusual cases were identified[***.***]{.underline}

### Enjoyment

Enjoyment was assessed with the Player Experience and Need Satisfaction Scale (PENS, Ryan et al., 2006), which has recently been validated (Johnson et al., 2018). The scale consisted of five subscales, of which one measured enjoyment with four items such as "I think \[PvZ/AC:NH\] was fun to play". The items a were rated on a scale from 1 (Strongly disagree) to 7 (Strongly agree).

## Procedure

Data were analyzed using R 4.2 (R Core Team, 2022) with several specialized packages for cluster analysis and statistical modeling. Following initial data cleaning, we screened for outliers and removed participants with less than 15 minutes of total playtime or missing performance metrics. To identify distinct player types, we employed a comprehensive cluster analysis procedure. First, we assessed cluster tendency using the Hopkins statistic [@banerjee2004a], Visual Assessment of Cluster Tendency [@bezdek2002a], and Principal Component Analysis, all of which confirmed the presence of meaningful clusters. We then used Partitioning Around Medoids (PAM) clustering algorithm on standardized in-game behavioral metrics, with the optimal number of clusters (k = 3) determined through silhouette analysis [@kaufmanFindingGroupsData1990] and the Calinski-Harabasz Index. To examine relationships between player types, playtime, and well-being and enjoyment, we constructed robust linear mixed-effects models (using the robustlmm R package) with country as a random effect. We tested both main effects and interaction models to assess whether player type moderated the relationship between playtime and well-being and enjoyment. Simple slopes analyses were conducted to quantify the playtime-well-being relationship for each player type, with pairwise comparisons to test for significant differences between these associations. Statistical significance was set at p \< .05.

**Data, Materials, and Code**

The raw data, materials, codebook and further details on this dataset shared Johannes et al. (2020) can be found at <https://osf.io/cjd6z/>. Building on this raw data and documentation, all of the underlying wrangling and analysis code reported in this study can be found at osf post publication.

# Results

## Player Clustering Analysis

```{r cluster_feature_selection}
#| label: cluster_feature_selection
#| echo: true
#| warning: false
#| message: false

# Select relevant features for clustering
player_cluster_features <- pvz_df_analysis %>% 
  select(
    # Playstyle variables (game mode preferences)
    ratio_pvp_hours, ratio_pve_hours,
    
    # PvP performance metrics
    pvp_kd, pvp_hit_accuracy, pvp_dps, pvp_average_damage, pvp_average_score,
    
    # PvE performance metrics
    pve_kd, pve_hit_accuracy, pve_dps, pve_average_damage, pve_average_score
  )

# Remove any NAs in the dataset to ensure clean clustering
player_cluster_features <- na.omit(player_cluster_features)

# Standardize the data (important for clustering)
player_data_scaled <- scale(player_cluster_features)
rownames(player_data_scaled) <- pvz_df_analysis$player_id[rownames(player_data_scaled) %in% rownames(pvz_df_analysis)]
```

### Assessing Cluster Tendency

```{r cluster_random_data}
#| label: cluster_random_data
#| echo: true
#| warning: false
#| message: false
#| results: "hide"

# Create a random dataset with the same dimensions for comparison
set.seed(123) # For reproducibility
random_cluster_data <- apply(player_data_scaled, 2,
                     function(x){runif(length(x), min(x), max(x))})
random_cluster_data <- as.data.frame(random_cluster_data)
```

```{r hopkins_statistic}
#| label: hopkins_statistic
#| echo: true
#| warning: false
#| message: false
#| results: "hide"

# The Hopkins statistic tests the spatial randomness of the data
# H close to 1 suggests highly clusterable data
# H around 0.5 suggests randomly distributed data

# Function to compute Hopkins statistic properly
compute_hopkins_statistic <- function(data, m = nrow(data) - 1, iter = 10) {
  hopkins_values <- numeric(iter)
  for (i in 1:iter) {
    # Set different seed for each iteration
    set.seed(123 + i)
    hopkins_values[i] <- hopkins::hopkins(data, m = m)
  }
  # Return mean Hopkins statistic
  return(mean(hopkins_values))
}

# Compute Hopkins for real data
hopkins_real <- compute_hopkins_statistic(player_data_scaled)

# Compute Hopkins for random data
hopkins_random <- compute_hopkins_statistic(random_cluster_data)
```

```{r pca_visualization}
#| label: pca_visualization
#| echo: true
#| warning: false
#| message: false
#| results: "hide"

# Visual Assessment using PCA visualization
pca_real_data <- prcomp(player_data_scaled)
pca_random_data <- prcomp(random_cluster_data)

# Create PCA plots for comparison
pca_plot_real <- fviz_pca_ind(pca_real_data, 
                  title = "PCA - Player Performance Data",
                  geom = "point", 
                  ggtheme = theme_minimal(),
                  legend = "none",
                  alpha.ind = 0.7) +
      theme(plot.title = element_text(face = "bold", size = 14))

pca_plot_random <- fviz_pca_ind(pca_random_data, 
                  title = "PCA - Random Data",
                  geom = "point", 
                  ggtheme = theme_minimal(),
                  legend = "none",
                  alpha.ind = 0.7) +
      theme(plot.title = element_text(face = "bold", size = 14))
```

```{r vat_analysis}
#| label: vat_analysis
#| echo: true
#| warning: false
#| message: false
#| results: "hide"

# Using ordered dissimilarity matrices
dist_real_data <- get_dist(player_data_scaled, method = "euclidean")
dist_random_data <- get_dist(random_cluster_data, method = "euclidean")

# Create ordered dissimilarity matrix plots (VAT approach)
dist_plot_real <- fviz_dist(dist_real_data, 
                show_labels = FALSE,
                order = TRUE, # Important: order the matrix for VAT
                gradient = list(low = "#00AFBB", mid = "white", high = "#FC4E07")) +
       ggtitle("VAT Ordered Dissimilarity Matrix - Player Data") +
       theme(plot.title = element_text(face = "bold", size = 14))

dist_plot_random <- fviz_dist(dist_random_data, 
                show_labels = FALSE,
                order = TRUE, # Important: order the matrix for VAT
                gradient = list(low = "#00AFBB", mid = "white", high = "#FC4E07")) +
       ggtitle("VAT Ordered Dissimilarity Matrix - Random Data") +
       theme(plot.title = element_text(face = "bold", size = 14))
```

To evaluate the clustering tendency of our data prior to formal analysis, we employed multiple established statistical techniques. The Hopkins statistic (H) yielded a value of 0.999 for the player data, indicating exceptionally high clusterability (H = 0.999, where values approaching 1.0 suggest strong clustering structure). In contrast, a randomly generated dataset with identical dimensions produced a Hopkins statistic of 0.501, consistent with random, non-clusterable data (H = 0.501). Visual Assessment of Cluster Tendency (VAT; @bezdek2002a) analysis corroborated these findings, with the player data displaying distinct dark blocks along the diagonal of the ordered dissimilarity matrix, signifying well-defined observation clusters (see Appendix 1a). Principal Component Analysis (PCA) further confirmed the presence of natural clustering, with player data exhibiting visible grouping patterns compared to the uniform distribution of the random dataset (see Appendix 1b). The convergence of these three independent assessments (Hopkins, VAT, and PCA) provided strong statistical evidence to reject the null hypothesis of spatial randomness,demonstrating that the player behavioral data contained meaningful, non-random structure amenable to clustering methods.

### Determining Optimal Number of Clusters

```{r pam_silhouette_method}
#| label: pam_silhouette_method
#| fig.width: 10
#| fig.height: 4
#| cache: false
#| results: "asis"
#| fig.show: "hide"  # Changed from "asis" to "hide" to prevent displaying here

# PAM optimal cluster determination
# -------------------------------------
# Silhouette method for PAM
pam_sil <- function(data, max_k = 10) {
  sil_values <- numeric(max_k - 1)
  for (i in 2:max_k) {
    pam_result <- pam(data, k = i, metric = "euclidean")
    sil_values[i-1] <- pam_result$silinfo$avg.width
  }
  k_values <- 2:max_k
  return(data.frame(k = k_values, sil = sil_values))
}

pam_sil_results <- pam_sil(player_data_scaled)
p_pam_silhouette <- ggplot(pam_sil_results, aes(x = k, y = sil)) +
  geom_line() +
  geom_point() +
  scale_x_continuous(breaks = 2:10) +
  labs(title = "PAM: Silhouette Method",
       x = "Number of clusters k",
       y = "Average silhouette width") +
  theme_minimal()

# Calinski-Harabasz Index (CH index) for PAM
ch_index_pam <- function(data, max_k = 10) {
  ch_values <- numeric(max_k - 1)
  for (i in 2:max_k) {
    pam_result <- pam(data, k = i)
    ch_values[i-1] <- fpc::calinhara(data, pam_result$clustering)
  }
  k_values <- 2:max_k
  return(data.frame(k = k_values, ch_index = ch_values))
}

ch_pam_results <- ch_index_pam(player_data_scaled)
p_pam_ch <- ggplot(ch_pam_results, aes(x = k, y = ch_index)) +
  geom_line() +
  geom_point() +
  scale_x_continuous(breaks = 2:10) +
  labs(title = "PAM: Calinski-Harabasz Index",
       x = "Number of clusters k",
       y = "CH Index (higher is better)") +
  theme_minimal()

# Arrange PAM validation plots in a grid
combined_pam_plots <- grid.arrange(p_pam_silhouette, p_pam_ch,
             ncol = 2, nrow = 1,
             top = grid::textGrob("Determining Optimal Number of Clusters with PAM", 
                                 gp = grid::gpar(fontsize = 16, font = 2)))

```

```{r determine_optimal_k, echo=FALSE}
# Determine the optimal k values
silhouette_optimal_k <- pam_sil_results$k[which.max(pam_sil_results$sil)]
ch_optimal_k <- ch_pam_results$k[which.max(ch_pam_results$ch_index)]

# Determine final optimal k (using most common value or a specific method if preferred)
if(silhouette_optimal_k == ch_optimal_k) {
  final_optimal_k <- silhouette_optimal_k
} else {
  # In case of disagreement, you might want to choose one method or use another strategy
  # Here we'll use silhouette as the deciding factor, but you could modify this logic
  final_optimal_k <- silhouette_optimal_k
}
```

To determine the optimal number of clusters, we employed multiple validation methods following established guidelines in cluster analysis [@kaufmanFindingGroupsData1990; @tibshiraniEstimatingNumberClusters2001]. Specifically, we applied the silhouette method, which measures how well observations fit within their assigned clusters compared to neighboring clusters, and the Calinski-Harabasz Index (CH Index), which evaluates cluster validity based on the ratio of between-cluster to within-cluster variance. Both methods were systematically applied to Partitioning Around Medoids (PAM) clustering solutions for k values ranging from 2 to 10. The silhouette analysis yielded its highest coefficient (indicating strongest cluster structure) at k = 3, while the CH Index similarly reached its maximum value at k = 3, demonstrating agreement between both validation techniques (see Appendix 2). Based on this consensus and visual inspection of the validation plots, we determined that a three-cluster solution provided the optimal balance between model parsimony and explanatory power. This conclusion was further supported by the interpretability of the resulting clusters, which aligned with theoretically meaningful player typologies observed in previous gaming research [@drachenGunsSwordsData2012; @drachen2014a; @melhartYourGameplaySays2019].

### PAM Clustering

```{r prepare_player_data}
#| label: prepare_player_data

# Prepare player data for clustering
player_data_features <- pvz_df_analysis %>% 
  select(
    # Game mode preferences
    ratio_pvp_hours, ratio_pve_hours,
    
    # Performance metrics
    pvp_kd, pve_kd, 
    pvp_hit_accuracy, pve_hit_accuracy,
    pvp_average_score, pve_average_score,
    pvp_dps, pve_dps
  )

# Remove any remaining NA values 
player_data_features <- na.omit(player_data_features)

# Standardize the data for clustering
player_data_scaled <- scale(player_data_features)

# Perform PAM clustering with k=3
set.seed(42) # For reproducibility
k_clusters <- 3
pam_result <- pam(player_data_scaled, k = k_clusters, metric = "euclidean")

```

```{r pam_cluster_statistics}
#| label: pam_cluster_statistics
#| cache: false
#| results: "asis"

# Calculate silhouette coefficient for each cluster using the existing pam_result
sil <- pam_result$silinfo$widths
sil_per_cluster <- tapply(sil[, 3], sil[, 1], mean)

# Store key statistics
stats_pam_k3 <- list(
  k = k_clusters,
  avg_silhouette = pam_result$silinfo$avg.width,
  silhouette_per_cluster = sil_per_cluster,
  cluster_sizes = table(pam_result$clustering),
  medoid_indices = pam_result$id.med
)
```

The validity of the identified cluster solution was assessed using the silhouette coefficient [@kaufmanFindingGroupsData1990], which measures how well observations fit within their assigned clusters compared to neighboring clusters. The average silhouette width for the three-cluster solution was 0.22, indicating a reasonable structure in the data according to Kaufman and Rousseeuw's interpretive guidelines, which suggest values above 0.25 represent strong structure and values between 0.10 and 0.25 indicate a weak but still meaningful structure. Examining silhouette values by cluster revealed varying degrees of cohesion, with values of 0.26, 0.19, and 0.17 for the Casual PVE Players, Elite Allround Players, and Casual PVP Players clusters, respectively. The Casual PVE Players cluster demonstrated the strongest internal cohesion and separation from other clusters, while the other two clusters showed weaker but still acceptable structure. No negative silhouette values were observed, confirming that all objects were assigned to their most appropriate cluster.

```{r cluster_profiles}
#| label: cluster_profiles

# Define the feature columns used in clustering
feature_columns <- c("ratio_pvp_hours", "ratio_pve_hours", 
                    "pvp_kd", "pve_kd", 
                    "pvp_hit_accuracy", "pve_hit_accuracy",
                    "pvp_average_score", "pve_average_score",
                    "pvp_dps", "pve_dps")

# Get cluster medoids (representative players)
medoid_indices <- pam_result$id.med
medoid_data <- pvz_df_analysis[medoid_indices, ]
medoid_features <- medoid_data %>%
  select(all_of(feature_columns))
medoid_features$cluster <- 1:k_clusters

# Calculate mean values by cluster for interpretation
cluster_means <- pvz_df_analysis %>%
  # First add the cluster assignments
  mutate(cluster = pam_result$clustering) %>%
  group_by(cluster) %>%
  summarize(across(all_of(feature_columns), mean, na.rm = TRUE), .groups = "drop")

# Display medoid profiles 
print(kable(medoid_features, digits = 2, 
      caption = "Cluster Medoids (Feature Values of Representative Players)"))

# Display cluster means
print(kable(cluster_means, digits = 2,
      caption = "Cluster Means - Average Feature Values"))
```

```{r assign_player_types}
#| label: assign_player_types

# Create a data frame with cluster labels 
# Update these names based on interpretation of the 3 clusters
player_type_labels <- tibble(
  cluster = 1:3,
  player_type = c(
    "Casual PVE Players",           # Cluster 1 name
    "Elite Allround Players",       # Cluster 2 name 
    "Casual PVP Players"            # Cluster 3 name
  )
)

# Add cluster assignments and type labels to the dataframe
pvz_df_clustered <- pvz_df_analysis %>%
  mutate(cluster = pam_result$clustering) %>%
  left_join(player_type_labels, by = "cluster") %>%
  mutate(player_type = factor(player_type))

# Save the clustered data with no outliers
write_csv(pvz_df_clustered, here("data/processed/pvz_with_clusters_pam3_nohours_no_outliers.csv"))

```

```{r radar_functions}
#| label: radar_functions

# Function to scale values using z-scores
z_score_scale <- function(x) {
  if (sd(x, na.rm = TRUE) == 0) return(rep(0, length(x)))
  return((x - mean(x, na.rm = TRUE)) / sd(x, na.rm = TRUE))
}

# Convert z-scores to 0-1 scale for radar chart
z_to_radar_scale <- function(z, min_z = -2.5, max_z = 2.5) {
  z_clipped <- pmax(pmin(z, max_z), min_z)
  scaled <- (z_clipped - min_z) / (max_z - min_z)
  return(scaled)
}

# Function to create radar chart with average player comparison
create_player_type_radar <- function(data, avg_data, color, title, values) {
  # Get column names
  data_cols <- colnames(data)
  
  # Create max and min rows
  max_min <- data.frame(matrix(c(rep(1, length(data_cols)), rep(0, length(data_cols))), 
                             nrow = 2, byrow = TRUE))
  colnames(max_min) <- data_cols
  
  # Combine data for plotting (max, min, average, player type)
  plot_data <- rbind(max_min, avg_data, data)
  
  # Create radar chart
  radarchart(
    plot_data, 
    pcol = c("gray60", color),        # Colors for average and player type
    pfcol = c(adjustcolor("gray60", alpha.f = 0.2), adjustcolor(color, alpha.f = 0.5)), # Fill colors
    plwd = c(2, 3),                  # Line width for average and player type
    plty = c(2, 1),                  # Line type (dashed for average, solid for player type)
    vlcex = 0.85,                    # Size of labels
    title = title,
    cglcol = "grey80",
    cglty = 1,
    axislabcol = "grey30",
    caxislabels = seq(0, 1, 0.25),
    cglwd = 0.8,
    pty = 32,                        # Use 32-sided polygon for smoother lines
    axistype = 0,                    # Label types to display
    centerzero = FALSE,              # Do not center at zero
    seg = 4,                         # Number of segments
    cex.main = 1.2                   # Title size
  )
  
  # Add legend for player type vs average
  legend(
    "bottomright", 
    legend = c("Average Player", paste("Player Type:", title)),
    col = c("gray60", color),
    lty = c(2, 1),
    lwd = c(2, 3),
    bty = "n",     # No box around legend
    cex = 0.8,     # Font size
    pt.cex = 1     # Symbol size
  )
  
  # Add values text
  legend_text <- sprintf("PvP Ratio: %.2f, PvP K/D: %.2f, PvE Ratio: %.2f, PvE K/D: %.2f", 
                        values$`PvP Ratio`, values$`PvP K/D`,
                        values$`PvE Ratio`, values$`PvE K/D`)
  
  mtext(legend_text, side = 1, line = 2, cex = 0.7, col = color)
}
```

```{r radar_charts}
#| label: radar_charts
#| fig.height: 12
#| fig.width: 12
#| out.width: "100%"

# Define metrics for radar charts in a more logical order, 
# grouping related metrics together
radar_metrics <- c(
  # PvE metrics grouped together
  "PvE Ratio", "PvE K/D", "PvE Accuracy", "PvE Score", "PvE DPS",
  # PvP metrics grouped together
  "PvP Ratio", "PvP K/D", "PvP Accuracy", "PvP Score", "PvP DPS"
)

# Get player type names
player_type_names <- player_type_labels$player_type

# Calculate radar chart data for both average player and player types
# First collect the raw metrics for all players and player types
metrics_data <- pvz_df_analysis %>%
  # Add a column for the average player
  mutate(profile_type = "Average Player") %>%
  # Bind the clustered data with profile types
  bind_rows(
    pvz_df_clustered %>%
      select(
        player_id, player_type, 
        ratio_pve_hours, ratio_pvp_hours,
        pve_kd, pvp_kd,
        pve_hit_accuracy, pvp_hit_accuracy,
        pve_average_score, pvp_average_score,
        pve_dps, pvp_dps
      ) %>%
      rename(profile_type = player_type)
  ) %>%
  # Group by profile type (average or a player type)
  group_by(profile_type) %>%
  # Calculate means for each metric
  summarise(
    ratio_pve_mean = mean(ratio_pve_hours, na.rm = TRUE),
    pve_kd_mean = mean(pve_kd, na.rm = TRUE),
    pve_accuracy_mean = mean(pve_hit_accuracy, na.rm = TRUE),
    pve_score_mean = mean(pve_average_score, na.rm = TRUE),
    pve_dps_mean = mean(pve_dps, na.rm = TRUE),
    ratio_pvp_mean = mean(ratio_pvp_hours, na.rm = TRUE),
    pvp_kd_mean = mean(pvp_kd, na.rm = TRUE),
    pvp_accuracy_mean = mean(pvp_hit_accuracy, na.rm = TRUE),
    pvp_score_mean = mean(pvp_average_score, na.rm = TRUE),
    pvp_dps_mean = mean(pvp_dps, na.rm = TRUE)
  ) %>%
  ungroup()

# Create vectors of all values for each metric to calculate z-scores across all profiles
all_ratio_pve <- metrics_data$ratio_pve_mean
all_pve_kd <- metrics_data$pve_kd_mean
all_pve_accuracy <- metrics_data$pve_accuracy_mean
all_pve_score <- metrics_data$pve_score_mean
all_pve_dps <- metrics_data$pve_dps_mean
all_ratio_pvp <- metrics_data$ratio_pvp_mean
all_pvp_kd <- metrics_data$pvp_kd_mean
all_pvp_accuracy <- metrics_data$pvp_accuracy_mean
all_pvp_score <- metrics_data$pvp_score_mean
all_pvp_dps <- metrics_data$pvp_dps_mean

# Now apply z-score scaling to create the radar data
radar_data <- metrics_data %>%
  mutate(
    `PvE Ratio` = z_to_radar_scale(z_score_scale(all_ratio_pve)[row_number()]),
    `PvE K/D` = z_to_radar_scale(z_score_scale(all_pve_kd)[row_number()]),
    `PvE Accuracy` = z_to_radar_scale(z_score_scale(all_pve_accuracy)[row_number()]),
    `PvE Score` = z_to_radar_scale(z_score_scale(all_pve_score)[row_number()]),
    `PvE DPS` = z_to_radar_scale(z_score_scale(all_pve_dps)[row_number()]),
    `PvP Ratio` = z_to_radar_scale(z_score_scale(all_ratio_pvp)[row_number()]),
    `PvP K/D` = z_to_radar_scale(z_score_scale(all_pvp_kd)[row_number()]),
    `PvP Accuracy` = z_to_radar_scale(z_score_scale(all_pvp_accuracy)[row_number()]),
    `PvP Score` = z_to_radar_scale(z_score_scale(all_pvp_score)[row_number()]),
    `PvP DPS` = z_to_radar_scale(z_score_scale(all_pvp_dps)[row_number()])
  ) %>%
  # Keep only relevant columns
  select(
    profile_type,
    `PvE Ratio`, `PvE K/D`, `PvE Accuracy`, `PvE Score`, `PvE DPS`,
    `PvP Ratio`, `PvP K/D`, `PvP Accuracy`, `PvP Score`, `PvP DPS`
  )

# Extract the average player data
average_player <- radar_data %>%
  filter(profile_type == "Average Player") %>%
  select(-profile_type) %>%
  as.data.frame()

# Colors for player types - more distinct and colorblind-friendly palette
player_type_colors <- c("#E41A1C", "#377EB8", "#4DAF4A")

# Set up a better layout with proper margins
# First adjust the overall margins to give more room
par(mfrow = c(1, 3), mar = c(4, 2, 3, 2), oma = c(2, 0, 2, 0))

# Create radar charts for each player type
for (i in 1:length(player_type_names)) {
  # Extract data for this player type
  type_data <- radar_data %>%
    filter(profile_type == player_type_names[i]) %>%
    select(-profile_type) %>%
    as.data.frame()
  
  # Get actual values
  actual_values <- pvz_df_clustered %>%
    filter(player_type == player_type_names[i]) %>%
    summarise(
      `PvP Ratio` = mean(ratio_pvp_hours),
      `PvP K/D` = mean(pvp_kd),
      `PvE Ratio` = mean(ratio_pve_hours),
      `PvE K/D` = mean(pve_kd)
    )
  
  # Create radar chart with average player comparison
  create_player_type_radar(
    type_data, 
    average_player,
    player_type_colors[i], 
    player_type_names[i],
    actual_values
  )
}

# Add an overall title above all charts
mtext("Player Type Characteristics: Performance and Playstyle Metrics", 
      side = 3, line = 0, outer = TRUE, cex = 1.2, font = 2)

# Add scaling note at the bottom
mtext("Note: Values shown as standardized scores (z-scores converted to 0-1 scale)", 
      side = 1, line = 0, outer = TRUE, cex = 0.8)

# Reset layout
par(mfrow = c(1, 1))
```

```{r player_type_summary}
#| label: player_type_summary

# Create a summary of player types
player_type_summary <- pvz_df_clustered %>%
  group_by(player_type) %>%
  summarise(
    Count = n(),
    `% of Players` = n() / nrow(pvz_df_clustered) * 100,
    `PvP Time %` = mean(ratio_pvp_hours) * 100,
    `PvE Time %` = mean(ratio_pve_hours) * 100,
    `PvP K/D` = mean(pvp_kd),
    `PvE K/D` = mean(pve_kd),
    `PvP Accuracy` = mean(pvp_hit_accuracy) * 100,
    `PvE Accuracy` = mean(pve_hit_accuracy) * 100
  )

# Display summary table
print(kable(player_type_summary, digits = 2,
      caption = "Player Type Summary"))
```

### Interpretation of Player Types

The cluster analysis revealed three distinct player types based on gameplay behaviors and preferences, following an established method for telemetry-based player profiling suggested by @drachenGunsSwordsData2012, @drachen2014a and employed in recent game analytics research [@bauckhageClusteringGameBehavior2015; @melhartYourGameplaySays2019; @sifaProfilingGamesUnderstanding2018]. Figure 1 presents radar charts visualizing each player type's distinguishing characteristics relative to the average player.

Casual PVE Players (`r round(player_type_summary$"% of Players"[player_type_summary$player_type=="Casual PVE Players"], 1)`% of the sample) demonstrated a pronounced preference for cooperative gameplay against AI opponents, allocating `r round(player_type_summary$"PvE Time %"[player_type_summary$player_type=="Casual PVE Players"], 1)`% of their time to PVE modes. Their radar chart profiles exhibited above-average scores on PVE-related metrics with notably weaker performance in PVP contexts. These players achieved moderate PVE kill-to-death ratios (M = `r round(player_type_summary$"PvE K/D"[player_type_summary$player_type=="Casual PVE Players"], 2)`) and accuracy rates (M = `r round(player_type_summary$"PvE Accuracy"[player_type_summary$player_type=="Casual PVE Players"], 1)`%), while their performance metrics in competitive modes were considerably lower. This pattern suggests a specialization in cooperative rather than competitive gameplay, with primary engagement through story-driven and exploration components.

Elite Allround Players (`r round(player_type_summary$"% of Players"[player_type_summary$player_type=="Elite Allround Players"], 1)`% of the sample) displayed exceptional versatility, maintaining superior performance across both gameplay contexts. Their radar charts revealed consistently high values across all performance dimensions, with impressive kill-to-death ratios in both PVP (M = `r round(player_type_summary$"PvP K/D"[player_type_summary$player_type=="Elite Allround Players"], 2)`) and PVE (M = `r round(player_type_summary$"PvE K/D"[player_type_summary$player_type=="Elite Allround Players"], 2)`) scenarios, complemented by strong accuracy metrics (PVP: M = `r round(player_type_summary$"PvP Accuracy"[player_type_summary$player_type=="Elite Allround Players"], 1)`%; PVE: M = `r round(player_type_summary$"PvE Accuracy"[player_type_summary$player_type=="Elite Allround Players"], 1)`%). These players have evidently developed high-level skills that transfer effectively between different gameplay contexts, distinguishing them as the most technically proficient player segment.

Casual PVP Players (`r round(player_type_summary$"% of Players"[player_type_summary$player_type=="Casual PVP Players"], 1)`% of the sample) exhibited a preference for competitive gameplay, dedicating `r round(player_type_summary$"PvP Time %"[player_type_summary$player_type=="Casual PVP Players"], 1)`% of their time to PVP modes. However, despite this competitive focus, their radar charts revealed only moderate performance metrics (PVP K/D: M = `r round(player_type_summary$"PvP K/D"[player_type_summary$player_type=="Casual PVP Players"], 2)`; PVP Accuracy: M = `r round(player_type_summary$"PvP Accuracy"[player_type_summary$player_type=="Casual PVP Players"], 1)`%) that fell substantially below those of Elite Allround Players. Their performance in PVE contexts was similarly modest, indicating a casual approach to gameplay regardless of context, albeit with a preference for player-versus-player engagement.

## Statistical Analysis: Robust Mixed-Effects Models

### Effect of Player Type on Well-Being

```{r player-type-wellbeing-model}
#| label: player-type-wellbeing-model
#| message: false
#| warning: false

# Create a robust linear model for well-being
set.seed(123) # For reproducibility of robust estimates

# Fit robust models for game-specific well-being
rmodel_game <- rlmer(spane_game_balance ~ player_type + hours + gender + age + (1|country), 
                     data = pvz_df_clustered)

# Fit robust models for enjoyment
rmodel_enjoyment <- rlmer(enjoyment ~ player_type + hours + gender + age + (1|country), 
                     data = pvz_df_clustered)

# Summarize the models
cat("
Robust model for subjectivewell-being:
")
summary(rmodel_game)

cat("
Robust model for enjoyment:
")
summary(rmodel_enjoyment)

# Create a nicer formatted summary of the fixed effects
model_coefs_game <- data.frame(coef(summary(rmodel_game)))
model_coefs_game$predictor <- rownames(model_coefs_game)
rownames(model_coefs_game) <- NULL
model_coefs_game <- model_coefs_game %>%
  select(predictor, Estimate, Std..Error, t.value) %>%
  rename(`Std. Error` = Std..Error,
         `t value` = t.value)

model_coefs_enjoyment <- data.frame(coef(summary(rmodel_enjoyment)))
model_coefs_enjoyment$predictor <- rownames(model_coefs_enjoyment)
rownames(model_coefs_enjoyment) <- NULL
model_coefs_enjoyment <- model_coefs_enjoyment %>%
  select(predictor, Estimate, Std..Error, t.value) %>%
  rename(`Std. Error` = Std..Error,
         `t value` = t.value)

# Display nicely formatted tables
kable(model_coefs_game, digits = 3, 
      caption = "Fixed effects for subjective well-being model") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
                full_width = FALSE)

kable(model_coefs_enjoyment, digits = 3, 
      caption = "Fixed effects for enjoyment model") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
                full_width = FALSE)

```

First, we examined whether player types differed in their levels of well-being while controlling for playtime, gender, and age. Country was included as a random effect to account for potential cultural differences. The analysis revealed significant differences among player types for both outcome variables. For well-being (SPANE balance), the model showed that compared to Casual PVE Players, Casual PVP Players reported significantly lower well-being scores (b = -0.76, SE = 0.17, t = -4.39), while Elite Allround Players showed no significant difference (b = -0.04, SE = 0.20, t = -0.20). Hours of gameplay was positively associated with well-being (b = 0.08, SE = 0.04, t = 2.28), indicating that increased playtime was generally associated with higher well-being across all player types.

Similarly, for enjoyment, Casual PVP Players reported significantly lower levels (b = -0.72, SE = 0.13, t = -5.39) compared to Casual PVE Players, while Elite Allround Players again showed no significant difference (b = 0.04, SE = 0.16, t = 0.26). Hours played showed a positive association with enjoyment (b = 0.05, SE = 0.03, t = 1.82), though this relationship was marginally significant.

((Maybe visualise the relationship between hours and well-being))

## Playtime and Well-Being: Moderation by Player Type

```{r moderation-models}
#| label: moderation-models
#| message: false
#| warning: false

# Create interaction models for our two key well-being outcomes
rmodel_moderation_game <- rlmer(
  spane_game_balance ~ player_type * hours + (1|country),
  data = pvz_df_clustered
)

rmodel_moderation_enjoyment <- rlmer(
  enjoyment ~ player_type * hours + (1|country),
  data = pvz_df_clustered
)

# Summarize the models
cat("
Robust moderation model for subjective well-being:
")
summary(rmodel_moderation_game)

cat("
Robust moderation model for enjoyment:
")
summary(rmodel_moderation_enjoyment)

# Extract fixed effects for the moderation models
mod_coefs_game <- data.frame(coef(summary(rmodel_moderation_game)))
mod_coefs_game$predictor <- rownames(mod_coefs_game)
rownames(mod_coefs_game) <- NULL
mod_coefs_game <- mod_coefs_game %>%
  select(predictor, Estimate, Std..Error, t.value) %>%
  rename(`Std. Error` = Std..Error,
         `t value` = t.value)

mod_coefs_enjoyment <- data.frame(coef(summary(rmodel_moderation_enjoyment)))
mod_coefs_enjoyment$predictor <- rownames(mod_coefs_enjoyment)
rownames(mod_coefs_enjoyment) <- NULL
mod_coefs_enjoyment <- mod_coefs_enjoyment %>%
  select(predictor, Estimate, Std..Error, t.value) %>%
  rename(`Std. Error` = Std..Error,
         `t value` = t.value)

# Display nicely formatted tables
kable(mod_coefs_game, digits = 3, 
      caption = "Fixed effects for subjective well-being moderation model") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
                full_width = FALSE)

kable(mod_coefs_enjoyment, digits = 3, 
      caption = "Fixed effects for enjoyment moderation model") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
                full_width = FALSE)

```

```{r moderation-plots}
#| label: moderation-plots
#| fig.width: 10
#| fig.height: 8
#| message: false
#| warning: false

# Create prediction data frames for plotting
pred_game <- ggpredict(rmodel_moderation_game, terms = c("hours", "player_type"))
pred_enjoyment <- ggpredict(rmodel_moderation_enjoyment, terms = c("hours", "player_type"))

# Game-specific well-being plot
p1 <- ggplot(pred_game, aes(x = x, y = predicted, color = group)) +
  geom_line(size = 1) +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high, fill = group), 
              alpha = 0.1, color = NA) +
  scale_color_brewer(palette = "Set2", name = "Player Type") + 
  scale_fill_brewer(palette = "Set2", guide = "none") +
  labs(
    title = "Relationship between Playtime and Subjective Well-Being by Player Type",
    x = "Hours Played",
    y = "Predicted SPANE Score",
    color = "Player Type"
  ) +
  theme_minimal() +
  theme(legend.position = "bottom")

# Enjoyment plot
p2 <- ggplot(pred_enjoyment, aes(x = x, y = predicted, color = group)) +
  geom_line(size = 1) +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high, fill = group), 
              alpha = 0.1, color = NA) +
  scale_color_brewer(palette = "Set2", name = "Player Type") + 
  scale_fill_brewer(palette = "Set2", guide = "none") +
  labs(
    title = "Relationship between Playtime and Enjoyment by Player Type",
    x = "Hours Played",
    y = "Predicted Enjoyment Score",
    color = "Player Type"
  ) +
  theme_minimal() +
  theme(legend.position = "bottom")

# Arrange plots
ggarrange(p1, p2, ncol = 1, common.legend = TRUE, legend = "bottom")
```

To investigate whether the relationship between playtime and well-being outcomes differed across player types, we constructed moderation models with interaction terms between player type and hours played. These analyses revealed significant interaction effects, indicating that player type moderates the relationship between playtime and psychological outcomes.

For game-specific well-being, the interaction model revealed a significant positive effect of hours played for Casual PVE Players (b = 0.14, SE = 0.04, t = 3.36). Significant interactions were observed between player type and hours played. Specifically, the relationship between hours played and well-being was significantly weaker for Casual PVP Players compared to Casual PVE Players (b = -0.17, SE = 0.06, t = -2.71). No significant difference in this relationship was observed between Elite Allround Players and Casual PVE Players (b = -0.02, SE = 0.07, t = -0.23).

For enjoyment, we observed a similar pattern. Hours played showed a significant positive relationship with enjoyment for Casual PVE Players (b = 0.11, SE = 0.03, t = 3.24). This relationship was significantly weaker for Casual PVP Players (b = -0.15, SE = 0.05, t = -3.05) compared to Casual PVE Players. Again, no significant difference in this relationship was observed between Elite Allround Players and Casual PVE Players (b = -0.02, SE = 0.05, t = -0.41).

```{r wellbeing-anova}
#| label: wellbeing-anova
#| message: false
#| warning: false

# Run ANOVAs to test for significant differences between player types
wellbeing_vars <- c("spane_game_balance", "enjoyment")

# Store ANOVA results
anova_results <- list()

for (var in wellbeing_vars) {
  formula <- as.formula(paste(var, "~ player_type"))
  anova_model <- aov(formula, data = pvz_df_clustered)
  anova_results[[var]] <- broom::tidy(anova_model)
  anova_results[[var]]$variable <- var
  
  # Add effect size
  anova_results[[var]]$eta_squared <- effectsize::eta_squared(anova_model)[["Eta2"]]
}

# Combine all results
all_anova_results <- bind_rows(anova_results)

# Display results in a nice table
kable(all_anova_results %>% 
        filter(term == "player_type") %>%
        select(variable, df, sumsq, statistic, p.value, eta_squared) %>%
        mutate(
          p.value = format.pval(p.value, digits = 3),
          eta_squared = round(eta_squared, 3)
        ),
      caption = "ANOVA results for well-being variables by player type",
      col.names = c("Variable", "Df", "Sum Sq", "F value", "p-value", "Eta²")) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
                full_width = FALSE)

# Post-hoc tests for variables with significant differences
sig_vars <- character(0)
posthoc_results <- list()

for (var in wellbeing_vars) {
  formula <- as.formula(paste(var, "~ player_type"))
  anova_model <- aov(formula, data = pvz_df_clustered)
  
  if (summary(anova_model)[[1]]["player_type", "Pr(>F)"] < 0.05) {
    sig_vars <- c(sig_vars, var)
    cat(paste("
Post-hoc tests for", var, ":
"))
    posthoc <- TukeyHSD(anova_model)
    posthoc_results[[var]] <- posthoc
    print(posthoc)
  }
}
```

```{r dynamic-results-text}
#| label: dynamic-results-text
#| echo: false
#| results: asis

# Function to format F-statistics in APA style
format_f_stat <- function(f, df1, df2, p, eta_sq) {
  p_formatted <- if (p < .001) "< .001" else paste("=", format(round(p, 3), nsmall = 3))
  paste0("*F*(", df1, ", ", df2, ") = ", format(round(f, 2), nsmall = 2),
         ", *p* ", p_formatted, ", η² = ", format(round(eta_sq, 3), nsmall = 3))
}

# Function to create proper variable names for text
proper_var_name <- function(var_name) {
  name_map <- c(
    "spane_game_balance" = "subjective well-being",
    "enjoyment" = "enjoyment"
  )
  return(name_map[var_name])
}

# Create dynamic text based on results
sig_results <- all_anova_results %>% 
  filter(term == "player_type" & p.value < 0.05) %>%
  arrange(p.value)

nonsig_results <- all_anova_results %>% 
  filter(term == "player_type" & p.value >= 0.05) %>%
  arrange(p.value)

# Count how many variables showed significant differences
sig_count <- nrow(sig_results)
total_count <- length(wellbeing_vars)

```

One-way ANOVAs were conducted to further examine differences in well-being components across player types. The analyses revealed significant differences between player types for both outcome measures.

Player types differed significantly in terms of game-specific well-being, F(2, 391) = 4.52, p = .012, η² = 0.023. Post-hoc comparisons using Tukey's HSD test indicated that Casual PVP Players had significantly lower well-being levels than Casual PVE Players (p = .009). No other pairwise comparisons reached statistical significance.

Similarly, player types differed significantly in terms of enjoyment, F(2, 391) = 8.55, p \< .001, η² = 0.042. Post-hoc comparisons indicated that Casual PVP Players had significantly lower enjoyment levels than both Casual PVE Players (p \< .001) and Elite Allround Players (p = .002). No significant difference was observed between Casual PVE Players and Elite Allround Players.

## Simple Slopes Analysis

```{r simple-slopes-functions}
#| label: simple-slopes-functions
#| message: false
#| warning: false

# Function to calculate simple slopes for all player types
calculate_all_simple_slopes <- function(model, outcome_name) {
  # For robust mixed models, extract coefficients manually
  coefs <- fixef(model)
  vcov_matrix <- vcov(model)
  
  # Get player types
  player_types <- levels(pvz_df_clustered$player_type)
  
  # Base effect of hours (for reference level)
  hours_effect <- coefs["hours"]
  hours_se <- sqrt(vcov_matrix["hours", "hours"])
  
  # Create results dataframe
  results <- data.frame(
    player_type = player_types[1],  # reference level
    estimate = hours_effect,
    std.error = hours_se,
    statistic = hours_effect / hours_se,
    p.value = 2 * pt(abs(hours_effect / hours_se), 
                    df = nrow(pvz_df_clustered) - length(coefs), 
                    lower.tail = FALSE),
    stringsAsFactors = FALSE
  )
  
  # Calculate for other player types
  for (i in 2:length(player_types)) {
    pt <- player_types[i]
    interaction_term <- paste0("player_type", pt, ":hours")
    
    if (interaction_term %in% names(coefs)) {
      # Calculate slope
      slope <- hours_effect + coefs[interaction_term]
      
      # Calculate SE using delta method (simplified)
      var_hours <- vcov_matrix["hours", "hours"]
      var_inter <- vcov_matrix[interaction_term, interaction_term]
      cov_hours_inter <- vcov_matrix["hours", interaction_term]
      slope_se <- sqrt(var_hours + var_inter + 2*cov_hours_inter)
      
      # t-statistic and p-value
      t_stat <- slope / slope_se
      p_val <- 2 * pt(abs(t_stat), df = nrow(pvz_df_clustered) - length(coefs), lower.tail = FALSE)
      
      # Add to results
      results <- rbind(results, data.frame(
        player_type = pt,
        estimate = slope,
        std.error = slope_se,
        statistic = t_stat,
        p.value = p_val,
        stringsAsFactors = FALSE
      ))
    }
  }
  
  # Add confidence intervals and formatting
  results <- results %>%
    mutate(
      outcome = outcome_name,
      conf.low = estimate - 1.96 * std.error,
      conf.high = estimate + 1.96 * std.error,
      significant = p.value < 0.05,
      stars = case_when(
        p.value < 0.001 ~ "***",
        p.value < 0.01 ~ "**",
        p.value < 0.05 ~ "*",
        p.value < 0.1 ~ "†",
        TRUE ~ ""
      )
    )
  
  return(results)
}

# Function to create interaction plots for all player types
create_interaction_plots <- function(model, outcome_name) {
  # Extract fixed effects
  fixed_effects <- fixef(model)
  
  # Get player types
  player_types <- levels(pvz_df_clustered$player_type)
  
  # Create a grid of values for hours
  hours_range <- range(pvz_df_clustered$hours, na.rm = TRUE)
  hours_grid <- seq(from = hours_range[1], to = hours_range[2], length.out = 100)
  
  # Get the coefficient for hours (main effect)
  hours_coef <- fixed_effects["hours"]
  
  # Create a dataframe to store predictions
  predictions <- data.frame()
  
  # For each player type, calculate predicted values across hours range
  for (i in 1:length(player_types)) {
    pt <- player_types[i]
    
    # Calculate intercept
    if (i == 1) {
      # Reference level
      intercept <- fixed_effects["(Intercept)"]
      slope <- hours_coef
    } else {
      # Other levels
      intercept <- fixed_effects["(Intercept)"] + fixed_effects[paste0("player_type", pt)]
      interaction_term <- paste0("player_type", pt, ":hours")
      slope <- hours_coef + fixed_effects[interaction_term]
    }
    
    # Calculate predictions
    pred <- intercept + slope * hours_grid
    
    # Add to predictions dataframe
    predictions <- rbind(predictions, data.frame(
      player_type = pt,
      hours = hours_grid,
      predicted = pred
    ))
  }
  
  # Create the plot
  p <- ggplot(predictions, aes(x = hours, y = predicted, color = player_type)) +
    geom_line(size = 1.2) +
    geom_ribbon(aes(ymin = predicted - 1.96 * 0.5, ymax = predicted + 1.96 * 0.5, fill = player_type), alpha = 0.2) +
    labs(
      title = paste("Interaction Plot for", outcome_name),
      x = "Hours Played",
      y = outcome_name,
      color = "Player Type",
      fill = "Player Type"
    ) +
    theme_minimal() +
    theme(
      plot.title = element_text(hjust = 0.5, face = "bold"),
      legend.position = "bottom",
      panel.grid.minor = element_blank()
    ) +
    scale_color_brewer(palette = "Set2") +
    scale_fill_brewer(palette = "Set2")
  
  return(p)
}

# Function to create Johnson-Neyman plot
create_jn_plot <- function(model, moderator = "player_type", predictor = "hours", outcome_name) {
  # For rlmerMod models, we'll use a custom approach
  # This function creates a simplified version of Johnson-Neyman plot
  
  # Extract fixed effects
  fixed_effects <- fixef(model)
  vcov_matrix <- vcov(model)
  
  # Get player types
  player_types <- levels(pvz_df_clustered$player_type)
  
  # Create a dataframe to store slopes and confidence intervals
  jn_data <- data.frame(
    player_type = player_types,
    slope = numeric(length(player_types)),
    se = numeric(length(player_types)),
    lower_ci = numeric(length(player_types)),
    upper_ci = numeric(length(player_types)),
    stringsAsFactors = FALSE
  )
  
  # Get the coefficient for hours (main effect)
  hours_coef <- fixed_effects["hours"]
  hours_se <- sqrt(vcov_matrix["hours", "hours"])
  
  # For each player type, calculate slope and CI
  for (i in 1:length(player_types)) {
    pt <- player_types[i]
    
    if (i == 1) {
      # Reference level
      jn_data$slope[i] <- hours_coef
      jn_data$se[i] <- hours_se
    } else {
      # Other levels
      interaction_term <- paste0("player_type", pt, ":hours")
      slope <- hours_coef + fixed_effects[interaction_term]
      
      # Calculate SE using delta method (simplified)
      var_hours <- vcov_matrix["hours", "hours"]
      var_inter <- vcov_matrix[interaction_term, interaction_term]
      cov_hours_inter <- vcov_matrix["hours", interaction_term]
      slope_se <- sqrt(var_hours + var_inter + 2*cov_hours_inter)
      
      jn_data$slope[i] <- slope
      jn_data$se[i] <- slope_se
    }
    
    # Calculate 95% confidence intervals
    jn_data$lower_ci[i] <- jn_data$slope[i] - 1.96 * jn_data$se[i]
    jn_data$upper_ci[i] <- jn_data$slope[i] + 1.96 * jn_data$se[i]
  }
  
  # Determine significance
  jn_data$significant <- (jn_data$lower_ci > 0) | (jn_data$upper_ci < 0)
  
  # Create the plot
  p <- ggplot(jn_data, aes(x = player_type, y = slope, color = significant)) +
    geom_point(size = 3) +
    geom_errorbar(aes(ymin = lower_ci, ymax = upper_ci), width = 0.2) +
    geom_hline(yintercept = 0, linetype = "dashed", color = "gray50") +
    labs(
      title = paste("Conditional Effect of", predictor, "on", outcome_name),
      subtitle = "by Player Type with 95% Confidence Intervals",
      x = "Player Type",
      y = paste("Effect of", predictor),
      color = "Significant"
    ) +
    theme_minimal() +
    theme(
      plot.title = element_text(hjust = 0.5, face = "bold"),
      plot.subtitle = element_text(hjust = 0.5),
      legend.position = "bottom",
      panel.grid.minor = element_blank()
    ) +
    scale_color_manual(values = c("gray50", "blue"))
  
  return(p)
}
```

```{r wellbeing-simple-slopes}
#| label: wellbeing-simple-slopes
#| message: false
#| warning: false

# Calculate simple slopes for game-specific well-being model
wellbeing_slopes <- calculate_all_simple_slopes(rmodel_moderation_game, "Game-Specific Well-being")

# Create a formatted table for game-specific well-being simple slopes
wellbeing_slopes_table <- wellbeing_slopes %>%
  mutate(
    estimate = round(estimate, 3),
    std.error = round(std.error, 3),
    statistic = round(statistic, 3),
    p.value = case_when(
      p.value < 0.001 ~ "< 0.001",
      TRUE ~ round(p.value, 3) %>% as.character()
    ),
    ci = paste0("[", round(conf.low, 3), ", ", round(conf.high, 3), "]")
  ) %>%
  select(player_type, estimate, std.error, statistic, p.value, ci, stars)

kable(wellbeing_slopes_table, 
      caption = "Simple Slopes Analysis for Game-Specific Well-being",
      col.names = c("Player Type", "Estimate", "SE", "t-value", "p-value", "95% CI", ""),
      align = "lrrrrl") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), 
                full_width = FALSE, position = "center")
```

```{r enjoyment-simple-slopes}
#| label: enjoyment-simple-slopes
#| message: false
#| warning: false

# Calculate simple slopes for enjoyment model
enjoyment_slopes <- calculate_all_simple_slopes(rmodel_moderation_enjoyment, "Intrinsic Enjoyment")

# Create a formatted table for enjoyment simple slopes
enjoyment_slopes_table <- enjoyment_slopes %>%
  mutate(
    estimate = round(estimate, 3),
    std.error = round(std.error, 3),
    statistic = round(statistic, 3),
    p.value = case_when(
      p.value < 0.001 ~ "< 0.001",
      TRUE ~ round(p.value, 3) %>% as.character()
    ),
    ci = paste0("[", round(conf.low, 3), ", ", round(conf.high, 3), "]")
  ) %>%
  select(player_type, estimate, std.error, statistic, p.value, ci, stars)

kable(enjoyment_slopes_table, 
      caption = "Simple Slopes Analysis for Intrinsic Enjoyment",
      col.names = c("Player Type", "Estimate", "SE", "t-value", "p-value", "95% CI", ""),
      align = "lrrrrl") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), 
                full_width = FALSE, position = "center")
```

```{r wellbeing-pairwise-comparisons}
#| label: wellbeing-pairwise-comparisons
#| message: false
#| warning: false

# Create pairwise comparison of slopes for game-specific well-being
wellbeing_pairs <- data.frame(player_type1 = character(),
                        player_type2 = character(),
                        difference = numeric(),
                        std.error = numeric(),
                        statistic = numeric(),
                        p.value = numeric(),
                        stringsAsFactors = FALSE)

player_types <- levels(pvz_df_clustered$player_type)
for (i in 1:(length(player_types)-1)) {
  for (j in (i+1):length(player_types)) {
    pt1 <- player_types[i]
    pt2 <- player_types[j]
    
    # Get slopes and standard errors
    slope1 <- wellbeing_slopes$estimate[wellbeing_slopes$player_type == pt1]
    slope2 <- wellbeing_slopes$estimate[wellbeing_slopes$player_type == pt2]
    se1 <- wellbeing_slopes$std.error[wellbeing_slopes$player_type == pt1]
    se2 <- wellbeing_slopes$std.error[wellbeing_slopes$player_type == pt2]
    
    # Calculate difference and standard error
    diff <- slope1 - slope2
    diff_se <- sqrt(se1^2 + se2^2)  # Simplified approach, assumes independence
    
    # Calculate test statistic and p-value
    t_stat <- diff / diff_se
    p_val <- 2 * pt(abs(t_stat), df = nrow(pvz_df_clustered) - 10, lower.tail = FALSE)  # Approximate df
    
    # Add to results
    wellbeing_pairs <- rbind(wellbeing_pairs, data.frame(
      player_type1 = pt1,
      player_type2 = pt2,
      difference = diff,
      std.error = diff_se,
      statistic = t_stat,
      p.value = p_val,
      stringsAsFactors = FALSE
    ))
  }
}

# Format and display pairwise comparison table for subjective well-being
wellbeing_pairs <- wellbeing_pairs %>%
  mutate(
    stars = case_when(
      p.value < 0.001 ~ "***",
      p.value < 0.01 ~ "**",
      p.value < 0.05 ~ "*", 
      p.value < 0.1 ~ "†",
      TRUE ~ ""
    ),
    difference = round(difference, 3),
    std.error = round(std.error, 3),
    statistic = round(statistic, 3),
    p.value = case_when(
      p.value < 0.001 ~ "< 0.001",
      TRUE ~ round(p.value, 3) %>% as.character()
    )
  )

kable(wellbeing_pairs, 
      caption = "Pairwise Comparison of Simple Slopes for Game-Specific Well-being",
      col.names = c("Player Type 1", "Player Type 2", "Difference", "SE", "t-value", "p-value", ""),
      align = "llrrrrl") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), 
                full_width = FALSE, position = "center")
```

```{r enjoyment-pairwise-comparisons}
#| label: enjoyment-pairwise-comparisons
#| message: false
#| warning: false

# Create pairwise comparison of slopes for enjoyment model
enjoyment_pairs <- data.frame(player_type1 = character(),
                              player_type2 = character(),
                              difference = numeric(),
                              std.error = numeric(),
                              statistic = numeric(),
                              p.value = numeric(),
                              stringsAsFactors = FALSE)

for (i in 1:(length(player_types)-1)) {
  for (j in (i+1):length(player_types)) {
    pt1 <- player_types[i]
    pt2 <- player_types[j]
    
    # Get slopes and standard errors
    slope1 <- enjoyment_slopes$estimate[enjoyment_slopes$player_type == pt1]
    slope2 <- enjoyment_slopes$estimate[enjoyment_slopes$player_type == pt2]
    se1 <- enjoyment_slopes$std.error[enjoyment_slopes$player_type == pt1]
    se2 <- enjoyment_slopes$std.error[enjoyment_slopes$player_type == pt2]
    
    # Calculate difference and standard error
    diff <- slope1 - slope2
    diff_se <- sqrt(se1^2 + se2^2)  # Simplified approach, assumes independence
    
    # Calculate test statistic and p-value
    t_stat <- diff / diff_se
    p_val <- 2 * pt(abs(t_stat), df = nrow(pvz_df_clustered) - 10, lower.tail = FALSE)  # Approximate df
    
    # Add to results
    enjoyment_pairs <- rbind(enjoyment_pairs, data.frame(
      player_type1 = pt1,
      player_type2 = pt2,
      difference = diff,
      std.error = diff_se,
      statistic = t_stat,
      p.value = p_val,
      stringsAsFactors = FALSE
    ))
  }
}

# Format and display pairwise comparison table for enjoyment model
enjoyment_pairs <- enjoyment_pairs %>%
  mutate(
    stars = case_when(
      p.value < 0.001 ~ "***",
      p.value < 0.01 ~ "**",
      p.value < 0.05 ~ "*", 
      p.value < 0.1 ~ "†",
      TRUE ~ ""
    ),
    difference = round(difference, 3),
    std.error = round(std.error, 3),
    statistic = round(statistic, 3),
    p.value = case_when(
      p.value < 0.001 ~ "< 0.001",
      TRUE ~ round(p.value, 3) %>% as.character()
    )
  )

kable(enjoyment_pairs, 
      caption = "Pairwise Comparison of Simple Slopes for Enjoyment",
      col.names = c("Player Type 1", "Player Type 2", "Difference", "SE", "t-value", "p-value", ""),
      align = "llrrrrl") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), 
                full_width = FALSE, position = "center")
```

## Interaction Plots

```{r interaction-plots}
#| label: interaction-plots
#| message: false
#| warning: false
#| fig.width: 12
#| fig.height: 6
#| fig.cap: "Interaction plots showing the relationship between hours played and outcomes across player types"

# Create scatter plots for total playtime vs. outcomes across player types with APA-style formatting
wellbeing_plot <- ggplot(pvz_df_clustered, aes(x = hours, y = spane_game_balance, color = player_type)) +
  geom_point(alpha = 0.6) +
  geom_smooth(method = "lm", se = TRUE, alpha = 0.2, fullrange = TRUE) +
  scale_color_brewer(palette = "Set2") +
  scale_y_continuous(limits = c(-6, 6), breaks = seq(-6, 6, 2)) +
  labs(
    x = "Total Hours Played",
    y = "SPANE Score",
    color = "Player Type"
  ) +
  # Apply APA-style formatting
  theme_minimal() +
  theme(
    plot.title = element_text(size = 12, face = "bold", hjust = 0),
    plot.subtitle = element_text(size = 10, hjust = 0),
    axis.title = element_text(size = 11),
    axis.text = element_text(size = 10),
    panel.grid.minor = element_blank(),
    panel.border = element_rect(fill = NA, color = "black", size = 0.5),
    panel.background = element_rect(fill = "white")
  )

enjoyment_plot <- ggplot(pvz_df_clustered, aes(x = hours, y = enjoyment, color = player_type)) +
  geom_point(alpha = 0.6) +
  geom_smooth(method = "lm", se = TRUE, alpha = 0.2, fullrange = TRUE) +
  scale_color_brewer(palette = "Set2") +
  scale_y_continuous(limits = c(1, 7), breaks = seq(1, 7, 1)) +
  labs(
    x = "Total Hours Played",
    y = "Enjoyment Score",
    color = "Player Type"
  ) +
  # Apply APA-style formatting
  theme_minimal() +
  theme(
    plot.title = element_text(size = 12, face = "bold", hjust = 0),
    plot.subtitle = element_text(size = 10, hjust = 0),
    axis.title = element_text(size = 11),
    axis.text = element_text(size = 10),
    panel.grid.minor = element_blank(),
    panel.border = element_rect(fill = NA, color = "black", size = 0.5),
    panel.background = element_rect(fill = "white")
  )

# Combine the plots with common legend at bottom
ggarrange(wellbeing_plot, enjoyment_plot, ncol = 2, common.legend = TRUE, legend = "bottom")
```

## Johnson-Neyman Plots

```{r johnson-neyman-plots}
#| label: johnson-neyman-plots
#| message: false
#| warning: false
#| fig.width: 10
#| fig.height: 6
#| fig.cap: "Johnson-Neyman plots showing conditional effects of hours played by player type on Game-Specific Well-being and Enjoyment"

# Load patchwork if not already loaded
if (!requireNamespace("patchwork", quietly = TRUE)) {
  library(patchwork)
}

# Create Johnson-Neyman plots with consistent formatting following APA standards
jn_wellbeing <- create_jn_plot(rmodel_moderation_game, outcome_name = "Game-Specific Well-being") +
  # Apply APA-style formatting
  theme_minimal() +
  theme(
    plot.title = element_text(size = 12, face = "bold", hjust = 0),
    plot.subtitle = element_text(size = 10, hjust = 0),
    axis.title = element_text(size = 11),
    axis.text = element_text(size = 10),
    legend.position = "none",
    panel.grid.minor = element_blank(),
    panel.border = element_rect(fill = NA, color = "black", size = 0.5),
    panel.background = element_rect(fill = "white")
  ) +
  # Modify labels to be more concise for combined plot
  labs(
    title = "A: Game-Specific Well-being",
    subtitle = NULL,
    y = "Effect of Hours Played"
  )

jn_enjoyment <- create_jn_plot(rmodel_moderation_enjoyment, outcome_name = "Enjoyment") +
  # Apply APA-style formatting
  theme_minimal() +
  theme(
    plot.title = element_text(size = 12, face = "bold", hjust = 0),
    plot.subtitle = element_text(size = 10, hjust = 0),
    axis.title = element_text(size = 11),
    axis.text = element_text(size = 10),
    legend.title = element_text(size = 10, face = "bold"),
    legend.text = element_text(size = 9),
    panel.grid.minor = element_blank(),
    panel.border = element_rect(fill = NA, color = "black", size = 0.5),
    panel.background = element_rect(fill = "white")
  ) +
  # Modify labels to be more concise for combined plot
  labs(
    title = "B: Enjoyment",
    subtitle = NULL,
    y = "Effect of Hours Played"
  )

# Combine the plots with shared legend at the bottom
combined_plot <- jn_wellbeing + jn_enjoyment + 
  plot_layout(ncol = 2, guides = "collect") & 
  theme(legend.position = "bottom",
        legend.title = element_text(size = 10, face = "bold"),
        legend.text = element_text(size = 9))

# Display the combined plot
combined_plot
```

To further understand the moderation effects, we conducted simple slopes analyses to examine how the relationship between hours played and well-being outcomes varied across the three player types.

For game-specific well-being, hours played had a significant positive association for Casual PVE Players (b = 0.142, SE = 0.042, p \< .001, 95% CI \[0.059, 0.225\]), indicating that increased playtime was associated with higher well-being for this player type. For Casual PVP Players, hours played did not significantly predict well-being (b = -0.032, SE = 0.048, p = .505, 95% CI \[-0.126, 0.062\]), suggesting that playtime had no reliable effect on well-being for these players. For Elite Allround Players, hours played showed a significant positive relationship with well-being (b = 0.126, SE = 0.053, p = .018, 95% CI \[0.022, 0.230\]), similar to the pattern observed for Casual PVE Players.

Pairwise comparisons of these slopes revealed significant differences between player types (see Appendix 1). The relationship between hours played and well-being was significantly stronger for Casual PVE Players compared to Casual PVP Players (difference = 0.174, SE = 0.064, p \< .001). Similarly, the slope for Elite Allround Players was significantly stronger than for Casual PVP Players (difference = 0.158, SE = 0.071, p = .027). No significant difference was found between the slopes for Casual PVE Players and Elite Allround Players (difference = 0.016, SE = 0.068, p = .815).

For enjoyment, a similar pattern emerged. Casual PVE Players showed a significant positive relationship between hours played and enjoyment (b = 0.108, SE = 0.033, p = .001, 95% CI \[0.043, 0.173\]). For Casual PVP Players, no significant relationship was found (b = -0.045, SE = 0.038, p = .237, 95% CI \[-0.120, 0.030\]). Elite Allround Players showed a significant positive relationship between hours played and enjoyment (b = 0.086, SE = 0.042, p = .041, 95% CI \[0.004, 0.168\]), though the magnitude was slightly smaller than for Casual PVE Players.

Pairwise comparisons of these slopes also revealed significant differences (see Appendix 1). The relationship between hours played and enjoyment was significantly stronger for Casual PVE Players compared to Casual PVP Players (difference = 0.153, SE = 0.050, p = .002). Similarly, the slope for Elite Allround Players was significantly stronger than for Casual PVP Players (difference = 0.131, SE = 0.057, p = .021). No significant difference was found between the slopes for Casual PVE Players and Elite Allround Players (difference = 0.022, SE = 0.053, p = .679).

### Visualization of Interaction Effects

The interaction plots (Figure 6) visually confirmed these differential relationships, showing positive slopes for Casual PVE Players and Elite Allround Players, and flat or slightly negative slopes for Casual PVP Players for both outcome variables. These visualizations highlighted that the relationship between playtime and well-being outcomes was moderated by player type.

The Johnson-Neyman plots (Figure 7) further illustrated these conditional effects by displaying the estimated effect of hours played on each outcome variable across player types, with confidence intervals. These plots confirmed that the effect of hours played was significantly different from zero for Casual PVE Players and Elite Allround Players, but not for Casual PVP Players.

The heatmap visualizations of pairwise differences in simple slopes (Appendix 1) provided an additional perspective on these findings. The heatmaps highlighted the significant differences between Casual PVP Players and the other two player types, with color gradients indicating the magnitude and direction of these differences.

# Discussion

The aims of this study were two-fold. First, to challenge the assumption of homogeneity amongst players by demonstrating the existence of distinct playing styles based on behavioral telemetry. Second, to highlight the importance of accounting for player-specific playing style differences in analyzing the psychosocial outcomes of video game play. Taken as a whole, these results demonstrate that even players of the same game do not have a homogenous playing style and that subgroups with distinct playing styles can be clearly identified by clustering in-game behavioral telemetry data. We also found a slight positive relationship between overall active playtime and well-being, confirming the findings of previous studies [@johannesVideoGamePlay2021]. Furthermore, we found that playing styles are related to well-being and that Casual PVE Players reported significantly higher well-being scores compared to Casual PVP Players. Most importantly, significant moderation effects for player types were found, demonstrating that the relationship between playtime and well-being varies according to a player's playing style.

Through PAM (Partitioning Around Medoids) clustering of behavioral telemetry data, three distinct playing styles were identified. The Casual PVE Players (who preferred cooperative gameplay against AI opponents), Elite Allround Players (who demonstrated high performance across both PVP and PVE contexts), and Casual PVP Players (who favored competitive gameplay but exhibited only moderate performance). These clusters are conceptually similar to player segments previously identified in other shooter games [@drachenGunsSwordsData2012; @melhartYourGameplaySays2019]. For instance, @drachenGunsSwordsData2012 identified player types in Battlefield 2: Bad Company 2 that parallel our Casual Players and Elite players. Similarly, @melhartYourGameplaySays2019 identified comparable profiles in Tom Clancy's The Division, including players focused on PVE content and those excelling in competitive play. The consistency of these patterns across different games suggests that player typologies may represent fundamental approaches to gameplay rather than game-specific behaviors. These findings challenge the often-made assumption of homogeneous play behavior and instead show that it is an uninformative oversimplification. This finding suggests that there is no one prototypical player and that generalizing over a player base of even a single game has very little ecological validity.

Our robust linear mixed-effects models revealed significant differences in well-being and enjoyment across player types. Compared to Casual PVE Players, Casual PVP Players reported significantly lower well-being scores (b = -0.76, SE = 0.17, t = -4.39) and enjoyment levels (b = -0.72, SE = 0.13, t = -5.39). Elite Allround Players showed no significant differences from Casual PVE Players on either outcome measure. These findings indicate that different styles of engagement with the same game are associated with different psychological experiences, highlighting the importance of considering playing style as a factor in gaming research.

More importantly, our interaction models demonstrated that player type significantly moderated the relationship between playtime and psychological outcomes. For Casual PVE Players, hours played showed a significant positive association with both well-being (b = 0.14, SE = 0.04, t = 3.36) and enjoyment (b = 0.11, SE = 0.03, t = 3.24). For Elite Allround Players, we observed similar positive relationships between hours played and both well-being (b = 0.13, SE = 0.05, t = 2.38) and enjoyment (b = 0.09, SE = 0.04, t = 2.05). However, for Casual PVP Players, hours played showed no significant relationship with either well-being (b = -0.03, SE = 0.05, t = -0.67) or enjoyment (b = -0.05, SE = 0.04, t = -1.18).

The simple slopes analyses further clarified these differential relationships. The effect of playtime on well-being and enjoyment was significantly stronger for both Casual PVE Players and Elite Allround Players compared to Casual PVP Players, with no significant difference between Casual PVE Players and Elite Allround Players. This pattern suggests that the psychological benefits of increased playtime may be contingent on how players engage with the game. Those who excel across different gameplay contexts (Elite Allround Players) or focus on cooperative gameplay (Casual PVE Players) appear to derive greater psychological benefits from increased playtime compared to those who prefer competitive play but perform moderately (Casual PVP Players).

These findings have important implications for the playtime paradigm that has dominated video game effects research. They demonstrate that the psychosocial outcomes of playing video games cannot be attributed to playtime alone without considering playing styles. The relationship between playtime and well-being is not uniform across all players but varies significantly based on how players engage with the game. This may help explain inconsistencies in previous research, where studies focusing on different player populations might have captured different segments of these varied relationships.

The stronger relationship between playtime and well-being for Elite Allround Players and Casual PVE Players might be explained through the lens of need satisfaction theory. Previous work has shown that player performance in video games positively affects psychological competence [@klimmt2009a; @przybylskiMotivationalModelVideo2010; @tamboriniMediaEnjoymentNeed2011], and competence is a strong predictor of increased subjective well-being [@ryanHappinessHumanPotentials2001]. Elite Allround Players, who demonstrate high skill levels across gameplay contexts, likely experience greater competence satisfaction during play. Similarly, Casual PVE Players may derive satisfaction from mastering cooperative challenges without the potentially frustrating competition against other skilled players that characterizes PVP gameplay.

## Limitations and Future Directions

While the study provided valuable insights, we must acknowledge several limitations. First, the results should be interpreted within the context of this specific game, player population, and timeframe. We lack evidence that playing styles remain consistent across different games or time periods. Unlike personality traits, playing styles may be contextual and fluid, varying based on the game environment and a player's development over time. Second, while our behavioral telemetry provided comprehensive metrics on core gameplay mechanics and performance, it captured only a subset of potential player behaviors. Notably absent were metrics on social interactions, an important aspect of multiplayer gaming experiences. Third, the cross-sectional nature of our data limits causal inferences about the relationship between playing styles, playtime, and well-being.

Future research should address these limitations by examining the stability of playing styles across different games and over time, incorporating metrics on social interactions and other qualitative aspects of gameplay, and employing longitudinal or experimental designs to clarify causal relationships. Additionally, investigating the mechanisms through which different playing styles might influence well-being would be valuable. For instance, studies could directly measure need satisfaction (competence, autonomy, and relatedness) across different player types to test whether these mediate the relationship between playing style and well-being outcomes.

## Conclusion

The rise of video games to become the dominant form of entertainment across the globe has raised important questions about the potential benefits and costs of digital play. Understanding how and to what extent playing video games influences the well-being of its players concerns players, caregivers, game developers, and policymakers alike. To date, the literature has mostly followed a playtime paradigm, measuring and recommending how specific limits on playtime might minimize negative effects on players. However, our findings suggest that relying on playtime alone is a misguided oversimplification. This study demonstrates that player types, derived from actual in-game behavioral data, significantly moderate the relationship between playtime and well-being outcomes. The same amount of playtime may be associated with different psychological outcomes depending on how players engage with the game. These findings call for a more nuanced approach to understanding the effects of gaming on well-being, one that considers not just how much people play, but how they play. As richer forms of play data become available through industry-academic collaborations, researchers should move beyond the playtime paradigm to develop more sophisticated models of player experience and its relationship to psychological outcomes.


# References

<!-- References will auto-populate in the refs div below -->

# Appendix

## Appendix 1a: Visual Assessment of Cluster Tendency (VAT)

```{r appendix_vat_plots}
#| label: appendix_vat_plots
#| fig.width: 10
#| fig.height: 8
#| echo: false
#| warning: false
#| message: false

# Display VAT plots individually
print(dist_plot_real)
print(dist_plot_random)
```

## Appendix 1b: PCA Visualizations for Cluster Tendency

```{r appendix_pca_plots}
#| label: appendix_pca_plots
#| fig.width: 10
#| fig.height: 8
#| echo: false
#| warning: false
#| message: false

# Display PCA plots individually
print(pca_plot_real)
print(pca_plot_random)
```

## Appendix 2: Cluster Validation Plots

```{r appendix_cluster_validation_plots}
#| label: appendix_cluster_validation_plots
#| fig.width: 10
#| fig.height: 6
#| echo: false
#| warning: false
#| message: false

# Display the cluster validation plots that were created earlier
print(combined_pam_plots)

# Also display individual plots for better clarity
cat("\n### Silhouette Analysis\n")
print(p_pam_silhouette)

cat("\n### Calinski-Harabasz Index\n")
print(p_pam_ch)
```

The plots above show the validation methods used to determine the optimal number of clusters. The Silhouette method (top left) measures how well observations fit within their assigned clusters compared to neighboring clusters, with higher values indicating better fit. The Calinski-Harabasz Index (top right) evaluates cluster validity based on the ratio of between-cluster to within-cluster variance, with higher values indicating better separation between clusters. Both methods indicated that k = 3 was the optimal number of clusters for our player data.

## Appendix 3: Heatmaps of Slope Differences

```{r slope-heatmaps}
#| label: slope-heatmaps
#| message: false
#| warning: false
#| fig.width: 10
#| fig.height: 8
#| fig.cap: "Heatmaps of pairwise differences in simple slopes across player types"

# Create heatmap of slope differences for subjective well-being
slope_diff_matrix_wellbeing <- matrix(0, nrow = length(player_types), ncol = length(player_types))
p_matrix_wellbeing <- matrix(1, nrow = length(player_types), ncol = length(player_types))

for (i in 1:length(player_types)) {
  for (j in 1:length(player_types)) {
    if (i != j) {
      # Find the correct pair in the pairwise comparison data
      if (i < j) {
        pair_row <- wellbeing_pairs[wellbeing_pairs$player_type1 == player_types[i] & 
                               wellbeing_pairs$player_type2 == player_types[j], ]
        slope_diff_matrix_wellbeing[i, j] <- pair_row$difference
        p_matrix_wellbeing[i, j] <- as.numeric(ifelse(pair_row$p.value == "< 0.001", 0.001, pair_row$p.value))
      } else {
        pair_row <- wellbeing_pairs[wellbeing_pairs$player_type1 == player_types[j] & 
                               wellbeing_pairs$player_type2 == player_types[i], ]
        slope_diff_matrix_wellbeing[i, j] <- -pair_row$difference  # Negate the difference
        p_matrix_wellbeing[i, j] <- as.numeric(ifelse(pair_row$p.value == "< 0.001", 0.001, pair_row$p.value))
      }
    }
  }
}

# Create a data frame for the heatmap
heatmap_data_wellbeing <- expand.grid(PlayerType1 = player_types, PlayerType2 = player_types)
heatmap_data_wellbeing$Difference <- as.vector(slope_diff_matrix_wellbeing)
heatmap_data_wellbeing$Significant <- as.vector(p_matrix_wellbeing) < 0.05

# Create the heatmap
heat_wellbeing <- ggplot(heatmap_data_wellbeing, aes(x = PlayerType1, y = PlayerType2, fill = Difference)) +
  geom_tile() +
  geom_text(aes(label = ifelse(PlayerType1 != PlayerType2, 
                              sprintf("%.3f%s", Difference, 
                                     ifelse(Significant, "*", "")), ""))) +
  scale_fill_gradient2(low = "blue", mid = "white", high = "red", midpoint = 0) +
  labs(
    title = "Heatmap of Slope Differences for Game-Specific Well-being",
    subtitle = "Positive values indicate row player type has stronger slope than column player type",
    x = "Player Type",
    y = "Player Type",
    fill = "Difference"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold"),
    plot.subtitle = element_text(hjust = 0.5),
    axis.text.x = element_text(angle = 45, hjust = 1)
  )

heat_wellbeing

# Create heatmap of slope differences for enjoyment model
slope_diff_matrix_enjoyment <- matrix(0, nrow = length(player_types), ncol = length(player_types))
p_matrix_enjoyment <- matrix(1, nrow = length(player_types), ncol = length(player_types))

for (i in 1:length(player_types)) {
  for (j in 1:length(player_types)) {
    if (i != j) {
      # Find the correct pair in the pairwise comparison data
      if (i < j) {
        pair_row <- enjoyment_pairs[enjoyment_pairs$player_type1 == player_types[i] & 
                                   enjoyment_pairs$player_type2 == player_types[j], ]
        slope_diff_matrix_enjoyment[i, j] <- pair_row$difference
        p_matrix_enjoyment[i, j] <- as.numeric(ifelse(pair_row$p.value == "< 0.001", 0.001, pair_row$p.value))
      } else {
        pair_row <- enjoyment_pairs[enjoyment_pairs$player_type1 == player_types[j] & 
                                   enjoyment_pairs$player_type2 == player_types[i], ]
        slope_diff_matrix_enjoyment[i, j] <- -pair_row$difference  # Negate the difference
        p_matrix_enjoyment[i, j] <- as.numeric(ifelse(pair_row$p.value == "< 0.001", 0.001, pair_row$p.value))
      }
    }
  }
}

# Create a data frame for the heatmap
heatmap_data_enjoyment <- expand.grid(PlayerType1 = player_types, PlayerType2 = player_types)
heatmap_data_enjoyment$Difference <- as.vector(slope_diff_matrix_enjoyment)
heatmap_data_enjoyment$Significant <- as.vector(p_matrix_enjoyment) < 0.05

# Create the heatmap
heat_enjoyment <- ggplot(heatmap_data_enjoyment, aes(x = PlayerType1, y = PlayerType2, fill = Difference)) +
  geom_tile() +
  geom_text(aes(label = ifelse(PlayerType1 != PlayerType2, 
                              sprintf("%.3f%s", Difference, 
                                     ifelse(Significant, "*", "")), ""))) +
  scale_fill_gradient2(low = "blue", mid = "white", high = "red", midpoint = 0) +
  labs(
    title = "Heatmap of Slope Differences for Intrinsic Enjoyment",
    subtitle = "Positive values indicate row player type has stronger slope than column player type",
    x = "Player Type",
    y = "Player Type",
    fill = "Difference"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold"),
    plot.subtitle = element_text(hjust = 0.5),
    axis.text.x = element_text(angle = 45, hjust = 1)
  )

heat_enjoyment
```