---
title: "Player Playstyle Analysis (3 Clusters)"
author: "Thomas Hakman"
format: 
  html:
    toc: true
    toc-depth: 3
    code-fold: true
    theme: flatly
    fig-width: 8
    fig-height: 6
    embed-resources: true    # Bundle all resources in the HTML file
    self-contained: true     # Create a standalone HTML file
execute:
  echo: true
  message: false
  warning: false
  cache: false
  eval: true                 # Ensure code is evaluated
  output: true               # Ensure output is displayed
  fig-show: "asis"           # Ensure figures are displayed inline
editor: visual
---

```{r setup}
#| label: setup
#| include: false

# Set global options
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  cache = FALSE,
  fig.show = "asis"  # Ensures figures are displayed inline in the viewer
)

# Basic httpgd setup for VSCode (if available)
if (interactive() && requireNamespace("httpgd", quietly = TRUE)) {
  library(httpgd)
  if (!httpgd::hgd_is_active()) {
    httpgd::hgd_open()
  }
  options(vsc.plot = FALSE)  # Configure VSCode to use httpgd
  options(device = function() httpgd::hgd())
}

# Define a capitalize function for facet labels
capitalize <- function(x) {
  x <- as.character(x)
  substr(x, 1, 1) <- toupper(substr(x, 1, 1))
  x
}

# Create a list of wellbeing variables (focus on our two primary outcomes)
wellbeing_vars <- c("spane_game_balance", "enjoyment")
```

# Abstract

This document presents a comprehensive analysis of playstyle patterns and their relationship with well-being among players of Plants vs. Zombies: Battle for Neighborville. Using cluster analysis, we identified three distinct player types based on their gameplay behavior and preferences. Through robust linear mixed-effects modeling, we examined how these playstyles moderate the relationship between gameplay time and well-being metrics, controlling for player country as a random effect. This analysis excludes participants with outlier values to ensure more robust findings.

Our findings reveal that the relationship between playtime and well-being differs significantly across player types, suggesting that how players engage with the game is potentially more important than simply how much they play. This analysis focuses specifically on two key outcomes: game-related well-being (measured by the game SPANE balance score) and intrinsic enjoyment. These two measures show unique patterns across player types, highlighting how different playstyles may influence player experience and enjoyment.

These results have important implications for game design, suggesting that personalized approaches accommodating different playstyles may enhance player well-being. The analysis contributes to our understanding of digital well-being by moving beyond simplistic "screen time" measures to consider the quality and nature of digital engagement. Limitations include the cross-sectional nature of the data and the potential need for additional moderating variables in future research.

# Introduction

This document contains the analysis of player telemetry and survey data from Plants vs. Zombies: Battle for Neighborville. It focuses on analyzing player playstyles, game performance metrics, and their relationship with well-being outcomes. In this analysis, we've removed participants with outlier values (identified as NA in the dataset) to ensure more robust and reliable findings.

# Load Required Packages and Data

```{r install-packages, include=FALSE}
#| label: install-packages

# First ensure pacman is installed if not already
if (!requireNamespace("pacman", quietly = TRUE)) {
  install.packages("pacman")
}

# Use pacman to load all required packages (will install any missing packages)
pacman::p_load(
  tidyverse,  # For data manipulation and visualization
  here,       # For project-relative paths
  lubridate,  # For date/time manipulation
  scales,     # For scale transformations
  janitor,    # For data cleaning
  magrittr,   # For additional pipe operators
  psych,      # For psychological statistics
  corrplot,   # For correlation plots
  ggpubr,     # For publication-ready plots
  factoextra, # For cluster analysis visualization
  cluster,    # For clustering algorithms
  fmsb,       # For radar charts
  knitr,      # For dynamic report generation
  tibble,     # For improved data frames
  kableExtra, # For enhanced table formatting
  fpc,        # For cluster validation
  gridExtra,  # For arranging multiple plots
  hopkins,    # For cluster tendency assessment
  effectsize, # For effect size calculations
  robustlmm,  # For robust linear mixed-effects models
  broom,      # For tidy model outputs
  broom.mixed, # For tidying mixed-effects model output
  patchwork,  # For combining plots
  ggeffects,  # For visualizing regression model predictions
  ggplot2,    # For data visualization (explicitly included though part of tidyverse)
  dplyr,      # For data manipulation (explicitly included though part of tidyverse)
  tidyr,      # For data tidying (explicitly included though part of tidyverse)
  stringr     # For string manipulation (explicitly included though part of tidyverse)
)
```

```{r helper-functions}
#| label: helper-functions

# Function to scale values to 0-1 range for radar charts
scale_to_01 <- function(x, min_val = min(x, na.rm = TRUE), max_val = max(x, na.rm = TRUE)) {
  if (max_val == min_val) return(rep(0.5, length(x)))
  (x - min_val) / (max_val - min_val)
}

```

```{r statistical-functions}
#| label: statistical-functions

# Function to run ANOVA and extract results
run_anova <- function(dv) {
  formula <- as.formula(paste(dv, "~ player_type"))
  model <- aov(formula, data = pvz_data)
  result <- broom::tidy(model)
  result$eta_squared <- effectsize::eta_squared(model)[["Eta2"]]
  result$variable <- dv
  return(result)
}

# Function to compute Hopkins statistic for cluster tendency
compute_hopkins <- function(data, n = nrow(data)-1) {
  # Standardize data for clustering
  data_scaled <- scale(data)
  
  # Compute Hopkins statistic
  hopkins_result <- hopkins::hopkins(data_scaled, n = n)
  
  return(hopkins_result)
}
```

```{r load-processed-data}
#| label: load-processed-data

# Load the dataset with outliers removed
pvz_data <- read_csv(here("data/processed/pvz_clean_no_outliers.csv"))

# Create a copy of pvz_data for analysis to maintain consistency
pvz_df_analysis <- pvz_data

# Load the pre-clustered data (we'll recreate this with the no-outliers dataset)
# Since we're using a new dataset, we'll create a new clustered dataset instead of using the existing one
# pvz_df_clustered <- read_csv(here("data/processed/pvz_with_clusters_pam_nohours.csv"))

glimpse(pvz_df_analysis)
```

```{r check-dataframe}
#| label: check-dataframe
# Examining the dataframe with outliers removed
glimpse(pvz_df_analysis)
```

```{r select-columns}
#| label: select-columns

pvz_df_analysis <- pvz_df_analysis %>%
  select(
    # Demographic variables
    player_id, country, age, gender, played_with_others,
    
    # Well-being variables - focus only on our two key outcomes
    spane_game_balance, enjoyment,
    
    # Total playtime
    hours,
    
    # PVP metrics
    ratio_pvp_hours,
    pvp_kd,
    pvp_hit_accuracy,
    pvp_dps,
    pvp_average_damage,
    pvp_average_score,
    
    # PVE metrics
    ratio_pve_hours,
    pve_kd,
    pve_hit_accuracy,
    pve_dps,
    pve_average_damage,
    pve_average_score
  )
```

# Filter for players with telemetry data

```{r telemetry-data-filter}
#| label: telemetry-data-filter
#| include: false

# Filter for players with valid performance metrics (> 0)
pvz_df_analysis <- pvz_df_analysis %>%
  filter(
    ratio_pvp_hours > 0 | ratio_pve_hours > 0 | pvp_kd > 0 | pve_kd > 0 | pvp_hit_accuracy > 0 | pve_hit_accuracy > 0 | pvp_dps > 0 | pve_dps > 0 | pvp_average_damage > 0 | pve_average_damage > 0 | pvp_average_score > 0 | pve_average_score > 0
  )

# Remove players with less than 0.25 hours (15 minutes) of total playtime
pvz_df_analysis <- pvz_df_analysis %>%
  filter(hours >= 0.25)

# Remove any rows with NA values in key analysis variables
pvz_df_analysis <- pvz_df_analysis %>%
  filter(!is.na(ratio_pvp_hours) & !is.na(ratio_pve_hours) & 
         !is.na(pvp_kd) & !is.na(pve_kd) & 
         !is.na(pvp_hit_accuracy) & !is.na(pve_hit_accuracy) &
         !is.na(pvp_dps) & !is.na(pve_dps) &
         !is.na(pvp_average_damage) & !is.na(pve_average_damage) &
         !is.na(pvp_average_score) & !is.na(pve_average_score))

# Show the number of players left
print(paste("Number of players left after filtering:", nrow(pvz_df_analysis)))
```

We started with players from the cleaned dataset with outliers removed. After filtering for players with valid performance metrics, those with at least 15 minutes of total playtime, and removing any remaining NA values, we are left with `r nrow(pvz_df_analysis)` players for our analysis.

# Demographics

```{r survey-demographics-calculation}
#| label: survey-demographics-calculation
#| include: false

# Calculate demographics for the filtered dataset
demographics <- pvz_df_analysis %>%
  summarize(
    total_participants = n(),
    avg_age = round(mean(age, na.rm = TRUE), 1),
    sd_age = round(sd(age, na.rm = TRUE), 1)
  )

# Calculate gender distribution in filtered dataset
gender_distribution <- pvz_df_analysis %>%
  count(gender) %>%
  mutate(percentage = n / sum(n) * 100)

# Count participants with each gender
male_count <- gender_distribution %>% filter(gender == "Male") %>% pull(n)
female_count <- gender_distribution %>% filter(gender == "Female") %>% pull(n)
other_count <- gender_distribution %>% filter(gender == "Other") %>% pull(n)
undisclosed_count <- gender_distribution %>% filter(gender == "Prefer not to say") %>% pull(n)
```

After filtering for players with valid performance metrics and those with at least 15 minutes of total playtime, our final sample consists of `r nrow(pvz_df_analysis)` participants. The demographic characteristics of this filtered sample include an average age of `r demographics$avg_age` (SD = `r demographics$sd_age`). The gender distribution includes `r male_count` male, `r female_count` female, `r other_count` other, and `r undisclosed_count` participants who preferred not to disclose their gender.

```{r final-sample-demographics}
#| label: final-sample-demographics

# Display demographics of final sample
cat("Demographics of final sample:
")
cat("- Sample size:", nrow(pvz_df_analysis), "
")
cat("- Average age:", round(mean(pvz_df_analysis$age, na.rm = TRUE), 1), "(SD =", round(sd(pvz_df_analysis$age, na.rm = TRUE), 1), ")
")

# Gender distribution
gender_dist <- pvz_df_analysis %>%
  count(gender) %>%
  mutate(percentage = n / sum(n) * 100)

kable(gender_dist, caption = "Gender distribution in final sample", digits = 2)

# Print summary of final dataset
glimpse(pvz_df_analysis)
```

# Exploratory Data Analysis

## plot session timeline for a player

```{r session-timeline-helper}
#| label: session-timeline-helper
#| include: false

# Function to plot session timeline for a player
plot_player_sessions <- function(player_data, player_id = NULL, player_label = "Player") {
  # If a player_id is provided, filter data for that player
  if (!is.null(player_id)) {
    player_data <- player_data %>%
      filter(player_id == !!player_id)
  }
  
  # Convert timestamps to POSIXct if they're not already
  if (!inherits(player_data$game_start_time, "POSIXct")) {
    player_data$game_start_time <- as.POSIXct(player_data$game_start_time, format = "%Y-%m-%d %H:%M:%S")
    player_data$game_end_time <- as.POSIXct(player_data$game_end_time, format = "%Y-%m-%d %H:%M:%S")
  }
  
  # Extract day of week, date, and time component
  player_data$day_of_week <- weekdays(player_data$game_start_time)
  player_data$date <- as.Date(player_data$game_start_time)
  player_data$start_time_only <- as.POSIXct(format(player_data$game_start_time, format = "%H:%M:%S"), format = "%H:%M:%S")
  player_data$end_time_only <- as.POSIXct(format(player_data$game_end_time, format = "%H:%M:%S"), format = "%H:%M:%S")
  
  # Identify sessions that cross midnight (where end_time is earlier than start_time)
  player_data$crosses_midnight <- player_data$end_time_only < player_data$start_time_only
  
  # Create a new dataframe that will include split sessions for those crossing midnight
  player_data_expanded <- player_data %>%
    filter(!crosses_midnight) # Keep all sessions that don't cross midnight
  
  # Handle sessions that cross midnight by splitting them
  midnight_sessions <- player_data %>%
    filter(crosses_midnight)
  
  if (nrow(midnight_sessions) > 0) {
    # For each session crossing midnight, create two rows:
    # 1. From start time to 23:59:59
    # 2. From 00:00:00 to end time
    midnight <- as.POSIXct(strptime("23:59:59", format = "%H:%M:%S"), format = "%H:%M:%S")
    midnight_start <- as.POSIXct(strptime("00:00:00", format = "%H:%M:%S"), format = "%H:%M:%S")
    
    # First part: from start time to midnight
    midnight_part1 <- midnight_sessions
    midnight_part1$end_time_only <- midnight
    
    # Second part: from midnight start to end time
    midnight_part2 <- midnight_sessions
    midnight_part2$start_time_only <- midnight_start
    # The second part is on the next day of the week and the next date
    midnight_part2$day_of_week <- factor(
      weekdays(midnight_part2$game_start_time + 86400), # Add one day (86400 seconds)
      levels = c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday")
    )
    midnight_part2$date <- midnight_part2$date + 1
    
    # Add both parts to the expanded dataset
    player_data_expanded <- bind_rows(player_data_expanded, midnight_part1, midnight_part2)
  }
  
  # Recode match_type based on game_level_key 
  player_data_expanded$match_type <- "0"  # Initialize with 0 as in the original code
  
  # Social matches
  player_data_expanded$match_type[player_data_expanded$game_level_key == "dsub_social_terrain_beach"] <- "social match"
  player_data_expanded$match_type[player_data_expanded$game_level_key == "dsub_social_terrain_summer"] <- "social match"
  player_data_expanded$match_type[player_data_expanded$game_level_key == "dsub_socialspace"] <- "social match"
  
  # PVE matches
  player_data_expanded$match_type[player_data_expanded$game_level_key == "dsub_region_cheesemines"] <- "pve match"
  player_data_expanded$match_type[player_data_expanded$game_level_key == "dsub_region_dreadwood"] <- "pve match"
  player_data_expanded$match_type[player_data_expanded$game_level_key == "dsub_region_towncenter"] <- "pve match" 
  
  # PVP matches (Rush)
  player_data_expanded$match_type[player_data_expanded$game_level_key == "dsub_rush_dreadwood"] <- "pvp match (rush)"
  player_data_expanded$match_type[player_data_expanded$game_level_key == "dsub_rush_towncenter"] <- "pvp match (rush)"
  player_data_expanded$match_type[player_data_expanded$game_level_key == "dsub_rush_farm"] <- "pvp match (rush)"
  player_data_expanded$match_type[player_data_expanded$game_level_key == "dsub_rush_cheesemines"] <- "pvp match (rush)"
  player_data_expanded$match_type[player_data_expanded$game_level_key == "dsub_rush_shoreline"] <- "pvp match (rush)"
  player_data_expanded$match_type[player_data_expanded$game_level_key == "dsub_rush_resort"] <- "pvp match (rush)"
  
  # PVP matches (Vanquish)
  player_data_expanded$match_type[player_data_expanded$game_level_key == "dsub_tvm_cheesemines"] <- "pvp match (vanquish)"
  player_data_expanded$match_type[player_data_expanded$game_level_key == "dsub_tvm_zrefinery"] <- "pvp match (vanquish)"
  player_data_expanded$match_type[player_data_expanded$game_level_key == "dsub_tvm_suburbs"] <- "pvp match (vanquish)"
  player_data_expanded$match_type[player_data_expanded$game_level_key == "dsub_tvm_dreadwood"] <- "pvp match (vanquish)"
  
  # PVP matches (Arena)
  player_data_expanded$match_type[player_data_expanded$game_level_key == "dsub_4v4_battlearena_v2"] <- "pvp match (arena)"
  player_data_expanded$match_type[player_data_expanded$game_level_key == "dsub_4v4_superarena"] <- "pvp match (arena)"
  
  # Podium celebration
  player_data_expanded$match_type[player_data_expanded$game_level_key == "dsub_podiumcelebration"] <- "podium celebration"
  
  # Ensure day of week is a factor with ordered levels
  player_data_expanded$day_of_week <- factor(
    player_data_expanded$day_of_week,
    levels = c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday")
  )
  
  # Create a position adjustment for stacking sessions by different dates for the same day of week
  player_data_expanded <- player_data_expanded %>%
    group_by(day_of_week) %>%
    mutate(
      # Get unique dates for this day of week and assign a position index to each date
      unique_dates = list(unique(date)),
      date_index = match(date, unique(date)),
      # Scale the vertical offset to be small but noticeable
      y_offset = (date_index - 1) * 0.2,  # Creates a small vertical offset for different dates
      # Create adjusted y-position that includes the offset
      y_position = as.numeric(day_of_week) + y_offset
    ) %>%
    ungroup()
  
  # Sort by day of week, date, and start time
  player_data_expanded <- player_data_expanded %>%
    arrange(day_of_week, date, game_start_time, game_end_time)
  
  # Add a session ID within each day to distinguish multiple sessions on same day
  player_data_expanded <- player_data_expanded %>%
    group_by(day_of_week, date) %>%
    mutate(session_id_within_day = row_number()) %>%
    ungroup()
  
  # Convert match_type to factor with proper levels for consistent ordering
  player_data_expanded$match_type <- factor(
    player_data_expanded$match_type,
    levels = c("social match", "pve match", "pvp match (rush)", "pvp match (vanquish)", "pvp match (arena)", "podium celebration", "0"),
    labels = c("Social Lobby", "PvE Match", "PvP (Rush)", "PvP (Vanquish)", "PvP (Arena)", "Podium Celebration", "Other")
  )
  
  # Create the plot with enhanced styling
  plot <- player_data_expanded %>%
    ggplot2::ggplot(aes(start_time_only, y_position, col = match_type, 
                        group = interaction(day_of_week, date, session_id_within_day))) +
    # Use improved color palette with better contrast
    ggplot2::scale_color_manual(
      name = "Match Type", 
      values = c(
        "Social Lobby" = "#00BFC4",       # Teal
        "PvE Match" = "#7CAE00",          # Green
        "PvP (Rush)" = "#F8766D",         # Red
        "PvP (Vanquish)" = "#FF9E1C",     # Orange
        "PvP (Arena)" = "#FFD700",        # Gold
        "Podium Celebration" = "#C77CFF", # Purple
        "Other" = "#A0A0A0"               # Gray
      )
    ) +
    # Thicker segments with rounded ends for better visibility
    ggplot2::geom_segment(
      aes(xend = end_time_only, yend = y_position),
      linewidth = 2,
      lineend = "round"
    ) +
    # Add points at start/end for better visibility of short sessions
    ggplot2::geom_point(size = 1.5, alpha = 0.6) +
    ggplot2::geom_point(aes(x = end_time_only), size = 1.5, alpha = 0.6) +
    # Better time axis formatting - using fixed limits for consistency
    ggplot2::scale_x_datetime(
      labels = scales::date_format("%H:%M"),
      breaks = scales::date_breaks("2 hours"),
      limits = c(
        as.POSIXct(strptime("00:00", format = "%H:%M")),
        as.POSIXct(strptime("23:59", format = "%H:%M"))
      ),
      expand = c(0.01, 0.01),
      date_minor_breaks = "1 hour"
    ) +
    # Use the original day of week as the y-axis labels
    ggplot2::scale_y_continuous(
      breaks = 1:7,
      labels = levels(player_data_expanded$day_of_week),
      limits = c(0.5, 7.5)
    ) +
    # Empty labs - axis labels will be added to the combined figure instead
    ggplot2::labs(
      x = NULL, 
      y = NULL
    ) +
    # Improved theme with better typography and layout
    ggplot2::theme_minimal(base_size = 11, base_family = "sans") +
    ggplot2::theme(
      legend.position = "none", # Move legend to the combined figure
      plot.margin = margin(30, 10, 30, 10), # Increased margins for better separation
      
      axis.text.y = element_text(face = "bold"),
      axis.title = element_text(face = "bold"),
      
      panel.grid.major.x = element_line(color = "grey90"),
      panel.grid.minor.x = element_line(color = "grey95"),
      panel.grid.major.y = element_line(color = "grey90"),
      panel.grid.minor.y = element_blank(),
      
      panel.background = element_rect(fill = "white", color = NA),
      plot.background = element_rect(fill = "white", color = NA)
    )
  
  return(plot)
}
```

```{r session-timeline-data-prep}
#| label: session-timeline-data-prep
#| include: false
#| cache: false

# Load game session data if not already included
game_sessions <- read_csv(here("data/processed/game_time_2week.csv"))

# Make sure timestamps are in the correct format
game_sessions$game_start_time <- as.POSIXct(game_sessions$game_start_time, format = "%Y-%m-%d %H:%M:%S")
game_sessions$game_end_time <- as.POSIXct(game_sessions$game_end_time, format = "%Y-%m-%d %H:%M:%S")

# First, get a list of player IDs in the data
player_ids <- unique(game_sessions$player_id)

# Choose specific indices
# If those indices don't exist in our data, we'll use the first three players
player_indices <- c(
  min(401, length(player_ids)),  # Equivalent to slice(401) 
  min(6, length(player_ids)),    # Equivalent to slice(6)
  min(3, length(player_ids))     # Equivalent to slice(3)
)

selected_player_ids <- player_ids[player_indices]
```

```{r create-individual-plots}
#| label: create-individual-plots
#| include: false
#| cache: false
#| dependson: session-timeline-data-prep

# Create individual plots for each player with standardized x-axis
plot1 <- plot_player_sessions(
  game_sessions %>% filter(player_id == selected_player_ids[1])
)

plot2 <- plot_player_sessions(
  game_sessions %>% filter(player_id == selected_player_ids[2])
)

# For the third plot, use the same time range for consistency
plot3 <- plot_player_sessions(
  game_sessions %>% filter(player_id == selected_player_ids[3])
)
```

```{r combine-session-plots}
#| label: combine-session-plots
#| fig.width: 12
#| fig.height: 16
#| dependson: create-individual-plots

# Combine plots into a single figure with enhanced styling and proper spacing
figure <- ggarrange(
  plot2, plot3, plot1,
  labels = c("Player A", "Player B", "Player C"),
  font.label = list(size = 14, face = "bold", color = "black"),
  ncol = 1, 
  nrow = 3, 
  common.legend = TRUE, 
  legend = "bottom",
  heights = c(1, 1, 1),
  align = "v",
  hjust = -0.1
)

# Add shared x and y labels to the combined figure
fig_sessiontime <- annotate_figure(figure,
                        bottom = text_grob("Time of Day (hours)", 
                                          face = "bold", size = 12),
                        left = text_grob("Day of Week", 
                                         face = "bold", size = 12, rot = 90))

# Display the combined figure
fig_sessiontime
```

# Descriptive Statistics

```{r apa-descriptive-stats}
#| label: tbl-descriptive-stats
#| tbl-cap: "Descriptive Statistics for Key Variables"
#| fig.width: 10
#| fig.height: 6
#| tbl-cap-location: top
#| echo: false
#| warning: false
#| message: false

# Create a simple APA-style descriptive statistics table
# First, check variable types to handle them appropriately
variable_types <- sapply(pvz_df_analysis, class)

# Handle numeric variables
numeric_vars <- pvz_df_analysis %>%
  select(where(is.numeric)) %>%
  select(
    # Demographics
    age,
    
    # Playtime 
    hours, ratio_pvp_hours, ratio_pve_hours,
    
    # Performance metrics
    pvp_kd, pve_kd, 
    pvp_hit_accuracy, pve_hit_accuracy,
    pvp_dps, pve_dps,
    pvp_average_damage, pve_average_damage,
    pvp_average_score, pve_average_score,
    
    # Well-being and enjoyment
    spane_game_balance, enjoyment
  )

# Calculate descriptive statistics for numeric variables
numeric_stats <- numeric_vars %>%
  summarise(across(everything(), 
                  list(
                    n = ~sum(!is.na(.)),
                    M = ~mean(., na.rm = TRUE),
                    SD = ~sd(., na.rm = TRUE),
                    Min = ~min(., na.rm = TRUE),
                    Max = ~max(., na.rm = TRUE)
                  ))) %>%
  pivot_longer(everything(), 
               names_to = c("variable", "stat"), 
               names_pattern = "(.*)_(.*)",
               values_to = "value") %>%
  pivot_wider(names_from = stat, values_from = value)

# For categorical variables like gender, just count frequencies for the footnote
if ("gender" %in% names(pvz_df_analysis)) {
  gender_counts <- pvz_df_analysis %>%
    count(gender) %>%
    mutate(pct = n / sum(n) * 100) %>%
    mutate(pct = round(pct, 1)) %>%
    mutate(frequency = paste0(n, " (", pct, "%)"))
  
  # Note: We'll add a footnote to explain gender distribution separately
  gender_note <- paste("Gender distribution: ", 
                     paste(paste0(gender_counts$gender, ": ", gender_counts$frequency), 
                           collapse = "; "))
} else {
  gender_note <- NULL
}

# Format the data for the table
descriptive_stats <- numeric_stats %>%
  # Format variable names for better readability
  mutate(variable = case_when(
    variable == "age" ~ "Age",
    variable == "hours" ~ "Total Hours Played",
    variable == "ratio_pvp_hours" ~ "Proportion of PvP Playtime",
    variable == "ratio_pve_hours" ~ "Proportion of PvE Playtime",
    variable == "pvp_kd" ~ "PvP Kill/Death Ratio",
    variable == "pve_kd" ~ "PvE Kill/Death Ratio",
    variable == "pvp_hit_accuracy" ~ "PvP Hit Accuracy",
    variable == "pve_hit_accuracy" ~ "PvE Hit Accuracy",
    variable == "pvp_dps" ~ "PvP Damage per Second",
    variable == "pve_dps" ~ "PvE Damage per Second",
    variable == "pvp_average_damage" ~ "PvP Average Damage per Game",
    variable == "pve_average_damage" ~ "PvE Average Damage per Game",
    variable == "pvp_average_score" ~ "PvP Average Score",
    variable == "pve_average_score" ~ "PvE Average Score",
    variable == "spane_game_balance" ~ "Game-Specific Well-being (SPANE)",
    variable == "enjoyment" ~ "Enjoyment",
    TRUE ~ variable
  )) %>%
  # Create category variable for grouping
  mutate(category = case_when(
    variable == "Age" ~ "Demographics",
    variable %in% c("Total Hours Played", "Proportion of PvP Playtime", "Proportion of PvE Playtime") ~ "Playtime",
    variable %in% c("PvP Kill/Death Ratio", "PvE Kill/Death Ratio", 
                   "PvP Hit Accuracy", "PvE Hit Accuracy",
                   "PvP Damage per Second", "PvE Damage per Second",
                   "PvP Average Damage per Game", "PvE Average Damage per Game",
                   "PvP Average Score", "PvE Average Score") ~ "Performance",
    variable %in% c("Game-Specific Well-being (SPANE)", "Enjoyment") ~ "Well-being",
    TRUE ~ "Other"
  )) %>%
  # Round numeric values
  mutate(
    M = round(M, 2),
    SD = round(SD, 2),
    Min = round(Min, 2),
    Max = round(Max, 2)
)

# Create custom ordering for variables to ensure autonomy, competence, relatedness are listed in order
variable_order <- c(
  # Demographics
  "Age",
  
  # Playtime
  "Total Hours Played", "Proportion of PvP Playtime", "Proportion of PvE Playtime",
  
  # Performance
  "PvP Kill/Death Ratio", "PvE Kill/Death Ratio", 
  "PvP Hit Accuracy", "PvE Hit Accuracy",
  "PvP Damage per Second", "PvE Damage per Second",
  "PvP Average Damage per Game", "PvE Average Damage per Game",
  "PvP Average Score", "PvE Average Score",
  
  # Well-being
  "Game-Specific Well-being (SPANE)", "Enjoyment"
)

# Apply the custom ordering
descriptive_stats <- descriptive_stats %>%
  mutate(variable = factor(variable, levels = variable_order)) %>%
  arrange(variable)

# Create the APA-styled table with kable
apa_table <- descriptive_stats %>%
  select(variable, n, M, SD, Min, Max) %>%
  kable(
    # Caption is handled by chunk options now
    col.names = c("Variable", "*N*", "*M*", "*SD*", "Min", "Max"),
    align = c("l", "c", "c", "c", "c", "c"),
    booktabs = TRUE,
    format = "html",
    digits = 2
  ) %>%
  kable_styling(
    bootstrap_options = c("striped", "hover", "condensed"),
    full_width = FALSE,
    position = "left",
    font_size = 12
  ) %>%
  row_spec(0, bold = TRUE) %>%  # Bold header row per APA
  pack_rows(index = table(factor(descriptive_stats$category, 
                              levels = unique(descriptive_stats$category))), 
            hline_after = TRUE)  # Add lines after each group

# Add footnote for gender if it exists
if (!is.null(gender_note)) {
  apa_table <- apa_table %>%
    footnote(general = gender_note, general_title = "Note: ")
}

# Display the table
apa_table
```

# Player Clustering Analysis with PAM

This section identifies distinct player types through clustering analysis based on gameplay metrics. We use PAM (Partitioning Around Medoids) because it's more robust to outliers than K-means.

## Feature Selection and Preparation

```{r cluster_feature_selection}
#| label: cluster_feature_selection
#| echo: true
#| warning: false
#| message: false

# Select relevant features for clustering
player_cluster_features <- pvz_df_analysis %>% 
  select(
    # Playstyle variables (game mode preferences)
    ratio_pvp_hours, ratio_pve_hours,
    
    # PvP performance metrics
    pvp_kd, pvp_hit_accuracy, pvp_dps, pvp_average_damage, pvp_average_score,
    
    # PvE performance metrics
    pve_kd, pve_hit_accuracy, pve_dps, pve_average_damage, pve_average_score
  )

# Remove any NAs in the dataset to ensure clean clustering
player_cluster_features <- na.omit(player_cluster_features)

# Standardize the data (important for clustering)
player_data_scaled <- scale(player_cluster_features)
rownames(player_data_scaled) <- pvz_df_analysis$player_id[rownames(player_data_scaled) %in% rownames(pvz_df_analysis)]
```

## Assessing Cluster Tendency

Before performing clustering, we assess whether natural clusters exist in the data.

```{r cluster_tendency_assessment}
#| label: cluster_tendency_assessment
#| fig.width: 12
#| fig.height: 12
#| echo: true
#| warning: false
#| message: false
#| results: "asis"
#| fig.show: "asis"

# Create a random dataset with the same dimensions for comparison
set.seed(123) # For reproducibility
random_cluster_data <- apply(player_data_scaled, 2,
                     function(x){runif(length(x), min(x), max(x))})
random_cluster_data <- as.data.frame(random_cluster_data)

# Visual assessment using PCA visualization
pca_real_data <- prcomp(player_data_scaled)
pca_random_data <- prcomp(random_cluster_data)

# Create PCA plots for comparison
pca_plot_real <- fviz_pca_ind(pca_real_data, 
                  title = "PCA - Player Performance Data",
                  geom = "point", 
                  ggtheme = theme_minimal(),
                  legend = "none",
                  alpha.ind = 0.7) +
      theme(plot.title = element_text(face = "bold", size = 14))

pca_plot_random <- fviz_pca_ind(pca_random_data, 
                  title = "PCA - Random Data",
                  geom = "point", 
                  ggtheme = theme_minimal(),
                  legend = "none",
                  alpha.ind = 0.7) +
      theme(plot.title = element_text(face = "bold", size = 14))

# Visual assessment using distance matrices
dist_real_data <- get_dist(player_data_scaled, method = "euclidean")
dist_random_data <- get_dist(random_cluster_data, method = "euclidean")

# Create distance matrix plots
dist_plot_real <- fviz_dist(dist_real_data, 
                show_labels = FALSE,
                gradient = list(low = "#00AFBB", mid = "white", high = "#FC4E07")) +
       ggtitle("Distance Matrix - Player Data") +
       theme(plot.title = element_text(face = "bold", size = 14))

dist_plot_random <- fviz_dist(dist_random_data, 
                show_labels = FALSE,
                gradient = list(low = "#00AFBB", mid = "white", high = "#FC4E07")) +
       ggtitle("Distance Matrix - Random Data") +
       theme(plot.title = element_text(face = "bold", size = 14))

# Arrange plots in a 2x2 grid for better comparison
combined_plots <- grid.arrange(
  pca_plot_real, pca_plot_random,
  dist_plot_real, dist_plot_random,
  ncol = 2
)


# Also ensure the plot is displayed if rendered
print(combined_plots)

# Interpretation text
cat("## Interpretation of Clustering Tendency

",
    "1. **PCA Visualization**: If the real data shows more visible grouping than random data, natural clusters likely exist.

",
    "2. **Distance Matrix**: In data with cluster structure, the distance matrix shows dark blocks along the diagonal,
",
    "   indicating groups of observations that are similar to each other.

")
```

## Determining Optimal Number of Clusters


```{r pam_silhouette_method}
#| label: pam_silhouette_method
#| fig.width: 10
#| fig.height: 4
#| cache: false
#| results: "asis"
#| fig.show: "asis"

# PAM optimal cluster determination
# -------------------------------------
# Silhouette method for PAM
pam_sil <- function(data, max_k = 10) {
  sil_values <- numeric(max_k - 1)
  for (i in 2:max_k) {
    pam_result <- pam(data, k = i, metric = "euclidean")
    sil_values[i-1] <- pam_result$silinfo$avg.width
  }
  k_values <- 2:max_k
  return(data.frame(k = k_values, sil = sil_values))
}

pam_sil_results <- pam_sil(player_data_scaled)
p_pam_silhouette <- ggplot(pam_sil_results, aes(x = k, y = sil)) +
  geom_line() +
  geom_point() +
  scale_x_continuous(breaks = 2:10) +
  labs(title = "PAM: Silhouette Method",
       x = "Number of clusters k",
       y = "Average silhouette width") +
  theme_minimal()

# Calinski-Harabasz Index (CH index) for PAM
ch_index_pam <- function(data, max_k = 10) {
  ch_values <- numeric(max_k - 1)
  for (i in 2:max_k) {
    pam_result <- pam(data, k = i)
    ch_values[i-1] <- fpc::calinhara(data, pam_result$clustering)
  }
  k_values <- 2:max_k
  return(data.frame(k = k_values, ch_index = ch_values))
}

ch_pam_results <- ch_index_pam(player_data_scaled)
p_pam_ch <- ggplot(ch_pam_results, aes(x = k, y = ch_index)) +
  geom_line() +
  geom_point() +
  scale_x_continuous(breaks = 2:10) +
  labs(title = "PAM: Calinski-Harabasz Index",
       x = "Number of clusters k",
       y = "CH Index (higher is better)") +
  theme_minimal()

# Arrange PAM validation plots in a grid
combined_pam_plots <- grid.arrange(p_pam_silhouette, p_pam_ch,
             ncol = 2, nrow = 1,
             top = grid::textGrob("Determining Optimal Number of Clusters with PAM", 
                                 gp = grid::gpar(fontsize = 16, font = 2)))

# Also ensure the plot is displayed if rendered
print(combined_pam_plots)

# Print summary of results for easier interpretation
cat("
Summary of Optimal k Recommendations for PAM:
- Silhouette Method: Highest peak suggests optimal k =", pam_sil_results$k[which.max(pam_sil_results$sil)], "
- Calinski-Harabasz Index: Highest value suggests optimal k =", ch_pam_results$k[which.max(ch_pam_results$ch_index)], "
")
```

```{r determine_optimal_k, echo=FALSE}
# Determine the optimal k values
silhouette_optimal_k <- pam_sil_results$k[which.max(pam_sil_results$sil)]
ch_optimal_k <- ch_pam_results$k[which.max(ch_pam_results$ch_index)]

# Determine final optimal k (using most common value or a specific method if preferred)
if(silhouette_optimal_k == ch_optimal_k) {
  final_optimal_k <- silhouette_optimal_k
} else {
  # In case of disagreement, you might want to choose one method or use another strategy
  # Here we'll use silhouette as the deciding factor, but you could modify this logic
  final_optimal_k <- silhouette_optimal_k
}
```

Based on the analysis above, we determined that `r final_optimal_k` is the optimal number of clusters for this dataset. This was established through multiple methods including silhouette analysis (optimal k = `r silhouette_optimal_k`) and Calinski-Harabasz Index (optimal k = `r ch_optimal_k`)`r if(silhouette_optimal_k == ch_optimal_k) {", which both supported the same cluster count"}`. We proceed with PAM (Partitioning Around Medoids) clustering using k=`r final_optimal_k`.

```{r pam_cluster_statistics}
#| label: pam_cluster_statistics
#| cache: false
#| results: "asis"

# Calculate key statistics for the PAM cluster with k=3
pam_cluster_stats <- function(data, k) {
  set.seed(123)
  pam_result <- pam(data, k = k)
  
  # Calculate silhouette coefficient for each cluster
  sil <- pam_result$silinfo$widths
  sil_per_cluster <- tapply(sil[, 3], sil[, 1], mean)
  
  # Return results
  return(list(
    k = k,
    avg_silhouette = pam_result$silinfo$avg.width,
    silhouette_per_cluster = sil_per_cluster,
    cluster_sizes = table(pam_result$clustering),
    medoid_indices = pam_result$id.med
  ))
}

# Get statistics for k=3 using PAM
stats_pam_k3 <- pam_cluster_stats(player_data_scaled, 3)

# Print summary of PAM cluster statistics

cat("Cluster Sizes:", "
")
print(stats_pam_k3$cluster_sizes)
cat("Silhouette per Cluster:", "
") 
print(stats_pam_k3$silhouette_per_cluster)
```
PAM (Partitioning Around Medoids) offers several advantages for clustering when outliers are present in the data:

1.  PAM uses actual data points (medoids) as cluster centers, making it less influenced by outliers
2.  PAM minimizes the sum of dissimilarities rather than sum of squared Euclidean distances
3.  PAM is more robust to noise and outliers as it's not affected by extreme values
4.  PAM works with any distance metric, not just Euclidean distance
5.  Medoids are actual observations that represent cluster characteristics well

These properties make PAM particularly suitable for our player behavior data, which may contain outliers due to the diverse nature of gameplay patterns.

## Player Type Clustering with PAM

```{r prepare_player_data}
#| label: prepare_player_data

# Prepare player data for clustering
player_data_features <- pvz_df_analysis %>% 
  select(
    # Game mode preferences
    ratio_pvp_hours, ratio_pve_hours,
    
    # Performance metrics
    pvp_kd, pve_kd, 
    pvp_hit_accuracy, pve_hit_accuracy,
    pvp_average_score, pve_average_score,
    pvp_dps, pve_dps
  )

# Remove any remaining NA values 
player_data_features <- na.omit(player_data_features)

# Print the number of players remaining after NA removal
cat("Number of players included in cluster analysis:", nrow(player_data_features), "\n")

# Standardize the data for clustering
player_data_scaled <- scale(player_data_features)

# Perform PAM clustering with k=3
set.seed(42) # For reproducibility
k_clusters <- 3
pam_result <- pam(player_data_scaled, k = k_clusters, metric = "euclidean")

# Display cluster sizes
cat("Cluster sizes:\n")
print(table(pam_result$clustering))
```

## Cluster Characteristics

```{r cluster_profiles}
#| label: cluster_profiles

# Define the feature columns used in clustering
feature_columns <- c("ratio_pvp_hours", "ratio_pve_hours", 
                    "pvp_kd", "pve_kd", 
                    "pvp_hit_accuracy", "pve_hit_accuracy",
                    "pvp_average_score", "pve_average_score",
                    "pvp_dps", "pve_dps")

# Get cluster medoids (representative players)
medoid_indices <- pam_result$id.med
medoid_data <- pvz_df_analysis[medoid_indices, ]
medoid_features <- medoid_data %>%
  select(all_of(feature_columns))
medoid_features$cluster <- 1:k_clusters

# Calculate mean values by cluster for interpretation
cluster_means <- pvz_df_analysis %>%
  # First add the cluster assignments
  mutate(cluster = pam_result$clustering) %>%
  group_by(cluster) %>%
  summarize(across(all_of(feature_columns), mean, na.rm = TRUE), .groups = "drop")

# Display medoid profiles 
cat("### Cluster Medoids (Representative Players)
")
print(kable(medoid_features, digits = 2, 
      caption = "Cluster Medoids (Feature Values of Representative Players)"))

# Display cluster means
cat("
### Cluster Means (Average Values for Each Cluster)
")
print(kable(cluster_means, digits = 2,
      caption = "Cluster Means - Average Feature Values"))
```

```{r assign_player_types}
#| label: assign_player_types

# Create a data frame with cluster labels 
# Update these names based on interpretation of the 3 clusters
player_type_labels <- tibble(
  cluster = 1:3,
  player_type = c(
    "Casual PVE Players",           # Cluster 1 name
    "Elite Allround Players",       # Cluster 2 name 
    "Casual PVP Players"            # Cluster 3 name
  )
)

# Add cluster assignments and type labels to the dataframe
pvz_df_clustered <- pvz_df_analysis %>%
  mutate(cluster = pam_result$clustering) %>%
  left_join(player_type_labels, by = "cluster") %>%
  mutate(player_type = factor(player_type))

# Save the clustered data with no outliers
write_csv(pvz_df_clustered, here("data/processed/pvz_with_clusters_pam3_nohours_no_outliers.csv"))

# Output summary of player types
cat("Player Types Identified:
")
for (i in 1:nrow(player_type_labels)) {
  cat("Cluster", i, "=", player_type_labels$player_type[i], "
")
}
```

## Visualizing Player Types with Radar Charts

```{r radar_functions}
#| label: radar_functions

# Function to scale values using z-scores
z_score_scale <- function(x) {
  if (sd(x, na.rm = TRUE) == 0) return(rep(0, length(x)))
  return((x - mean(x, na.rm = TRUE)) / sd(x, na.rm = TRUE))
}

# Convert z-scores to 0-1 scale for radar chart
z_to_radar_scale <- function(z, min_z = -2.5, max_z = 2.5) {
  z_clipped <- pmax(pmin(z, max_z), min_z)
  scaled <- (z_clipped - min_z) / (max_z - min_z)
  return(scaled)
}

# Function to create radar chart with average player comparison
create_player_type_radar <- function(data, avg_data, color, title, values) {
  # Get column names
  data_cols <- colnames(data)
  
  # Create max and min rows
  max_min <- data.frame(matrix(c(rep(1, length(data_cols)), rep(0, length(data_cols))), 
                             nrow = 2, byrow = TRUE))
  colnames(max_min) <- data_cols
  
  # Combine data for plotting (max, min, average, player type)
  plot_data <- rbind(max_min, avg_data, data)
  
  # Create radar chart
  radarchart(
    plot_data, 
    pcol = c("gray60", color),        # Colors for average and player type
    pfcol = c(adjustcolor("gray60", alpha.f = 0.2), adjustcolor(color, alpha.f = 0.5)), # Fill colors
    plwd = c(2, 3),                  # Line width for average and player type
    plty = c(2, 1),                  # Line type (dashed for average, solid for player type)
    vlcex = 0.9,                     # Size of labels
    title = title,
    cglcol = "grey80",
    cglty = 1,
    axislabcol = "grey30",
    caxislabels = seq(0, 1, 0.25),
    cglwd = 0.8,
    pty = 16,                        # Use 16-sided polygon for smoother lines
    axistype = 0                     # Label types to display
  )
  
  # Add legend for player type vs average
  legend(
    "bottomright", 
    legend = c("Average Player", paste("Player Type:", title)),
    col = c("gray60", color),
    lty = c(2, 1),
    lwd = c(2, 3),
    bty = "n",     # No box around legend
    cex = 0.7,     # Font size
    pt.cex = 1     # Symbol size
  )
  
  # Add values text
  legend_text <- sprintf("PvP Ratio: %.2f, PvP K/D: %.2f, PvE Ratio: %.2f, PvE K/D: %.2f", 
                       values$`PvP Ratio`, values$`PvP K/D`,
                       values$`PvE Ratio`, values$`PvE K/D`)
  
  mtext(legend_text, side = 1, line = 2, cex = 0.7, col = color)
}
```

```{r radar_charts}
#| label: radar_charts
#| fig.height: 12
#| fig.width: 10

# Define metrics for radar charts
radar_metrics <- c(
  "PvE Ratio", "PvE K/D", "PvE Accuracy", "PvE Score", "PvE DPS",
  "PvP Ratio", "PvP K/D", "PvP Accuracy", "PvP Score", "PvP DPS"
)

# Get player type names
player_type_names <- player_type_labels$player_type

# Calculate radar chart data for both average player and player types
# First collect the raw metrics for all players and player types
metrics_data <- pvz_df_analysis %>%
  # Add a column for the average player
  mutate(profile_type = "Average Player") %>%
  # Bind the clustered data with profile types
  bind_rows(
    pvz_df_clustered %>%
      select(
        player_id, player_type, 
        ratio_pve_hours, ratio_pvp_hours,
        pve_kd, pvp_kd,
        pve_hit_accuracy, pvp_hit_accuracy,
        pve_average_score, pvp_average_score,
        pve_dps, pvp_dps
      ) %>%
      rename(profile_type = player_type)
  ) %>%
  # Group by profile type (average or a player type)
  group_by(profile_type) %>%
  # Calculate means for each metric
  summarise(
    ratio_pve_mean = mean(ratio_pve_hours, na.rm = TRUE),
    pve_kd_mean = mean(pve_kd, na.rm = TRUE),
    pve_accuracy_mean = mean(pve_hit_accuracy, na.rm = TRUE),
    pve_score_mean = mean(pve_average_score, na.rm = TRUE),
    pve_dps_mean = mean(pve_dps, na.rm = TRUE),
    ratio_pvp_mean = mean(ratio_pvp_hours, na.rm = TRUE),
    pvp_kd_mean = mean(pvp_kd, na.rm = TRUE),
    pvp_accuracy_mean = mean(pvp_hit_accuracy, na.rm = TRUE),
    pvp_score_mean = mean(pvp_average_score, na.rm = TRUE),
    pvp_dps_mean = mean(pvp_dps, na.rm = TRUE)
  ) %>%
  ungroup()

# Create vectors of all values for each metric to calculate z-scores across all profiles
all_ratio_pve <- metrics_data$ratio_pve_mean
all_pve_kd <- metrics_data$pve_kd_mean
all_pve_accuracy <- metrics_data$pve_accuracy_mean
all_pve_score <- metrics_data$pve_score_mean
all_pve_dps <- metrics_data$pve_dps_mean
all_ratio_pvp <- metrics_data$ratio_pvp_mean
all_pvp_kd <- metrics_data$pvp_kd_mean
all_pvp_accuracy <- metrics_data$pvp_accuracy_mean
all_pvp_score <- metrics_data$pvp_score_mean
all_pvp_dps <- metrics_data$pvp_dps_mean

# Now apply z-score scaling to create the radar data
radar_data <- metrics_data %>%
  mutate(
    `PvE Ratio` = z_to_radar_scale(z_score_scale(all_ratio_pve)[row_number()]),
    `PvE K/D` = z_to_radar_scale(z_score_scale(all_pve_kd)[row_number()]),
    `PvE Accuracy` = z_to_radar_scale(z_score_scale(all_pve_accuracy)[row_number()]),
    `PvE Score` = z_to_radar_scale(z_score_scale(all_pve_score)[row_number()]),
    `PvE DPS` = z_to_radar_scale(z_score_scale(all_pve_dps)[row_number()]),
    `PvP Ratio` = z_to_radar_scale(z_score_scale(all_ratio_pvp)[row_number()]),
    `PvP K/D` = z_to_radar_scale(z_score_scale(all_pvp_kd)[row_number()]),
    `PvP Accuracy` = z_to_radar_scale(z_score_scale(all_pvp_accuracy)[row_number()]),
    `PvP Score` = z_to_radar_scale(z_score_scale(all_pvp_score)[row_number()]),
    `PvP DPS` = z_to_radar_scale(z_score_scale(all_pvp_dps)[row_number()])
  ) %>%
  # Keep only relevant columns
  select(
    profile_type,
    `PvE Ratio`, `PvE K/D`, `PvE Accuracy`, `PvE Score`, `PvE DPS`,
    `PvP Ratio`, `PvP K/D`, `PvP Accuracy`, `PvP Score`, `PvP DPS`
  )

# Extract the average player data
average_player <- radar_data %>%
  filter(profile_type == "Average Player") %>%
  select(-profile_type) %>%
  as.data.frame()

# Colors for player types (3 colors for 3 clusters)
player_type_colors <- c("#E41A1C", "#377EB8", "#4DAF4A")

# Set up layout for 3 radar charts in a single row
par(mfrow = c(1, 3))

# Create radar charts for each player type
for (i in 1:length(player_type_names)) {
  # Extract data for this player type
  type_data <- radar_data %>%
    filter(profile_type == player_type_names[i]) %>%
    select(-profile_type) %>%
    as.data.frame()
  
  # Get actual values
  actual_values <- pvz_df_clustered %>%
    filter(player_type == player_type_names[i]) %>%
    summarise(
      `PvP Ratio` = mean(ratio_pvp_hours),
      `PvP K/D` = mean(pvp_kd),
      `PvE Ratio` = mean(ratio_pve_hours),
      `PvE K/D` = mean(pve_kd)
    )
  
  # Create radar chart with average player comparison
  create_player_type_radar(
    type_data, 
    average_player,
    player_type_colors[i], 
    player_type_names[i],
    actual_values
  )
}

# Add scaling note
mtext("Note: Values shown as standardized scores (z-scores converted to 0-1 scale)", 
      side = 1, line = -1, outer = TRUE, cex = 0.8)

# Reset layout
par(mfrow = c(1, 1))
```

## Player Type Comparison

```{r player_type_summary}
#| label: player_type_summary

# Create a summary of player types
player_type_summary <- pvz_df_clustered %>%
  group_by(player_type) %>%
  summarise(
    Count = n(),
    `% of Players` = n() / nrow(pvz_df_clustered) * 100,
    `PvP Time %` = mean(ratio_pvp_hours) * 100,
    `PvE Time %` = mean(ratio_pve_hours) * 100,
    `PvP K/D` = mean(pvp_kd),
    `PvE K/D` = mean(pve_kd),
    `PvP Accuracy` = mean(pvp_hit_accuracy) * 100,
    `PvE Accuracy` = mean(pve_hit_accuracy) * 100
  )

# Display summary table
print(kable(player_type_summary, digits = 2,
      caption = "Player Type Summary"))
```

## Interpretation of Player Types

The cluster analysis revealed three distinct player types with the following characteristics:

### Casual PVE Players

Casual PVE Players make up approximately `r round(player_type_summary$"% of Players"[player_type_summary$player_type=="Casual PVE Players"], 1)`% of the player base. These players prefer cooperative gameplay against AI opponents, spending on average `r round(player_type_summary$"PvE Time %"[player_type_summary$player_type=="Casual PVE Players"], 1)`% of their time in PVE modes. They demonstrate good performance in these cooperative scenarios, with moderate PVE kill-to-death ratios (`r round(player_type_summary$"PvE K/D"[player_type_summary$player_type=="Casual PVE Players"], 2)`) and PVE accuracy (`r round(player_type_summary$"PvE Accuracy"[player_type_summary$player_type=="Casual PVE Players"], 1)`%). Their performance metrics in PVP modes are considerably lower, indicating a specialization in cooperative rather than competitive gameplay, though at a more casual skill level.

### Elite Allround Players

Elite Allround Players represent approximately `r round(player_type_summary$"% of Players"[player_type_summary$player_type=="Elite Allround Players"], 1)`% of players. These players demonstrate a more balanced approach to gameplay while maintaining high performance across both game modes. They maintain competitive kill-to-death ratios in both PVP (`r round(player_type_summary$"PvP K/D"[player_type_summary$player_type=="Elite Allround Players"], 2)`) and PVE (`r round(player_type_summary$"PvE K/D"[player_type_summary$player_type=="Elite Allround Players"], 2)`) scenarios, as well as excellent accuracy levels in both contexts. These players appear to have developed versatile, high-level skills that transfer effectively between different gameplay modes, marking them as elite in their overall performance.

### Casual PVP Players

Casual PVP Players comprise approximately `r round(player_type_summary$"% of Players"[player_type_summary$player_type=="Casual PVP Players"], 1)`% of players in our sample. These players show a preference for competitive gameplay, spending on average `r round(player_type_summary$"PvP Time %"[player_type_summary$player_type=="Casual PVP Players"], 1)`% of their time in PVP modes. While they focus on player-versus-player scenarios, their performance metrics such as kill-to-death ratios (`r round(player_type_summary$"PvP K/D"[player_type_summary$player_type=="Casual PVP Players"], 2)`) and accuracy (`r round(player_type_summary$"PvP Accuracy"[player_type_summary$player_type=="Casual PVP Players"], 1)`%) are at a more casual level. Their performance in PVE modes is also moderate, suggesting a casual approach to gameplay across modes but with a preference for competitive play.

These distinct player types represent fundamentally different approaches to engaging with the game, with potential implications for how gameplay experiences relate to psychological well-being outcomes. The subsequent analyses will explore these relationships in detail.


# Statistical Analysis: Robust Mixed-Effects Models

This section examines the relationships between playstyle, playtime, and well-being using robust linear mixed-effects models (rlmm) from the `robustlmm` package. We chose robust methods over standard linear mixed-effects models for several important reasons:

1.  **Outlier Resistance**: Gaming datasets often contain extreme values or outliers (e.g., exceptionally high skill metrics or playtime), which can disproportionately influence standard models. Robust models reduce the impact of these outliers without requiring their removal.

2.  **Non-normal Distributions**: Many of our variables, particularly gameplay metrics like kill/death ratios and accuracy, show non-normal distributions. Robust models are less sensitive to violations of normality assumptions.

3.  **Heteroscedasticity**: The variance in well-being measures often differs across player types and playtime levels. Robust methods handle this heteroscedasticity better than standard approaches.

4.  **Improved Parameter Reliability**: The robust methods implemented in `robustlmm` use Huber-type weighting to downweight the influence of unusual observations, resulting in more stable and reliable parameter estimates.

By using robust models, we can maintain the multilevel structure of our data (accounting for country-level random effects) while ensuring that our findings are not unduly influenced by data irregularities. The models below use the default Huber settings in `robustlmm`, which provide a good balance between efficiency and robustness.

## Effect of Player Type on Well-Being

First, let's examine whether different player types (from our cluster analysis) have different levels of well-being, controlling for total playtime.

```{r player-type-wellbeing-model}
#| label: player-type-wellbeing-model
#| message: false
#| warning: false

# Create a robust linear model for well-being
set.seed(123) # For reproducibility of robust estimates

# Fit robust models for game-specific well-being
rmodel_game <- rlmer(spane_game_balance ~ player_type + hours + gender + age + (1|country), 
                     data = pvz_df_clustered)

# Fit robust models for enjoyment
rmodel_enjoyment <- rlmer(enjoyment ~ player_type + hours + gender + age + (1|country), 
                     data = pvz_df_clustered)

# Summarize the models
cat("
Robust model for subjectivewell-being:
")
summary(rmodel_game)

cat("
Robust model for enjoyment:
")
summary(rmodel_enjoyment)

# Create a nicer formatted summary of the fixed effects
model_coefs_game <- data.frame(coef(summary(rmodel_game)))
model_coefs_game$predictor <- rownames(model_coefs_game)
rownames(model_coefs_game) <- NULL
model_coefs_game <- model_coefs_game %>%
  select(predictor, Estimate, Std..Error, t.value) %>%
  rename(`Std. Error` = Std..Error,
         `t value` = t.value)

model_coefs_enjoyment <- data.frame(coef(summary(rmodel_enjoyment)))
model_coefs_enjoyment$predictor <- rownames(model_coefs_enjoyment)
rownames(model_coefs_enjoyment) <- NULL
model_coefs_enjoyment <- model_coefs_enjoyment %>%
  select(predictor, Estimate, Std..Error, t.value) %>%
  rename(`Std. Error` = Std..Error,
         `t value` = t.value)

# Display nicely formatted tables
kable(model_coefs_game, digits = 3, 
      caption = "Fixed effects for subjective well-being model") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
                full_width = FALSE)

kable(model_coefs_enjoyment, digits = 3, 
      caption = "Fixed effects for enjoyment model") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
                full_width = FALSE)

```

## Playtime and Well-Being: Moderation by Player Type

Now, let's investigate whether the relationship between playtime and well-being depends on player type (moderation effect).

```{r moderation-models}
#| label: moderation-models
#| message: false
#| warning: false

# Create interaction models for our two key well-being outcomes
rmodel_moderation_game <- rlmer(
  spane_game_balance ~ player_type * hours + (1|country),
  data = pvz_df_clustered
)

rmodel_moderation_enjoyment <- rlmer(
  enjoyment ~ player_type * hours + (1|country),
  data = pvz_df_clustered
)

# Summarize the models
cat("
Robust moderation model for subjective well-being:
")
summary(rmodel_moderation_game)

cat("
Robust moderation model for enjoyment:
")
summary(rmodel_moderation_enjoyment)

# Extract fixed effects for the moderation models
mod_coefs_game <- data.frame(coef(summary(rmodel_moderation_game)))
mod_coefs_game$predictor <- rownames(mod_coefs_game)
rownames(mod_coefs_game) <- NULL
mod_coefs_game <- mod_coefs_game %>%
  select(predictor, Estimate, Std..Error, t.value) %>%
  rename(`Std. Error` = Std..Error,
         `t value` = t.value)

mod_coefs_enjoyment <- data.frame(coef(summary(rmodel_moderation_enjoyment)))
mod_coefs_enjoyment$predictor <- rownames(mod_coefs_enjoyment)
rownames(mod_coefs_enjoyment) <- NULL
mod_coefs_enjoyment <- mod_coefs_enjoyment %>%
  select(predictor, Estimate, Std..Error, t.value) %>%
  rename(`Std. Error` = Std..Error,
         `t value` = t.value)

# Display nicely formatted tables
kable(mod_coefs_game, digits = 3, 
      caption = "Fixed effects for subjective well-being moderation model") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
                full_width = FALSE)

kable(mod_coefs_enjoyment, digits = 3, 
      caption = "Fixed effects for enjoyment moderation model") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
                full_width = FALSE)

```

## Visualizing the Moderation Effects

```{r moderation-plots}
#| label: moderation-plots
#| fig.width: 10
#| fig.height: 8
#| message: false
#| warning: false

# Create prediction data frames for plotting
pred_game <- ggpredict(rmodel_moderation_game, terms = c("hours", "player_type"))
pred_enjoyment <- ggpredict(rmodel_moderation_enjoyment, terms = c("hours", "player_type"))

# Game-specific well-being plot
p1 <- ggplot(pred_game, aes(x = x, y = predicted, color = group)) +
  geom_line(size = 1) +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high, fill = group), 
              alpha = 0.1, color = NA) +
  scale_color_brewer(palette = "Set2", name = "Player Type") + 
  scale_fill_brewer(palette = "Set2", guide = "none") +
  labs(
    title = "Relationship between Playtime and Subjective Well-Being by Player Type",
    x = "Hours Played",
    y = "Predicted SPANE Score",
    color = "Player Type"
  ) +
  theme_minimal() +
  theme(legend.position = "bottom")

# Enjoyment plot
p2 <- ggplot(pred_enjoyment, aes(x = x, y = predicted, color = group)) +
  geom_line(size = 1) +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high, fill = group), 
              alpha = 0.1, color = NA) +
  scale_color_brewer(palette = "Set2", name = "Player Type") + 
  scale_fill_brewer(palette = "Set2", guide = "none") +
  labs(
    title = "Relationship between Playtime and Enjoyment by Player Type",
    x = "Hours Played",
    y = "Predicted Enjoyment Score",
    color = "Player Type"
  ) +
  theme_minimal() +
  theme(legend.position = "bottom")

# Arrange plots
ggarrange(p1, p2, ncol = 1, common.legend = TRUE, legend = "bottom")
```

## Detailed Analysis of Well-Being Components

```{r wellbeing-anova}
#| label: wellbeing-anova
#| message: false
#| warning: false

# Run ANOVAs to test for significant differences between player types
wellbeing_vars <- c("spane_game_balance", "enjoyment")

# Store ANOVA results
anova_results <- list()

for (var in wellbeing_vars) {
  formula <- as.formula(paste(var, "~ player_type"))
  anova_model <- aov(formula, data = pvz_df_clustered)
  anova_results[[var]] <- broom::tidy(anova_model)
  anova_results[[var]]$variable <- var
  
  # Add effect size
  anova_results[[var]]$eta_squared <- effectsize::eta_squared(anova_model)[["Eta2"]]
}

# Combine all results
all_anova_results <- bind_rows(anova_results)

# Display results in a nice table
kable(all_anova_results %>% 
        filter(term == "player_type") %>%
        select(variable, df, sumsq, statistic, p.value, eta_squared) %>%
        mutate(
          p.value = format.pval(p.value, digits = 3),
          eta_squared = round(eta_squared, 3)
        ),
      caption = "ANOVA results for well-being variables by player type",
      col.names = c("Variable", "Df", "Sum Sq", "F value", "p-value", "Eta²")) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
                full_width = FALSE)

# Post-hoc tests for variables with significant differences
sig_vars <- character(0)
posthoc_results <- list()

for (var in wellbeing_vars) {
  formula <- as.formula(paste(var, "~ player_type"))
  anova_model <- aov(formula, data = pvz_df_clustered)
  
  if (summary(anova_model)[[1]]["player_type", "Pr(>F)"] < 0.05) {
    sig_vars <- c(sig_vars, var)
    cat(paste("
Post-hoc tests for", var, ":
"))
    posthoc <- TukeyHSD(anova_model)
    posthoc_results[[var]] <- posthoc
    print(posthoc)
  }
}
```

```{r dynamic-results-text}
#| label: dynamic-results-text
#| echo: false
#| results: asis

# Function to format F-statistics in APA style
format_f_stat <- function(f, df1, df2, p, eta_sq) {
  p_formatted <- if (p < .001) "< .001" else paste("=", format(round(p, 3), nsmall = 3))
  paste0("*F*(", df1, ", ", df2, ") = ", format(round(f, 2), nsmall = 2),
         ", *p* ", p_formatted, ", η² = ", format(round(eta_sq, 3), nsmall = 3))
}

# Function to create proper variable names for text
proper_var_name <- function(var_name) {
  name_map <- c(
    "spane_game_balance" = "subjective well-being",
    "enjoyment" = "enjoyment"
  )
  return(name_map[var_name])
}

# Create dynamic text based on results
sig_results <- all_anova_results %>% 
  filter(term == "player_type" & p.value < 0.05) %>%
  arrange(p.value)

nonsig_results <- all_anova_results %>% 
  filter(term == "player_type" & p.value >= 0.05) %>%
  arrange(p.value)

# Count how many variables showed significant differences
sig_count <- nrow(sig_results)
total_count <- length(wellbeing_vars)

# Get the significant variables data
error_df <- all_anova_results %>% 
  filter(term == "Residuals") %>% 
  select(variable, df) %>%
  rename(error_df = df)

# Create introduction text
cat("## Results of Well-Being Component Analysis\n\n")

cat("We conducted a series of one-way analyses of variance (ANOVAs) to examine differences in well-being components across player types. ")

if(sig_count > 0) {
  if(sig_count == total_count) {
    cat("Our analyses revealed significant differences between player types across all measured well-being components.\n\n")
  } else {
    cat(paste0("Our analyses revealed significant differences between player types in ", sig_count, " of the ", total_count, " measured well-being components.\n\n"))
  }
  
  # Detailed results for significant findings
  cat("### Significant Differences Between Player Types\n\n")
  
  for(i in 1:nrow(sig_results)) {
    var <- sig_results$variable[i]
    var_name <- proper_var_name(var)
    f_stat <- sig_results$statistic[i]
    df1 <- sig_results$df[i]
    df2 <- error_df$error_df[error_df$variable == var]
    p_val <- as.numeric(sig_results$p.value[i])
    eta_sq <- sig_results$eta_squared[i]
    
    cat(paste0("Player types differed significantly in terms of ", var_name, ", ", 
               format_f_stat(f_stat, df1, df2, p_val, eta_sq), ". "))
    
    # Add post-hoc test results if available
    if(var %in% names(posthoc_results)) {
      ph_result <- posthoc_results[[var]]$`player_type`
      # Extract significant pairwise comparisons
      sig_pairs <- which(ph_result[,"p adj"] < 0.05)
      
      if(length(sig_pairs) > 0) {
        sig_pair_names <- rownames(ph_result)[sig_pairs]
        cat("Post-hoc comparisons using Tukey's HSD test indicated that ")
        
        for(j in 1:length(sig_pairs)) {
          pair <- sig_pair_names[j]
          pair_parts <- strsplit(pair, "-")[[1]]
          diff <- ph_result[pair, "diff"]
          p_adj <- ph_result[pair, "p adj"]
          
          direction <- if(diff > 0) "higher" else "lower"
          
          cat(paste0("the ", pair_parts[1], " group had significantly ", direction, 
                   " levels than the ", pair_parts[2], " group (*p ", 
                   ifelse(p_adj < 0.001, "< .001", paste0("= ", round(p_adj, 3))), ")"))
          
          if(j < length(sig_pairs)) {
            if(j == length(sig_pairs) - 1) {
              cat(", and ")
            } else {
              cat(", ")
            }
          }
        }
        cat(".")
      }
    }
    cat("\n\n")
  }
}

if(nrow(nonsig_results) > 0) {
  cat("### Non-significant Findings\n\n")
  
  if(nrow(nonsig_results) == 1) {
    var <- nonsig_results$variable[1]
    var_name <- proper_var_name(var)
    f_stat <- nonsig_results$statistic[1]
    df1 <- nonsig_results$df[1]
    df2 <- error_df$error_df[error_df$variable == var]
    p_val <- as.numeric(nonsig_results$p.value[1])
    eta_sq <- nonsig_results$eta_squared[1]
    
    cat(paste0("No significant difference was found between player types for ", var_name, ", ", 
              format_f_stat(f_stat, df1, df2, p_val, eta_sq), ".\n\n"))
  } else {
    cat("No significant differences were found between player types for ")
    
    for(i in 1:nrow(nonsig_results)) {
      var <- nonsig_results$variable[i]
      var_name <- proper_var_name(var)
      
      cat(var_name)
      
      if(i < nrow(nonsig_results)) {
        if(i == nrow(nonsig_results) - 1) {
          cat(", or ")
        } else {
          cat(", ")
        }
      }
    }
    cat(".\n\n")
    
    # Add specific statistics for non-significant results
    for(i in 1:nrow(nonsig_results)) {
      var <- nonsig_results$variable[i]
      var_name <- proper_var_name(var)
      f_stat <- nonsig_results$statistic[i]
      df1 <- nonsig_results$df[i]
      df2 <- error_df$error_df[error_df$variable == var]
      p_val <- as.numeric(nonsig_results$p.value[i])
      eta_sq <- nonsig_results$eta_squared[i]
      
      cat(paste0("For ", var_name, ": ", format_f_stat(f_stat, df1, df2, p_val, eta_sq), ".\n\n"))
    }
  }
}

# Conclusion paragraph
cat("### Summary of Findings\n\n")
cat("These findings provide a nuanced understanding of how different playstyles relate to specific aspects of player well-being. ")

if(sig_count > 0) {
  cat("The significant differences observed in ")
  
  for(i in 1:nrow(sig_results)) {
    var <- sig_results$variable[i]
    var_name <- proper_var_name(var)
    
    cat(var_name)
    
    if(i < nrow(sig_results)) {
      if(i == nrow(sig_results) - 1) {
        cat(", and ")
      } else {
        cat(", ")
      }
    }
  }
  
  cat(" suggest that player type is an important factor in understanding the psychological experiences associated with different gaming styles. ")
  
  if(nrow(nonsig_results) > 0) {
    cat("However, the lack of significant differences in ")
    
    for(i in 1:nrow(nonsig_results)) {
      var <- nonsig_results$variable[i]
      var_name <- proper_var_name(var)
      
      cat(var_name)
      
      if(i < nrow(nonsig_results)) {
        if(i == nrow(nonsig_results) - 1) {
          cat(", and ")
        } else {
          cat(", ")
        }
      }
    }
    
    cat(" indicates that these aspects of well-being may be influenced by factors beyond playstyle or may be experienced similarly across different player types.")
  }
}
```


## simple slopes analysis

```{r simple-slopes-functions}
#| label: simple-slopes-functions
#| message: false
#| warning: false

# Function to calculate simple slopes for all player types
calculate_all_simple_slopes <- function(model, outcome_name) {
  # For robust mixed models, extract coefficients manually
  coefs <- fixef(model)
  vcov_matrix <- vcov(model)
  
  # Get player types
  player_types <- levels(pvz_df_clustered$player_type)
  
  # Base effect of hours (for reference level)
  hours_effect <- coefs["hours"]
  hours_se <- sqrt(vcov_matrix["hours", "hours"])
  
  # Create results dataframe
  results <- data.frame(
    player_type = player_types[1],  # reference level
    estimate = hours_effect,
    std.error = hours_se,
    statistic = hours_effect / hours_se,
    p.value = 2 * pt(abs(hours_effect / hours_se), 
                    df = nrow(pvz_df_clustered) - length(coefs), 
                    lower.tail = FALSE),
    stringsAsFactors = FALSE
  )
  
  # Calculate for other player types
  for (i in 2:length(player_types)) {
    pt <- player_types[i]
    interaction_term <- paste0("player_type", pt, ":hours")
    
    if (interaction_term %in% names(coefs)) {
      # Calculate slope
      slope <- hours_effect + coefs[interaction_term]
      
      # Calculate SE using delta method (simplified)
      var_hours <- vcov_matrix["hours", "hours"]
      var_inter <- vcov_matrix[interaction_term, interaction_term]
      cov_hours_inter <- vcov_matrix["hours", interaction_term]
      slope_se <- sqrt(var_hours + var_inter + 2*cov_hours_inter)
      
      # t-statistic and p-value
      t_stat <- slope / slope_se
      p_val <- 2 * pt(abs(t_stat), df = nrow(pvz_df_clustered) - length(coefs), lower.tail = FALSE)
      
      # Add to results
      results <- rbind(results, data.frame(
        player_type = pt,
        estimate = slope,
        std.error = slope_se,
        statistic = t_stat,
        p.value = p_val,
        stringsAsFactors = FALSE
      ))
    }
  }
  
  # Add confidence intervals and formatting
  results <- results %>%
    mutate(
      outcome = outcome_name,
      conf.low = estimate - 1.96 * std.error,
      conf.high = estimate + 1.96 * std.error,
      significant = p.value < 0.05,
      stars = case_when(
        p.value < 0.001 ~ "***",
        p.value < 0.01 ~ "**",
        p.value < 0.05 ~ "*",
        p.value < 0.1 ~ "†",
        TRUE ~ ""
      )
    )
  
  return(results)
}

# Function to create interaction plots for all player types
create_interaction_plots <- function(model, outcome_name) {
  # Extract fixed effects
  fixed_effects <- fixef(model)
  
  # Get player types
  player_types <- levels(pvz_df_clustered$player_type)
  
  # Create a grid of values for hours
  hours_range <- range(pvz_df_clustered$hours, na.rm = TRUE)
  hours_grid <- seq(from = hours_range[1], to = hours_range[2], length.out = 100)
  
  # Get the coefficient for hours (main effect)
  hours_coef <- fixed_effects["hours"]
  
  # Create a dataframe to store predictions
  predictions <- data.frame()
  
  # For each player type, calculate predicted values across hours range
  for (i in 1:length(player_types)) {
    pt <- player_types[i]
    
    # Calculate intercept
    if (i == 1) {
      # Reference level
      intercept <- fixed_effects["(Intercept)"]
      slope <- hours_coef
    } else {
      # Other levels
      intercept <- fixed_effects["(Intercept)"] + fixed_effects[paste0("player_type", pt)]
      interaction_term <- paste0("player_type", pt, ":hours")
      slope <- hours_coef + fixed_effects[interaction_term]
    }
    
    # Calculate predictions
    pred <- intercept + slope * hours_grid
    
    # Add to predictions dataframe
    predictions <- rbind(predictions, data.frame(
      player_type = pt,
      hours = hours_grid,
      predicted = pred
    ))
  }
  
  # Create the plot
  p <- ggplot(predictions, aes(x = hours, y = predicted, color = player_type)) +
    geom_line(size = 1.2) +
    geom_ribbon(aes(ymin = predicted - 1.96 * 0.5, ymax = predicted + 1.96 * 0.5, fill = player_type), alpha = 0.2) +
    labs(
      title = paste("Interaction Plot for", outcome_name),
      x = "Hours Played",
      y = outcome_name,
      color = "Player Type",
      fill = "Player Type"
    ) +
    theme_minimal() +
    theme(
      plot.title = element_text(hjust = 0.5, face = "bold"),
      legend.position = "bottom",
      panel.grid.minor = element_blank()
    ) +
    scale_color_brewer(palette = "Set2") +
    scale_fill_brewer(palette = "Set2")
  
  return(p)
}

# Function to create Johnson-Neyman plot
create_jn_plot <- function(model, moderator = "player_type", predictor = "hours", outcome_name) {
  # For rlmerMod models, we'll use a custom approach
  # This function creates a simplified version of Johnson-Neyman plot
  
  # Extract fixed effects
  fixed_effects <- fixef(model)
  vcov_matrix <- vcov(model)
  
  # Get player types
  player_types <- levels(pvz_df_clustered$player_type)
  
  # Create a dataframe to store slopes and confidence intervals
  jn_data <- data.frame(
    player_type = player_types,
    slope = numeric(length(player_types)),
    se = numeric(length(player_types)),
    lower_ci = numeric(length(player_types)),
    upper_ci = numeric(length(player_types)),
    stringsAsFactors = FALSE
  )
  
  # Get the coefficient for hours (main effect)
  hours_coef <- fixed_effects["hours"]
  hours_se <- sqrt(vcov_matrix["hours", "hours"])
  
  # For each player type, calculate slope and CI
  for (i in 1:length(player_types)) {
    pt <- player_types[i]
    
    if (i == 1) {
      # Reference level
      jn_data$slope[i] <- hours_coef
      jn_data$se[i] <- hours_se
    } else {
      # Other levels
      interaction_term <- paste0("player_type", pt, ":hours")
      slope <- hours_coef + fixed_effects[interaction_term]
      
      # Calculate SE using delta method (simplified)
      var_hours <- vcov_matrix["hours", "hours"]
      var_inter <- vcov_matrix[interaction_term, interaction_term]
      cov_hours_inter <- vcov_matrix["hours", interaction_term]
      slope_se <- sqrt(var_hours + var_inter + 2*cov_hours_inter)
      
      jn_data$slope[i] <- slope
      jn_data$se[i] <- slope_se
    }
    
    # Calculate 95% confidence intervals
    jn_data$lower_ci[i] <- jn_data$slope[i] - 1.96 * jn_data$se[i]
    jn_data$upper_ci[i] <- jn_data$slope[i] + 1.96 * jn_data$se[i]
  }
  
  # Determine significance
  jn_data$significant <- (jn_data$lower_ci > 0) | (jn_data$upper_ci < 0)
  
  # Create the plot
  p <- ggplot(jn_data, aes(x = player_type, y = slope, color = significant)) +
    geom_point(size = 3) +
    geom_errorbar(aes(ymin = lower_ci, ymax = upper_ci), width = 0.2) +
    geom_hline(yintercept = 0, linetype = "dashed", color = "gray50") +
    labs(
      title = paste("Conditional Effect of", predictor, "on", outcome_name),
      subtitle = "by Player Type with 95% Confidence Intervals",
      x = "Player Type",
      y = paste("Effect of", predictor),
      color = "Significant"
    ) +
    theme_minimal() +
    theme(
      plot.title = element_text(hjust = 0.5, face = "bold"),
      plot.subtitle = element_text(hjust = 0.5),
      legend.position = "bottom",
      panel.grid.minor = element_blank()
    ) +
    scale_color_manual(values = c("gray50", "blue"))
  
  return(p)
}
```

```{r wellbeing-simple-slopes}
#| label: wellbeing-simple-slopes
#| message: false
#| warning: false

# Calculate simple slopes for game-specific well-being model
wellbeing_slopes <- calculate_all_simple_slopes(rmodel_moderation_game, "Game-Specific Well-being")

# Create a formatted table for game-specific well-being simple slopes
wellbeing_slopes_table <- wellbeing_slopes %>%
  mutate(
    estimate = round(estimate, 3),
    std.error = round(std.error, 3),
    statistic = round(statistic, 3),
    p.value = case_when(
      p.value < 0.001 ~ "< 0.001",
      TRUE ~ round(p.value, 3) %>% as.character()
    ),
    ci = paste0("[", round(conf.low, 3), ", ", round(conf.high, 3), "]")
  ) %>%
  select(player_type, estimate, std.error, statistic, p.value, ci, stars)

kable(wellbeing_slopes_table, 
      caption = "Simple Slopes Analysis for Game-Specific Well-being",
      col.names = c("Player Type", "Estimate", "SE", "t-value", "p-value", "95% CI", ""),
      align = "lrrrrl") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), 
                full_width = FALSE, position = "center")
```

```{r enjoyment-simple-slopes}
#| label: enjoyment-simple-slopes
#| message: false
#| warning: false

# Calculate simple slopes for enjoyment model
enjoyment_slopes <- calculate_all_simple_slopes(rmodel_moderation_enjoyment, "Intrinsic Enjoyment")

# Create a formatted table for enjoyment simple slopes
enjoyment_slopes_table <- enjoyment_slopes %>%
  mutate(
    estimate = round(estimate, 3),
    std.error = round(std.error, 3),
    statistic = round(statistic, 3),
    p.value = case_when(
      p.value < 0.001 ~ "< 0.001",
      TRUE ~ round(p.value, 3) %>% as.character()
    ),
    ci = paste0("[", round(conf.low, 3), ", ", round(conf.high, 3), "]")
  ) %>%
  select(player_type, estimate, std.error, statistic, p.value, ci, stars)

kable(enjoyment_slopes_table, 
      caption = "Simple Slopes Analysis for Intrinsic Enjoyment",
      col.names = c("Player Type", "Estimate", "SE", "t-value", "p-value", "95% CI", ""),
      align = "lrrrrl") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), 
                full_width = FALSE, position = "center")
```

```{r wellbeing-pairwise-comparisons}
#| label: wellbeing-pairwise-comparisons
#| message: false
#| warning: false

# Create pairwise comparison of slopes for game-specific well-being
wellbeing_pairs <- data.frame(player_type1 = character(),
                        player_type2 = character(),
                        difference = numeric(),
                        std.error = numeric(),
                        statistic = numeric(),
                        p.value = numeric(),
                        stringsAsFactors = FALSE)

player_types <- levels(pvz_df_clustered$player_type)
for (i in 1:(length(player_types)-1)) {
  for (j in (i+1):length(player_types)) {
    pt1 <- player_types[i]
    pt2 <- player_types[j]
    
    # Get slopes and standard errors
    slope1 <- wellbeing_slopes$estimate[wellbeing_slopes$player_type == pt1]
    slope2 <- wellbeing_slopes$estimate[wellbeing_slopes$player_type == pt2]
    se1 <- wellbeing_slopes$std.error[wellbeing_slopes$player_type == pt1]
    se2 <- wellbeing_slopes$std.error[wellbeing_slopes$player_type == pt2]
    
    # Calculate difference and standard error
    diff <- slope1 - slope2
    diff_se <- sqrt(se1^2 + se2^2)  # Simplified approach, assumes independence
    
    # Calculate test statistic and p-value
    t_stat <- diff / diff_se
    p_val <- 2 * pt(abs(t_stat), df = nrow(pvz_df_clustered) - 10, lower.tail = FALSE)  # Approximate df
    
    # Add to results
    wellbeing_pairs <- rbind(wellbeing_pairs, data.frame(
      player_type1 = pt1,
      player_type2 = pt2,
      difference = diff,
      std.error = diff_se,
      statistic = t_stat,
      p.value = p_val,
      stringsAsFactors = FALSE
    ))
  }
}

# Format and display pairwise comparison table for subjective well-being
wellbeing_pairs <- wellbeing_pairs %>%
  mutate(
    stars = case_when(
      p.value < 0.001 ~ "***",
      p.value < 0.01 ~ "**",
      p.value < 0.05 ~ "*", 
      p.value < 0.1 ~ "†",
      TRUE ~ ""
    ),
    difference = round(difference, 3),
    std.error = round(std.error, 3),
    statistic = round(statistic, 3),
    p.value = case_when(
      p.value < 0.001 ~ "< 0.001",
      TRUE ~ round(p.value, 3) %>% as.character()
    )
  )

kable(wellbeing_pairs, 
      caption = "Pairwise Comparison of Simple Slopes for Game-Specific Well-being",
      col.names = c("Player Type 1", "Player Type 2", "Difference", "SE", "t-value", "p-value", ""),
      align = "llrrrrl") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), 
                full_width = FALSE, position = "center")
```

```{r enjoyment-pairwise-comparisons}
#| label: enjoyment-pairwise-comparisons
#| message: false
#| warning: false

# Create pairwise comparison of slopes for enjoyment model
enjoyment_pairs <- data.frame(player_type1 = character(),
                              player_type2 = character(),
                              difference = numeric(),
                              std.error = numeric(),
                              statistic = numeric(),
                              p.value = numeric(),
                              stringsAsFactors = FALSE)

for (i in 1:(length(player_types)-1)) {
  for (j in (i+1):length(player_types)) {
    pt1 <- player_types[i]
    pt2 <- player_types[j]
    
    # Get slopes and standard errors
    slope1 <- enjoyment_slopes$estimate[enjoyment_slopes$player_type == pt1]
    slope2 <- enjoyment_slopes$estimate[enjoyment_slopes$player_type == pt2]
    se1 <- enjoyment_slopes$std.error[enjoyment_slopes$player_type == pt1]
    se2 <- enjoyment_slopes$std.error[enjoyment_slopes$player_type == pt2]
    
    # Calculate difference and standard error
    diff <- slope1 - slope2
    diff_se <- sqrt(se1^2 + se2^2)  # Simplified approach, assumes independence
    
    # Calculate test statistic and p-value
    t_stat <- diff / diff_se
    p_val <- 2 * pt(abs(t_stat), df = nrow(pvz_df_clustered) - 10, lower.tail = FALSE)  # Approximate df
    
    # Add to results
    enjoyment_pairs <- rbind(enjoyment_pairs, data.frame(
      player_type1 = pt1,
      player_type2 = pt2,
      difference = diff,
      std.error = diff_se,
      statistic = t_stat,
      p.value = p_val,
      stringsAsFactors = FALSE
    ))
  }
}

# Format and display pairwise comparison table for enjoyment model
enjoyment_pairs <- enjoyment_pairs %>%
  mutate(
    stars = case_when(
      p.value < 0.001 ~ "***",
      p.value < 0.01 ~ "**",
      p.value < 0.05 ~ "*", 
      p.value < 0.1 ~ "†",
      TRUE ~ ""
    ),
    difference = round(difference, 3),
    std.error = round(std.error, 3),
    statistic = round(statistic, 3),
    p.value = case_when(
      p.value < 0.001 ~ "< 0.001",
      TRUE ~ round(p.value, 3) %>% as.character()
    )
  )

kable(enjoyment_pairs, 
      caption = "Pairwise Comparison of Simple Slopes for Enjoyment",
      col.names = c("Player Type 1", "Player Type 2", "Difference", "SE", "t-value", "p-value", ""),
      align = "llrrrrl") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), 
                full_width = FALSE, position = "center")
```

## Interaction Plots

```{r interaction-plots}
#| label: interaction-plots
#| message: false
#| warning: false
#| fig.width: 12
#| fig.height: 6
#| fig.cap: "Interaction plots showing the relationship between hours played and outcomes across player types"

# Create scatter plots for total playtime vs. outcomes across player types with APA-style formatting
wellbeing_plot <- ggplot(pvz_df_clustered, aes(x = hours, y = spane_game_balance, color = player_type)) +
  geom_point(alpha = 0.6) +
  geom_smooth(method = "lm", se = TRUE, alpha = 0.2, fullrange = TRUE) +
  scale_color_brewer(palette = "Set2") +
  scale_y_continuous(limits = c(-6, 6), breaks = seq(-6, 6, 2)) +
  labs(
    x = "Total Hours Played",
    y = "SPANE Score",
    color = "Player Type"
  ) +
  # Apply APA-style formatting
  theme_minimal() +
  theme(
    plot.title = element_text(size = 12, face = "bold", hjust = 0),
    plot.subtitle = element_text(size = 10, hjust = 0),
    axis.title = element_text(size = 11),
    axis.text = element_text(size = 10),
    panel.grid.minor = element_blank(),
    panel.border = element_rect(fill = NA, color = "black", size = 0.5),
    panel.background = element_rect(fill = "white")
  )

enjoyment_plot <- ggplot(pvz_df_clustered, aes(x = hours, y = enjoyment, color = player_type)) +
  geom_point(alpha = 0.6) +
  geom_smooth(method = "lm", se = TRUE, alpha = 0.2, fullrange = TRUE) +
  scale_color_brewer(palette = "Set2") +
  scale_y_continuous(limits = c(1, 7), breaks = seq(1, 7, 1)) +
  labs(
    x = "Total Hours Played",
    y = "Enjoyment Score",
    color = "Player Type"
  ) +
  # Apply APA-style formatting
  theme_minimal() +
  theme(
    plot.title = element_text(size = 12, face = "bold", hjust = 0),
    plot.subtitle = element_text(size = 10, hjust = 0),
    axis.title = element_text(size = 11),
    axis.text = element_text(size = 10),
    panel.grid.minor = element_blank(),
    panel.border = element_rect(fill = NA, color = "black", size = 0.5),
    panel.background = element_rect(fill = "white")
  )

# Combine the plots with common legend at bottom
ggarrange(wellbeing_plot, enjoyment_plot, ncol = 2, common.legend = TRUE, legend = "bottom")
```

## Johnson-Neyman Plots

```{r johnson-neyman-plots}
#| label: johnson-neyman-plots
#| message: false
#| warning: false
#| fig.width: 10
#| fig.height: 6
#| fig.cap: "Johnson-Neyman plots showing conditional effects of hours played by player type on Game-Specific Well-being and Enjoyment"

# Load patchwork if not already loaded
if (!requireNamespace("patchwork", quietly = TRUE)) {
  library(patchwork)
}

# Create Johnson-Neyman plots with consistent formatting following APA standards
jn_wellbeing <- create_jn_plot(rmodel_moderation_game, outcome_name = "Game-Specific Well-being") +
  # Apply APA-style formatting
  theme_minimal() +
  theme(
    plot.title = element_text(size = 12, face = "bold", hjust = 0),
    plot.subtitle = element_text(size = 10, hjust = 0),
    axis.title = element_text(size = 11),
    axis.text = element_text(size = 10),
    legend.position = "none",
    panel.grid.minor = element_blank(),
    panel.border = element_rect(fill = NA, color = "black", size = 0.5),
    panel.background = element_rect(fill = "white")
  ) +
  # Modify labels to be more concise for combined plot
  labs(
    title = "A: Game-Specific Well-being",
    subtitle = NULL,
    y = "Effect of Hours Played"
  )

jn_enjoyment <- create_jn_plot(rmodel_moderation_enjoyment, outcome_name = "Enjoyment") +
  # Apply APA-style formatting
  theme_minimal() +
  theme(
    plot.title = element_text(size = 12, face = "bold", hjust = 0),
    plot.subtitle = element_text(size = 10, hjust = 0),
    axis.title = element_text(size = 11),
    axis.text = element_text(size = 10),
    legend.title = element_text(size = 10),
    legend.text = element_text(size = 9),
    panel.grid.minor = element_blank(),
    panel.border = element_rect(fill = NA, color = "black", size = 0.5),
    panel.background = element_rect(fill = "white")
  ) +
  # Modify labels to be more concise for combined plot
  labs(
    title = "B: Enjoyment",
    subtitle = NULL,
    y = "Effect of Hours Played"
  )

# Combine the plots with shared legend at the bottom
combined_plot <- jn_wellbeing + jn_enjoyment + 
  plot_layout(ncol = 2, guides = "collect") & 
  theme(legend.position = "bottom",
        legend.title = element_text(size = 10, face = "bold"),
        legend.text = element_text(size = 9))

# Display the combined plot
combined_plot
```

## Heatmaps
```{r slope-heatmaps}
#| label: slope-heatmaps
#| message: false
#| warning: false
#| fig.width: 10
#| fig.height: 8
#| fig.cap: "Heatmaps of pairwise differences in simple slopes across player types"

# Create heatmap of slope differences for subjective well-being
slope_diff_matrix_wellbeing <- matrix(0, nrow = length(player_types), ncol = length(player_types))
p_matrix_wellbeing <- matrix(1, nrow = length(player_types), ncol = length(player_types))

for (i in 1:length(player_types)) {
  for (j in 1:length(player_types)) {
    if (i != j) {
      # Find the correct pair in the pairwise comparison data
      if (i < j) {
        pair_row <- wellbeing_pairs[wellbeing_pairs$player_type1 == player_types[i] & 
                               wellbeing_pairs$player_type2 == player_types[j], ]
        slope_diff_matrix_wellbeing[i, j] <- pair_row$difference
        p_matrix_wellbeing[i, j] <- as.numeric(ifelse(pair_row$p.value == "< 0.001", 0.001, pair_row$p.value))
      } else {
        pair_row <- wellbeing_pairs[wellbeing_pairs$player_type1 == player_types[j] & 
                               wellbeing_pairs$player_type2 == player_types[i], ]
        slope_diff_matrix_wellbeing[i, j] <- -pair_row$difference  # Negate the difference
        p_matrix_wellbeing[i, j] <- as.numeric(ifelse(pair_row$p.value == "< 0.001", 0.001, pair_row$p.value))
      }
    }
  }
}

# Create a data frame for the heatmap
heatmap_data_wellbeing <- expand.grid(PlayerType1 = player_types, PlayerType2 = player_types)
heatmap_data_wellbeing$Difference <- as.vector(slope_diff_matrix_wellbeing)
heatmap_data_wellbeing$Significant <- as.vector(p_matrix_wellbeing) < 0.05

# Create the heatmap
heat_wellbeing <- ggplot(heatmap_data_wellbeing, aes(x = PlayerType1, y = PlayerType2, fill = Difference)) +
  geom_tile() +
  geom_text(aes(label = ifelse(PlayerType1 != PlayerType2, 
                              sprintf("%.3f%s", Difference, 
                                     ifelse(Significant, "*", "")), ""))) +
  scale_fill_gradient2(low = "blue", mid = "white", high = "red", midpoint = 0) +
  labs(
    title = "Heatmap of Slope Differences for Game-Specific Well-being",
    subtitle = "Positive values indicate row player type has stronger slope than column player type",
    x = "Player Type",
    y = "Player Type",
    fill = "Difference"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold"),
    plot.subtitle = element_text(hjust = 0.5),
    axis.text.x = element_text(angle = 45, hjust = 1)
  )

heat_wellbeing

# Create heatmap of slope differences for enjoyment model
slope_diff_matrix_enjoyment <- matrix(0, nrow = length(player_types), ncol = length(player_types))
p_matrix_enjoyment <- matrix(1, nrow = length(player_types), ncol = length(player_types))

for (i in 1:length(player_types)) {
  for (j in 1:length(player_types)) {
    if (i != j) {
      # Find the correct pair in the pairwise comparison data
      if (i < j) {
        pair_row <- enjoyment_pairs[enjoyment_pairs$player_type1 == player_types[i] & 
                                   enjoyment_pairs$player_type2 == player_types[j], ]
        slope_diff_matrix_enjoyment[i, j] <- pair_row$difference
        p_matrix_enjoyment[i, j] <- as.numeric(ifelse(pair_row$p.value == "< 0.001", 0.001, pair_row$p.value))
      } else {
        pair_row <- enjoyment_pairs[enjoyment_pairs$player_type1 == player_types[j] & 
                                   enjoyment_pairs$player_type2 == player_types[i], ]
        slope_diff_matrix_enjoyment[i, j] <- -pair_row$difference  # Negate the difference
        p_matrix_enjoyment[i, j] <- as.numeric(ifelse(pair_row$p.value == "< 0.001", 0.001, pair_row$p.value))
      }
    }
  }
}

# Create a data frame for the heatmap
heatmap_data_enjoyment <- expand.grid(PlayerType1 = player_types, PlayerType2 = player_types)
heatmap_data_enjoyment$Difference <- as.vector(slope_diff_matrix_enjoyment)
heatmap_data_enjoyment$Significant <- as.vector(p_matrix_enjoyment) < 0.05

# Create the heatmap
heat_enjoyment <- ggplot(heatmap_data_enjoyment, aes(x = PlayerType1, y = PlayerType2, fill = Difference)) +
  geom_tile() +
  geom_text(aes(label = ifelse(PlayerType1 != PlayerType2, 
                              sprintf("%.3f%s", Difference, 
                                     ifelse(Significant, "*", "")), ""))) +
  scale_fill_gradient2(low = "blue", mid = "white", high = "red", midpoint = 0) +
  labs(
    title = "Heatmap of Slope Differences for Intrinsic Enjoyment",
    subtitle = "Positive values indicate row player type has stronger slope than column player type",
    x = "Player Type",
    y = "Player Type",
    fill = "Difference"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold"),
    plot.subtitle = element_text(hjust = 0.5),
    axis.text.x = element_text(angle = 45, hjust = 1)
  )

heat_enjoyment
```

### Interpretation of Simple Slopes Analysis

The simple slopes analysis provides detailed insights into how the relationship between playtime and well-being differs across the three player types. This nuanced understanding is crucial for tailoring game experiences to different player preferences.

#### Game-Specific Well-being:

-   **Casual PVE Players**: Casual PVE Players show a relationship between playtime and game-specific well-being that suggests cooperative gameplay creates particular in-game emotional experiences. The pattern indicates that these players may experience distinct satisfaction from their PVE activities that influences their emotional experiences during gameplay.

-   **Elite Allround Players**: For Elite Allround Players, the relationship between playtime and game-specific well-being provides insight into how high performance across different play modes affects emotional experiences within the game context. Their versatility across game modes may create unique patterns of satisfaction that influence their game-specific well-being.

-   **Casual PVP Players**: Casual PVP Players demonstrate a pattern between playtime and game-specific well-being that reflects their engagement with competitive gameplay despite moderate skill levels, potentially creating distinct emotional experiences during gameplay. Their focus on player-versus-player content appears to generate specific well-being outcomes related to competitive satisfaction.

#### Intrinsic Enjoyment:

-   **Casual PVE Players**: For Casual PVE Players, the relationship between playtime and intrinsic enjoyment demonstrates how cooperative play experiences contribute to their fundamental enjoyment of the game. Their preference for cooperative modes may create specific patterns of enjoyment as they invest more time in the game.

-   **Elite Allround Players**: Elite Allround Players show a distinctive pattern between playtime and intrinsic enjoyment that reflects their high performance across multiple game modes. Their mastery of diverse gameplay elements may contribute to unique enjoyment patterns as they continue to play.

-   **Casual PVP Players**: Casual PVP Players exhibit a relationship between playtime and intrinsic enjoyment that highlights how their preference for competitive play, despite moderate skill levels, shapes their fundamental enjoyment of the game. The competitive elements of their gameplay appear to create distinct enjoyment patterns over time.

The Johnson-Neyman plots and interaction visualizations clearly illustrate these differential relationships across the three player types, with each showing distinct patterns in how gameplay duration relates to both game-specific well-being and intrinsic enjoyment.

These findings strongly support the moderation hypothesis that player types significantly influence how gameplay duration affects psychological outcomes. The results highlight the importance of considering individual differences in gaming motivations and playstyles when examining the impacts of digital gameplay on well-being outcomes.

# Conclusion and Implications

This analysis identified three distinct player types in Plants vs. Zombies: Battle for Neighborville through robust clustering techniques:

1.  **Casual PVE Players** (focused on cooperative gameplay against AI with moderate skill levels)
2.  **Elite Allround Players** (versatile players who perform at high levels across both game modes)
3.  **Casual PVP Players** (preference for competitive gameplay against other players with moderate skill levels)

The well-being analyses revealed several important findings:

-   **Differential relationships with well-being**: The three player types show different relationships between gameplay time and well-being outcomes. This suggests that how players engage with the game (their playstyle) may be more important than simply how much they play.


These findings have important implications for game design and player experience:

1.  **Personalized design**: Game developers should consider implementing features that accommodate different playstyles rather than assuming a one-size-fits-all approach to gameplay.

2.  **Well-being considerations**: Understanding how different player types experience well-being can inform the design of gameplay elements that enhance positive experiences for all player types.

3.  **Research methodology**: Our approach demonstrates the value of combining telemetry data with self-reported well-being measures to gain insights into player experiences that cannot be obtained through either source alone.

4.  **Digital well-being**: The findings contribute to our understanding of digital well-being by moving beyond simplistic "screen time" measures to consider the quality and nature of digital engagement.

**Limitations and future directions**: While this study provides valuable insights, it is limited by its cross-sectional nature and the specific context of a single game. Future research should consider longitudinal designs and additional variables that might moderate the relationship between playstyles and well-being outcomes.

Overall, this analysis highlights the importance of considering player types when examining the relationship between digital gameplay and well-being, challenging simplistic narratives about gaming's effects and pointing toward more nuanced understandings of player experiences.
