---
title: "Not All Play is Equal: In-game Player Behaviour Predicts Wellbeing Differences"
shorttitle: "In-game player behaviour and wellbeing"
running-head: "Playing Style and Wellbeing"
author:
  - name: Thomas Hakman
    corresponding: true
    orcid: 0009-0009-8292-2482
    email: thomas.hakman@oii.ox.ac.uk
    affiliation:
      - ref: 1
    roles:
      - conceptualization
      - data curation
      - investigation
      - methodology
      - formal analysis
      - visualization
      - writing
      - editing
  - name: Matti Vuorre
    corresponding: false
    orcid: 0000-0001-5052-066X
    email: mjvuorre@uvt.nl
    affiliations:
      - ref: 2
    roles:
      - data curation
      - methodology
      - formal analysis
      - supervision
      - editing
  - name: Andrew K. Przybylski
    corresponding: false
    orcid: 0000-0001-5547-2185
    email: andy.przybylski@oii.ox.ac.uk
    affiliations:
      - ref: 1
    roles:
      - conceptualization
      - funding acquisition
      - project administration
      - supervision
      - editing
affiliations:
  - id: 1
    name: University of Oxford
    department: Oxford Internet Institute
  - id: 2
    name: Tilburg University
    department: Department of Social Psychology
abstract: |
  Large-scale studies using digital trace data consistently find that playtime alone is not meaningfully associated with wellbeing. Yet these studies typically capture playtime across platforms and titles without examining what happens inside the game, the granular in-game behaviours that might differentiate players’ experiences. The Integrated Model of Player Experience posits that the mechanics and contexts players encounter during play shape the resulting experience and its downstream psychological effects, implying that players who engage with the same game in qualitatively different ways may report different wellbeing outcomes. We tested this using telemetry-driven profiling of session-level behavioural data from 412 players of the online shooter Plants vs. Zombies: Battle for Neighborville. Cluster analysis on 10 standardised performance and mode-preference metrics identified three distinct playstyles characterised by: (A) explorative, (B) high-performing, and (C) competitive playstyles. Explorative players reported greater wellbeing and enjoyment compared to competitive players. However, we found no evidence that playstyle moderated the relationship between playtime and either outcome. These findings indicate that how players engage with a game’s mechanics and modes is associated with different average levels of wellbeing, but we know little about how such playstyles might alter the relationship between the amount of play and wellbeing. 
categories: [video games, playstyle, wellbeing, play experience, telemetry]
authornote: |
  Correspondence concerning this article should be addressed to Thomas Hakman, Oxford Internet Institute, University of Oxford. Email: thomas.hakman@oii.ox.ac.uk
date: last-modified
bibliography: bibliography.bib
csl: apa.csl
prefer-html: true
execute:
  echo: false
  warning: false
  message: false
  cache: false
  freeze: false
params:
  refit_models: true
  rerun_clustering: true
format:
  html:
    toc: true
    toc-location: left
    theme: cosmo
    code-fold: true
    html-math-method: katex
    embed-resources: true
  preprint-typst:
    keep-typ: true
    citeproc: true
    include-before-body:
      - text: "#show <refs>: set par(hanging-indent: 1.4em)"
  preprint-docx: default
knitr:
  opts_chunk:
    dev: png
    fig.format: png
    dpi: 300
---

```{r}
#| label: setup
#| include: false
knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE,
  fig.align = "center"
)

# Toggle model refitting via Quarto param
eval_models <- isTRUE(params$refit_models)
eval_clustering <- isTRUE(params$rerun_clustering)

# Ensure output directory exists for cached models
if (!dir.exists("output/models")) {
  dir.create("output/models", recursive = TRUE)
}
```

# Introduction

The debate about whether video games harm or improve wellbeing has persisted for decades, yet empirical findings remain mixed, constantly changing with the methodological approaches used to study them. Earlier studies relying on players' retrospective estimates of their own playtime and behaviour produced contradictory findings. Several studies reported negative associations between self-reported playtime and measures of mental health or wellbeing [@fergusonLackConsensusScholars2019; @liuDigitalCommunicationMedia2019; @mathurFindingCommonGround2019], with some reporting negative associations only among highly engaged players [@coldercarrasVideoGamingHyperconnected2017; @przybylski2017a]. Whilst others found positive associations [@halbrookWhenHowVideo2019; @villaniVideogamesEmotionRegulation2018], with some reporting those only for moderate usage [@przybylski2017a] or specific game genres such as MMOs [@raithMassivelyMultiplayerOnline2021]. To overcome the well-documented inaccuracy of self-reported measures [@johannesVideoGamePlay2021; @parrySystematicReviewMetaanalysis2021a], later work utilised digital trace data (i.e., objective behavioural logs collected directly from game platforms). These studies demonstrated that the association between the average player's quantity of playtime and their wellbeing are small and unlikely to be practically meaningful [@ballouOpenPlayLongitudinal2025; @egamiCausalEffectVideo2024; @johannesVideoGamePlay2021; @larrieuHowGamingMotives2023; @vuorreTimeSpentPlaying2022].

These small, mixed, and often null results do not necessarily imply that playing video games is irrelevant to wellbeing. Rather, they suggest that the most measured dimension of play, its quantity, might not be the most important factor. In a large-scale analysis of adult Nintendo players, playtime was indeed not meaningfully related to multiple wellbeing indicators, but players' subjective evaluations of gaming's life-fit (e.g. the perceived harmful or beneficial value of gaming across various life domains outside of play) were [@ballouPerceivedValueVideo2025]. The underlying problem is that, conceptually, 'video game play' does not capture a single phenomenon, but instead heterogeneous exposures such as 'playing with friends', 'succeeding or failing at the game's mechanics', or 'playing to alleviate study-related stress'. More generally, playtime can be thought to consist of context-centred exposures (e.g., when and with whom one plays), game-centred exposures (e.g., game modes, mechanics, and business models), and player-centred exposures (e.g., playstyles, motivations, and player experience) [@ballouHowVideoGames2025]. When studies collapse these layers into a single playtime variable, they conflate qualitatively distinct exposures and risk both theoretical ambiguity and empirical inconsistency [@ballouHowVideoGames2025].

This highlights a crucial gap: while existing digital trace research is a mile wide in its scope of game platforms and genres, it is only an inch deep in actual player experiences and behaviors: Existing studies capture playtime across platforms and titles, but not what happens inside the game, the granular in-game behaviours and experiences that might explain why some players benefit from play while others do not. Quantity alone is insufficient; two players can accumulate the same playtime yet have markedly different player experiences. To reconcile the mixed historical findings with the contemporary null results, research needs to start modelling the quality of play with in-game digital trace data. Nevertheless, how researchers should best study in-game player experience empirically without reverting to subjective self-reports remains an open question.

One theoretically grounded operationalisation comes from the Integrated Model of Player experience [IMP; @elsonMoreStoriesButtons2014], which identifies three categories of game characteristics that shape the playing experience: narrative (story and setting), mechanics (rules, interaction demands, and feedback), and context (the social and situational embedding). Where user-centric theories such as self-determination theory [@ryanIntrinsicMotivationVideo2006] and uses and gratifications [@ruggieroUsesGratificationsTheory2000] focus primarily on motivations and player states, the IMP is game-centric and focuses on how interaction with specific in-game characteristics during play alters the play experience and subsequent wellbeing outcomes [@elsonMoreStoriesButtons2014]. In this sense, the IMP opens the "black box" of in-game behaviour by connecting the game characteristics players encounter during play to the outcomes that follow from it, thereby guiding empirically testable hypotheses about these relationships [@elsonMoreStoriesButtons2014]. Crucially, the IMP positions mechanics as the foundational element of player experience; without mechanics there is no play, whereas narrative and context contribute situationally. A conceptually meaningful operationalisation of play quality should therefore incorporate digital trace data on mechanics-relevant indicators (e.g., performance metrics), contextual indicators (e.g., mode selection), and narrative indicators (e.g., dialogue engagement) where appropriate.

Within game analytics research, a similar categorisation process is known as telemetry-driven profiling, where collections of in-game behaviours, events, and experiences are condensed into behavioural profiles, or playstyles, that describe how players engaged with a game during the measurement period [@drachen2014a; @ravariInvestigatingRelationPlaying2022]. Playstyles are distinct from player types. Player types refer to stable individual characteristics assessed via self-reports, as in Bartle's taxonomy [@bartleHeartsClubsDiamonds1996], which categorizes players as Achievers, Explorers, Socializers, or Killers based on their preferences for achievement versus exploration and interaction with the game world versus other players; Yee's motivation model [@yeeMotivationsPlayOnline2006], which identifies three primary components of player motivation (i.e., achievement, social, and immersion); and the Hexad framework [@tondelloGamificationUserTypes2016], which proposes six gamification user types grounded in self-determination theory. The utility of such typologies has, however, been called into question by recent evidence of their temporal instability. @santosConsistencyGamificationUser2023 found that 72% of participants changed their player type classification over six months, suggesting that self-reported types may not reliably reflect actual in-game behaviour.

In contrast, telemetry-derived playstyles capture emergent behavioural patterns as they occur, providing a more ecologically valid representation of how players engage with games. As such, playstyles offer a quantifiable, data-driven way to characterise how players pursue a game's objectives [@sifaProfilingGamesUnderstanding2018]. Previous work has used telemetry-driven profiling to examine the relationship between playstyles and personality [@beanVideoGamersPersonality2016; @vanlankveldGamesPersonalityProfiling2011], motivation [@melhartYourGameplaySays2019; @yeeGamerMotivationProfile2016], national culture [@ravariInvestigatingRelationPlaying2022], and individual and team performance [@jiangWideDeepMaverick2021]. However, whether and how playstyles relate to wellbeing and enjoyment has remained untested, in large part because datasets combining granular session-level behavioural telemetry with validated measures of wellbeing are exceedingly rare [@norwoodBuildingCollaborativeDigital2026].

The present study addresses three questions by leveraging one such dataset, collected through an industry-academia collaboration [@johannesVideoGamePlay2021]. First, can distinct playstyles be identified from in-game behavioural telemetry within a single game title? Second, do players with different playstyles report different levels of wellbeing or enjoyment? Third, does the association between playtime and wellbeing or enjoyment differ between playstyles?





```{r}
#| label: install-packages
#| include: false
#| cache: false

# First ensure pacman is installed if not already
if (!requireNamespace("pacman", quietly = TRUE)) {
  install.packages("pacman")
}

# Use pacman to load all required packages (will install any missing packages)
pacman::p_load(
  here,         # For project-relative paths
  lubridate,    # For date/time manipulation
  scales,       # For scale transformations
  janitor,      # For data cleaning
  magrittr,     # For additional pipe operators
  psych,        # For psychological statistics
  corrplot,     # For correlation plots
  ggpubr,       # For publication-ready plots
  factoextra,   # For cluster analysis visualization
  cluster,      # For clustering algorithms
  fmsb,         # For radar charts
  knitr,        # For dynamic report generation
  fpc,          # For cluster validation
  gridExtra,    # For arranging multiple plots
  hopkins,      # For cluster tendency assessment
  effectsize,   # For effect size calculations
  robustlmm,    # For robust linear mixed-effects models
  broom,        # For tidy model outputs
  broom.mixed,  # For tidying mixed-effects model output
  patchwork,    # For combining plots
  ggeffects,    # For visualizing regression model predictions
  grid,         # For grid graphics (needed for tableGrob)
  tinytable,    # For publication-quality tables
  modelsummary, # For regression model tables
  report,       # For automated statistical reporting
  tidyverse     # For data manipulation and visualization (loaded last to avoid masking)
)

# Load helper functions
source(here("helpers.R"))

# Colourblind-safe Okabe-Ito palette for the three playstyles
playstyle_colors <- c(
  "Casual PvP Players"      = "#E69F00",
  "Casual PvE Players"      = "#56B4E9",
  "Elite Allround Players"  = "#009E73"
)
playstyle_scale_color <- scale_color_manual(values = playstyle_colors)
playstyle_scale_fill  <- scale_fill_manual(values = playstyle_colors)
```

```{r}
#| label: helper-functions
#| code-summary: "Load helper functions"

# Helper functions are now sourced from helpers.R in the install-packages chunk
# This chunk is kept for any analysis-specific functions not in helpers.R

# Function to run ANOVA and extract results
run_anova <- function(dv) {
  formula <- as.formula(paste(dv, "~ player_type"))
  model <- aov(formula, data = pvz_data)
  result <- broom::tidy(model)
  result$eta_squared <- effectsize::eta_squared(model)[["Eta2"]]
  result$variable <- dv
  return(result)
}

# Mean helper for mode-specific telemetry metrics:
# treat missing values as 0 and always return a finite scalar.
mode_mean <- function(x) {
  mean(coalesce(x, 0), na.rm = TRUE)
}

# Safe lookup helper for inline reporting from summary tables.
type_stat <- function(tbl, type_label, column_name, digits = 2) {
  value <- tbl %>%
    filter(.data$player_type == type_label) %>%
    pull(.data[[column_name]]) %>%
    first(default = NA_real_)

  if (is.na(value) || is.nan(value) || is.infinite(value)) {
    value <- 0
  }

  sprintf(paste0("%.", digits, "f"), value)
}
```

```{r}
#| label: load-processed-data

# Load the dataset with outliers removed
pvz_data <- read_csv(here("data/pvz_clean_final.csv"))

# Create a copy of pvz_data for analysis to maintain consistency
pvz_df_analysis <- pvz_data

```

```{r}
#| label: select-columns

pvz_df_analysis <- pvz_df_analysis %>%
  select(
    # Demographic variables
    player_id, country, age, gender, played_with_others,
    
    # Wellbeing variables - focus only on our two key outcomes
    spane_game_balance, enjoyment,
    
    # Total playtime
    hours,
    
    # PvP metrics
    ratio_pvp_hours,
    pvp_kd,
    pvp_hit_accuracy,
    pvp_dps,
    pvp_average_damage,
    pvp_average_score,
    
    # PvE metrics
    ratio_pve_hours,
    pve_kd,
    pve_hit_accuracy,
    pve_dps,
    pve_average_damage,
    pve_average_score
  )
```

```{r}
#| label: telemetry-data-filter
#| include: false

# Keep players who engaged in at least one combat mode (PvP or PvE)
# All 10 clustering features are checked: hours > 0 AND non-missing performance
# metrics for at least one mode. NAs in the other mode are set to 0 before clustering (see replace_na in cluster_feature_selection chunk).
combat_metrics <- c("kd", "hit_accuracy", "dps", "average_damage", "average_score")
pvz_df_analysis <- pvz_df_analysis %>%
  filter(
    (ratio_pvp_hours > 0 & if_all(paste0("pvp_", combat_metrics), ~ !is.na(.x))) |
    (ratio_pve_hours > 0 & if_all(paste0("pve_", combat_metrics), ~ !is.na(.x)))
  )

# Show the number of players left
print(paste("Number of players left after filtering:", nrow(pvz_df_analysis)))
```

# Methods

In this study, we analysed data reported in @johannesVideoGamePlay2021, who collected digital trace data and survey responses of 518 Plants versus Zombies: Battle for Neighborville players.

## Plants versus Zombies: Battle for Neighborville

```{r}
#| label: survey-demographics-calculation
#| include: false
#| cache: false

# Calculate demographics for the filtered dataset
demographics <- pvz_df_analysis %>%
  summarize(
    total_participants = n(),
    avg_age = round(mean(age, na.rm = TRUE), 1),
    sd_age = round(sd(age, na.rm = TRUE), 1)
  )

# Normalize gender labels; treat missing/unexpected labels as undisclosed
gender_distribution <- pvz_df_analysis %>%
  mutate(
    gender_clean = case_when(
      str_trim(as.character(gender)) == "Male" ~ "Male",
      str_trim(as.character(gender)) == "Female" ~ "Female",
      str_trim(as.character(gender)) == "Other" ~ "Other",
      str_trim(as.character(gender)) == "Prefer not to say" ~ "Prefer not to say",
      TRUE ~ "Prefer not to say"
    )
  ) %>%
  count(gender_clean, name = "n") %>%
  mutate(percentage = n / sum(n) * 100)

count_or_zero <- function(df, label) {
  value <- df %>% filter(gender_clean == label) %>% pull(n)
  if (length(value) == 0) 0L else as.integer(value)
}

# Count participants with each gender category for manuscript text
male_count <- count_or_zero(gender_distribution, "Male")
female_count <- count_or_zero(gender_distribution, "Female")
other_count <- count_or_zero(gender_distribution, "Other")
undisclosed_count <- count_or_zero(gender_distribution, "Prefer not to say")
```

```{r}
#| label: final-sample-demographics
#| include: false
#| cache: false

# Display demographics of final sample
cat("Demographics of final sample:
")
cat("- Sample size:", nrow(pvz_df_analysis), "
")
cat("- Average age:", round(mean(pvz_df_analysis$age, na.rm = TRUE), 1), "(SD =", round(sd(pvz_df_analysis$age, na.rm = TRUE), 1), ")
")

# Gender distribution
gender_dist <- pvz_df_analysis %>%
  mutate(
    gender = case_when(
      str_trim(as.character(gender)) == "Male" ~ "Male",
      str_trim(as.character(gender)) == "Female" ~ "Female",
      str_trim(as.character(gender)) == "Other" ~ "Other",
      str_trim(as.character(gender)) == "Prefer not to say" ~ "Prefer not to say",
      TRUE ~ "Prefer not to say"
    )
  ) %>%
  count(gender) %>%
  mutate(percentage = n / sum(n) * 100)
```

Plants versus Zombies: Battle for Neighborville (PvZ) is an online third-person multiplayer shooter game released by Electronic Arts (EA) in October 2019. Players can choose from 23 customizable gameplay classes, each with unique weapons, abilities, and movement styles. The game offers players the ability to compete against each other in various competitive Player-versus-Player (PvP) modes, including objective-based matches, team deathmatches, and 2v2 arena matches. The game also features several Player-versus-Environment (PvE) open zones, which feature storylines with quests and allow players to explore, find collectables, and complete side-missions. Players can earn experience points and in-game currencies, using the former to indicate progression and the latter to upgrade and buy new characters. The game was played on PC, PS4, Xbox One, and Nintendo Switch [@electronicArtsPlantsVsZombiesBattle2019].

## Participants

As detailed in @johannesVideoGamePlay2021, participants were recruited by Electronic Arts directly from their active player base via email invitations sent in two waves: the first in early August 2020 to 50,000 players, and the second in late September 2020 to 200,000 players. In total, 518 PvZ players completed the wellbeing survey (approximately 0.21% response rate). For details on the full dataset, which also includes Animal Crossing: New Horizons players, see @johannesVideoGamePlay2021. Because our research questions concern mechanics-based playstyles, we applied several inclusion criteria to ensure the analytic sample consisted of players who actively engaged with the game's core mechanics. We first excluded participants without matching telemetry data (n = 48). We then excluded 57 participants without telemetry across any activity (PvP or PvE). These players spent time in the game's social hub, an idle space for navigation and character customisation, but did not participate in any gameplay, and thus had no behavioural data from which to derive playstyle profiles. The final analytic sample consisted of `r nrow(pvz_df_analysis)` combat-active players (*M* age = `r demographics$avg_age`, *SD* = `r demographics$sd_age`; `r male_count` male, `r female_count` female, `r other_count` other, `r undisclosed_count` undisclosed), representing `r sprintf('%.1f', (nrow(pvz_df_analysis) / 518) * 100)`% of the original 518 respondents. Following the methodology from the original paper, values more than 6 standard deviations from a variable's mean were replaced with NA on a case-by-case basis [@johannesVideoGamePlay2021]. In this dataset, this procedure replaced 87 values (across 48 variables; 26 participants affected) and did not reduce the final analytic sample.


## Measures

### Playtime

The first measure we calculated is playtime, yet determining this is already complex. Simply measuring the total time an account has been online does not distinguish between active and idle time. Idle time is when the player’s account is online but not actively engaging with the objectives or mechanics of the game. Idle time can happen when a player’s account is online, but they are away from the keyboard. Other forms of idle time are spending time in home screens, customisation menus, lobbies, and social spaces. In this study, active playtime was calculated by aggregating session data. Sessions represent the different levels, game modes, or worlds a player can be present in. Every player is always in a session and can only be in one session at a time; however, multiple players could be in the same session. For example, when a player launches PvZ, they enter the hub world (where they can select what type of level or mode to play). Entering the hub world starts the session timer, which ends as they choose a level or leave to join a match, which then starts a new session timer for that level. In total, the `r nrow(pvz_df_analysis)` players in our sample played 36,621 sessions, of which 25,937 happened in the two weeks prior to a player taking the survey. These sessions took place across 19 different maps that each facilitated different forms of gameplay, such as Player-versus-Player (PvP; 13,594 (51.27%) of sessions), 4,854 (18.31%) Player-versus-Environment (PvE), and 7,489 (28.24%) Social Hub maps. After exclusion of the social hub sessions, this left 17,505 total sessions, 12,445 PvP sessions (71.09%) and 4,509 PvE sessions (25.76%). On visual inspection, players had distinctive preferences for which game mode they chose to play and when (@fig-session-timeline).

```{r}
#| label: session-timeline-helper
#| include: false

# Function to plot session timeline for a player
plot_player_sessions <- function(player_data, player_id = NULL, player_label = "Player") {
  # If a player_id is provided, filter data for that player
  if (!is.null(player_id)) {
    player_data <- player_data %>%
      filter(player_id == !!player_id)
  }
  
  # Convert timestamps to POSIXct if they're not already
  if (!inherits(player_data$game_start_time, "POSIXct")) {
    player_data$game_start_time <- as.POSIXct(player_data$game_start_time, format = "%Y-%m-%d %H:%M:%S")
    player_data$game_end_time <- as.POSIXct(player_data$game_end_time, format = "%Y-%m-%d %H:%M:%S")
  }
  
  # Extract day of week, date, and time component
  player_data$day_of_week <- weekdays(player_data$game_start_time)
  player_data$date <- as.Date(player_data$game_start_time)
  player_data$start_time_only <- as.POSIXct(format(player_data$game_start_time, format = "%H:%M:%S"), format = "%H:%M:%S")
  player_data$end_time_only <- as.POSIXct(format(player_data$game_end_time, format = "%H:%M:%S"), format = "%H:%M:%S")
  
  # Identify sessions that cross midnight (where end_time is earlier than start_time)
  player_data$crosses_midnight <- player_data$end_time_only < player_data$start_time_only
  
  # Create a new dataframe that will include split sessions for those crossing midnight
  player_data_expanded <- player_data %>%
    filter(!crosses_midnight) # Keep all sessions that don't cross midnight
  
  # Handle sessions that cross midnight by splitting them
  midnight_sessions <- player_data %>%
    filter(crosses_midnight)
  
  if (nrow(midnight_sessions) > 0) {
    # For each session crossing midnight, create two rows:
    # 1. From start time to 23:59:59
    # 2. From 00:00:00 to end time
    midnight <- as.POSIXct(strptime("23:59:59", format = "%H:%M:%S"), format = "%H:%M:%S")
    midnight_start <- as.POSIXct(strptime("00:00:00", format = "%H:%M:%S"), format = "%H:%M:%S")
    
    # First part: from start time to midnight
    midnight_part1 <- midnight_sessions
    midnight_part1$end_time_only <- midnight
    
    # Second part: from midnight start to end time
    midnight_part2 <- midnight_sessions
    midnight_part2$start_time_only <- midnight_start
    # The second part is on the next day of the week and the next date
    midnight_part2$day_of_week <- factor(
      weekdays(midnight_part2$game_start_time + 86400), # Add one day (86400 seconds)
      levels = c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday")
    )
    midnight_part2$date <- midnight_part2$date + 1
    
    # Add both parts to the expanded dataset
    player_data_expanded <- bind_rows(player_data_expanded, midnight_part1, midnight_part2)
  }
  
  # Recode match_type based on game_level_key prefix
  player_data_expanded <- player_data_expanded %>%
    mutate(
      match_type = case_when(
        str_starts(game_level_key, "dsub_social")  ~ "social match",
        str_starts(game_level_key, "dsub_region")  ~ "pve match",
        str_starts(game_level_key, "dsub_rush")    ~ "pvp match (rush)",
        str_starts(game_level_key, "dsub_tvm")     ~ "pvp match (vanquish)",
        str_starts(game_level_key, "dsub_4v4")     ~ "pvp match (arena)",
        game_level_key == "dsub_podiumcelebration"  ~ "podium celebration",
        .default = "0"
      )
    )
  
  # Ensure day of week is a factor with ordered levels
  player_data_expanded$day_of_week <- factor(
    player_data_expanded$day_of_week,
    levels = c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday")
  )
  
  # Create a position adjustment for stacking sessions by different dates for the same day of week
  player_data_expanded <- player_data_expanded %>%
    group_by(day_of_week) %>%
    mutate(
      unique_dates = list(unique(date)),
      date_index = match(date, unique(date)),
      y_offset = (date_index - 1) * 0.2,
      y_position = as.numeric(day_of_week) + y_offset
    ) %>%
    ungroup()
  
  # Sort by day of week, date, and start time
  player_data_expanded <- player_data_expanded %>%
    arrange(day_of_week, date, game_start_time, game_end_time)
  
  # Add a session ID within each day to distinguish multiple sessions on same day
  player_data_expanded <- player_data_expanded %>%
    group_by(day_of_week, date) %>%
    mutate(session_id_within_day = row_number()) %>%
    ungroup()
  
  # Convert match_type to factor with proper levels for consistent ordering
  player_data_expanded$match_type <- factor(
    player_data_expanded$match_type,
    levels = c("social match", "pve match", "pvp match (rush)", "pvp match (vanquish)", "pvp match (arena)", "podium celebration", "0"),
    labels = c("Social Lobby", "PvE Match", "PvP (Rush)", "PvP (Vanquish)", "PvP (Arena)", "Podium Celebration", "Other")
  )
  
  plot <- player_data_expanded %>%
    ggplot(aes(start_time_only, y_position, col = match_type, 
               group = interaction(day_of_week, date, session_id_within_day))) +
    scale_color_manual(
      name = "Match Type", 
      values = c(
        "Social Lobby" = "#56B4E9",
        "PvE Match" = "#009E73",
        "PvP (Rush)" = "#E69F00",
        "PvP (Vanquish)" = "#D55E00",
        "PvP (Arena)" = "#CC79A7",
        "Podium Celebration" = "#0072B2",
        "Other" = "#999999"
      )
    ) +
    geom_segment(
      aes(xend = end_time_only, yend = y_position),
      linewidth = 2.5,
      lineend = "round"
    ) +
    geom_point(size = 1.5, alpha = 0.6) +
    geom_point(aes(x = end_time_only), size = 1.5, alpha = 0.6) +
    scale_x_datetime(
      labels = scales::date_format("%H:%M"),
      breaks = scales::date_breaks("3 hours"),
      limits = c(
        as.POSIXct(strptime("00:00", format = "%H:%M")),
        as.POSIXct(strptime("23:59", format = "%H:%M"))
      ),
      expand = c(0.01, 0.01)
    ) +
    scale_y_continuous(
      breaks = 1:7,
      labels = levels(player_data_expanded$day_of_week),
      limits = c(0.5, 7.4),
      expand = c(0, 0)
    ) +
    labs(
      x = NULL, 
      y = NULL
    ) +
    theme_minimal(base_size = 14) +
    theme(
      legend.position = "none",
      plot.margin = margin(10, 10, 10, 10),
      
      axis.text.y = element_text(size = 12, face = "bold"),
      axis.text.x = element_text(size = 11),
      axis.title = element_text(size = 13, face = "bold"),
      
      panel.grid.major.x = element_line(color = "grey90"),
      panel.grid.minor = element_blank(),
      panel.grid.major.y = element_line(color = "grey90"),
      
      panel.background = element_rect(fill = "white", color = NA),
      plot.background = element_rect(fill = "white", color = NA)
    )
  
  return(plot)
}
```

```{r}
#| label: session-timeline-data-prep
#| include: false
#| cache: false

# Load game session data if not already included
game_sessions <- read_csv(here("data/processed/game_time_2week.csv"))

# Make sure timestamps are in the correct format
game_sessions$game_start_time <- as.POSIXct(game_sessions$game_start_time, format = "%Y-%m-%d %H:%M:%S")
game_sessions$game_end_time <- as.POSIXct(game_sessions$game_end_time, format = "%Y-%m-%d %H:%M:%S")

# First, get a list of player IDs in the data
player_ids <- unique(game_sessions$player_id)

# Choose specific indices
# If those indices don't exist in our data, we'll use the first three players
player_indices <- c(
  min(401, length(player_ids)),  # Equivalent to slice(401) 
  min(6, length(player_ids)),    # Equivalent to slice(6)
  min(3, length(player_ids))     # Equivalent to slice(3)
)

selected_player_ids <- player_ids[player_indices]
```

```{r}
#| label: create-individual-plots
#| include: false
#| cache: false

# Create individual plots for each player with standardized x-axis
plot1 <- plot_player_sessions(
  game_sessions %>% filter(player_id == selected_player_ids[1])
)

plot2 <- plot_player_sessions(
  game_sessions %>% filter(player_id == selected_player_ids[2])
)

# For the third plot, use the same time range for consistency
plot3 <- plot_player_sessions(
  game_sessions %>% filter(player_id == selected_player_ids[3])
)
```

```{r}
#| label: fig-session-timeline
#| fig-cap: "Session timelines for two example players showing gaming activity across the two-week study period. Each horizontal bar represents a gaming session, with colours indicating match type (Social Lobby, PvE Match, PvP Rush/Vanquish/Arena). Each player's panel contains two rows corresponding to the two weeks of telemetry data, collapsed onto a single weekly timeline to facilitate comparison of play patterns across weeks."
#| fig-alt: "Two-panel Gantt-style chart with days of the week (Monday to Sunday) on the vertical axis and time of day (00:00 to 23:59) on the horizontal axis. Player A's panel shows predominantly green PvE sessions clustered in evening hours on a few days, with short scattered sessions. Player B's panel shows longer, more frequent sessions spanning multiple match types including orange and red PvP modes alongside green PvE, distributed more evenly across the week and across a wider range of hours. The contrast illustrates how two players can exhibit markedly different temporal patterns and mode preferences within the same two-week window."
#| fig-width: 12
#| fig-height: 8
#| code-summary: "Session timeline visualization"

figure <- ggarrange(
  plot2, plot1,
  labels = c("Player A", "Player B"),
  font.label = list(size = 14, face = "bold", color = "black"),
  ncol = 1, 
  nrow = 2, 
  common.legend = TRUE, 
  legend = "bottom",
  heights = c(1, 1),
  align = "v",
  hjust = -0.1
)

# Add shared x and y labels to the combined figure
fig_sessiontime <- annotate_figure(figure,
                        bottom = text_grob("Time of Day (hours)", 
                                          face = "bold", size = 14),
                        left = text_grob("Day of Week", 
                                         face = "bold", size = 14, rot = 90))

# Display the combined figure
fig_sessiontime
```

### Behavioural telemetry

The raw telemetry contained session-based records of in-game events for each player. From these records, we selected and computed metrics only if they had been previously validated as informative measures of player behaviour and performance in shooter games [@drachenGunsSwordsData2012; @lugrinImmersiveFPSGames2013; @melhartYourGameplaySays2019]. Two principles guided the computation of the behavioural metrics. First, all performance metrics were constructed to be independent of total playtime. Common cumulative indicators such as ‘player level’ or ‘total kills’ were deliberately excluded because they conflate skill with playtime; a player's level may simply be higher because they played more, not because they performed better. Second, every metric was computed separately for PvP and PvE modes, because the two contexts differ fundamentally in difficulty, objectives, and opponent type; eliminating a computer-controlled enemy in PvE is not comparable to eliminating a matchmade opponent of similar rank in PvP. Pooling across modes would obscure meaningful behavioural differences.
This resulted in 12 behavioural metrics plus one overall playtime measure (for descriptive statistics, see @tbl-descriptive-stats). Mode ratio captured the proportion of active playtime each player devoted to PvP and PvE modes, reflecting game-mode preference. Kill/death ratio indexed combat effectiveness as the number of eliminations per death, with higher values indicating greater survivability and lethality. Hit accuracy measures the proportion of shots that connect with a target, reflecting aiming precision. Damage per second captured sustained offensive output, indicating how efficiently a player dealt damage over time. Average damage per game reflected total offensive contribution within a match, encompassing both direct combat and area-of-effect output. Average score per game provided a composite performance indicator incorporating objectives completed, assists, and other in-game scoring events beyond raw combat. Each of these five performance metrics were computed separately for PvP and PvE, yielding 10 mode-specific indicators alongside the two mode-ratio variables.

```{r}
#| label: tbl-descriptive-stats
#| tbl-cap: "Descriptive Statistics for Key Variables"
#| fig.width: 10
#| fig.height: 6
#| tbl-cap-location: top
#| echo: false
#| warning: false
#| message: false
#| cache: false

# Create a simple APA-style descriptive statistics table
# First, check variable types to handle them appropriately
variable_types <- sapply(pvz_df_analysis, class)

# Handle numeric variables
numeric_vars <- pvz_df_analysis %>%
  select(where(is.numeric)) %>%
  select(
    # Demographics
    age,
    
    # Playtime 
    hours, ratio_pvp_hours, ratio_pve_hours,
    
    # Performance metrics
    pvp_kd, pve_kd, 
    pvp_hit_accuracy, pve_hit_accuracy,
    pvp_dps, pve_dps,
    pvp_average_damage, pve_average_damage,
    pvp_average_score, pve_average_score,
    
    # Wellbeing and enjoyment
    spane_game_balance, enjoyment
  )

# Calculate descriptive statistics for numeric variables
numeric_stats <- numeric_vars %>%
  summarise(across(everything(), 
                  list(
                    n = ~sum(!is.na(.)),
                    M = ~mean(., na.rm = TRUE),
                    SD = ~sd(., na.rm = TRUE),
                    Min = ~min(., na.rm = TRUE),
                    Max = ~max(., na.rm = TRUE)
                  ))) %>%
  pivot_longer(everything(), 
               names_to = c("variable", "stat"), 
               names_pattern = "(.*)_(.*)",
               values_to = "value") %>%
  pivot_wider(names_from = stat, values_from = value)

# For categorical variables like gender, just count frequencies for the footnote
if ("gender" %in% names(pvz_df_analysis)) {
  gender_counts <- pvz_df_analysis %>%
    mutate(
      gender_clean = case_when(
        str_trim(as.character(gender)) == "Male" ~ "Male",
        str_trim(as.character(gender)) == "Female" ~ "Female",
        str_trim(as.character(gender)) == "Other" ~ "Other",
        str_trim(as.character(gender)) == "Prefer not to say" ~ "Prefer not to say",
        TRUE ~ "Prefer not to say"
      )
    ) %>%
    count(gender_clean, name = "n") %>%
    mutate(gender = gender_clean) %>%
    select(gender, n) %>%
    mutate(pct = n / sum(n) * 100) %>%
    mutate(pct = round(pct, 1)) %>%
    mutate(frequency = paste0(n, " [", pct, "]"))
  
  # Note: We'll add a footnote to explain gender distribution separately
  # Using brackets instead of parentheses to avoid LaTeX issues
  gender_note <- paste("Gender distribution:", 
                     paste(paste0(gender_counts$gender, ": ", gender_counts$frequency), 
                           collapse = ", "))
} else {
  gender_note <- NULL
}

# Format the data for the table
descriptive_stats <- numeric_stats %>%
  # Format variable names for better readability
  mutate(variable = case_when(
    variable == "age" ~ "Age",
    variable == "hours" ~ "Total Hours Played",
    variable == "ratio_pvp_hours" ~ "Proportion of PvP Playtime",
    variable == "ratio_pve_hours" ~ "Proportion of PvE Playtime",
    variable == "pvp_kd" ~ "PvP Kill/Death Ratio",
    variable == "pve_kd" ~ "PvE Kill/Death Ratio",
    variable == "pvp_hit_accuracy" ~ "PvP Hit Accuracy",
    variable == "pve_hit_accuracy" ~ "PvE Hit Accuracy",
    variable == "pvp_dps" ~ "PvP Damage per Second",
    variable == "pve_dps" ~ "PvE Damage per Second",
    variable == "pvp_average_damage" ~ "PvP Average Damage per Game",
    variable == "pve_average_damage" ~ "PvE Average Damage per Game",
    variable == "pvp_average_score" ~ "PvP Average Score",
    variable == "pve_average_score" ~ "PvE Average Score",
    variable == "spane_game_balance" ~ "Game-Specific Wellbeing (SPANE)",
    variable == "enjoyment" ~ "Enjoyment",
    TRUE ~ variable
  )) %>%
  # Create category variable for grouping
  mutate(category = case_when(
    variable == "Age" ~ "Demographics",
    variable %in% c("Total Hours Played", "Proportion of PvP Playtime", "Proportion of PvE Playtime") ~ "Playtime",
    variable %in% c("PvP Kill/Death Ratio", "PvE Kill/Death Ratio", 
                   "PvP Hit Accuracy", "PvE Hit Accuracy",
                   "PvP Damage per Second", "PvE Damage per Second",
                   "PvP Average Damage per Game", "PvE Average Damage per Game",
                   "PvP Average Score", "PvE Average Score") ~ "Performance",
    variable %in% c("Game-Specific Wellbeing (SPANE)", "Enjoyment") ~ "Wellbeing",
    TRUE ~ "Other"
  )) %>%
  # Round numeric values
  mutate(
    M = round(M, 2),
    SD = round(SD, 2),
    Min = round(Min, 2),
    Max = round(Max, 2)
)

# Create custom ordering for variables to ensure autonomy, competence, relatedness are listed in order
variable_order <- c(
  # Demographics
  "Age",
  
  # Playtime
  "Total Hours Played", "Proportion of PvP Playtime", "Proportion of PvE Playtime",
  
  # Performance
  "PvP Kill/Death Ratio", "PvE Kill/Death Ratio", 
  "PvP Hit Accuracy", "PvE Hit Accuracy",
  "PvP Damage per Second", "PvE Damage per Second",
  "PvP Average Damage per Game", "PvE Average Damage per Game",
  "PvP Average Score", "PvE Average Score",
  
  # Wellbeing
  "Game-Specific Wellbeing (SPANE)", "Enjoyment"
)

# Apply the custom ordering
descriptive_stats <- descriptive_stats %>%
  mutate(variable = factor(variable, levels = variable_order)) %>%
  arrange(variable)

# Get sample size for the note
sample_n <- nrow(pvz_df_analysis)

# Create the table with tinytable
descriptive_stats |>
  select(Variable = variable, M, SD, Min, Max) |>
  tt(
    caption = "Descriptive Statistics for Key Variables",
    notes = paste0("N = ", sample_n, ". ", gender_note)
  ) |>
  format_tt(j = 2:5, digits = 2)
```


### Wellbeing

We assessed affective wellbeing with the Scale of Positive and Negative Experiences [SPANE; @dienerNewWellbeingMeasures2010; @dienerAdvancesOpenQuestions2018]. Respondents were prompted to reflect on how they felt over the past two weeks and rated how often they experienced each of six positive (positive, good, pleasant, happy, joyful, contented) and six negative (negative, bad, unpleasant, sad, afraid, angry) feelings on a scale from 1 (Very rarely or never) to 7 (Very often or always). An overall wellbeing balance score was computed by subtracting the mean negative score from the mean positive score, yielding a possible range of −6 to +6, where higher values indicate more positive affect. 

### Enjoyment

Enjoyment was assessed with four items from the Player Experience and Need Satisfaction scale [PENS; @ryanIntrinsicMotivationVideo2006], validated by @johnsonValidationPlayerExperience2018. The full PENS comprises subscales measuring autonomy, competence, relatedness, enjoyment, and extrinsic motivation; we used only the four-item enjoyment subscale, which captures hedonic appreciation of the play experience. Respondents were asked to reflect on their experience playing PvZ over the past two weeks and rated items such as "I think PvZ was fun to play" on a scale from 1 (Strongly disagree) to 7 (Strongly agree).

## Data analysis

Data were analyzed in R version 4.5.1 [@rCoreTeamRLanguageEnvironment2025]. Key packages included `fpc` for cluster validation indices [@hennig2020a], `factoextra` for multivariate visualization [@kassambara2020a], and `robustlmm` for robust mixed-effects estimation [@kollerRobustlmmRPackage2016].

Following initial data cleaning, all survey responses were screened for straightlining (e.g. when respondents give identical answers to all items) to identify records with low data quality commonly found in online surveys [@leinerTooFastToo2019], however, no unusual cases were identified. Afterwards, the telemetry records were matched to the same two-week window referenced by the survey items, and analyses were conducted on active players with sufficient mode-specific telemetry.

To test whether clustering was appropriate before fitting any cluster model, we evaluated cluster tendency using three complementary approaches. First, we computed the Hopkins statistic [@banerjee2004a], which compares distances in observed data to distances expected under spatial randomness; values near 0.5 indicate random structure, whereas larger values indicate clusterable structure. Second, we used Visual Assessment of Cluster Tendency (VAT) [@bezdek2002a], which reorders the dissimilarity matrix so potential clusters appear as darker diagonal blocks. Third, we examined Principal Component Analysis (PCA) projections to assess whether separable groupings were visible in low-dimensional space relative to random reference data. Together, these diagnostics indicated non-random structure suitable for clustering.

We then identified play styles with Partitioning Around Medoids (PAM) [@kaufmanFindingGroupsData1990] on standardized in-game behavioral features. PAM was selected over k-means because it is less sensitive to extreme points and defines each cluster by an observed medoid (a real player profile), which improves robustness and interpretability for telemetry-derived behavioral profiles. Features were z-standardized before clustering, and missing mode-specific performance values were set to zero to encode no activity in that mode rather than dropping cases. The number of clusters was determined by maximizing the Calinski-Harabasz index across k = 2 to 10 [@tibshiraniEstimatingNumberClusters2001], which selected a three-cluster solution; we then checked internal coherence with silhouette widths [@kaufmanFindingGroupsData1990].

To examine associations between play style, playtime, and outcomes, we fit robust linear mixed-effects models [@kollerRobustlmmRPackage2016]. Robust estimation was preferred over standard maximum-likelihood mixed models because telemetry-derived playtime and performance metrics are characteristically right-skewed and prone to influential observations; a small number of highly engaged players can exert disproportionate leverage on ordinary least-squares estimates. Robust estimation downweights such influential cases automatically through iteratively reweighted least squares, yielding coefficient estimates that better reflect the bulk of the sample without requiring ad hoc outlier exclusion rules that introduce researcher degrees of freedom. This approach also maintains methodological consistency with our clustering strategy: PAM was selected precisely because it is less sensitive to extreme points, and using a robust estimator at the modelling stage carries that same design philosophy through the analytic pipeline. All models included random intercepts for country to account for baseline between-country differences; random slopes were not estimated because the number of countries was too small to support the additional parameters reliably. Outcomes were treated as continuous and approximately normally distributed, and playtime was entered on its raw scale. Approximate *p*-values were derived from normal-distribution reference of *t*-statistics, and 95% confidence intervals were computed as *b* ± 1.96 × SE. We used simple-slopes analyses to estimate within-play-style playtime slopes and pairwise slope-difference tests to compare those slopes across play styles. Statistical significance was set at *p* < .05.

## Data, Materials, and Code

The present study was not preregistered. The data and code reported in this study can be found at <<<<<LINK OSF/ZENODO>>>>>. The raw data, materials, codebook and further details on the dataset shared by @johannesVideoGamePlay2021 can be found at <https://osf.io/cjd6z/>.

# Results

## RQ1. Player Clustering Analysis

```{r}
#| label: cluster_feature_selection
#| warning: false
#| message: false

# Select relevant features for clustering
player_cluster_features <- pvz_df_analysis %>% 
  select(
    # Playstyle variables (game mode preferences)
    ratio_pvp_hours, ratio_pve_hours,
    
    # PvP performance metrics
    pvp_kd, pvp_hit_accuracy, pvp_dps, pvp_average_damage, pvp_average_score,
    
    # PvE performance metrics
    pve_kd, pve_hit_accuracy, pve_dps, pve_average_damage, pve_average_score
  ) %>%
  # Replace NAs with 0s - makes sense because missing performance = no activity in that mode
  replace_na(list(
    ratio_pvp_hours = 0, ratio_pve_hours = 0,
    pvp_kd = 0, pvp_hit_accuracy = 0, pvp_dps = 0, 
    pvp_average_damage = 0, pvp_average_score = 0,
    pve_kd = 0, pve_hit_accuracy = 0, pve_dps = 0,
    pve_average_damage = 0, pve_average_score = 0
  ))

# Standardize the data (important for clustering)
player_data_scaled <- scale(player_cluster_features)
rownames(player_data_scaled) <- pvz_df_analysis$player_id[rownames(player_data_scaled) %in% rownames(pvz_df_analysis)]
```

### Assessing Cluster Tendency

```{r}
#| label: cluster_random_data
#| warning: false
#| message: false
#| results: "hide"

# Create a random dataset with the same dimensions for comparison
set.seed(123) # For reproducibility
random_cluster_data <- apply(player_data_scaled, 2,
                     function(x){runif(length(x), min(x), max(x))})
random_cluster_data <- as.data.frame(random_cluster_data)
```

```{r}
#| label: hopkins_statistic
#| warning: false
#| message: false
#| results: "hide"

# Create meaningful features for clustering analysis
# Focus on performance metrics that differentiate playstyles
meaningful_features <- pvz_df_analysis |>
  mutate(
    # Only include players with some gameplay data
    has_pvp = !is.na(pvp_kd) & ratio_pvp_hours > 0,
    has_pve = !is.na(pve_kd) & ratio_pve_hours > 0
  ) |>
  filter(has_pvp | has_pve) |>  # At least one mode
  select(
    ratio_pvp_hours, 
    pvp_kd, 
    pvp_hit_accuracy,
    pve_kd, 
    pve_hit_accuracy
  ) |>
  # Replace NAs with 0 for missing performance metrics
  # (missing performance = no activity in that mode)
  replace_na(list(
    ratio_pvp_hours = 0,
    pvp_kd = 0, 
    pvp_hit_accuracy = 0,
    pve_kd = 0, 
    pve_hit_accuracy = 0
  ))

# Standardize the features
meaningful_scaled <- scale(meaningful_features)

# Calculate Hopkins statistic using manual implementation 
# (more reliable than hopkins package for constrained gaming data)
set.seed(123)
hopkins_results <- replicate(10, {
  manual_hopkins(meaningful_scaled, m = 30)
}, simplify = TRUE)

# Calculate final statistics for manuscript reporting
hopkins_mean <- mean(hopkins_results)
hopkins_sd <- sd(hopkins_results)
hopkins_ci_lower <- hopkins_mean - 1.96 * hopkins_sd
hopkins_ci_upper <- hopkins_mean + 1.96 * hopkins_sd

# Display results
cat("Hopkins Statistic Results:\n")
cat("Sample size:", nrow(meaningful_features), "observations\n")
cat("Mean Hopkins statistic (m=30):", round(hopkins_mean, 4), "\n")
cat("95% CI: [", round(hopkins_ci_lower, 4), ",", round(hopkins_ci_upper, 4), "]\n")

# Interpretation
if(hopkins_mean > 0.75) {
  cat("Interpretation: Strong evidence of clustering tendency (H > 0.75)\n")
} else if(hopkins_mean > 0.5) {
  cat("Interpretation: Moderate clustering tendency\n") 
} else {
  cat("Interpretation: Data appears randomly distributed\n")
}

# Store final values for reporting in results section
hopkins_final <- hopkins_mean
hopkins_ci_final <- c(hopkins_ci_lower, hopkins_ci_upper)

# Validation: Test on random data to ensure method works correctly
set.seed(456)
random_test_data <- matrix(rnorm(nrow(meaningful_features) * ncol(meaningful_features)), 
                          nrow = nrow(meaningful_features), 
                          ncol = ncol(meaningful_features))
random_scaled <- scale(random_test_data)

hopkins_random_check <- mean(replicate(5, {
  manual_hopkins(random_scaled, m = 30)
}))

cat("\nValidation - Hopkins on random data:", round(hopkins_random_check, 4))
cat(" (should be ~0.5)\n")
```

```{r}
#| label: pca_visualization
#| warning: false
#| message: false
#| results: "hide"

# Visual Assessment using PCA visualization
pca_real_data <- prcomp(player_data_scaled)
pca_random_data <- prcomp(random_cluster_data)

# Create PCA plots for comparison
pca_plot_real <- fviz_pca_ind(pca_real_data, 
                  title = "PCA - Player Performance Data",
                  geom = "point", 
                  ggtheme = theme_minimal(),
                  legend = "none",
                  alpha.ind = 0.7) +
      theme(plot.title = element_text(face = "bold", size = 14))

pca_plot_random <- fviz_pca_ind(pca_random_data, 
                  title = "PCA - Random Data",
                  geom = "point", 
                  ggtheme = theme_minimal(),
                  legend = "none",
                  alpha.ind = 0.7) +
      theme(plot.title = element_text(face = "bold", size = 14))
```

```{r}
#| label: vat_analysis
#| echo: false
#| warning: false
#| message: false
#| results: "hide"

# Using ordered dissimilarity matrices
dist_real_data <- get_dist(player_data_scaled, method = "euclidean")
dist_random_data <- get_dist(random_cluster_data, method = "euclidean")

# Create ordered dissimilarity matrix plots (VAT approach)
dist_plot_real <- fviz_dist(dist_real_data, 
                show_labels = FALSE,
                order = TRUE, # Important: order the matrix for VAT
                gradient = list(low = "#00AFBB", mid = "white", high = "#FC4E07")) +
       ggtitle("VAT Ordered Dissimilarity Matrix - Player Data") +
       theme(plot.title = element_text(face = "bold", size = 14))

dist_plot_random <- fviz_dist(dist_random_data, 
                show_labels = FALSE,
                order = TRUE, # Important: order the matrix for VAT
                gradient = list(low = "#00AFBB", mid = "white", high = "#FC4E07")) +
       ggtitle("VAT Ordered Dissimilarity Matrix - Random Data") +
       theme(plot.title = element_text(face = "bold", size = 14))
```

The Hopkins statistic computed from our player behavioural data was high (`r round(hopkins_final, 4)`, 95% CI [`r round(hopkins_ci_final[1], 4)`, `r round(hopkins_ci_final[2], 4)`]), substantially exceeding the 0.70 threshold and indicating strong cluster tendency. As a validity check, a random dataset with identical dimensions produced a substantially lower value (`r round(hopkins_random_check, 2)`), closer to the 0.5 benchmark expected under spatial randomness. VAT plots showed clear dark diagonal blocks for the player data, consistent with separable groups, while the random reference data showed no comparable block structure (Appendix 1; @bezdek2002a). PCA visualisations converged on the same conclusion, with visible grouping in the player data but not in the random comparison. The convergence of these three independent assessments (Hopkins, VAT, and PCA) provided strong statistical evidence to reject the null hypothesis of spatial randomness, demonstrating that the player behavioural data contained meaningful, non-random structure amenable to clustering methods.


### PAM Clustering

```{r}
#| label: pam_silhouette_method
#| fig.width: 10
#| fig.height: 4
#| cache: false
#| results: "asis"
#| fig.show: "hide"  # Changed from "asis" to "hide" to prevent displaying here

# PAM optimal cluster determination
# -------------------------------------
# Calinski-Harabasz Index (CH index) for PAM
ch_index_pam <- function(data, max_k = 10) {
  ch_values <- numeric(max_k - 1)
  for (i in 2:max_k) {
    pam_result <- pam(data, k = i)
    ch_values[i-1] <- fpc::calinhara(data, pam_result$clustering)
  }
  k_values <- 2:max_k
  return(data.frame(k = k_values, ch_index = ch_values))
}

ch_pam_results <- ch_index_pam(player_data_scaled)
p_pam_ch <- ggplot(ch_pam_results, aes(x = k, y = ch_index)) +
  geom_line() +
  geom_point() +
  scale_x_continuous(breaks = 2:10) +
  labs(title = "Calinski-Harabasz Index",
       x = "Number of clusters k",
       y = "CH Index (higher is better)") +
  theme_minimal()

# Display the plot
p_pam_ch
```

```{r}
#| label: determine_optimal_k
#| echo: false
# Determine the optimal k value based on CH Index
ch_optimal_k <- ch_pam_results$k[which.max(ch_pam_results$ch_index)]
final_optimal_k <- ch_optimal_k
```

To select the number of clusters, we evaluated PAM solutions for k = 2 to 10 using the Calinski-Harabasz (CH) index, where higher values indicate better separation relative to within-cluster dispersion [@kaufmanFindingGroupsData1990; @tibshiraniEstimatingNumberClusters2001]. The CH index peaked at k = `r final_optimal_k` (Appendix 2), supporting a three-cluster solution as the best trade-off between fit and parsimony.

```{r}
#| label: prepare_player_data

# Prepare player data for clustering
player_data_features <- pvz_df_analysis %>% 
  select(
    # Game mode preferences
    ratio_pvp_hours, ratio_pve_hours,
    
    # Performance metrics
    pvp_kd, pve_kd, 
    pvp_hit_accuracy, pve_hit_accuracy,
    pvp_average_score, pve_average_score,
    pvp_dps, pve_dps
  ) %>%
  # Replace NAs with 0s instead of removing them
  replace_na(list(
    ratio_pvp_hours = 0, ratio_pve_hours = 0,
    pvp_kd = 0, pve_kd = 0,
    pvp_hit_accuracy = 0, pve_hit_accuracy = 0,
    pvp_average_score = 0, pve_average_score = 0,
    pvp_dps = 0, pve_dps = 0
  ))

# Standardize the data for clustering
player_data_scaled <- scale(player_data_features)

# Perform PAM clustering with k=3
set.seed(42) # For reproducibility
k_clusters <- 3
pam_result <- pam(player_data_scaled, k = k_clusters, metric = "euclidean")

```

```{r}
#| label: pam_cluster_statistics
#| cache: false
#| include: false

# Calculate silhouette coefficient for each cluster using the existing pam_result
sil <- pam_result$silinfo$widths
sil_per_cluster <- tapply(sil[, 3], sil[, 1], mean)

# Store key statistics
stats_pam_k3 <- list(
  k = k_clusters,
  avg_silhouette = pam_result$silinfo$avg.width,
  silhouette_per_cluster = sil_per_cluster,
  cluster_sizes = table(pam_result$clustering),
  medoid_indices = pam_result$id.med
)
```

Having selected k = 3 on the basis of the CH index, we assessed the internal validity of this solution using silhouette analysis [@kaufmanFindingGroupsData1990], which quantifies how well each observation fits its assigned cluster relative to the nearest neighbouring cluster. The average silhouette width was `r round(stats_pam_k3$avg_silhouette, 2)`. According to Kaufman and Rousseeuw's interpretive guidelines, values above 0.25 indicate strong structure and values between 0.10 and 0.25 indicate weak but still meaningful structure; the overall value thus falls at the boundary, suggesting reasonable separation. Per-cluster silhouette widths ranged from `r round(min(stats_pam_k3$silhouette_per_cluster), 2)` to `r round(max(stats_pam_k3$silhouette_per_cluster), 2)`, with one cluster exceeding the 0.25 threshold and the remaining two showing acceptable cohesion. No negative silhouette values were observed, confirming that every observation was assigned to its most appropriate cluster. The substantive interpretation of these three clusters is presented below.




```{r}
#| label: cluster_profiles
#| include: false

# Define the feature columns used in clustering
feature_columns <- c("ratio_pvp_hours", "ratio_pve_hours", 
                    "pvp_kd", "pve_kd", 
                    "pvp_hit_accuracy", "pve_hit_accuracy",
                    "pvp_average_score", "pve_average_score",
                    "pvp_dps", "pve_dps")

# Get cluster medoids (representative players)
medoid_indices <- pam_result$id.med
medoid_data <- pvz_df_analysis[medoid_indices, ]
medoid_features <- medoid_data %>%
  select(all_of(feature_columns))
medoid_features$cluster <- 1:k_clusters

# Calculate mean values by cluster for interpretation
cluster_means <- pvz_df_analysis %>%
  # First add the cluster assignments
  mutate(cluster = pam_result$clustering) %>%
  group_by(cluster) %>%
  summarize(across(all_of(feature_columns), mean, na.rm = TRUE), .groups = "drop")

# Display medoid profiles 
medoid_features |>
  tt(caption = "Cluster Medoids (Feature Values of Representative Players)") |>
  format_tt(digits = 2)

# Display cluster means
cluster_means |>
  tt(caption = "Cluster Means - Average Feature Values") |>
  format_tt(digits = 2)
```

```{r}
#| label: assign_player_types
#| code-summary: "Assign player type labels to clusters"

# Create a data frame with cluster labels 
# Update these names based on interpretation of the 3 clusters
player_type_labels <- tibble(
  cluster = 1:3,
  player_type = c(
    "Casual PvP Players",           # Cluster 1 name
    "Casual PvE Players",       # Cluster 2 name 
    "Elite Allround Players"            # Cluster 3 name
  )
)

# Add cluster assignments and type labels to the dataframe
pvz_df_clustered <- pvz_df_analysis %>%
  mutate(cluster = pam_result$clustering) %>%
  left_join(player_type_labels, by = "cluster") %>%
  mutate(player_type = factor(player_type))

# Save the clustered data with no outliers
write_csv(pvz_df_clustered, here("data/processed/pvz_with_clusters_pam3_nohours_no_outliers.csv"))

```

```{r}
#| label: compute-key-statistics
#| code-summary: "Compute key statistics for inline reporting"

# Calculate player type summary statistics for inline reporting
player_type_summary <- pvz_df_clustered %>%
  group_by(player_type) %>%
  summarise(
    n = n(),
    pct = n() / nrow(pvz_df_clustered) * 100,
    pvp_time_pct = mode_mean(ratio_pvp_hours) * 100,
    pve_time_pct = mode_mean(ratio_pve_hours) * 100,
    pvp_kd = mode_mean(pvp_kd),
    pve_kd = mode_mean(pve_kd),
    pvp_accuracy = mode_mean(pvp_hit_accuracy) * 100,
    pve_accuracy = mode_mean(pve_hit_accuracy) * 100,
    wellbeing_mean = mean(spane_game_balance, na.rm = TRUE),
    enjoyment_mean = mean(enjoyment, na.rm = TRUE),
    .groups = "drop"
  )

# Extract model coefficients for inline reporting
get_model_coef <- function(model, term) {
  coefs <- coef(summary(model))
  if (term %in% rownames(coefs)) {
    return(list(
      b = coefs[term, "Estimate"],
      se = coefs[term, "Std. Error"],
      t = coefs[term, "t value"]
    ))
  }
  return(NULL)
}

# Store key coefficients after models are fit (will be populated after model fitting)
```

```{r}
#| label: radar_functions

# Function to scale values using z-scores
z_score_scale <- function(x) {
  if (sd(x, na.rm = TRUE) == 0) return(rep(0, length(x)))
  return((x - mean(x, na.rm = TRUE)) / sd(x, na.rm = TRUE))
}

# Convert z-scores to 0-1 scale for radar chart
z_to_radar_scale <- function(z, min_z = -2.5, max_z = 2.5) {
  z_clipped <- pmax(pmin(z, max_z), min_z)
  scaled <- (z_clipped - min_z) / (max_z - min_z)
  return(scaled)
}

# Function to create radar chart with average player comparison
create_player_type_radar <- function(data, avg_data, color, title) {
  data_cols <- colnames(data)
  
  max_min <- data.frame(matrix(c(rep(1, length(data_cols)), rep(0, length(data_cols))), 
                             nrow = 2, byrow = TRUE))
  colnames(max_min) <- data_cols
  
  plot_data <- rbind(max_min, avg_data, data)
  
  radarchart(
    plot_data, 
    pcol = c("gray50", color),
    pfcol = c(adjustcolor("gray50", alpha.f = 0.15), adjustcolor(color, alpha.f = 0.4)),
    plwd = c(2.5, 3.5),              
    plty = c(2, 1),
    vlcex = 1.3,                     
    title = title,
    cglcol = "grey70",
    cglty = 1,
    axislabcol = "grey20",           
    caxislabels = seq(0, 1, 0.25),
    cglwd = 1,                       
    pty = 32,
    axistype = 0,                    
    centerzero = FALSE,
    seg = 4,                         
    cex.main = 1.6                   
  )
}
```

```{r}
#| label: player_type_summary
#| include: false

# Create a summary of player types
player_type_summary <- pvz_df_clustered %>%
  group_by(player_type) %>%
  summarise(
    Count = n(),
    `% of Players` = n() / nrow(pvz_df_clustered) * 100,
    `PvP Time %` = mode_mean(ratio_pvp_hours) * 100,
    `PvE Time %` = mode_mean(ratio_pve_hours) * 100,
    `PvP K/D` = mode_mean(pvp_kd),
    `PvE K/D` = mode_mean(pve_kd),
    `PvP Accuracy` = mode_mean(pvp_hit_accuracy) * 100,
    `PvE Accuracy` = mode_mean(pve_hit_accuracy) * 100
  )
```

```{r}
#| label: player_type_inline_values
#| include: false

# Keep inline text simple: pull one row per player type, then named scalars
pt_pve <- player_type_summary %>% filter(player_type == "Casual PvE Players")
pt_elite <- player_type_summary %>% filter(player_type == "Elite Allround Players")
pt_pvp <- player_type_summary %>% filter(player_type == "Casual PvP Players")

pve_pct_players <- round(pt_pve$`% of Players`, 1)
pve_time <- round(pt_pve$`PvE Time %`, 1)
pve_pve_kd <- round(pt_pve$`PvE K/D`, 2)
pve_pve_acc <- round(pt_pve$`PvE Accuracy`, 1)

elite_pct_players <- round(pt_elite$`% of Players`, 1)
elite_pvp_kd <- round(pt_elite$`PvP K/D`, 2)
elite_pve_kd <- round(pt_elite$`PvE K/D`, 2)
elite_pvp_acc <- round(pt_elite$`PvP Accuracy`, 1)
elite_pve_acc <- round(pt_elite$`PvE Accuracy`, 1)

pvp_pct_players <- round(pt_pvp$`% of Players`, 1)
pvp_time <- round(pt_pvp$`PvP Time %`, 1)
pvp_pvp_kd <- round(pt_pvp$`PvP K/D`, 2)
pvp_pvp_acc <- round(pt_pvp$`PvP Accuracy`, 1)
```

### Interpretation of Clusters

Following established approaches to telemetry-based player profiling [@bauckhageClusteringGameBehavior2015; @drachenGunsSwordsData2012; @drachen2014a; @melhartYourGameplaySays2019; @sifaProfilingGamesUnderstanding2018], we labelled the three clusters on the basis of their dominant mode preferences and performance levels (@fig-radar-charts). The labels reflect two dimensions that emerged from the data: the game mode players predominantly engaged with (PvE vs. PvP) and their overall performance level (casual vs. elite).

Casual PvP Players (`r pvp_pct_players`% of the sample) devoted the majority of their time to competitive PvP modes (`r pvp_time`%), yet their performance was moderate (PvP K/D: M = `r pvp_pvp_kd`; PvP Accuracy: M = `r pvp_pvp_acc`%), falling well below Elite Allround Players. Performance in PvE was similarly modest. This pattern indicates a preference for competitive engagement that is not accompanied by correspondingly high skill, suggesting a casual orientation to play within a competitive context.

Casual PvE Players (`r pve_pct_players`% of the sample) allocated the majority of their playtime to cooperative PvE modes (`r pve_time`%), achieving moderate performance within that context (PvE K/D: M = `r pve_pve_kd`; PvE Accuracy: M = `r pve_pve_acc`%) while showing markedly lower metrics in PvP. This profile suggests a specialisation in story-driven exploration and cooperative play over competitive engagement.

Elite Allround Players (`r elite_pct_players`% of the sample), the largest cluster, maintained high performance across both gameplay contexts, with kill-to-death ratios and accuracy rates substantially above the sample mean in both PvP (K/D: M = `r elite_pvp_kd`; Accuracy: M = `r elite_pvp_acc`%) and PvE (K/D: M = `r elite_pve_kd`; Accuracy: M = `r elite_pve_acc`%). Rather than specialising in one mode, these players engaged broadly and performed well regardless of context, distinguishing them as the most proficient segment.

```{r}
#| label: radar-caption-values
#| include: false

# Calculate summary values for each player type for the figure caption
radar_caption_data <- pvz_df_clustered %>%
  group_by(player_type) %>%
  summarise(
    pvp_ratio = mode_mean(ratio_pvp_hours),
    pve_ratio = mode_mean(ratio_pve_hours),
    pvp_kd = mode_mean(pvp_kd),
    pve_kd = mode_mean(pve_kd),
    .groups = "drop"
  )

radar_caption <- paste0(

"Radar charts showing performance and mode selection for each playstyle relative to the playstyle of the average player (dashed line). Values are standardized mean scores (z-scores converted to 0-1 scale). ",
"Mean values: Casual PvE Players (PvP Ratio: ", type_stat(radar_caption_data, "Casual PvE Players", "pvp_ratio", 2),
", PvE Ratio: ", type_stat(radar_caption_data, "Casual PvE Players", "pve_ratio", 2),
", PvP K/D: ", type_stat(radar_caption_data, "Casual PvE Players", "pvp_kd", 2),
", PvE K/D: ", type_stat(radar_caption_data, "Casual PvE Players", "pve_kd", 2), "); ",
"Elite Allround Players (PvP Ratio: ", type_stat(radar_caption_data, "Elite Allround Players", "pvp_ratio", 2),
", PvE Ratio: ", type_stat(radar_caption_data, "Elite Allround Players", "pve_ratio", 2),
", PvP K/D: ", type_stat(radar_caption_data, "Elite Allround Players", "pvp_kd", 2),
", PvE K/D: ", type_stat(radar_caption_data, "Elite Allround Players", "pve_kd", 2), "); ",
"Casual PvP Players (PvP Ratio: ", type_stat(radar_caption_data, "Casual PvP Players", "pvp_ratio", 2),
", PvE Ratio: ", type_stat(radar_caption_data, "Casual PvP Players", "pve_ratio", 2),
", PvP K/D: ", type_stat(radar_caption_data, "Casual PvP Players", "pvp_kd", 2),
", PvE K/D: ", type_stat(radar_caption_data, "Casual PvP Players", "pve_kd", 2), "). ",
"PvE = Player versus Environment; PvP = Player versus Player; K/D = Kill/Death ratio; DPS = Damage per second."
)
```

```{r}
#| label: fig-radar-charts
#| fig-cap: !expr radar_caption
#| fig-alt: "Three radar charts arranged in a 2-by-2 grid with a shared legend in the fourth cell. Each chart plots 10 performance metrics (PvE Ratio, PvE K/D, PvE Accuracy, PvE Score, PvE DPS, PvP Ratio, PvP K/D, PvP Accuracy, PvP Score, PvP DPS) on radial axes scaled from 0 to 1. A dashed grey polygon marks the average player profile. The Casual PvE Players chart shows high values on all PvE metrics and low values on PvP metrics, creating a lopsided shape skewed toward PvE. The Elite Allround Players chart shows uniformly high values across both PvE and PvP metrics, forming a large, roughly symmetric polygon that exceeds the average on nearly every axis. The Casual PvP Players chart shows elevated PvP Ratio but otherwise below-average values on most performance metrics, with especially low PvE scores. The legend identifies the dashed grey line as Average Player and three coloured filled polygons as the three play styles."
#| fig-height: 12
#| fig-width: 12
#| out-width: "100%"
#| code-summary: "Player type radar chart visualization"

# Define metrics for radar charts in a more logical order, 
# grouping related metrics together
radar_metrics <- c(
  # PvE metrics grouped together
  "PvE Ratio", "PvE K/D", "PvE Accuracy", "PvE Score", "PvE DPS",
  # PvP metrics grouped together
  "PvP Ratio", "PvP K/D", "PvP Accuracy", "PvP Score", "PvP DPS"
)

# Get player type names
player_type_names <- player_type_labels$player_type

# Calculate radar chart data for both average player and player types
# First collect the raw metrics for all players and player types
metrics_data <- pvz_df_analysis %>%
  # Add a column for the average player
  mutate(profile_type = "Average Player") %>%
  # Bind the clustered data with profile types
  bind_rows(
    pvz_df_clustered %>%
      select(
        player_id, player_type, 
        ratio_pve_hours, ratio_pvp_hours,
        pve_kd, pvp_kd,
        pve_hit_accuracy, pvp_hit_accuracy,
        pve_average_score, pvp_average_score,
        pve_dps, pvp_dps
      ) %>%
      rename(profile_type = player_type)
  ) %>%
  # Group by profile type (average or a player type)
  group_by(profile_type) %>%
  # Calculate means for each metric
  summarise(
    ratio_pve_mean = mode_mean(ratio_pve_hours),
    pve_kd_mean = mode_mean(pve_kd),
    pve_accuracy_mean = mode_mean(pve_hit_accuracy),
    pve_score_mean = mode_mean(pve_average_score),
    pve_dps_mean = mode_mean(pve_dps),
    ratio_pvp_mean = mode_mean(ratio_pvp_hours),
    pvp_kd_mean = mode_mean(pvp_kd),
    pvp_accuracy_mean = mode_mean(pvp_hit_accuracy),
    pvp_score_mean = mode_mean(pvp_average_score),
    pvp_dps_mean = mode_mean(pvp_dps)
  ) %>%
  ungroup()

# Create vectors of all values for each metric to calculate z-scores across all profiles
all_ratio_pve <- metrics_data$ratio_pve_mean
all_pve_kd <- metrics_data$pve_kd_mean
all_pve_accuracy <- metrics_data$pve_accuracy_mean
all_pve_score <- metrics_data$pve_score_mean
all_pve_dps <- metrics_data$pve_dps_mean
all_ratio_pvp <- metrics_data$ratio_pvp_mean
all_pvp_kd <- metrics_data$pvp_kd_mean
all_pvp_accuracy <- metrics_data$pvp_accuracy_mean
all_pvp_score <- metrics_data$pvp_score_mean
all_pvp_dps <- metrics_data$pvp_dps_mean

# Now apply z-score scaling to create the radar data
radar_data <- metrics_data %>%
  mutate(
    `PvE Ratio` = z_to_radar_scale(z_score_scale(all_ratio_pve)[row_number()]),
    `PvE K/D` = z_to_radar_scale(z_score_scale(all_pve_kd)[row_number()]),
    `PvE Accuracy` = z_to_radar_scale(z_score_scale(all_pve_accuracy)[row_number()]),
    `PvE Score` = z_to_radar_scale(z_score_scale(all_pve_score)[row_number()]),
    `PvE DPS` = z_to_radar_scale(z_score_scale(all_pve_dps)[row_number()]),
    `PvP Ratio` = z_to_radar_scale(z_score_scale(all_ratio_pvp)[row_number()]),
    `PvP K/D` = z_to_radar_scale(z_score_scale(all_pvp_kd)[row_number()]),
    `PvP Accuracy` = z_to_radar_scale(z_score_scale(all_pvp_accuracy)[row_number()]),
    `PvP Score` = z_to_radar_scale(z_score_scale(all_pvp_score)[row_number()]),
    `PvP DPS` = z_to_radar_scale(z_score_scale(all_pvp_dps)[row_number()])
  ) %>%
  # Keep only relevant columns
  select(
    profile_type,
    `PvE Ratio`, `PvE K/D`, `PvE Accuracy`, `PvE Score`, `PvE DPS`,
    `PvP Ratio`, `PvP K/D`, `PvP Accuracy`, `PvP Score`, `PvP DPS`
  )

# Extract the average player data
average_player <- radar_data %>%
  filter(profile_type == "Average Player") %>%
  select(-profile_type) %>%
  as.data.frame()

player_type_colors <- unname(playstyle_colors)

layout(matrix(c(1, 2, 3, 4), nrow = 2, byrow = TRUE))
par(mar = c(2, 5, 4, 5), oma = c(1, 0, 0, 0), xpd = TRUE)

for (i in seq_along(player_type_names)) {
  type_data <- radar_data %>%
    filter(profile_type == player_type_names[i]) %>%
    select(-profile_type) %>%
    as.data.frame()
  
  create_player_type_radar(
    type_data, 
    average_player,
    player_type_colors[i], 
    player_type_names[i]
  )
}

# Panel 4: shared legend
par(mar = c(0, 0, 0, 0))
plot.new()
legend(
  "center",
  legend = c("Average Player", player_type_names),
  col = c("gray50", player_type_colors),
  lty = c(2, 1, 1, 1),
  lwd = c(2.5, 3.5, 3.5, 3.5),
  fill = c(adjustcolor("gray50", alpha.f = 0.15),
           adjustcolor(player_type_colors[1], alpha.f = 0.4),
           adjustcolor(player_type_colors[2], alpha.f = 0.4),
           adjustcolor(player_type_colors[3], alpha.f = 0.4)),
  border = c("gray50", player_type_colors),
  bty = "n",
  cex = 1.5,
  pt.cex = 1.4
)

par(mfrow = c(1, 1))
```

## RQ2. Wellbeing differences between player types

```{r}
#| label: player-type-wellbeing-model
#| code-summary: "Fit robust mixed-effects models for wellbeing outcomes"
#| message: false
#| warning: false
#| include: false

# Create a robust linear model for wellbeing
set.seed(123) # For reproducibility of robust estimates

# Check if models should be refit or loaded from cache
if (eval_models || !model_exists("rmodel_game") || !model_exists("rmodel_enjoyment")) {
  # Fit robust models for game-specific wellbeing
  rmodel_game <- rlmer(spane_game_balance ~ player_type + hours + gender + age + (1|country), 
                       data = pvz_df_clustered)
  
  # Fit robust models for enjoyment
  rmodel_enjoyment <- rlmer(enjoyment ~ player_type + hours + gender + age + (1|country), 
                       data = pvz_df_clustered)
  
  # Save models to cache
  save_model(rmodel_game, "rmodel_game")
  save_model(rmodel_enjoyment, "rmodel_enjoyment")
} else {
  # Load cached models
  rmodel_game <- load_model("rmodel_game")
  rmodel_enjoyment <- load_model("rmodel_enjoyment")
}

# Display models side-by-side using modelsummary
modelsummary(
  list("Wellbeing (SPANE)" = rmodel_game, "Enjoyment" = rmodel_enjoyment),
  estimate = "{estimate} ({std.error})",
  statistic = "t = {statistic}",
  gof_omit = "AIC|BIC|RMSE|Log",
  coef_rename = c(
    "(Intercept)" = "Intercept",
    "player_typeElite Allround Players" = "Elite Allround Players",
    "player_typeCasual PvP Players" = "Casual PvP Players",
    "hours" = "Hours Played",
    "genderMale" = "Gender (Male)",
    "genderOther" = "Gender (Other)",
    "genderPrefer not to say" = "Gender (Undisclosed)",
    "age" = "Age"
  ),
  title = "Robust Linear Mixed-Effects Models Predicting Wellbeing and Enjoyment",
  notes = "Reference category for play style: Casual PvE Players. Reference category for gender: Female. Country included as random effect."
)

# Extract coefficients for inline reporting
coefs_game <- coef(summary(rmodel_game))
coefs_enjoyment <- coef(summary(rmodel_enjoyment))

# Compute approximate marginal and conditional R² (Nakagawa & Schielzeth, 2013)
compute_r2 <- function(model) {
  fe_var <- var(predict(model, re.form = NA))
  re_var <- as.numeric(VarCorr(model)$country)
  resid_var <- sigma(model)^2
  total_var <- fe_var + re_var + resid_var
  c(marginal = fe_var / total_var, conditional = (fe_var + re_var) / total_var)
}

r2_game <- compute_r2(rmodel_game)
r2_enjoyment <- compute_r2(rmodel_enjoyment)

# Format coefficient with 95% CI and approximate p-value for inline reporting
fmt_coef <- function(coefs, row, prefix = "b") {
  est <- coefs[row, "Estimate"]
  se <- coefs[row, "Std. Error"]
  t_val <- coefs[row, "t value"]
  ci_lo <- est - 1.96 * se
  ci_hi <- est + 1.96 * se
  p_val <- 2 * pnorm(abs(t_val), lower.tail = FALSE)
  p_str <- if (p_val < .001) "< .001" else sprintf("= %.3f", p_val)
  sprintf("*%s* = %.2f, 95%% CI [%.2f, %.2f], *t* = %.2f, *p* %s",
          prefix, est, ci_lo, ci_hi, t_val, p_str)
}
```

For each outcome (game-specific wellbeing and enjoyment), we estimated a main-effects model predicting the outcome from player type, hours played, gender, and age, with random intercepts for country (outcome ~ player_type + hours + gender + age + (1 | country); see Appendix 3). The fixed effects accounted for a modest share of outcome variance (wellbeing: marginal *R*² = `r sprintf("%.3f", r2_game["marginal"])`, conditional *R*² = `r sprintf("%.3f", r2_game["conditional"])`; enjoyment: marginal *R*² = `r sprintf("%.3f", r2_enjoyment["marginal"])`, conditional *R*² = `r sprintf("%.3f", r2_enjoyment["conditional"])`). Casual PvE Players served as the reference category. Relative to these players, both Casual PvP Players (`r fmt_coef(coefs_game, "player_typeCasual PvP Players")`) and Elite Allround Players (`r fmt_coef(coefs_game, "player_typeElite Allround Players")`) reported significantly lower game-specific wellbeing. For enjoyment, Casual PvP Players again scored lower than Casual PvE Players (`r fmt_coef(coefs_enjoyment, "player_typeCasual PvP Players")`), whereas the difference between Elite Allround and Casual PvE Players was smaller and not statistically reliable (`r fmt_coef(coefs_enjoyment, "player_typeElite Allround Players")`) (see @fig-anova-boxplots).

```{r}
#| label: fig-anova-boxplots
#| fig-cap: "Distribution of wellbeing outcomes by play style. Panel A shows game-specific wellbeing (SPANE balance score); Panel B shows enjoyment. Boxes represent interquartile ranges with median lines; whiskers extend to 1.5 × IQR. Individual points show outliers."
#| fig-alt: "Side-by-side boxplots in two panels for three play styles (Casual PvE Players, Elite Allround Players, Casual PvP Players). Panel A (Game-Specific Wellbeing, SPANE balance score on Y-axis) shows Casual PvE Players with the highest median around 2, Elite Allround Players slightly lower near 1, and Casual PvP Players lowest near 0. All three distributions overlap substantially. Panel B (Enjoyment score on Y-axis) shows a similar ordering: Casual PvE Players highest with a median near 5.5, Elite Allround Players near 5, and Casual PvP Players lowest near 4.5. Semi-transparent jittered points overlay each box, revealing that most data clusters in the upper half of the score range for all groups."
#| fig-width: 10
#| fig-height: 5
#| message: false
#| warning: false

# Create boxplot for game-specific wellbeing
p_wellbeing <- ggplot(pvz_df_clustered, aes(x = player_type, y = spane_game_balance, fill = player_type)) +
  geom_boxplot(alpha = 0.7, outlier.shape = 21) +
  geom_jitter(width = 0.2, alpha = 0.2, size = 1) +
  playstyle_scale_fill +
  labs(
    title = "A. Game-Specific Wellbeing",
    x = NULL,
    y = "SPANE Balance Score"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    legend.position = "none",
    plot.title = element_text(size = 15, face = "bold"),
    axis.title = element_text(size = 13),
    axis.text = element_text(size = 12),
    axis.text.x = element_text(angle = 15, hjust = 1, size = 12),
    panel.grid.major.x = element_blank()
  )

# Create boxplot for enjoyment
p_enjoyment <- ggplot(pvz_df_clustered, aes(x = player_type, y = enjoyment, fill = player_type)) +
  geom_boxplot(alpha = 0.7, outlier.shape = 21) +
  geom_jitter(width = 0.2, alpha = 0.2, size = 1) +
  playstyle_scale_fill +
  labs(
    title = "B. Enjoyment",
    x = NULL,
    y = "Enjoyment Score"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    legend.position = "none",
    plot.title = element_text(size = 15, face = "bold"),
    axis.title = element_text(size = 13),
    axis.text = element_text(size = 12),
    axis.text.x = element_text(angle = 15, hjust = 1, size = 12),
    panel.grid.major.x = element_blank()
  )

# Combine plots
p_wellbeing + p_enjoyment
```

## RQ3. Does Playstyle Moderate the Playtime–Wellbeing Relationship?

```{r}
#| label: moderation-models
#| code-summary: "Fit moderation models (play style × hours interaction)"
#| message: false
#| warning: false
#| include: false

# Check if models should be refit or loaded from cache
if (eval_models || !model_exists("rmodel_moderation_game") || !model_exists("rmodel_moderation_enjoyment")) {
  # Create interaction models for our two key wellbeing outcomes
  rmodel_moderation_game <- rlmer(
    spane_game_balance ~ player_type * hours + gender + age + (1|country),
    data = pvz_df_clustered
  )

  rmodel_moderation_enjoyment <- rlmer(
    enjoyment ~ player_type * hours + gender + age + (1|country),
    data = pvz_df_clustered
  )
  
  # Save models to cache
  save_model(rmodel_moderation_game, "rmodel_moderation_game")
  save_model(rmodel_moderation_enjoyment, "rmodel_moderation_enjoyment")
} else {
  # Load cached models
  rmodel_moderation_game <- load_model("rmodel_moderation_game")
  rmodel_moderation_enjoyment <- load_model("rmodel_moderation_enjoyment")
}

# Display moderation models side-by-side using modelsummary
modelsummary(
  list("Wellbeing (SPANE)" = rmodel_moderation_game, "Enjoyment" = rmodel_moderation_enjoyment),
  estimate = "{estimate} ({std.error})",
  statistic = "t = {statistic}",
  gof_omit = "AIC|BIC|RMSE|Log",
  coef_rename = c(
    "(Intercept)" = "Intercept",
    "player_typeElite Allround Players" = "Elite Allround Players",
    "player_typeCasual PvP Players" = "Casual PvP Players",
    "hours" = "Hours Played",
    "player_typeElite Allround Players:hours" = "Elite Allround x Hours",
    "player_typeCasual PvP Players:hours" = "Casual PvP x Hours"
  ),
  title = "Moderation Models: Play Style x Hours Interaction",
  notes = "Reference category for play style: Casual PvE Players. Country included as random effect."
)

# Extract coefficients for inline reporting
coefs_mod_game <- coef(summary(rmodel_moderation_game))
coefs_mod_enjoyment <- coef(summary(rmodel_moderation_enjoyment))

r2_mod_game <- compute_r2(rmodel_moderation_game)
r2_mod_enjoyment <- compute_r2(rmodel_moderation_enjoyment)
```

To test whether the playtime–wellbeing association differed across playstyles, we extended the main-effects models by adding player type × hours interaction terms, retaining gender and age as covariates (outcome ~ player_type × hours + gender + age + (1 | country); see Appendix 4). Adding the interaction terms did not meaningfully increase variance explained (wellbeing: marginal *R*² = `r sprintf("%.3f", r2_mod_game["marginal"])`; enjoyment: marginal *R*² = `r sprintf("%.3f", r2_mod_enjoyment["marginal"])`). In the main-effects models, hours played was positively associated with both outcomes (wellbeing: `r fmt_coef(coefs_game, "hours")`; enjoyment: `r fmt_coef(coefs_enjoyment, "hours")`). Each interaction coefficient represents the difference in the playtime slope for a given playstyle relative to Casual PvE Players. None of the interaction terms reached significance: for wellbeing, neither Casual PvP Players (`r fmt_coef(coefs_mod_game, "player_typeCasual PvP Players:hours", "Δb")`) nor Elite Allround Players (`r fmt_coef(coefs_mod_game, "player_typeElite Allround Players:hours", "Δb")`) had a playtime slope that differed from the reference group. The same held for enjoyment (Casual PvP: `r fmt_coef(coefs_mod_enjoyment, "player_typeCasual PvP Players:hours", "Δb")`; Elite Allround: `r fmt_coef(coefs_mod_enjoyment, "player_typeElite Allround Players:hours", "Δb")`). Playstyle differences in intercepts remained in the interaction models, particularly for Casual PvP Players (wellbeing: `r fmt_coef(coefs_mod_game, "player_typeCasual PvP Players")`; enjoyment: `r fmt_coef(coefs_mod_enjoyment, "player_typeCasual PvP Players")`).

```{r}
#| label: wellbeing-anova
#| message: false
#| warning: false
#| include: false

# Run ANOVAs to test for significant differences between player types
wellbeing_vars <- c("spane_game_balance", "enjoyment")

# Store ANOVA results and models
anova_results <- list()
anova_models <- list()

for (var in wellbeing_vars) {
  formula <- as.formula(paste(var, "~ player_type"))
  anova_model <- aov(formula, data = pvz_df_clustered)
  anova_models[[var]] <- anova_model
  anova_results[[var]] <- broom::tidy(anova_model)
  anova_results[[var]]$variable <- var
  
  # Add effect size
  anova_results[[var]]$eta_squared <- effectsize::eta_squared(anova_model)[["Eta2"]]
  
  # Add residual df
  anova_results[[var]]$df_residual <- summary(anova_model)[[1]]["Residuals", "Df"]
}

# Combine all results
all_anova_results <- bind_rows(anova_results)

# Display results in a nice table
all_anova_results |> 
  filter(term == "player_type") |>
  select(Variable = variable, df, SS = sumsq, F = statistic, p = p.value, `η²` = eta_squared) |>
  mutate(
    p = case_when(
      p < 0.001 ~ "< .001^†^",
      p < 0.05 ~ paste0(sprintf("%.3f", round(p, 3)), "^†^"),
      TRUE ~ sprintf("%.3f", round(p, 3))
    ),
    `η²` = round(`η²`, 3)
  )

# Post-hoc tests for variables with significant differences
posthoc_results <- list()

for (var in wellbeing_vars) {
  if (summary(anova_models[[var]])[[1]]["player_type", "Pr(>F)"] < 0.05) {
    posthoc <- TukeyHSD(anova_models[[var]])
    posthoc_results[[var]] <- broom::tidy(posthoc)
  }
}

# Extract key statistics for inline reporting
wellbeing_anova <- all_anova_results |> 
  filter(term == "player_type" & variable == "spane_game_balance")
enjoyment_anova <- all_anova_results |> 
  filter(term == "player_type" & variable == "enjoyment")

# Extract post-hoc p-values
wellbeing_posthoc <- posthoc_results[["spane_game_balance"]]
enjoyment_posthoc <- posthoc_results[["enjoyment"]]

# Get specific comparisons (TukeyHSD tidy output uses 'contrast' column)
wb_pvp_pve <- wellbeing_posthoc |> 
  filter(grepl("Casual PvP.*Casual PvE|Casual PvE.*Casual PvP", contrast))
enj_pvp_pve <- enjoyment_posthoc |> 
  filter(grepl("Casual PvP.*Casual PvE|Casual PvE.*Casual PvP", contrast))
enj_pvp_elite <- enjoyment_posthoc |> 
  filter(grepl("Casual PvP.*Elite|Elite.*Casual PvP", contrast))
```

```{r}
#| label: dynamic-results-text
#| echo: false

# Function to format p-value for inline text
format_p_inline <- function(p) {
  if (length(p) == 0 || is.na(p[1])) return("N/A")
  p <- p[1]  # Take first value if vector
  if (p < .001) return("< .001")
  sprintf("= %.3f", p)
}
```

These playstyle differences were corroborated by one-way ANOVAs (wellbeing: *F*(`r wellbeing_anova$df`, `r wellbeing_anova$df_residual`) = `r sprintf("%.2f", wellbeing_anova$statistic)`, *p* `r format_p_inline(wellbeing_anova$p.value)`, η² = `r sprintf("%.3f", wellbeing_anova$eta_squared)`; enjoyment: *F*(`r enjoyment_anova$df`, `r enjoyment_anova$df_residual`) = `r sprintf("%.2f", enjoyment_anova$statistic)`, *p* `r format_p_inline(enjoyment_anova$p.value)`, η² = `r sprintf("%.3f", enjoyment_anova$eta_squared)`) and Tukey HSD post-hoc tests, which confirmed that Casual PvP Players scored lower than Casual PvE Players on both wellbeing (*p* `r format_p_inline(wb_pvp_pve$adj.p.value)`) and enjoyment (*p* `r format_p_inline(enj_pvp_pve$adj.p.value)`), and lower than Elite Allround Players on enjoyment (*p* `r format_p_inline(enj_pvp_elite$adj.p.value)`).


```{r}
#| label: simple-slopes-functions
#| message: false
#| warning: false
#| include: false

# Function to calculate simple slopes for all player types
calculate_all_simple_slopes <- function(model, outcome_name) {
  # For robust mixed models, extract coefficients manually
  coefs <- fixef(model)
  vcov_matrix <- vcov(model)
  
  # Get player types
  player_types <- levels(pvz_df_clustered$player_type)
  
  # Base effect of hours (for reference level)
  hours_effect <- coefs["hours"]
  hours_se <- sqrt(vcov_matrix["hours", "hours"])
  
  # Create results dataframe
  results <- data.frame(
    player_type = player_types[1],  # reference level
    estimate = hours_effect,
    std.error = hours_se,
    statistic = hours_effect / hours_se,
    p.value = 2 * pt(abs(hours_effect / hours_se), 
                    df = nrow(pvz_df_clustered) - length(coefs), 
                    lower.tail = FALSE),
    stringsAsFactors = FALSE
  )
  
  # Calculate for other player types
  for (i in 2:length(player_types)) {
    pt <- player_types[i]
    interaction_term <- paste0("player_type", pt, ":hours")
    
    if (interaction_term %in% names(coefs)) {
      # Calculate slope
      slope <- hours_effect + coefs[interaction_term]
      
      # Calculate SE using delta method (simplified)
      var_hours <- vcov_matrix["hours", "hours"]
      var_inter <- vcov_matrix[interaction_term, interaction_term]
      cov_hours_inter <- vcov_matrix["hours", interaction_term]
      slope_se <- sqrt(var_hours + var_inter + 2*cov_hours_inter)
      
      # t-statistic and p-value
      t_stat <- slope / slope_se
      p_val <- 2 * pt(abs(t_stat), df = nrow(pvz_df_clustered) - length(coefs), lower.tail = FALSE)
      
      # Add to results
      results <- rbind(results, data.frame(
        player_type = pt,
        estimate = slope,
        std.error = slope_se,
        statistic = t_stat,
        p.value = p_val,
        stringsAsFactors = FALSE
      ))
    }
  }
  
  # Add confidence intervals and formatting
  results <- results %>%
    mutate(
      outcome = outcome_name,
      conf.low = estimate - 1.96 * std.error,
      conf.high = estimate + 1.96 * std.error,
      significant = p.value < 0.05,
      sig = case_when(
        p.value < 0.001 ~ "^†^",
        p.value < 0.01 ~ "^†^",
        p.value < 0.05 ~ "^†^",
        TRUE ~ ""
      )
    )
  
  return(results)
}

# Function to create interaction plots for all player types
create_interaction_plots <- function(model, outcome_name) {
  # Extract fixed effects
  fixed_effects <- fixef(model)
  
  # Get player types
  player_types <- levels(pvz_df_clustered$player_type)
  
  # Create a grid of values for hours
  hours_range <- range(pvz_df_clustered$hours, na.rm = TRUE)
  hours_grid <- seq(from = hours_range[1], to = hours_range[2], length.out = 100)
  
  # Get the coefficient for hours (main effect)
  hours_coef <- fixed_effects["hours"]
  
  # Create a dataframe to store predictions
  predictions <- data.frame()
  
  # For each player type, calculate predicted values across hours range
  for (i in 1:length(player_types)) {
    pt <- player_types[i]
    
    # Calculate intercept
    if (i == 1) {
      # Reference level
      intercept <- fixed_effects["(Intercept)"]
      slope <- hours_coef
    } else {
      # Other levels
      intercept <- fixed_effects["(Intercept)"] + fixed_effects[paste0("player_type", pt)]
      interaction_term <- paste0("player_type", pt, ":hours")
      slope <- hours_coef + fixed_effects[interaction_term]
    }
    
    # Calculate predictions
    pred <- intercept + slope * hours_grid
    
    # Add to predictions dataframe
    predictions <- rbind(predictions, data.frame(
      player_type = pt,
      hours = hours_grid,
      predicted = pred
    ))
  }
  
  # Create the plot
  p <- ggplot(predictions, aes(x = hours, y = predicted, color = player_type)) +
    geom_line(size = 1.2) +
    geom_ribbon(aes(ymin = predicted - 1.96 * 0.5, ymax = predicted + 1.96 * 0.5, fill = player_type), alpha = 0.2) +
    labs(
      title = paste("Interaction Plot for", outcome_name),
      x = "Hours Played",
      y = outcome_name,
      color = "Player Type",
      fill = "Player Type"
    ) +
    theme_minimal() +
    theme(
      plot.title = element_text(hjust = 0.5, face = "bold"),
      legend.position = "bottom",
      panel.grid.minor = element_blank()
    ) +
    playstyle_scale_color +
    playstyle_scale_fill
  
  return(p)
}

# Function to create Johnson-Neyman plot
create_jn_plot <- function(model, moderator = "player_type", predictor = "hours", outcome_name) {
  # For rlmerMod models, we'll use a custom approach
  # This function creates a simplified version of Johnson-Neyman plot
  
  # Extract fixed effects
  fixed_effects <- fixef(model)
  vcov_matrix <- vcov(model)
  
  # Get player types
  player_types <- levels(pvz_df_clustered$player_type)
  
  # Create a dataframe to store slopes and confidence intervals
  jn_data <- data.frame(
    player_type = player_types,
    slope = numeric(length(player_types)),
    se = numeric(length(player_types)),
    lower_ci = numeric(length(player_types)),
    upper_ci = numeric(length(player_types)),
    stringsAsFactors = FALSE
  )
  
  # Get the coefficient for hours (main effect)
  hours_coef <- fixed_effects["hours"]
  hours_se <- sqrt(vcov_matrix["hours", "hours"])
  
  # For each player type, calculate slope and CI
  for (i in 1:length(player_types)) {
    pt <- player_types[i]
    
    if (i == 1) {
      # Reference level
      jn_data$slope[i] <- hours_coef
      jn_data$se[i] <- hours_se
    } else {
      # Other levels
      interaction_term <- paste0("player_type", pt, ":hours")
      slope <- hours_coef + fixed_effects[interaction_term]
      
      # Calculate SE using delta method (simplified)
      var_hours <- vcov_matrix["hours", "hours"]
      var_inter <- vcov_matrix[interaction_term, interaction_term]
      cov_hours_inter <- vcov_matrix["hours", interaction_term]
      slope_se <- sqrt(var_hours + var_inter + 2*cov_hours_inter)
      
      jn_data$slope[i] <- slope
      jn_data$se[i] <- slope_se
    }
    
    # Calculate 95% confidence intervals
    jn_data$lower_ci[i] <- jn_data$slope[i] - 1.96 * jn_data$se[i]
    jn_data$upper_ci[i] <- jn_data$slope[i] + 1.96 * jn_data$se[i]
  }
  
  # Determine significance
  jn_data$significant <- (jn_data$lower_ci > 0) | (jn_data$upper_ci < 0)
  
  # Create the plot
  p <- ggplot(jn_data, aes(x = player_type, y = slope, color = significant)) +
    geom_point(size = 3) +
    geom_errorbar(aes(ymin = lower_ci, ymax = upper_ci), width = 0.2) +
    geom_hline(yintercept = 0, linetype = "dashed", color = "gray50") +
    labs(
      title = paste("Conditional Effect of", predictor, "on", outcome_name),
      subtitle = "by Player Type with 95% Confidence Intervals",
      x = "Player Type",
      y = paste("Effect of", predictor),
      color = "Significant"
    ) +
    theme_minimal() +
    theme(
      plot.title = element_text(hjust = 0.5, face = "bold"),
      plot.subtitle = element_text(hjust = 0.5),
      legend.position = "bottom",
      panel.grid.minor = element_blank()
    ) +
    scale_color_manual(values = c("gray50", "blue"))
  
  return(p)
}
```

```{r}
#| label: wellbeing-simple-slopes
#| message: false
#| warning: false
#| include: false

# Calculate simple slopes for game-specific wellbeing model
wellbeing_slopes <- calculate_all_simple_slopes(rmodel_moderation_game, "Game-Specific Wellbeing")

# Create a formatted table for game-specific wellbeing simple slopes
wellbeing_slopes_table <- wellbeing_slopes %>%
  mutate(
    estimate = round(estimate, 3),
    std.error = round(std.error, 3),
    statistic = round(statistic, 3),
    p.value = case_when(
      p.value < 0.001 ~ paste0("< .001", sig),
      p.value < 0.05 ~ paste0(sprintf("%.3f", round(p.value, 3)), sig),
      TRUE ~ sprintf("%.3f", round(p.value, 3))
    ),
    ci = paste0("[", round(conf.low, 3), ", ", round(conf.high, 3), "]")
  ) %>%
  select(`Player Type` = player_type, b = estimate, SE = std.error, t = statistic, p = p.value, `95% CI` = ci)
```

```{r}
#| label: enjoyment-simple-slopes
#| message: false
#| warning: false
#| include: false

# Calculate simple slopes for enjoyment model
enjoyment_slopes <- calculate_all_simple_slopes(rmodel_moderation_enjoyment, "Intrinsic Enjoyment")

# Create a formatted table for enjoyment simple slopes
enjoyment_slopes_table <- enjoyment_slopes %>%
  mutate(
    estimate = round(estimate, 3),
    std.error = round(std.error, 3),
    statistic = round(statistic, 3),
    p.value = case_when(
      p.value < 0.001 ~ paste0("< .001", sig),
      p.value < 0.05 ~ paste0(sprintf("%.3f", round(p.value, 3)), sig),
      TRUE ~ sprintf("%.3f", round(p.value, 3))
    ),
    ci = paste0("[", round(conf.low, 3), ", ", round(conf.high, 3), "]")
  ) %>%
  select(`Player Type` = player_type, b = estimate, SE = std.error, t = statistic, p = p.value, `95% CI` = ci)
```

```{r}
#| label: wellbeing-pairwise-comparisons
#| message: false
#| warning: false
#| include: false

# Create pairwise comparison of slopes for game-specific wellbeing
wellbeing_pairs <- data.frame(player_type1 = character(),
                        player_type2 = character(),
                        difference = numeric(),
                        std.error = numeric(),
                        statistic = numeric(),
                        p.value = numeric(),
                        stringsAsFactors = FALSE)

player_types <- levels(pvz_df_clustered$player_type)
for (i in 1:(length(player_types)-1)) {
  for (j in (i+1):length(player_types)) {
    pt1 <- player_types[i]
    pt2 <- player_types[j]
    
    # Get slopes and standard errors
    slope1 <- wellbeing_slopes$estimate[wellbeing_slopes$player_type == pt1]
    slope2 <- wellbeing_slopes$estimate[wellbeing_slopes$player_type == pt2]
    se1 <- wellbeing_slopes$std.error[wellbeing_slopes$player_type == pt1]
    se2 <- wellbeing_slopes$std.error[wellbeing_slopes$player_type == pt2]
    
    # Calculate difference and standard error
    diff <- slope1 - slope2
    diff_se <- sqrt(se1^2 + se2^2)  # Simplified approach, assumes independence
    
    # Calculate test statistic and p-value
    t_stat <- diff / diff_se
    p_val <- 2 * pt(abs(t_stat), df = nrow(pvz_df_clustered) - 10, lower.tail = FALSE)  # Approximate df
    
    # Add to results
    wellbeing_pairs <- rbind(wellbeing_pairs, data.frame(
      player_type1 = pt1,
      player_type2 = pt2,
      difference = diff,
      std.error = diff_se,
      statistic = t_stat,
      p.value = p_val,
      stringsAsFactors = FALSE
    ))
  }
}

# Format and display pairwise comparison table for subjective wellbeing
wellbeing_pairs <- wellbeing_pairs %>%
  mutate(
    sig = case_when(
      p.value < 0.001 ~ "^†^",
      p.value < 0.01 ~ "^†^",
      p.value < 0.05 ~ "^†^", 
      TRUE ~ ""
    ),
    difference = round(difference, 3),
    std.error = round(std.error, 3),
    statistic = round(statistic, 3),
    p.value = case_when(
      p.value < 0.001 ~ paste0("< .001", sig),
      p.value < 0.05 ~ paste0(sprintf("%.3f", round(p.value, 3)), sig),
      TRUE ~ sprintf("%.3f", round(p.value, 3))
    )
  ) |>
  rename(`Type 1` = player_type1, `Type 2` = player_type2, Diff = difference, SE = std.error, t = statistic, p = p.value)
```

```{r}
#| label: enjoyment-pairwise-comparisons
#| message: false
#| warning: false
#| include: false

# Create pairwise comparison of slopes for enjoyment model
enjoyment_pairs <- data.frame(player_type1 = character(),
                              player_type2 = character(),
                              difference = numeric(),
                              std.error = numeric(),
                              statistic = numeric(),
                              p.value = numeric(),
                              stringsAsFactors = FALSE)

for (i in 1:(length(player_types)-1)) {
  for (j in (i+1):length(player_types)) {
    pt1 <- player_types[i]
    pt2 <- player_types[j]
    
    # Get slopes and standard errors
    slope1 <- enjoyment_slopes$estimate[enjoyment_slopes$player_type == pt1]
    slope2 <- enjoyment_slopes$estimate[enjoyment_slopes$player_type == pt2]
    se1 <- enjoyment_slopes$std.error[enjoyment_slopes$player_type == pt1]
    se2 <- enjoyment_slopes$std.error[enjoyment_slopes$player_type == pt2]
    
    # Calculate difference and standard error
    diff <- slope1 - slope2
    diff_se <- sqrt(se1^2 + se2^2)  # Simplified approach, assumes independence
    
    # Calculate test statistic and p-value
    t_stat <- diff / diff_se
    p_val <- 2 * pt(abs(t_stat), df = nrow(pvz_df_clustered) - 10, lower.tail = FALSE)  # Approximate df
    
    # Add to results
    enjoyment_pairs <- rbind(enjoyment_pairs, data.frame(
      player_type1 = pt1,
      player_type2 = pt2,
      difference = diff,
      std.error = diff_se,
      statistic = t_stat,
      p.value = p_val,
      stringsAsFactors = FALSE
    ))
  }
}

# Format and display pairwise comparison table for enjoyment model
enjoyment_pairs <- enjoyment_pairs %>%
  mutate(
    sig = case_when(
      p.value < 0.001 ~ "^†††^",
      p.value < 0.01 ~ "^††^",
      p.value < 0.05 ~ "^†^", 
      TRUE ~ ""
    ),
    difference = round(difference, 3),
    std.error = round(std.error, 3),
    statistic = round(statistic, 3),
    p.value = case_when(
      p.value < 0.001 ~ paste0("< .001", sig),
      p.value < 0.05 ~ paste0(sprintf("%.3f", round(p.value, 3)), sig),
      TRUE ~ sprintf("%.3f", round(p.value, 3))
    )
  ) |>
  rename(`Type 1` = player_type1, `Type 2` = player_type2, Diff = difference, SE = std.error, t = statistic, p = p.value)
```

```{r}
#| label: fig-interaction-plots
#| fig-cap: "Interaction plots showing the relationship between hours played and psychological outcomes across play styles. Left panel shows game-specific wellbeing (SPANE balance score); right panel shows enjoyment. Lines represent linear regression slopes with 95% confidence intervals. Casual PvE Players and Elite Allround Players show positive associations between playtime and outcomes, while Casual PvP Players show no significant relationship."
#| fig-alt: "Two scatter plots side by side with total hours played on the horizontal axis. The left panel plots SPANE balance score (range negative 6 to 6) and the right panel plots enjoyment score (range 1 to 7). In each panel, individual data points are coloured by play style and overlaid with linear regression lines and shaded 95 percent confidence bands. In both panels, Casual PvE Players (blue) and Elite Allround Players (green) show upward-sloping lines indicating positive associations between playtime and outcomes. Casual PvP Players (orange) show a nearly flat line in both panels, indicating no meaningful association. Confidence bands are widest at higher hours values where data are sparser. Most data points cluster below 20 hours of total play."
#| fig-width: 12
#| fig-height: 6
#| code-summary: "Interaction plots for moderation analysis"
#| message: false
#| warning: false

# Create scatter plots for total playtime vs. outcomes across player types with APA-style formatting
wellbeing_plot <- ggplot(pvz_df_clustered, aes(x = hours, y = spane_game_balance, color = player_type)) +
  geom_point(alpha = 0.6) +
  geom_smooth(method = "lm", se = TRUE, alpha = 0.2, fullrange = TRUE) +
  playstyle_scale_color +
  scale_y_continuous(breaks = seq(-6, 6, 2)) +
  coord_cartesian(ylim = c(-6, 6)) +
  labs(
    x = "Total Hours Played",
    y = "SPANE Score",
    color = "Play Style"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    axis.title = element_text(size = 13),
    axis.text = element_text(size = 12),
    legend.text = element_text(size = 12),
    legend.title = element_text(size = 13),
    panel.grid.minor = element_blank(),
    panel.grid.major.x = element_blank()
  )

enjoyment_plot <- ggplot(pvz_df_clustered, aes(x = hours, y = enjoyment, color = player_type)) +
  geom_point(alpha = 0.6) +
  geom_smooth(method = "lm", se = TRUE, alpha = 0.2, fullrange = TRUE) +
  playstyle_scale_color +
  scale_y_continuous(breaks = seq(1, 7, 1)) +
  coord_cartesian(ylim = c(1, 7)) +
  labs(
    x = "Total Hours Played",
    y = "Enjoyment Score",
    color = "Play Style"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    axis.title = element_text(size = 13),
    axis.text = element_text(size = 12),
    legend.text = element_text(size = 12),
    legend.title = element_text(size = 13),
    panel.grid.minor = element_blank(),
    panel.grid.major.x = element_blank()
  )

ggarrange(wellbeing_plot, enjoyment_plot, ncol = 2, common.legend = TRUE, legend = "bottom")
```

```{r}
#| label: simple-slopes-inline-values
#| echo: false

# Extract simple slopes values for inline reporting
# Wellbeing slopes
wb_pve <- wellbeing_slopes |> filter(player_type == "Casual PvE Players")
wb_pvp <- wellbeing_slopes |> filter(player_type == "Casual PvP Players")
wb_elite <- wellbeing_slopes |> filter(player_type == "Elite Allround Players")

# Enjoyment slopes  
enj_pve <- enjoyment_slopes |> filter(player_type == "Casual PvE Players")
enj_pvp <- enjoyment_slopes |> filter(player_type == "Casual PvP Players")
enj_elite <- enjoyment_slopes |> filter(player_type == "Elite Allround Players")

# Wellbeing pairwise comparisons (need numeric version before formatting)
wb_pairs_numeric <- data.frame(player_type1 = character(),
                               player_type2 = character(),
                               difference = numeric(),
                               std.error = numeric(),
                               statistic = numeric(),
                               p.value = numeric(),
                               stringsAsFactors = FALSE)

for (i in 1:(length(player_types)-1)) {
  for (j in (i+1):length(player_types)) {
    pt1 <- player_types[i]
    pt2 <- player_types[j]
    slope1 <- wellbeing_slopes$estimate[wellbeing_slopes$player_type == pt1]
    slope2 <- wellbeing_slopes$estimate[wellbeing_slopes$player_type == pt2]
    se1 <- wellbeing_slopes$std.error[wellbeing_slopes$player_type == pt1]
    se2 <- wellbeing_slopes$std.error[wellbeing_slopes$player_type == pt2]
    diff <- slope1 - slope2
    diff_se <- sqrt(se1^2 + se2^2)
    t_stat <- diff / diff_se
    p_val <- 2 * pt(abs(t_stat), df = nrow(pvz_df_clustered) - 6, lower.tail = FALSE)
    wb_pairs_numeric <- bind_rows(wb_pairs_numeric, data.frame(
      player_type1 = pt1, player_type2 = pt2,
      difference = diff, std.error = diff_se, statistic = t_stat, p.value = p_val
    ))
  }
}

# Enjoyment pairwise comparisons
enj_pairs_numeric <- data.frame(player_type1 = character(),
                                player_type2 = character(),
                                difference = numeric(),
                                std.error = numeric(),
                                statistic = numeric(),
                                p.value = numeric(),
                                stringsAsFactors = FALSE)

for (i in 1:(length(player_types)-1)) {
  for (j in (i+1):length(player_types)) {
    pt1 <- player_types[i]
    pt2 <- player_types[j]
    slope1 <- enjoyment_slopes$estimate[enjoyment_slopes$player_type == pt1]
    slope2 <- enjoyment_slopes$estimate[enjoyment_slopes$player_type == pt2]
    se1 <- enjoyment_slopes$std.error[enjoyment_slopes$player_type == pt1]
    se2 <- enjoyment_slopes$std.error[enjoyment_slopes$player_type == pt2]
    diff <- slope1 - slope2
    diff_se <- sqrt(se1^2 + se2^2)
    t_stat <- diff / diff_se
    p_val <- 2 * pt(abs(t_stat), df = nrow(pvz_df_clustered) - 6, lower.tail = FALSE)
    enj_pairs_numeric <- bind_rows(enj_pairs_numeric, data.frame(
      player_type1 = pt1, player_type2 = pt2,
      difference = diff, std.error = diff_se, statistic = t_stat, p.value = p_val
    ))
  }
}

# Helper function to format p-values for inline text
fmt_p <- function(p) {
  if (p < 0.001) return("< .001")
  sprintf("= %.3f", p)
}
```

To further probe the interaction, simple slopes analyses decomposed the playtime–outcome association within each playstyle (@fig-interaction-plots). Only Elite Allround Players showed a significant positive slope for both wellbeing (b = `r round(wb_elite$estimate, 3)`, *p* `r fmt_p(wb_elite$p.value)`) and enjoyment (b = `r round(enj_elite$estimate, 3)`, *p* `r fmt_p(enj_elite$p.value)`). Slopes for Casual PvE Players (wellbeing: b = `r round(wb_pve$estimate, 3)`, *p* `r fmt_p(wb_pve$p.value)`; enjoyment: b = `r round(enj_pve$estimate, 3)`, *p* `r fmt_p(enj_pve$p.value)`) and Casual PvP Players (wellbeing: b = `r round(wb_pvp$estimate, 3)`, *p* `r fmt_p(wb_pvp$p.value)`; enjoyment: b = `r round(enj_pvp$estimate, 3)`, *p* `r fmt_p(enj_pvp$p.value)`) were not significant. Critically, pairwise comparisons of slopes did not provide evidence that the playtime–outcome association differed between any pair of playstyles (all *p*s > .05), though the present sample may lack the power to detect small slope differences. Taken together, these results indicate that playstyle is associated with different average levels of wellbeing and enjoyment, but the data do not support the conclusion that the playtime–outcome relationship itself varies meaningfully across playstyles.

# Discussion

The objectives of this study were two-fold. First, we aimed to complement the breadth of platform-level playtime research with the depth of granular in-game playstyle analysis. Through telemetry-driven profiling, we aimed to demonstrate how players' in-game behavioural heterogeneity can be modelled within a single game. Second, informed by the Integrated Model of Player Experience, we wanted to explore how players' interaction with the game mechanics and preferences for different in-game modes shape their play experience and subsequent wellbeing, giving us more insight into the conditions under which and for whom video game play might support or undermine wellbeing.

As such, this study addressed three research questions. First, whether distinct playstyles can be identified from in-game behavioural telemetry within a single game. Second, whether players with different playstyles report different levels of wellbeing and enjoyment. Third, whether the association between playtime and wellbeing outcomes differs across playstyles. In short, the findings support the first two but not the third. Three distinct playstyles were identified from behavioural telemetry and labelled to reflect two dimensions that emerged from the data: (1) the game mode of preference (PvE vs. PvP) and (2) their overall performance level (casual vs. elite). These playstyles were associated with meaningful differences in wellbeing and enjoyment, but we found no evidence that they moderated the relationship between playtime and wellbeing or enjoyment. The implications and limitations of these findings will be discussed in detail below.

The participants in the first cluster, named "Casual PvE Players", allocated the majority of their playtime to cooperative PvE modes, achieving moderate performance within that context while showing markedly lower metrics in PvP. This playstyle suggests a preference for story-driven exploration and cooperative play over competitive engagement, paralleling the "Adventurer" type identified by @melhartYourGameplaySays2019 in *Tom Clancy's The Division*, who similarly focused on solo and cooperative mission content rather than competitive play. We hypothesise that these players adopt a more recreation-oriented approach to play, gravitating toward narrative content and cooperative challenges that offer immediate enjoyment without the sustained pressure of competitive evaluation. Within the IMP framework [@elsonMoreStoriesButtons2014], this pattern reflects a preference for game contexts in which narrative and cooperative mechanics are foregrounded over adversarial mechanics, potentially affording greater perceived autonomy over how to engage with the game's content and a lower threshold for competence satisfaction. Consistent with this interpretation, Casual PvE Players reported the highest game-specific wellbeing and enjoyment of the three groups. This finding is consistent with @melhartYourGameplaySays2019's observation that cooperative and exploratory play features were among the strongest predictors of need satisfaction in *The Division*.

The players in the second cluster, called "Elite Allround Players", formed the largest cluster. They maintained high performance across all gameplay contexts, with kill-to-death ratios and accuracy rates substantially above the sample mean in both PvP and PvE. Rather than specialising in one mode, these players engaged broadly and performed well regardless of context, distinguishing them as the most proficient segment. This playstyle is conceptually consistent with the "Elite" type in @melhartYourGameplaySays2019's analysis of *The Division*, characterised by top-tier progression metrics, and with the high-performing player segments identified by @drachenGunsSwordsData2012 in *Battlefield 2: Bad Company 2*. The breadth of their engagement suggests that mode preference alone does not define this group; rather, it is the consistent pursuit of high performance that characterises their play, consistent with a mastery- or achievement-oriented orientation in which players seek to develop and exercise skill across the full range of available game content. Despite their superior performance, however, Elite Allround Players reported lower wellbeing and enjoyment than Casual PvE Players, though higher enjoyment than Casual PvP Players. This pattern is noteworthy, prior work suggests that competence satisfaction positively predicts wellbeing in gaming contexts [@klimmt2009a; @przybylskiMotivationalModelVideo2010; @tamboriniMediaEnjoymentNeed2011; @ryanHappinessHumanPotentials2001], however if competence alone drove wellbeing, the most skilled players should report the highest wellbeing. Our findings do not support this, suggesting that the context in which competence is exercised matters as well, consistent with the IMP's emphasis on both mechanics and context as jointly shaping player experience [@elsonMoreStoriesButtons2014]. Potentially, Elite Allround Players' broad engagement across both cooperative and competitive modes may dilute the wellbeing benefits that cooperative PvE environments appear to afford. Furthermore, their high skill may buffer against the competence frustration that competitive environments can impose on less proficient players, though without direct measurement, this remains speculative.

Finally, the players in the third cluster, called "Casual PvP Players", devoted the majority of their time to competitive PvP modes, yet their performance was moderate, falling well below that of Elite Allround Players. Their performance in PvE was similarly modest. This pattern indicates a preference for competitive engagement that is not accompanied by correspondingly high skill, suggesting a casual orientation to play within a competitive context. A comparable playstyle appears in @drachenGunsSwordsData2012's analysis of *Battlefield 2: Bad Company 2*, where a segment of players engaged frequently in competitive modes without achieving high performance metrics, and bears some resemblance to the "Social Dark Zone Player" in @melhartYourGameplaySays2019's *Division* typology, who preferred competitive areas but were distinguished more by their social engagement than their skill. These players were potentially drawn to the multi-player and competitive dynamics of PvP modes rather than to the mastery of competitive mechanics per se. This combination of competitive exposure and moderate skill may create a less optimal play experience. Competitive PvP environments are characterised by direct performance feedback, social comparison, and outcome uncertainty, conditions that could frustrate competence needs when players lack the skill to succeed consistently [@ballouHowVideoGames2025]. Consistent with this reasoning, Casual PvP Players reported the lowest wellbeing and enjoyment of the three groups. The finding that competitive engagement at moderate skill levels is associated with the poorest psychological outcomes aligns with evidence that need frustration in gaming contexts can undermine wellbeing [@ballouHowVideoGames2025], and suggests that the experiential costs of competitive play may not be offset by the social or excitement-related benefits that draw these players to PvP modes.

Lastly, we found no evidence that playstyle moderates the association between playtime and wellbeing or enjoyment. Although Elite Allround Players showed the only significant positive simple slope, this did not translate into a significant difference from the slopes of other groups. These findings align closely with @johannesVideoGamePlay2021, who tested whether self-reported player experiences moderated the playtime-wellbeing association in the same game, *Plants vs. Zombies: Battle for Neighborville*. They found that neither need satisfaction, nor enjoyment, nor extrinsic motivation significantly interacted with playtime in predicting wellbeing, concluding that player experience contributes to wellbeing independently rather than altering how playtime relates to it. Our study extends this result by testing moderation through an entirely different operationalisation, objectively measured behavioural playstyles derived from telemetry, rather than self-reported motivational experiences. That neither subjective player experience nor objective behavioural profile moderates the playtime-wellbeing association in this game strengthens the case that playtime alone is not meaningfully associated with wellbeing for most players [@ballouOpenPlayLongitudinal2025; @johannesVideoGamePlay2021; @vuorreTimeSpentPlaying2022].

It is worth noting, however, that the absence of moderation in our data does not constitute evidence that playtime is wholly irrelevant to wellbeing under all conditions. The playtime-wellbeing relationship is subject to a wide range of time-varying confounds, including physical health, grief, work stress, and leisure availability, that can bias observed associations in either direction [@ballouIllHealthGrief2025]. Our cross-sectional design and limited set of covariates leave open the possibility that such confounds obscure true differential effects. Moreover, our study examines only one game; it remains possible that playstyle-playtime interactions exist in games with greater mechanical diversity, stronger competitive ranking systems, or more pronounced social features. The null moderation finding should therefore be interpreted as specific to the present context rather than as a general claim about the impossibility of such effects.

What the data do show is that the playtime paradigm is incomplete in a critical respect; it misses qualitative differences in play that are associated with meaningful variation in average wellbeing. Playtime treats all minutes of play as equivalent, but players who devote their time to cooperative PvE gameplay report higher wellbeing than those who spend comparable time in competitive PvP modes. This between-group difference in levels, rather than in slopes, is the primary empirical contribution of the present study. In this respect, our results complement rather than contradict platform-level research; playtime may indeed be uninformative as a predictor of wellbeing, but the behavioural composition of that playtime is not.

## Limitations & Future Work

Several limitations constrain the interpretation of these findings and point toward productive directions for future research. First, the study examines a single game with a specific player population recruited by the developer, and we lack evidence that the identified playstyles generalise to other games, genres, or player populations. Unlike personality traits, playstyles may be contextual and fluid, varying with game design and a player's development over time.

Second, while our behavioural telemetry provided comprehensive metrics on core gameplay mechanics and performance, it captured only a subset of potential in-game behaviours. Notably absent were indicators of social interaction, which is an important dimension of multiplayer gaming experiences that the IMP identifies as part of the contextual layer of player experience [@elsonMoreStoriesButtons2014]. In this study, sessions spent in the game's social hub were considered idle time and thus excluded from the playstyle analysis because these sessions generated no mechanics-relevant telemetry on which to cluster. Yet recent work has shown that ostensibly idle or non-interactive periods in games, including socialising between matches and exploring social spaces, can serve experiential functions such as community-building and strategic preparation that shape the broader play experience [@tepponenUnderstandingWaitingVideo2025]. Future work could address this gap by extending the playstyle profiling framework to incorporate social and communicative dimensions of gameplay. Gesture-based digital trace data, such as the 'ping' communication systems common in multiplayer games, have been demonstrated to be related to game context, team dynamics, and performance outcomes [@zhengUnderstandingPlayersGestureBased2023].

Third, the cross-sectional design precludes causal inference. The observed associations between playstyle and wellbeing may reflect selection effects (e.g., players with higher baseline wellbeing may gravitate toward cooperative play) rather than the influence of play behaviour on outcomes. This concern is not merely hypothetical; a wide range of time-varying confounds, from physical health conditions to work stress, plausibly affect both what players do in games and how they feel [@ballouIllHealthGrief2025]. Our design cannot disentangle whether cooperative PvE play promotes wellbeing, whether wellbeing promotes cooperative play, or whether both are driven by unmeasured third variables.

## Conclusion

This study demonstrates that players of the same game engage with it in qualitatively different ways and that these differences are associated with meaningful variation in wellbeing and enjoyment. However, we found no evidence that playstyle moderates the playtime-wellbeing relationship. These findings contribute to the growing body on digital trace research in two respects. First, they provide empirical support for the view that gameplay constitutes heterogeneous exposures with distinct psychological correlates, even within a single title. Second, they demonstrate a practical methodology, telemetry-driven profiling, for incorporating in-game behavioural depth into wellbeing research. At the same time, the null moderation results caution against overclaiming; the present data show that what players do is associated with how they feel, but not that different styles of play change how playtime itself relates to wellbeing. As richer forms of in-game data become available through industry-academic partnerships, researchers should dig deeper and continue to develop more granular, theory-driven approaches to understanding player experience and its relationship to psychological outcomes.

## Acknowledgements

Claude (Sonnet/Opus 4.6; Anthropic) was used to assist with code development and minor text edits during the preparation of this manuscript. All AI-assisted contributions were reviewed, verified, and, where necessary, further revised by members of the author team, who maintain full responsibility for the final content.

## Author contributions
Conceptualization: TH, AKP;
Data curation: TH, MV;
Formal analysis: TH, MV;
Funding acquisition: AKP;
Investigation: TH;
Methodology: TH, MV;
Project administration: AKP;
Supervision: MV, AKP;
Visualization: TH;
Writing – original draft: TH;
Writing – review & editing: TH, MV, AKP


## Data Availability

All data, materials, and code related to the dataset and this manuscript are available under {{< meta data-license >}} at <https://github.com/Thomhak/playstyle-article.git>.

## Funding

[funding]

## Conflicts of Interest

The authors declare no competing financial, intellectual, or institutional interests relevant to this research. The disclosures below are provided in the interest of full transparency.
No authors have received funding or compensation from the tech companies or other industries connected to the work presented here.
MV’s work is or has been funded by the European Research Council, the Economic and Social Research Council, Tilburg University, and the University of Oxford’s John Fell Fund. MV has provided in-kind consultations and served as a nonpaid panel member for Meta and K-Games. AKP has served in advisory or expert roles for the Sync Digital Wellbeing Program (2022–2024), the Google Expert Advisory on Youth and Technology (2025), the OpenAI Expert Council on Wellbeing and AI (2025), and UK Department for Science, Innovation and Technology–funded research on children's smartphone and social media use (University of Cambridge, 2025). He is a member of the UK Department for Culture, Media & Sport College of Experts. Any industry fees are directly donated to charity, and his research is conducted in accordance with University of Oxford academic integrity policies.



# References

::: {#refs}
:::

```{=typst}
#show: appendix.with()
```

# Appendix {.appendix}

## Appendix 1: Cluster Tendency Assessment

```{r}
#| label: appendix_cluster_tendency
#| fig.width: 12
#| fig.height: 10
#| echo: false
#| warning: false
#| message: false
#| fig.cap: "Cluster tendency assessment comparing player data (left column) to random data (right column). Top row: Visual Assessment of Cluster Tendency (VAT) ordered dissimilarity matrices. Bottom row: PCA visualizations. The player data shows clear clustering structure (dark blocks in VAT, visible groupings in PCA) while random data shows no such patterns."
#| fig.alt: "Two-by-two grid of four panels. Top-left (A, VAT Player Data): a heatmap of the reordered distance matrix showing three distinct dark-coloured square blocks along the diagonal, indicating three natural groups. Top-right (B, VAT Random Data): a similar heatmap for uniformly random data showing no block structure, with distances distributed evenly as a smooth gradient. Bottom-left (C, PCA Player Data): a two-dimensional scatter plot of the first two principal components with points coloured by cluster assignment, revealing three visually separable groups. Bottom-right (D, PCA Random Data): the same PCA projection for random data showing a single diffuse cloud with no discernible groupings."

# Combine all four plots into a 2x2 grid
grid.arrange(
  dist_plot_real + labs(title = "A: VAT - Player Data"),
  dist_plot_random + labs(title = "B: VAT - Random Data"),
  pca_plot_real + labs(title = "C: PCA - Player Data"),
  pca_plot_random + labs(title = "D: PCA - Random Data"),
  ncol = 2, nrow = 2
)
```

## Appendix 2: Cluster Validation Plots

```{r}
#| label: appendix_cluster_validation_plots
#| fig.width: 6
#| fig.height: 4
#| echo: false
#| warning: false
#| message: false
#| fig.cap: "Cluster validation using the Calinski-Harabasz Index across k = 2 to 10."
#| fig.alt: "Line chart with number of clusters k (2 through 10) on the horizontal axis and the Calinski-Harabasz Index on the vertical axis. The line peaks sharply at k equals 3, then declines steadily as k increases, indicating that three clusters provide the best ratio of between-cluster to within-cluster variance for this dataset."

# Display CH-index plot in appendix
p_pam_ch
```

The plot above shows the validation method used to determine the optimal number of clusters. The Calinski-Harabasz Index evaluates cluster validity based on the ratio of between-cluster to within-cluster variance, with higher values indicating better separation between clusters. The index indicated that k = 3 was the optimal number of clusters for our player data.

## Appendix 3: Robust Mixed-Model Coefficients

```{r}
#| label: appendix_model_coefficients
#| echo: false
#| warning: false
#| message: false

fmt_p_table <- function(p) {
  case_when(
    is.na(p) ~ "NA",
    p < 0.001 ~ "< .001",
    TRUE ~ sprintf("%.3f", p)
  )
}

tidy_rlmer_table <- function(model, outcome, model_name) {
  as.data.frame(coef(summary(model))) %>%
    rownames_to_column("term") %>%
    as_tibble() %>%
    rename(
      estimate = Estimate,
      std_error = `Std. Error`,
      t_value = `t value`
    ) %>%
    mutate(
      outcome = outcome,
      model = model_name,
      p_approx = 2 * pnorm(-abs(t_value))
    )
}

term_map <- c(
  "(Intercept)" = "Intercept",
  "player_typeCasual PvP Players" = "Casual PvP vs Casual PvE",
  "player_typeElite Allround Players" = "Elite Allround vs Casual PvE",
  "hours" = "Hours played",
  "genderMale" = "Gender: Male (vs Female)",
  "genderOther" = "Gender: Other (vs Female)",
  "genderPrefer not to say" = "Gender: Undisclosed (vs Female)",
  "age" = "Age",
  "player_typeCasual PvP Players:hours" = "Casual PvP x Hours",
  "player_typeElite Allround Players:hours" = "Elite Allround x Hours"
)

bind_rows(
  tidy_rlmer_table(rmodel_game, "Wellbeing", "Main effects"),
  tidy_rlmer_table(rmodel_enjoyment, "Enjoyment", "Main effects"),
  tidy_rlmer_table(rmodel_moderation_game, "Wellbeing", "Moderation"),
  tidy_rlmer_table(rmodel_moderation_enjoyment, "Enjoyment", "Moderation")
) %>%
  mutate(term = recode(term, !!!term_map, .default = term)) %>%
  filter(term != "Intercept") %>%
  transmute(
    Model = model,
    Outcome = outcome,
    Term = term,
    b = round(estimate, 3),
    SE = round(std_error, 3),
    t = round(t_value, 2),
    p = fmt_p_table(p_approx)
  ) %>%
  tt(caption = "Fixed-effect coefficients from robust linear mixed-effects models (rlmer) with random intercepts for country. Approximate p-values derived from normal-approximation of t-statistics.") %>%
  format_tt(digits = 3)
```

This table presents the full coefficient estimates for all four robust mixed-effects models. ANOVA omnibus tests, Tukey HSD post-hoc comparisons, simple slopes, and pairwise slope-difference tests are reported in the main text.
