---
title: "Not All Play is Equal: In-game Player Behaviour Predicts Wellbeing Differences"
shorttitle: "In-game player behaviour and wellbeing"
running-head: "Playing Style and Well-being"
author:
  - name: Thomas Hakman
    corresponding: true
    orcid: 0009-0009-8292-2482
    email: thomas.hakman@oii.ox.ac.uk
    affiliation:
      - ref: 1
    roles:
      - conceptualization
      - data curation
      - investigation
      - methodology
      - formal analysis
      - visualization
      - writing
      - editing
  - name: Matti Vuorre
    corresponding: false
    orcid: 0000-0001-5052-066X
    email: mjvuorre@uvt.nl
    affiliations:
      - ref: 2
    roles:
      - data curation
      - methodology
      - formal analysis
      - supervision
      - editing
  - name: Andrew K Przybylski
    corresponding: false
    orcid: 0000-0001-5547-2185
    email: andy.przybylski@oii.ox.ac.uk
    affiliations:
      - ref: 1
    roles:
      - conceptualization
      - funding acquisition
      - project administration
      - supervision
      - editing
affiliations:
  - id: 1
    name: University of Oxford
    department: Oxford Internet Institute
  - id: 2
    name: Tilburg University
    department: Department of Social Psychology
abstract: |
  Large-scale studies using digital trace data consistently find that playtime alone is not meaningfully associated with wellbeing, suggesting that the quality of play may matter more than the quantity. Yet these studies typically capture playtime across platforms and titles without examining what happens inside the game, the granular in-game behaviours that might differentiate players' experiences. The Integrated Model of Player Experience (IMP) posits that the mechanics and contexts players encounter during play shape the resulting experience and its downstream psychological effects, implying that players who engage with the same game in qualitatively different ways may report different wellbeing outcomes. We tested this using telemetry-driven profiling of session-level behavioural data from 413 players of the online shooter Plants vs. Zombies: Battle for Neighborville. Cluster analysis on 10 standardised performance and mode-preference metrics identified three distinct playstyles characterised by cooperative, high-performing generalist, and competitive playstyles. Players clustered as cooperative-focused reported greater wellbeing and enjoyment than competition-focused players. However, we found no evidence that playstyle moderated the relationship between playtime and either outcome. These findings indicate that how players engage with a game's mechanics and modes is associated with different average levels of wellbeing, but does not appear to alter the relationship between the amount of play and wellbeing. We discuss implications for moving beyond playtime-centric approaches toward in-game behavioural methods for understanding games and wellbeing.
categories: [video games, playing style, well-being, performance, telemetry]
authornote: |
  Correspondence concerning this article should be addressed to Thomas Hakman, Oxford Internet Institute, University of Oxford. Email: thomas.hakman@oii.ox.ac.uk
date: last-modified
bibliography: bibliography.bib
csl: apa.csl
prefer-html: true
execute:
  echo: false
  warning: false
  message: false
  cache: false
  freeze: false
params:
  refit_models: true
  rerun_clustering: true
format:
  html:
    toc: true
    toc-location: left
    theme: cosmo
    code-fold: true
    html-math-method: katex
    embed-resources: true
  preprint-typst:
    keep-typ: true
  preprint-docx: default
knitr:
  opts_chunk:
    dev: png
    fig.format: png
    dpi: 300
---

```{r}
#| label: setup
#| include: false
knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE,
  fig.align = "center"
)

# Toggle model refitting via Quarto param
eval_models <- isTRUE(params$refit_models)
eval_clustering <- isTRUE(params$rerun_clustering)

# Ensure output directory exists for cached models
if (!dir.exists("output/models")) {
  dir.create("output/models", recursive = TRUE)
}
```

# Introduction

The debate about whether video games harm or improve wellbeing has persisted for decades, yet empirical findings remain mixed, constantly changing with the methods used to study them. Earlier studies relying on players' retrospective estimates of their own playtime and behaviour produced contradictory findings. Several studies reported negative associations between self-reported playtime and measures of mental health or wellbeing [@fergusonLackConsensusScholars2019; @liuDigitalCommunicationMedia2019; @mathurFindingCommonGround2019], with some reporting negative associations only among highly engaged players [@coldercarrasVideoGamingHyperconnected2017; @przybylski2017a]. Whilst others found positive associations [@halbrookWhenHowVideo2019; @villaniVideogamesEmotionRegulation2018], with some reporting those only for moderate usage [@przybylski2017a] or specific game genres such as MMOs [@raithMassivelyMultiplayerOnline2021]. To overcome the well-documented inaccuracy of self-reported measures [@johannesVideoGamePlay2021; @parrySystematicReviewMetaanalysis2021a], later work utilised large-scale digital trace data (i.e., objective behavioural logs collected directly from game platforms). These studies demonstrated that, for most players, the overall associations between the average player's quantity of playtime and their wellbeing are small and unlikely to be practically meaningful [@ballouOpenPlayLongitudinal2025; @egamiCausalEffectVideo2024; @johannesVideoGamePlay2021; @larrieuHowGamingMotives2023; @vuorreTimeSpentPlaying2022].

These null results do not imply that playing video games is irrelevant to wellbeing. Rather, they suggest that the most measured dimension of play, its quantity, might not be the determining factor. Large-scale analysis of adult Nintendo players corroborated this, whilst they found that playtime was not meaningfully related to multiple wellbeing indicators, it did show that players' subjective evaluations of gaming's life-fit were [@ballouPerceivedValueVideo2025]. The underlying problem is that, conceptually, there is no such thing as 'video game play', but only exposures such as 'playing with friends', 'playing competitively', 'succeeding or failing at the game's mechanics', or 'playing to alleviate study-related stress'. That is, playtime consists of heterogeneous exposures, including context-centered exposures (e.g., when and with whom one plays), game-centered exposures (e.g., game modes, mechanics, and business models), and player-centered exposures (e.g., playstyles, motivations, and player experience) [@ballouHowVideoGames2025]. When studies collapse these layers into a single playtime variable, they conflate qualitatively distinct exposures and risk both theoretical ambiguity and empirical inconsistency [@ballouHowVideoGames2025].

This highlights a crucial gap: while existing digital trace research is a mile wide, it is only an inch deep. They capture playtime across platforms and titles, but not what happens inside the game, the granular in-game behaviours and experiences that might explain why some players benefit from play while others do not. Thus, to reconcile the mixed historical findings with the contemporary null results, research needs to start modelling the quality of play with in-game digital trace data because quantity alone is insufficient; two players can accumulate the same playtime yet have markedly different player experiences. While this makes intuitive sense, how researchers can study in-game player experience empirically without reverting to subjective self-reports remains an open question.

One theoretically grounded operationalisation comes from the Integrated Model of Player experience [IMP; @elsonMoreStoriesButtons2014], which identifies three categories of game characteristics that shape the playing experience: narrative (story and setting), mechanics (rules, interaction demands, and feedback), and context (the social and situational embedding). Where user-centric theories such as self-determination theory [@ryanIntrinsicMotivationVideo2006] and uses and gratifications [@ruggieroUsesGratificationsTheory2000] focus primarily on motivations and player states, the IMP is game-centric and focuses on how interaction with specific in-game characteristics during play alters the play experience and subsequent wellbeing outcomes [@elsonMoreStoriesButtons2014]. In this sense, the IMP opens the "black box" of in-game behaviour by connecting the game characteristics players encounter during play to the outcomes that follow from it, thereby guiding empirically testable hypotheses about these relationships [@elsonMoreStoriesButtons2014]. Crucially, the IMP positions mechanics as the foundational element of player experience; without mechanics there is no play, whereas narrative and context contribute situationally. A conceptually meaningful operationalisation of play quality should therefore incorporate digital trace data on mechanics-relevant indicators (e.g., performance metrics), contextual indicators (e.g., mode selection), and narrative indicators (e.g., dialogue engagement) where appropriate.

Within game analytics research, this process is known as telemetry-driven profiling, where collections of in-game behaviours, events, and experiences are condensed into behavioural profiles, or playstyles, that describe how players engaged with a game during the measurement period [@drachen2014a; @ravariInvestigatingRelationPlaying2022]. Playstyles are distinct from player types. Player types refer to stable individual characteristics assessed via self-reports, as in Bartle's taxonomy [@bartleHeartsClubsDiamonds1996], which categorizes players as Achievers, Explorers, Socializers, or Killers based on their preferences for achievement versus exploration and interaction with the game world versus other players; Yee's motivation model [@yeeMotivationsPlayOnline2006], which identifies three primary components of player motivation (i.e., achievement, social, and immersion); and the Hexad framework [@tondelloGamificationUserTypes2016], which proposes six gamification user types grounded in self-determination theory. The utility of such typologies has, however, been called into question by recent evidence of their temporal instability. @santosConsistencyGamificationUser2023 found that 72% of participants changed their player type classification over six months, suggesting that self-reported types may not reliably reflect actual in-game behaviour.

In contrast, telemetry-derived playstyles capture emergent behavioural patterns as they occur, providing a more ecologically valid representation of how players engage with games. As such, playstyles offer a quantifiable, data-driven way to characterise how players pursue a game's objectives [@sifaProfilingGamesUnderstanding2018]. Previous work has used telemetry-driven profiling to examine the relationship between playstyles and personality [@beanVideoGamersPersonality2016; @vanlankveldGamesPersonalityProfiling2011], motivation [@melhartYourGameplaySays2019; @yeeGamerMotivationProfile2016], national culture [@ravariInvestigatingRelationPlaying2022], and individual and team performance [@jiangWideDeepMaverick2021]. However, whether and how playstyles relate to wellbeing and enjoyment has remained untested, in large part because datasets combining granular session-level behavioural telemetry with validated measures of mental health are exceedingly rare [@norwoodBuildingCollaborativeDigital2026].

The present study addresses this question by leveraging one such rare dataset, collected through an industry-academia collaboration [@johannesVideoGamePlay2021]. We address three research questions. First, can distinct playstyles be identified from in-game behavioural telemetry within a single game title? Second, do players with different playstyles report different levels of game-specific well-being or enjoyment? Third, does the association between playtime and well-being or enjoyment differ between playstyles?





```{r}
#| label: install-packages
#| include: false
#| cache: false

# First ensure pacman is installed if not already
if (!requireNamespace("pacman", quietly = TRUE)) {
  install.packages("pacman")
}

# Use pacman to load all required packages (will install any missing packages)
pacman::p_load(
  here,         # For project-relative paths
  lubridate,    # For date/time manipulation
  scales,       # For scale transformations
  janitor,      # For data cleaning
  magrittr,     # For additional pipe operators
  psych,        # For psychological statistics
  corrplot,     # For correlation plots
  ggpubr,       # For publication-ready plots
  factoextra,   # For cluster analysis visualization
  cluster,      # For clustering algorithms
  fmsb,         # For radar charts
  knitr,        # For dynamic report generation
  fpc,          # For cluster validation
  gridExtra,    # For arranging multiple plots
  hopkins,      # For cluster tendency assessment
  effectsize,   # For effect size calculations
  robustlmm,    # For robust linear mixed-effects models
  broom,        # For tidy model outputs
  broom.mixed,  # For tidying mixed-effects model output
  patchwork,    # For combining plots
  ggeffects,    # For visualizing regression model predictions
  grid,         # For grid graphics (needed for tableGrob)
  tinytable,    # For publication-quality tables
  modelsummary, # For regression model tables
  report,       # For automated statistical reporting
  tidyverse     # For data manipulation and visualization (loaded last to avoid masking)
)

# Load helper functions
source(here("helpers.R"))
```

```{r}
#| label: helper-functions
#| code-summary: "Load helper functions"

# Helper functions are now sourced from helpers.R in the install-packages chunk
# This chunk is kept for any analysis-specific functions not in helpers.R

# Function to run ANOVA and extract results
run_anova <- function(dv) {
  formula <- as.formula(paste(dv, "~ player_type"))
  model <- aov(formula, data = pvz_data)
  result <- broom::tidy(model)
  result$eta_squared <- effectsize::eta_squared(model)[["Eta2"]]
  result$variable <- dv
  return(result)
}

# Mean helper for mode-specific telemetry metrics:
# treat missing values as 0 and always return a finite scalar.
mode_mean <- function(x) {
  mean(coalesce(x, 0), na.rm = TRUE)
}

# Safe lookup helper for inline reporting from summary tables.
type_stat <- function(tbl, type_label, column_name, digits = 2) {
  value <- tbl %>%
    filter(.data$player_type == type_label) %>%
    pull(.data[[column_name]]) %>%
    first(default = NA_real_)

  if (is.na(value) || is.nan(value) || is.infinite(value)) {
    value <- 0
  }

  sprintf(paste0("%.", digits, "f"), value)
}
```

```{r}
#| label: load-processed-data

# Load the dataset with outliers removed
pvz_data <- read_csv(here("data/pvz_clean_final.csv"))

# Create a copy of pvz_data for analysis to maintain consistency
pvz_df_analysis <- pvz_data

```

```{r}
#| label: select-columns

pvz_df_analysis <- pvz_df_analysis %>%
  select(
    # Demographic variables
    player_id, country, age, gender, played_with_others,
    
    # Well-being variables - focus only on our two key outcomes
    spane_game_balance, enjoyment,
    
    # Total playtime
    hours,
    
    # PVP metrics
    ratio_pvp_hours,
    pvp_kd,
    pvp_hit_accuracy,
    pvp_dps,
    pvp_average_damage,
    pvp_average_score,
    
    # PVE metrics
    ratio_pve_hours,
    pve_kd,
    pve_hit_accuracy,
    pve_dps,
    pve_average_damage,
    pve_average_score
  )
```

```{r}
#| label: telemetry-data-filter
#| include: false

# Keep players who engaged in at least one combat mode (PVP or PVE)
# All 10 clustering features are checked: hours > 0 AND non-missing performance
# metrics for at least one mode. NAs in the other mode are set to 0 before clustering (see replace_na in cluster_feature_selection chunk).
combat_metrics <- c("kd", "hit_accuracy", "dps", "average_damage", "average_score")
pvz_df_analysis <- pvz_df_analysis %>%
  filter(
    (ratio_pvp_hours > 0 & if_all(paste0("pvp_", combat_metrics), ~ !is.na(.x))) |
    (ratio_pve_hours > 0 & if_all(paste0("pve_", combat_metrics), ~ !is.na(.x)))
  )

# Show the number of players left
print(paste("Number of players left after filtering:", nrow(pvz_df_analysis)))
```

# Methods

In this study, we analysed data reported in Johannes et al. (2021), who collected digital trace data and survey responses of 518 Plants versus Zombies: Battle for Neighbourville players.

## Plants versus Zombies: Battle for Neighbourville

```{r}
#| label: survey-demographics-calculation
#| include: false
#| cache: false

# Calculate demographics for the filtered dataset
demographics <- pvz_df_analysis %>%
  summarize(
    total_participants = n(),
    avg_age = round(mean(age, na.rm = TRUE), 1),
    sd_age = round(sd(age, na.rm = TRUE), 1)
  )

# Normalize gender labels; treat missing/unexpected labels as undisclosed
gender_distribution <- pvz_df_analysis %>%
  mutate(
    gender_clean = case_when(
      str_trim(as.character(gender)) == "Male" ~ "Male",
      str_trim(as.character(gender)) == "Female" ~ "Female",
      str_trim(as.character(gender)) == "Other" ~ "Other",
      str_trim(as.character(gender)) == "Prefer not to say" ~ "Prefer not to say",
      TRUE ~ "Prefer not to say"
    )
  ) %>%
  count(gender_clean, name = "n") %>%
  mutate(percentage = n / sum(n) * 100)

count_or_zero <- function(df, label) {
  value <- df %>% filter(gender_clean == label) %>% pull(n)
  if (length(value) == 0) 0L else as.integer(value)
}

# Count participants with each gender category for manuscript text
male_count <- count_or_zero(gender_distribution, "Male")
female_count <- count_or_zero(gender_distribution, "Female")
other_count <- count_or_zero(gender_distribution, "Other")
undisclosed_count <- count_or_zero(gender_distribution, "Prefer not to say")
```

```{r}
#| label: final-sample-demographics
#| include: false
#| cache: false

# Display demographics of final sample
cat("Demographics of final sample:
")
cat("- Sample size:", nrow(pvz_df_analysis), "
")
cat("- Average age:", round(mean(pvz_df_analysis$age, na.rm = TRUE), 1), "(SD =", round(sd(pvz_df_analysis$age, na.rm = TRUE), 1), ")
")

# Gender distribution
gender_dist <- pvz_df_analysis %>%
  mutate(
    gender = case_when(
      str_trim(as.character(gender)) == "Male" ~ "Male",
      str_trim(as.character(gender)) == "Female" ~ "Female",
      str_trim(as.character(gender)) == "Other" ~ "Other",
      str_trim(as.character(gender)) == "Prefer not to say" ~ "Prefer not to say",
      TRUE ~ "Prefer not to say"
    )
  ) %>%
  count(gender) %>%
  mutate(percentage = n / sum(n) * 100)
```

Plants versus Zombies: Battle for Neighbourville (PvZ) is an online multiplayer First-person shooter (FPS) game released by Electronic Arts (EA) in October 2019. The gameplay is third-person, with an emphasis on shooter-style combat and exploration. Players can play 23 customizable gameplay classes that each have unique weapons, abilities, and forms of movement. The game offers the players the ability to compete against each other in various competitive Player-versus-Player (PVP) modes, including objective-based matches, team deathmatches, 2v2 arena matches, etc. The game also features several Player-versus-Environment (PVE) open zones, which feature storylines with quests and allow players to explore, find collectables, and complete side-missions. Players can earn experience points and in-game currencies, using the former to indicate progression and the latter to upgrade and buy new characters. The game was played on PC, PS4, Xbox One, and Nintendo Switch (@electronicArtsPlantsVsZombiesBattle2019).

## Participants

As detailed in @johannesVideoGamePlay2021, participants were recruited by Electronic Arts directly from their active player base via email invitations sent in two waves: the first in early August 2020 to 50,000 players, and the second in late September 2020 to 200,000 players. In total, 518 PvZ players completed the well-being survey (approximately 0.21% response rate). For details on the full dataset, which also includes Animal Crossing: New Horizons players, see @johannesVideoGamePlay2021.
Because our research questions concern mechanics-based playstyles, we applied several inclusion criteria to ensure the analytic sample consisted of players who actively engaged with the game's core mechanics. We first excluded participants without matching telemetry data (n = 48). We then excluded 57 participants whose telemetry contained no performance metrics across any activity (PvP or PvE). These players spent time in the game's social hub, an idle space for navigation and character customisation, but did not participate in any gameplay, and thus had no behavioural data from which to derive playstyle profiles. Social-hub-only players differed markedly from combat-active players in engagement: they logged substantially less playtime (Mdn = 0.12 vs. 4.82 hours; Mann–Whitney U = 1,156, p < .001) and fewer sessions (Mdn = 1 vs. 28; U = 589, p < .001), though survey response quality did not differ between groups (straightlining rates: 1.8% vs. 1.5%, χ² = 0, p = 1.0). The final analytic sample consisted of `r nrow(pvz_df_analysis)` combat-active players (M age = `r demographics$avg_age`, SD = `r demographics$sd_age`; `r male_count` male, `r female_count` female, `r other_count` other, `r undisclosed_count` undisclosed), representing `r sprintf('%.1f', (nrow(pvz_df_analysis) / 518) * 100)`% of the original 518 respondents.
Following the methodology from the original paper, values more than 6 standard deviations from a variable mean were replaced with NA on a case-by-case basis (@johannesVideoGamePlay2021). In this dataset, this procedure replaced 87 values (across 48 variables; 26 participants affected) and did not reduce the final analytic sample.
The final analytic sample consisted of `r nrow(pvz_df_analysis)` combat-active players (M age = `r demographics$avg_age`, SD = `r demographics$sd_age`; `r male_count` male, `r female_count` female, `r other_count` other, `r undisclosed_count` undisclosed), representing `r sprintf('%.1f', (nrow(pvz_df_analysis) / 518) * 100)`% of the original 518 respondents.


## Measures

### Playtime

The first measure afforded by these digital trace data is playtime, yet determining this is already complex. Simply measuring the total time an account has been online does not distinguish between active and idle time. Idle time is when the player’s account is online but not actively engaging with the objectives or mechanics of the game. Idle time can happen when a player’s computer is logged onto their account, but they are away from the keyboard. Other forms of idle time are spending time in home screens, customization menus, lobbies, and social spaces.
In this study, active playtime was calculated by aggregating session data. Sessions represent the different levels, game modes, or worlds a player can be present in. Every player is always in a session and could only be in one session at a time, however, multiple players could be in the same session. For example, when a player launches PvZ, they enter the hub world (where they can select what type of level or mode to play), entering the hub world starts the session timer, which ends as they choose a level or leave to join a match, which then starts a new session timer for that level.
In total, the `r nrow(pvz_df_analysis)` players in our sample played 36,621 sessions, of which 25,937 happened in the two weeks prior to a player taking the survey. These sessions took place across 19 different maps that each facilitated different forms of gameplay, such as Player-versus-Player (PvP; 13,594 (51.27%) of sessions), 4,854 (18.31%) Player-versus-Environment (PVE), and 7,489 (28.24%) Social Hub maps. In PvP maps players battle against each other in various competitive modes such as objective-based matches, team deathmatches, and 2v2 arena matches. In PVE maps players explore open worlds that are designed for adventure with story-driven quests, side missions, and hidden collectables. The Social Hub served as an idle space for navigation and customization between matches and was excluded. This left 17,505 total sessions, 12,445 PVP sessions (71.09%) and 4,509 PVE sessions (25.76%). Players had clear preferences for which game mode they chose to play (@fig-session-timeline), which already indicated clear differences in playing styles.

```{r}
#| label: session-timeline-helper
#| include: false

# Function to plot session timeline for a player
plot_player_sessions <- function(player_data, player_id = NULL, player_label = "Player") {
  # If a player_id is provided, filter data for that player
  if (!is.null(player_id)) {
    player_data <- player_data %>%
      filter(player_id == !!player_id)
  }
  
  # Convert timestamps to POSIXct if they're not already
  if (!inherits(player_data$game_start_time, "POSIXct")) {
    player_data$game_start_time <- as.POSIXct(player_data$game_start_time, format = "%Y-%m-%d %H:%M:%S")
    player_data$game_end_time <- as.POSIXct(player_data$game_end_time, format = "%Y-%m-%d %H:%M:%S")
  }
  
  # Extract day of week, date, and time component
  player_data$day_of_week <- weekdays(player_data$game_start_time)
  player_data$date <- as.Date(player_data$game_start_time)
  player_data$start_time_only <- as.POSIXct(format(player_data$game_start_time, format = "%H:%M:%S"), format = "%H:%M:%S")
  player_data$end_time_only <- as.POSIXct(format(player_data$game_end_time, format = "%H:%M:%S"), format = "%H:%M:%S")
  
  # Identify sessions that cross midnight (where end_time is earlier than start_time)
  player_data$crosses_midnight <- player_data$end_time_only < player_data$start_time_only
  
  # Create a new dataframe that will include split sessions for those crossing midnight
  player_data_expanded <- player_data %>%
    filter(!crosses_midnight) # Keep all sessions that don't cross midnight
  
  # Handle sessions that cross midnight by splitting them
  midnight_sessions <- player_data %>%
    filter(crosses_midnight)
  
  if (nrow(midnight_sessions) > 0) {
    # For each session crossing midnight, create two rows:
    # 1. From start time to 23:59:59
    # 2. From 00:00:00 to end time
    midnight <- as.POSIXct(strptime("23:59:59", format = "%H:%M:%S"), format = "%H:%M:%S")
    midnight_start <- as.POSIXct(strptime("00:00:00", format = "%H:%M:%S"), format = "%H:%M:%S")
    
    # First part: from start time to midnight
    midnight_part1 <- midnight_sessions
    midnight_part1$end_time_only <- midnight
    
    # Second part: from midnight start to end time
    midnight_part2 <- midnight_sessions
    midnight_part2$start_time_only <- midnight_start
    # The second part is on the next day of the week and the next date
    midnight_part2$day_of_week <- factor(
      weekdays(midnight_part2$game_start_time + 86400), # Add one day (86400 seconds)
      levels = c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday")
    )
    midnight_part2$date <- midnight_part2$date + 1
    
    # Add both parts to the expanded dataset
    player_data_expanded <- bind_rows(player_data_expanded, midnight_part1, midnight_part2)
  }
  
  # Recode match_type based on game_level_key prefix
  player_data_expanded <- player_data_expanded %>%
    mutate(
      match_type = case_when(
        str_starts(game_level_key, "dsub_social")  ~ "social match",
        str_starts(game_level_key, "dsub_region")  ~ "pve match",
        str_starts(game_level_key, "dsub_rush")    ~ "pvp match (rush)",
        str_starts(game_level_key, "dsub_tvm")     ~ "pvp match (vanquish)",
        str_starts(game_level_key, "dsub_4v4")     ~ "pvp match (arena)",
        game_level_key == "dsub_podiumcelebration"  ~ "podium celebration",
        .default = "0"
      )
    )
  
  # Ensure day of week is a factor with ordered levels
  player_data_expanded$day_of_week <- factor(
    player_data_expanded$day_of_week,
    levels = c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday")
  )
  
  # Create a position adjustment for stacking sessions by different dates for the same day of week
  player_data_expanded <- player_data_expanded %>%
    group_by(day_of_week) %>%
    mutate(
      # Get unique dates for this day of week and assign a position index to each date
      unique_dates = list(unique(date)),
      date_index = match(date, unique(date)),
      # Scale the vertical offset to be small but noticeable
      y_offset = (date_index - 1) * 0.2,  # Creates a small vertical offset for different dates
      # Create adjusted y-position that includes the offset
      y_position = as.numeric(day_of_week) + y_offset
    ) %>%
    ungroup()
  
  # Sort by day of week, date, and start time
  player_data_expanded <- player_data_expanded %>%
    arrange(day_of_week, date, game_start_time, game_end_time)
  
  # Add a session ID within each day to distinguish multiple sessions on same day
  player_data_expanded <- player_data_expanded %>%
    group_by(day_of_week, date) %>%
    mutate(session_id_within_day = row_number()) %>%
    ungroup()
  
  # Convert match_type to factor with proper levels for consistent ordering
  player_data_expanded$match_type <- factor(
    player_data_expanded$match_type,
    levels = c("social match", "pve match", "pvp match (rush)", "pvp match (vanquish)", "pvp match (arena)", "podium celebration", "0"),
    labels = c("Social Lobby", "PvE Match", "PvP (Rush)", "PvP (Vanquish)", "PvP (Arena)", "Podium Celebration", "Other")
  )
  
  # Create the plot with enhanced styling
  plot <- player_data_expanded %>%
    ggplot(aes(start_time_only, y_position, col = match_type, 
               group = interaction(day_of_week, date, session_id_within_day))) +
    # Use improved color palette with better contrast
    scale_color_manual(
      name = "Match Type", 
      values = c(
        "Social Lobby" = "#00BFC4",       # Teal
        "PvE Match" = "#7CAE00",          # Green
        "PvP (Rush)" = "#F8766D",         # Red
        "PvP (Vanquish)" = "#FF9E1C",     # Orange
        "PvP (Arena)" = "#FFD700",        # Gold
        "Podium Celebration" = "#C77CFF", # Purple
        "Other" = "#A0A0A0"               # Gray
      )
    ) +
    # Thicker segments with rounded ends for better visibility
    geom_segment(
      aes(xend = end_time_only, yend = y_position),
      linewidth = 2,
      lineend = "round"
    ) +
    # Add points at start/end for better visibility of short sessions
    geom_point(size = 1.5, alpha = 0.6) +
    geom_point(aes(x = end_time_only), size = 1.5, alpha = 0.6) +
    # Better time axis formatting - using fixed limits for consistency
    scale_x_datetime(
      labels = scales::date_format("%H:%M"),
      breaks = scales::date_breaks("2 hours"),
      limits = c(
        as.POSIXct(strptime("00:00", format = "%H:%M")),
        as.POSIXct(strptime("23:59", format = "%H:%M"))
      ),
      expand = c(0.01, 0.01),
      date_minor_breaks = "1 hour"
    ) +
    # Use the original day of week as the y-axis labels
    scale_y_continuous(
      breaks = 1:7,
      labels = levels(player_data_expanded$day_of_week),
      limits = c(0.5, 7.5)
    ) +
    # Empty labs - axis labels will be added to the combined figure instead
    labs(
      x = NULL, 
      y = NULL
    ) +
    # Improved theme with better typography and layout
    theme_minimal(base_size = 11, base_family = "sans") +
    theme(
      legend.position = "none", # Move legend to the combined figure
      plot.margin = margin(30, 10, 30, 10), # Increased margins for better separation
      
      axis.text.y = element_text(face = "bold"),
      axis.title = element_text(face = "bold"),
      
      panel.grid.major.x = element_line(color = "grey90"),
      panel.grid.minor.x = element_line(color = "grey95"),
      panel.grid.major.y = element_line(color = "grey90"),
      panel.grid.minor.y = element_blank(),
      
      panel.background = element_rect(fill = "white", color = NA),
      plot.background = element_rect(fill = "white", color = NA)
    )
  
  return(plot)
}
```

```{r}
#| label: session-timeline-data-prep
#| include: false
#| cache: false

# Load game session data if not already included
game_sessions <- read_csv(here("data/processed/game_time_2week.csv"))

# Make sure timestamps are in the correct format
game_sessions$game_start_time <- as.POSIXct(game_sessions$game_start_time, format = "%Y-%m-%d %H:%M:%S")
game_sessions$game_end_time <- as.POSIXct(game_sessions$game_end_time, format = "%Y-%m-%d %H:%M:%S")

# First, get a list of player IDs in the data
player_ids <- unique(game_sessions$player_id)

# Choose specific indices
# If those indices don't exist in our data, we'll use the first three players
player_indices <- c(
  min(401, length(player_ids)),  # Equivalent to slice(401) 
  min(6, length(player_ids)),    # Equivalent to slice(6)
  min(3, length(player_ids))     # Equivalent to slice(3)
)

selected_player_ids <- player_ids[player_indices]
```

```{r}
#| label: create-individual-plots
#| include: false
#| cache: false

# Create individual plots for each player with standardized x-axis
plot1 <- plot_player_sessions(
  game_sessions %>% filter(player_id == selected_player_ids[1])
)

plot2 <- plot_player_sessions(
  game_sessions %>% filter(player_id == selected_player_ids[2])
)

# For the third plot, use the same time range for consistency
plot3 <- plot_player_sessions(
  game_sessions %>% filter(player_id == selected_player_ids[3])
)
```

```{r}
#| label: fig-session-timeline
#| fig-cap: "Session timelines for three example players showing gaming activity across the two-week study period. Each horizontal bar represents a gaming session, with colors indicating match type (Social Lobby, PvE Match, PvP Rush/Vanquish/Arena). Each player's panel contains two rows corresponding to the two weeks of telemetry data, collapsed onto a single weekly timeline to facilitate comparison of play patterns across weeks. The three players illustrate the diversity of playing patterns observed in the data."
#| fig-width: 12
#| fig-height: 16
#| code-summary: "Session timeline visualization"

# Combine plots into a single figure with enhanced styling and proper spacing
figure <- ggarrange(
  plot2, plot3, plot1,
  labels = c("Player A", "Player B", "Player C"),
  font.label = list(size = 14, face = "bold", color = "black"),
  ncol = 1, 
  nrow = 3, 
  common.legend = TRUE, 
  legend = "bottom",
  heights = c(1, 1, 1),
  align = "v",
  hjust = -0.1
)

# Add shared x and y labels to the combined figure
fig_sessiontime <- annotate_figure(figure,
                        bottom = text_grob("Time of Day (hours)", 
                                          face = "bold", size = 12),
                        left = text_grob("Day of Week", 
                                         face = "bold", size = 12, rot = 90))

# Display the combined figure
fig_sessiontime
```

### Behavioural telemetry

The raw telemetry contained session-based records of game events for each player. From these records, we selected and computed metrics only if they had been previously validated as informative measures of player behaviour and performance in shooter games [@drachenGunsSwordsData2012; @lugrinImmersiveFPSGames2013; @melhartYourGameplaySays2019]. Two principles guided the computation of the behavioural metrics. First, all performance metrics were constructed to be independent of total playtime. Common cumulative indicators such as ‘player level’ or ‘total kills’ were deliberately excluded because they conflate skill with playtime; a player's level may simply be higher because they played more, not because they performed better. Second, every metric was computed separately for PvP and PvE modes, because the two contexts differ fundamentally in difficulty, objectives, and opponent type; eliminating a computer-controlled enemy in PvE is not comparable to eliminating a matchmade opponent of similar rank in PvP. Pooling across modes would obscure meaningful behavioural differences.
This resulted in 12 behavioural metrics plus one overall playtime measure. Mode ratio captured the proportion of active playtime each player devoted to PvP and PvE modes, reflecting game-mode preference. Kill/death ratio indexed combat effectiveness as the number of eliminations per death, with higher values indicating greater survivability and lethality. Hit accuracy measures the proportion of shots that connect with a target, reflecting aiming precision. Damage per second captured sustained offensive output, indicating how efficiently a player dealt damage over time. Average damage per game reflected total offensive contribution within a match, encompassing both direct combat and area-of-effect output. Average score per game provided a composite performance indicator incorporating objectives completed, assists, and other in-game scoring events beyond raw combat. Each of these five performance metrics was computed separately for PvP and PvE, yielding 10 mode-specific indicators alongside the two mode-ratio variables. See @tbl-descriptive-stats for descriptive statistics.

```{r}
#| label: tbl-descriptive-stats
#| tbl-cap: "Descriptive Statistics for Key Variables"
#| fig.width: 10
#| fig.height: 6
#| tbl-cap-location: top
#| echo: false
#| warning: false
#| message: false
#| cache: false

# Create a simple APA-style descriptive statistics table
# First, check variable types to handle them appropriately
variable_types <- sapply(pvz_df_analysis, class)

# Handle numeric variables
numeric_vars <- pvz_df_analysis %>%
  select(where(is.numeric)) %>%
  select(
    # Demographics
    age,
    
    # Playtime 
    hours, ratio_pvp_hours, ratio_pve_hours,
    
    # Performance metrics
    pvp_kd, pve_kd, 
    pvp_hit_accuracy, pve_hit_accuracy,
    pvp_dps, pve_dps,
    pvp_average_damage, pve_average_damage,
    pvp_average_score, pve_average_score,
    
    # Well-being and enjoyment
    spane_game_balance, enjoyment
  )

# Calculate descriptive statistics for numeric variables
numeric_stats <- numeric_vars %>%
  summarise(across(everything(), 
                  list(
                    n = ~sum(!is.na(.)),
                    M = ~mean(., na.rm = TRUE),
                    SD = ~sd(., na.rm = TRUE),
                    Min = ~min(., na.rm = TRUE),
                    Max = ~max(., na.rm = TRUE)
                  ))) %>%
  pivot_longer(everything(), 
               names_to = c("variable", "stat"), 
               names_pattern = "(.*)_(.*)",
               values_to = "value") %>%
  pivot_wider(names_from = stat, values_from = value)

# For categorical variables like gender, just count frequencies for the footnote
if ("gender" %in% names(pvz_df_analysis)) {
  gender_counts <- pvz_df_analysis %>%
    mutate(
      gender_clean = case_when(
        str_trim(as.character(gender)) == "Male" ~ "Male",
        str_trim(as.character(gender)) == "Female" ~ "Female",
        str_trim(as.character(gender)) == "Other" ~ "Other",
        str_trim(as.character(gender)) == "Prefer not to say" ~ "Prefer not to say",
        TRUE ~ "Prefer not to say"
      )
    ) %>%
    count(gender_clean, name = "n") %>%
    mutate(gender = gender_clean) %>%
    select(gender, n) %>%
    mutate(pct = n / sum(n) * 100) %>%
    mutate(pct = round(pct, 1)) %>%
    mutate(frequency = paste0(n, " [", pct, "]"))
  
  # Note: We'll add a footnote to explain gender distribution separately
  # Using brackets instead of parentheses to avoid LaTeX issues
  gender_note <- paste("Gender distribution:", 
                     paste(paste0(gender_counts$gender, ": ", gender_counts$frequency), 
                           collapse = ", "))
} else {
  gender_note <- NULL
}

# Format the data for the table
descriptive_stats <- numeric_stats %>%
  # Format variable names for better readability
  mutate(variable = case_when(
    variable == "age" ~ "Age",
    variable == "hours" ~ "Total Hours Played",
    variable == "ratio_pvp_hours" ~ "Proportion of PvP Playtime",
    variable == "ratio_pve_hours" ~ "Proportion of PvE Playtime",
    variable == "pvp_kd" ~ "PvP Kill/Death Ratio",
    variable == "pve_kd" ~ "PvE Kill/Death Ratio",
    variable == "pvp_hit_accuracy" ~ "PvP Hit Accuracy",
    variable == "pve_hit_accuracy" ~ "PvE Hit Accuracy",
    variable == "pvp_dps" ~ "PvP Damage per Second",
    variable == "pve_dps" ~ "PvE Damage per Second",
    variable == "pvp_average_damage" ~ "PvP Average Damage per Game",
    variable == "pve_average_damage" ~ "PvE Average Damage per Game",
    variable == "pvp_average_score" ~ "PvP Average Score",
    variable == "pve_average_score" ~ "PvE Average Score",
    variable == "spane_game_balance" ~ "Game-Specific Well-being (SPANE)",
    variable == "enjoyment" ~ "Enjoyment",
    TRUE ~ variable
  )) %>%
  # Create category variable for grouping
  mutate(category = case_when(
    variable == "Age" ~ "Demographics",
    variable %in% c("Total Hours Played", "Proportion of PvP Playtime", "Proportion of PvE Playtime") ~ "Playtime",
    variable %in% c("PvP Kill/Death Ratio", "PvE Kill/Death Ratio", 
                   "PvP Hit Accuracy", "PvE Hit Accuracy",
                   "PvP Damage per Second", "PvE Damage per Second",
                   "PvP Average Damage per Game", "PvE Average Damage per Game",
                   "PvP Average Score", "PvE Average Score") ~ "Performance",
    variable %in% c("Game-Specific Well-being (SPANE)", "Enjoyment") ~ "Well-being",
    TRUE ~ "Other"
  )) %>%
  # Round numeric values
  mutate(
    M = round(M, 2),
    SD = round(SD, 2),
    Min = round(Min, 2),
    Max = round(Max, 2)
)

# Create custom ordering for variables to ensure autonomy, competence, relatedness are listed in order
variable_order <- c(
  # Demographics
  "Age",
  
  # Playtime
  "Total Hours Played", "Proportion of PvP Playtime", "Proportion of PvE Playtime",
  
  # Performance
  "PvP Kill/Death Ratio", "PvE Kill/Death Ratio", 
  "PvP Hit Accuracy", "PvE Hit Accuracy",
  "PvP Damage per Second", "PvE Damage per Second",
  "PvP Average Damage per Game", "PvE Average Damage per Game",
  "PvP Average Score", "PvE Average Score",
  
  # Well-being
  "Game-Specific Well-being (SPANE)", "Enjoyment"
)

# Apply the custom ordering
descriptive_stats <- descriptive_stats %>%
  mutate(variable = factor(variable, levels = variable_order)) %>%
  arrange(variable)

# Get sample size for the note
sample_n <- nrow(pvz_df_analysis)

# Create the table with tinytable
descriptive_stats |>
  select(Variable = variable, M, SD, Min, Max) |>
  tt(
    caption = "Descriptive Statistics for Key Variables",
    notes = paste0("N = ", sample_n, ". ", gender_note)
  ) |>
  format_tt(j = 2:5, digits = 2)
```


### Well-being

We assessed affective well-being with the Scale of Positive and Negative Experiences [SPANE; @dienerNewWellbeingMeasures2010; @dienerAdvancesOpenQuestions2018]. Respondents were prompted to reflect on how playing PvZ had made them feel over the past two weeks and rated how often they experienced each of six positive (positive, good, pleasant, happy, joyful, contented) and six negative (negative, bad, unpleasant, sad, afraid, angry) feelings on a scale from 1 (Very rarely or never) to 7 (Very often or always). An overall well-being balance score was computed by subtracting the mean negative score from the mean positive score, yielding a possible range of −6 to +6, where higher values indicate more positive affect associated with play. All responses were screened for straightlining (e.g. when respondents give identical answers to all items) to identify records with low data quality commonly found in online surveys (@leinerTooFastToo2019), however, no unusual cases were identified.

### Enjoyment

Enjoyment was assessed with four items from the Player Experience and Need Satisfaction scale [PENS; @ryanIntrinsicMotivationVideo2006], validated by @johnsonValidationPlayerExperience2018. The full PENS comprises subscales measuring autonomy, competence, relatedness, enjoyment, and extrinsic motivation; we used only the four-item enjoyment subscale, which captures hedonic appreciation of the play experience. Respondents were asked to reflect on their experience playing PvZ over the past two weeks and rated items such as "I think PvZ was fun to play" on a scale from 1 (Strongly disagree) to 7 (Strongly agree).

## Data analysis

Data were analyzed in R version 4.5.1 [@rCoreTeamRLanguageEnvironment2025]. Key packages included `fpc` for cluster validation indices [@hennig2020a], `factoextra` for multivariate visualization [@kassambara2020a], and `robustlmm` for robust mixed-effects estimation [@kollerRobustlmmRPackage2016]. Following initial data cleaning, telemetry records were matched to the same two-week window referenced by the survey items, and analyses were conducted on combat-active players with sufficient mode-specific telemetry.

To test whether clustering was appropriate before fitting any cluster model, we evaluated cluster tendency using three complementary approaches. First, we computed the Hopkins statistic [@banerjee2004a], which compares distances in observed data to distances expected under spatial randomness; values near 0.5 indicate random structure, whereas larger values indicate clusterable structure. Second, we used Visual Assessment of Cluster Tendency (VAT) [@bezdek2002a], which reorders the dissimilarity matrix so potential clusters appear as darker diagonal blocks. Third, we examined Principal Component Analysis (PCA) projections to assess whether separable groupings were visible in low-dimensional space relative to random reference data. Together, these diagnostics indicated non-random structure suitable for clustering.

We then identified play styles with Partitioning Around Medoids (PAM) [@kaufmanFindingGroupsData1990] on standardized in-game behavioral features. PAM was selected over k-means because it is less sensitive to extreme points and defines each cluster by an observed medoid (a real player profile), which improves robustness and interpretability for telemetry-derived behavioral profiles. Features were z-standardized before clustering, and missing mode-specific performance values were set to zero to encode no activity in that mode rather than dropping cases. The number of clusters was determined by maximizing the Calinski-Harabasz index across k = 2 to 10 [@tibshiraniEstimatingNumberClusters2001], which selected a three-cluster solution; we then checked internal coherence with silhouette widths [@kaufmanFindingGroupsData1990].

To examine associations between play style, playtime, and outcomes, we fit robust linear mixed-effects models with random intercepts for country to account for baseline between-country differences in outcome levels. For both outcomes (game-specific well-being and enjoyment), we first estimated main-effects models: outcome ~ player type + hours played + gender + age + (1 | country). To address moderation (RQ3), we then estimated interaction models: outcome ~ player type * hours played + (1 | country). Outcomes were treated as continuous and approximately normally distributed, and playtime hours was modeled on its raw scale. We used simple-slopes analyses to estimate within-play-style playtime slopes and pairwise slope-difference tests to compare those slopes across play styles. Statistical significance was set at p \< .05.

## Data, Materials, and Code

The raw data, materials, codebook and further details on this dataset shared by @johannesVideoGamePlay2021 can be found at <https://osf.io/cjd6z/>. Building on this raw data and documentation, all of the underlying wrangling and analysis code reported in this study can be found at OSF upon publication.

# Results

## Player Clustering Analysis

```{r}
#| label: cluster_feature_selection
#| warning: false
#| message: false

# Select relevant features for clustering
player_cluster_features <- pvz_df_analysis %>% 
  select(
    # Playstyle variables (game mode preferences)
    ratio_pvp_hours, ratio_pve_hours,
    
    # PvP performance metrics
    pvp_kd, pvp_hit_accuracy, pvp_dps, pvp_average_damage, pvp_average_score,
    
    # PvE performance metrics
    pve_kd, pve_hit_accuracy, pve_dps, pve_average_damage, pve_average_score
  ) %>%
  # Replace NAs with 0s - makes sense because missing performance = no activity in that mode
  replace_na(list(
    ratio_pvp_hours = 0, ratio_pve_hours = 0,
    pvp_kd = 0, pvp_hit_accuracy = 0, pvp_dps = 0, 
    pvp_average_damage = 0, pvp_average_score = 0,
    pve_kd = 0, pve_hit_accuracy = 0, pve_dps = 0,
    pve_average_damage = 0, pve_average_score = 0
  ))

# Standardize the data (important for clustering)
player_data_scaled <- scale(player_cluster_features)
rownames(player_data_scaled) <- pvz_df_analysis$player_id[rownames(player_data_scaled) %in% rownames(pvz_df_analysis)]
```

### Assessing Cluster Tendency

```{r}
#| label: cluster_random_data
#| warning: false
#| message: false
#| results: "hide"

# Create a random dataset with the same dimensions for comparison
set.seed(123) # For reproducibility
random_cluster_data <- apply(player_data_scaled, 2,
                     function(x){runif(length(x), min(x), max(x))})
random_cluster_data <- as.data.frame(random_cluster_data)
```

```{r}
#| label: hopkins_statistic
#| warning: false
#| message: false
#| results: "hide"

# Create meaningful features for clustering analysis
# Focus on performance metrics that differentiate playstyles
meaningful_features <- pvz_df_analysis |>
  mutate(
    # Only include players with some gameplay data
    has_pvp = !is.na(pvp_kd) & ratio_pvp_hours > 0,
    has_pve = !is.na(pve_kd) & ratio_pve_hours > 0
  ) |>
  filter(has_pvp | has_pve) |>  # At least one mode
  select(
    ratio_pvp_hours, 
    pvp_kd, 
    pvp_hit_accuracy,
    pve_kd, 
    pve_hit_accuracy
  ) |>
  # Replace NAs with 0 for missing performance metrics
  # (missing performance = no activity in that mode)
  replace_na(list(
    ratio_pvp_hours = 0,
    pvp_kd = 0, 
    pvp_hit_accuracy = 0,
    pve_kd = 0, 
    pve_hit_accuracy = 0
  ))

# Standardize the features
meaningful_scaled <- scale(meaningful_features)

# Calculate Hopkins statistic using manual implementation 
# (more reliable than hopkins package for constrained gaming data)
set.seed(123)
hopkins_results <- replicate(10, {
  manual_hopkins(meaningful_scaled, m = 30)
}, simplify = TRUE)

# Calculate final statistics for manuscript reporting
hopkins_mean <- mean(hopkins_results)
hopkins_sd <- sd(hopkins_results)
hopkins_ci_lower <- hopkins_mean - 1.96 * hopkins_sd
hopkins_ci_upper <- hopkins_mean + 1.96 * hopkins_sd

# Display results
cat("Hopkins Statistic Results:\n")
cat("Sample size:", nrow(meaningful_features), "observations\n")
cat("Mean Hopkins statistic (m=30):", round(hopkins_mean, 4), "\n")
cat("95% CI: [", round(hopkins_ci_lower, 4), ",", round(hopkins_ci_upper, 4), "]\n")

# Interpretation
if(hopkins_mean > 0.75) {
  cat("Interpretation: Strong evidence of clustering tendency (H > 0.75)\n")
} else if(hopkins_mean > 0.5) {
  cat("Interpretation: Moderate clustering tendency\n") 
} else {
  cat("Interpretation: Data appears randomly distributed\n")
}

# Store final values for reporting in results section
hopkins_final <- hopkins_mean
hopkins_ci_final <- c(hopkins_ci_lower, hopkins_ci_upper)

# Validation: Test on random data to ensure method works correctly
set.seed(456)
random_test_data <- matrix(rnorm(nrow(meaningful_features) * ncol(meaningful_features)), 
                          nrow = nrow(meaningful_features), 
                          ncol = ncol(meaningful_features))
random_scaled <- scale(random_test_data)

hopkins_random_check <- mean(replicate(5, {
  manual_hopkins(random_scaled, m = 30)
}))

cat("\nValidation - Hopkins on random data:", round(hopkins_random_check, 4))
cat(" (should be ~0.5)\n")
```

```{r}
#| label: pca_visualization
#| warning: false
#| message: false
#| results: "hide"

# Visual Assessment using PCA visualization
pca_real_data <- prcomp(player_data_scaled)
pca_random_data <- prcomp(random_cluster_data)

# Create PCA plots for comparison
pca_plot_real <- fviz_pca_ind(pca_real_data, 
                  title = "PCA - Player Performance Data",
                  geom = "point", 
                  ggtheme = theme_minimal(),
                  legend = "none",
                  alpha.ind = 0.7) +
      theme(plot.title = element_text(face = "bold", size = 14))

pca_plot_random <- fviz_pca_ind(pca_random_data, 
                  title = "PCA - Random Data",
                  geom = "point", 
                  ggtheme = theme_minimal(),
                  legend = "none",
                  alpha.ind = 0.7) +
      theme(plot.title = element_text(face = "bold", size = 14))
```

```{r}
#| label: vat_analysis
#| echo: false
#| warning: false
#| message: false
#| results: "hide"

# Using ordered dissimilarity matrices
dist_real_data <- get_dist(player_data_scaled, method = "euclidean")
dist_random_data <- get_dist(random_cluster_data, method = "euclidean")

# Create ordered dissimilarity matrix plots (VAT approach)
dist_plot_real <- fviz_dist(dist_real_data, 
                show_labels = FALSE,
                order = TRUE, # Important: order the matrix for VAT
                gradient = list(low = "#00AFBB", mid = "white", high = "#FC4E07")) +
       ggtitle("VAT Ordered Dissimilarity Matrix - Player Data") +
       theme(plot.title = element_text(face = "bold", size = 14))

dist_plot_random <- fviz_dist(dist_random_data, 
                show_labels = FALSE,
                order = TRUE, # Important: order the matrix for VAT
                gradient = list(low = "#00AFBB", mid = "white", high = "#FC4E07")) +
       ggtitle("VAT Ordered Dissimilarity Matrix - Random Data") +
       theme(plot.title = element_text(face = "bold", size = 14))
```

Before fitting cluster models, we evaluated whether the data showed non-random structure. The Hopkins statistic [@banerjee2004a] assesses cluster tendency by comparing the observed data structure to what would be expected under spatial randomness. Specifically, it measures the ratio of distances from a sample of data points to their nearest neighbors within the actual dataset, relative to distances from uniformly distributed random points to their nearest neighbors in the actual data. The statistic ranges from 0 to 1, where values near 0.5 indicate spatial randomness (no meaningful cluster structure), values substantially below 0.5 suggest regular spacing, and values substantially above 0.5 indicate clusterable structure. Values above 0.70 are generally considered strong evidence of cluster tendency, with higher values reflecting more pronounced separation between natural groupings in the feature space.

The Hopkins statistic computed from our player behavioral data was high (`r round(hopkins_final, 4)`, 95% CI [`r round(hopkins_ci_final[1], 4)`, `r round(hopkins_ci_final[2], 4)`]), substantially exceeding the 0.70 threshold and indicating strong cluster tendency. As a validity check, a random dataset with identical dimensions produced a substantially lower value (`r round(hopkins_random_check, 4)`) closer to the 0.5 benchmark expected under spatial randomness. VAT plots showed clear dark diagonal blocks for the player data, consistent with separable groups, while the random reference data showed no comparable block structure (Appendix 1; @bezdek2002a). PCA visualizations converged on the same conclusion, with visible grouping in the player data but not in the random comparison. The convergence of these three independent assessments (Hopkins, VAT, and PCA) provided strong statistical evidence to reject the null hypothesis of spatial randomness, demonstrating that the player behavioral data contained meaningful, non-random structure amenable to clustering methods.


### Determining Optimal Number of Clusters

```{r}
#| label: pam_silhouette_method
#| fig.width: 10
#| fig.height: 4
#| cache: false
#| results: "asis"
#| fig.show: "hide"  # Changed from "asis" to "hide" to prevent displaying here

# PAM optimal cluster determination
# -------------------------------------
# Calinski-Harabasz Index (CH index) for PAM
ch_index_pam <- function(data, max_k = 10) {
  ch_values <- numeric(max_k - 1)
  for (i in 2:max_k) {
    pam_result <- pam(data, k = i)
    ch_values[i-1] <- fpc::calinhara(data, pam_result$clustering)
  }
  k_values <- 2:max_k
  return(data.frame(k = k_values, ch_index = ch_values))
}

ch_pam_results <- ch_index_pam(player_data_scaled)
p_pam_ch <- ggplot(ch_pam_results, aes(x = k, y = ch_index)) +
  geom_line() +
  geom_point() +
  scale_x_continuous(breaks = 2:10) +
  labs(title = "Calinski-Harabasz Index",
       x = "Number of clusters k",
       y = "CH Index (higher is better)") +
  theme_minimal()

# Display the plot
p_pam_ch
```

```{r}
#| label: determine_optimal_k
#| echo: false
# Determine the optimal k value based on CH Index
ch_optimal_k <- ch_pam_results$k[which.max(ch_pam_results$ch_index)]
final_optimal_k <- ch_optimal_k
```

To select the number of clusters, we evaluated PAM solutions for k = 2 to 10 using the Calinski-Harabasz (CH) index, where higher values indicate better separation relative to within-cluster dispersion [@kaufmanFindingGroupsData1990; @tibshiraniEstimatingNumberClusters2001]. The CH index peaked at k = `r final_optimal_k` (Appendix 2), supporting a three-cluster solution as the best trade-off between fit and parsimony. This solution also yielded interpretable profiles consistent with prior telemetry-based playstyle research [@drachenGunsSwordsData2012; @drachen2014a; @melhartYourGameplaySays2019].

### PAM Clustering

```{r}
#| label: prepare_player_data

# Prepare player data for clustering
player_data_features <- pvz_df_analysis %>% 
  select(
    # Game mode preferences
    ratio_pvp_hours, ratio_pve_hours,
    
    # Performance metrics
    pvp_kd, pve_kd, 
    pvp_hit_accuracy, pve_hit_accuracy,
    pvp_average_score, pve_average_score,
    pvp_dps, pve_dps
  ) %>%
  # Replace NAs with 0s instead of removing them
  replace_na(list(
    ratio_pvp_hours = 0, ratio_pve_hours = 0,
    pvp_kd = 0, pve_kd = 0,
    pvp_hit_accuracy = 0, pve_hit_accuracy = 0,
    pvp_average_score = 0, pve_average_score = 0,
    pvp_dps = 0, pve_dps = 0
  ))

# Standardize the data for clustering
player_data_scaled <- scale(player_data_features)

# Perform PAM clustering with k=3
set.seed(42) # For reproducibility
k_clusters <- 3
pam_result <- pam(player_data_scaled, k = k_clusters, metric = "euclidean")

```

```{r}
#| label: pam_cluster_statistics
#| cache: false
#| include: false

# Calculate silhouette coefficient for each cluster using the existing pam_result
sil <- pam_result$silinfo$widths
sil_per_cluster <- tapply(sil[, 3], sil[, 1], mean)

# Store key statistics
stats_pam_k3 <- list(
  k = k_clusters,
  avg_silhouette = pam_result$silinfo$avg.width,
  silhouette_per_cluster = sil_per_cluster,
  cluster_sizes = table(pam_result$clustering),
  medoid_indices = pam_result$id.med
)
```

Having selected k = 3 on the basis of the CH index, we assessed the internal validity of this solution using silhouette analysis [@kaufmanFindingGroupsData1990], which quantifies how well each observation fits its assigned cluster relative to the nearest neighbouring cluster. The average silhouette width was `r round(stats_pam_k3$avg_silhouette, 2)`. According to Kaufman and Rousseeuw's interpretive guidelines, values above 0.25 indicate strong structure and values between 0.10 and 0.25 indicate weak but still meaningful structure; the overall value thus falls at the boundary, suggesting reasonable though not crisp separation. Per-cluster silhouette widths ranged from `r round(min(stats_pam_k3$silhouette_per_cluster), 2)` to `r round(max(stats_pam_k3$silhouette_per_cluster), 2)`, with one cluster exceeding the 0.25 threshold and the remaining two showing acceptable cohesion. No negative silhouette values were observed, confirming that every observation was assigned to its most appropriate cluster. The substantive interpretation of these three clusters is presented below.




```{r}
#| label: cluster_profiles
#| include: false

# Define the feature columns used in clustering
feature_columns <- c("ratio_pvp_hours", "ratio_pve_hours", 
                    "pvp_kd", "pve_kd", 
                    "pvp_hit_accuracy", "pve_hit_accuracy",
                    "pvp_average_score", "pve_average_score",
                    "pvp_dps", "pve_dps")

# Get cluster medoids (representative players)
medoid_indices <- pam_result$id.med
medoid_data <- pvz_df_analysis[medoid_indices, ]
medoid_features <- medoid_data %>%
  select(all_of(feature_columns))
medoid_features$cluster <- 1:k_clusters

# Calculate mean values by cluster for interpretation
cluster_means <- pvz_df_analysis %>%
  # First add the cluster assignments
  mutate(cluster = pam_result$clustering) %>%
  group_by(cluster) %>%
  summarize(across(all_of(feature_columns), mean, na.rm = TRUE), .groups = "drop")

# Display medoid profiles 
medoid_features |>
  tt(caption = "Cluster Medoids (Feature Values of Representative Players)") |>
  format_tt(digits = 2)

# Display cluster means
cluster_means |>
  tt(caption = "Cluster Means - Average Feature Values") |>
  format_tt(digits = 2)
```

```{r}
#| label: assign_player_types
#| code-summary: "Assign player type labels to clusters"

# Create a data frame with cluster labels 
# Update these names based on interpretation of the 3 clusters
player_type_labels <- tibble(
  cluster = 1:3,
  player_type = c(
    "Casual PVP Players",           # Cluster 1 name
    "Casual PVE Players",       # Cluster 2 name 
    "Elite Allround Players"            # Cluster 3 name
  )
)

# Add cluster assignments and type labels to the dataframe
pvz_df_clustered <- pvz_df_analysis %>%
  mutate(cluster = pam_result$clustering) %>%
  left_join(player_type_labels, by = "cluster") %>%
  mutate(player_type = factor(player_type))

# Save the clustered data with no outliers
write_csv(pvz_df_clustered, here("data/processed/pvz_with_clusters_pam3_nohours_no_outliers.csv"))

```

```{r}
#| label: compute-key-statistics
#| code-summary: "Compute key statistics for inline reporting"

# Calculate player type summary statistics for inline reporting
player_type_summary <- pvz_df_clustered %>%
  group_by(player_type) %>%
  summarise(
    n = n(),
    pct = n() / nrow(pvz_df_clustered) * 100,
    pvp_time_pct = mode_mean(ratio_pvp_hours) * 100,
    pve_time_pct = mode_mean(ratio_pve_hours) * 100,
    pvp_kd = mode_mean(pvp_kd),
    pve_kd = mode_mean(pve_kd),
    pvp_accuracy = mode_mean(pvp_hit_accuracy) * 100,
    pve_accuracy = mode_mean(pve_hit_accuracy) * 100,
    wellbeing_mean = mean(spane_game_balance, na.rm = TRUE),
    enjoyment_mean = mean(enjoyment, na.rm = TRUE),
    .groups = "drop"
  )

# Extract model coefficients for inline reporting
get_model_coef <- function(model, term) {
  coefs <- coef(summary(model))
  if (term %in% rownames(coefs)) {
    return(list(
      b = coefs[term, "Estimate"],
      se = coefs[term, "Std. Error"],
      t = coefs[term, "t value"]
    ))
  }
  return(NULL)
}

# Store key coefficients after models are fit (will be populated after model fitting)
```

```{r}
#| label: radar_functions

# Function to scale values using z-scores
z_score_scale <- function(x) {
  if (sd(x, na.rm = TRUE) == 0) return(rep(0, length(x)))
  return((x - mean(x, na.rm = TRUE)) / sd(x, na.rm = TRUE))
}

# Convert z-scores to 0-1 scale for radar chart
z_to_radar_scale <- function(z, min_z = -2.5, max_z = 2.5) {
  z_clipped <- pmax(pmin(z, max_z), min_z)
  scaled <- (z_clipped - min_z) / (max_z - min_z)
  return(scaled)
}

# Function to create radar chart with average player comparison
create_player_type_radar <- function(data, avg_data, color, title) {
  # Get column names
  data_cols <- colnames(data)
  
  # Create max and min rows
  max_min <- data.frame(matrix(c(rep(1, length(data_cols)), rep(0, length(data_cols))), 
                             nrow = 2, byrow = TRUE))
  colnames(max_min) <- data_cols
  
  # Combine data for plotting (max, min, average, player type)
  plot_data <- rbind(max_min, avg_data, data)
  
  # Create radar chart
  radarchart(
    plot_data, 
    pcol = c("gray50", color),        # Colors for average and player type (darker gray)
    pfcol = c(adjustcolor("gray50", alpha.f = 0.15), adjustcolor(color, alpha.f = 0.4)), # Fill colors
    plwd = c(2.5, 3.5),              
    plty = c(2, 1),                  # Line type (dashed for average, solid for player type)
    vlcex = 1.1,                     
    title = title,
    cglcol = "grey70",
    cglty = 1,
    axislabcol = "grey20",           
    caxislabels = seq(0, 1, 0.25),
    cglwd = 1,                       
    pty = 32,                        # Use 32-sided polygon for smoother lines
    axistype = 0,                    
    centerzero = FALSE,              # Do not center at zero
    seg = 4,                         
    cex.main = 1.4                   
  )
  
  # Add legend for player type vs average
  legend(
    "bottomright", 
    legend = c("Average Player", title),
    col = c("gray50", color),
    lty = c(2, 1),
    lwd = c(2.5, 3.5),
    bty = "n",     # No box around legend
    cex = 1.0,     # Larger font size
    pt.cex = 1.2   # Larger symbol size
  )
}
```

```{r}
#| label: player_type_summary
#| include: false

# Create a summary of player types
player_type_summary <- pvz_df_clustered %>%
  group_by(player_type) %>%
  summarise(
    Count = n(),
    `% of Players` = n() / nrow(pvz_df_clustered) * 100,
    `PvP Time %` = mode_mean(ratio_pvp_hours) * 100,
    `PvE Time %` = mode_mean(ratio_pve_hours) * 100,
    `PvP K/D` = mode_mean(pvp_kd),
    `PvE K/D` = mode_mean(pve_kd),
    `PvP Accuracy` = mode_mean(pvp_hit_accuracy) * 100,
    `PvE Accuracy` = mode_mean(pve_hit_accuracy) * 100
  )
```

```{r}
#| label: player_type_inline_values
#| include: false

# Keep inline text simple: pull one row per player type, then named scalars
pt_pve <- player_type_summary %>% filter(player_type == "Casual PVE Players")
pt_elite <- player_type_summary %>% filter(player_type == "Elite Allround Players")
pt_pvp <- player_type_summary %>% filter(player_type == "Casual PVP Players")

pve_pct_players <- round(pt_pve$`% of Players`, 1)
pve_time <- round(pt_pve$`PvE Time %`, 1)
pve_pve_kd <- round(pt_pve$`PvE K/D`, 2)
pve_pve_acc <- round(pt_pve$`PvE Accuracy`, 1)

elite_pct_players <- round(pt_elite$`% of Players`, 1)
elite_pvp_kd <- round(pt_elite$`PvP K/D`, 2)
elite_pve_kd <- round(pt_elite$`PvE K/D`, 2)
elite_pvp_acc <- round(pt_elite$`PvP Accuracy`, 1)
elite_pve_acc <- round(pt_elite$`PvE Accuracy`, 1)

pvp_pct_players <- round(pt_pvp$`% of Players`, 1)
pvp_time <- round(pt_pvp$`PvP Time %`, 1)
pvp_pvp_kd <- round(pt_pvp$`PvP K/D`, 2)
pvp_pvp_acc <- round(pt_pvp$`PvP Accuracy`, 1)
```

### Interpretation of Playstyles

Following established approaches to telemetry-based player profiling [@drachenGunsSwordsData2012; @drachen2014a; @bauckhageClusteringGameBehavior2015; @melhartYourGameplaySays2019; @sifaProfilingGamesUnderstanding2018], we labelled the three clusters on the basis of their dominant mode preferences and performance levels (@fig-radar-charts). The labels reflect two dimensions that emerged from the data: the game mode players predominantly engaged with (PVE vs. PVP) and their overall performance level (casual vs. elite).

Casual PVE Players (`r pve_pct_players`% of the sample) allocated the majority of their playtime to cooperative PVE modes (`r pve_time`%), achieving moderate performance within that context (PVE K/D: M = `r pve_pve_kd`; PVE Accuracy: M = `r pve_pve_acc`%) while showing markedly lower metrics in PVP. This profile suggests a specialisation in story-driven exploration and cooperative play over competitive engagement.

Elite Allround Players (`r elite_pct_players`% of the sample), the largest cluster, maintained high performance across both gameplay contexts, with kill-to-death ratios and accuracy rates substantially above the sample mean in both PVP (K/D: M = `r elite_pvp_kd`; Accuracy: M = `r elite_pvp_acc`%) and PVE (K/D: M = `r elite_pve_kd`; Accuracy: M = `r elite_pve_acc`%). Rather than specialising in one mode, these players engaged broadly and performed well regardless of context, distinguishing them as the most proficient segment.

Casual PVP Players (`r pvp_pct_players`% of the sample) devoted the majority of their time to competitive PVP modes (`r pvp_time`%), yet their performance was moderate (PVP K/D: M = `r pvp_pvp_kd`; PVP Accuracy: M = `r pvp_pvp_acc`%), falling well below Elite Allround Players. Performance in PVE was similarly modest. This pattern indicates a preference for competitive engagement that is not accompanied by correspondingly high skill, suggesting a casual orientation to play within a competitive context.

```{r}
#| label: radar-caption-values
#| include: false

# Calculate summary values for each player type for the figure caption
radar_caption_data <- pvz_df_clustered %>%
  group_by(player_type) %>%
  summarise(
    pvp_ratio = mode_mean(ratio_pvp_hours),
    pve_ratio = mode_mean(ratio_pve_hours),
    pvp_kd = mode_mean(pvp_kd),
    pve_kd = mode_mean(pve_kd),
    .groups = "drop"
  )

radar_caption <- paste0(

"Radar charts showing performance and playstyle profiles for each player type relative to the average player (dashed line). Values are standardized scores (z-scores converted to 0-1 scale). ",
"Mean values: Casual PVE Players (PvP Ratio: ", type_stat(radar_caption_data, "Casual PVE Players", "pvp_ratio", 2),
", PvE Ratio: ", type_stat(radar_caption_data, "Casual PVE Players", "pve_ratio", 2),
", PvP K/D: ", type_stat(radar_caption_data, "Casual PVE Players", "pvp_kd", 2),
", PvE K/D: ", type_stat(radar_caption_data, "Casual PVE Players", "pve_kd", 2), "); ",
"Elite Allround Players (PvP Ratio: ", type_stat(radar_caption_data, "Elite Allround Players", "pvp_ratio", 2),
", PvE Ratio: ", type_stat(radar_caption_data, "Elite Allround Players", "pve_ratio", 2),
", PvP K/D: ", type_stat(radar_caption_data, "Elite Allround Players", "pvp_kd", 2),
", PvE K/D: ", type_stat(radar_caption_data, "Elite Allround Players", "pve_kd", 2), "); ",
"Casual PVP Players (PvP Ratio: ", type_stat(radar_caption_data, "Casual PVP Players", "pvp_ratio", 2),
", PvE Ratio: ", type_stat(radar_caption_data, "Casual PVP Players", "pve_ratio", 2),
", PvP K/D: ", type_stat(radar_caption_data, "Casual PVP Players", "pvp_kd", 2),
", PvE K/D: ", type_stat(radar_caption_data, "Casual PVP Players", "pve_kd", 2), "). ",
"PvE = Player versus Environment; PvP = Player versus Player; K/D = Kill/Death ratio; DPS = Damage per second."
)
```

```{r}
#| label: fig-radar-charts
#| fig-cap: !expr radar_caption
#| fig-height: 6
#| fig-width: 14
#| out-width: "100%"
#| code-summary: "Player type radar chart visualization"

# Define metrics for radar charts in a more logical order, 
# grouping related metrics together
radar_metrics <- c(
  # PvE metrics grouped together
  "PvE Ratio", "PvE K/D", "PvE Accuracy", "PvE Score", "PvE DPS",
  # PvP metrics grouped together
  "PvP Ratio", "PvP K/D", "PvP Accuracy", "PvP Score", "PvP DPS"
)

# Get player type names
player_type_names <- player_type_labels$player_type

# Calculate radar chart data for both average player and player types
# First collect the raw metrics for all players and player types
metrics_data <- pvz_df_analysis %>%
  # Add a column for the average player
  mutate(profile_type = "Average Player") %>%
  # Bind the clustered data with profile types
  bind_rows(
    pvz_df_clustered %>%
      select(
        player_id, player_type, 
        ratio_pve_hours, ratio_pvp_hours,
        pve_kd, pvp_kd,
        pve_hit_accuracy, pvp_hit_accuracy,
        pve_average_score, pvp_average_score,
        pve_dps, pvp_dps
      ) %>%
      rename(profile_type = player_type)
  ) %>%
  # Group by profile type (average or a player type)
  group_by(profile_type) %>%
  # Calculate means for each metric
  summarise(
    ratio_pve_mean = mode_mean(ratio_pve_hours),
    pve_kd_mean = mode_mean(pve_kd),
    pve_accuracy_mean = mode_mean(pve_hit_accuracy),
    pve_score_mean = mode_mean(pve_average_score),
    pve_dps_mean = mode_mean(pve_dps),
    ratio_pvp_mean = mode_mean(ratio_pvp_hours),
    pvp_kd_mean = mode_mean(pvp_kd),
    pvp_accuracy_mean = mode_mean(pvp_hit_accuracy),
    pvp_score_mean = mode_mean(pvp_average_score),
    pvp_dps_mean = mode_mean(pvp_dps)
  ) %>%
  ungroup()

# Create vectors of all values for each metric to calculate z-scores across all profiles
all_ratio_pve <- metrics_data$ratio_pve_mean
all_pve_kd <- metrics_data$pve_kd_mean
all_pve_accuracy <- metrics_data$pve_accuracy_mean
all_pve_score <- metrics_data$pve_score_mean
all_pve_dps <- metrics_data$pve_dps_mean
all_ratio_pvp <- metrics_data$ratio_pvp_mean
all_pvp_kd <- metrics_data$pvp_kd_mean
all_pvp_accuracy <- metrics_data$pvp_accuracy_mean
all_pvp_score <- metrics_data$pvp_score_mean
all_pvp_dps <- metrics_data$pvp_dps_mean

# Now apply z-score scaling to create the radar data
radar_data <- metrics_data %>%
  mutate(
    `PvE Ratio` = z_to_radar_scale(z_score_scale(all_ratio_pve)[row_number()]),
    `PvE K/D` = z_to_radar_scale(z_score_scale(all_pve_kd)[row_number()]),
    `PvE Accuracy` = z_to_radar_scale(z_score_scale(all_pve_accuracy)[row_number()]),
    `PvE Score` = z_to_radar_scale(z_score_scale(all_pve_score)[row_number()]),
    `PvE DPS` = z_to_radar_scale(z_score_scale(all_pve_dps)[row_number()]),
    `PvP Ratio` = z_to_radar_scale(z_score_scale(all_ratio_pvp)[row_number()]),
    `PvP K/D` = z_to_radar_scale(z_score_scale(all_pvp_kd)[row_number()]),
    `PvP Accuracy` = z_to_radar_scale(z_score_scale(all_pvp_accuracy)[row_number()]),
    `PvP Score` = z_to_radar_scale(z_score_scale(all_pvp_score)[row_number()]),
    `PvP DPS` = z_to_radar_scale(z_score_scale(all_pvp_dps)[row_number()])
  ) %>%
  # Keep only relevant columns
  select(
    profile_type,
    `PvE Ratio`, `PvE K/D`, `PvE Accuracy`, `PvE Score`, `PvE DPS`,
    `PvP Ratio`, `PvP K/D`, `PvP Accuracy`, `PvP Score`, `PvP DPS`
  )

# Extract the average player data
average_player <- radar_data %>%
  filter(profile_type == "Average Player") %>%
  select(-profile_type) %>%
  as.data.frame()

# Colors for player types - more distinct and colorblind-friendly palette
player_type_colors <- c("#E41A1C", "#377EB8", "#4DAF4A")

# Set up a better layout with proper margins for larger text
# Increased bottom margin for legend text
par(mfrow = c(1, 3), mar = c(5, 2, 4, 2), oma = c(2, 0, 2, 0))

# Create radar charts for each player type
for (i in 1:length(player_type_names)) {
  # Extract data for this player type
  type_data <- radar_data %>%
    filter(profile_type == player_type_names[i]) %>%
    select(-profile_type) %>%
    as.data.frame()
  
  # Create radar chart with average player comparison
  create_player_type_radar(
    type_data, 
    average_player,
    player_type_colors[i], 
    player_type_names[i]
  )
}
# Reset layout
par(mfrow = c(1, 1))
```

## Well-being differences between player types 

```{r}
#| label: player-type-wellbeing-model
#| code-summary: "Fit robust mixed-effects models for well-being outcomes"
#| message: false
#| warning: false
#| include: false

# Create a robust linear model for well-being
set.seed(123) # For reproducibility of robust estimates

# Check if models should be refit or loaded from cache
if (eval_models || !model_exists("rmodel_game") || !model_exists("rmodel_enjoyment")) {
  # Fit robust models for game-specific well-being
  rmodel_game <- rlmer(spane_game_balance ~ player_type + hours + gender + age + (1|country), 
                       data = pvz_df_clustered)
  
  # Fit robust models for enjoyment
  rmodel_enjoyment <- rlmer(enjoyment ~ player_type + hours + gender + age + (1|country), 
                       data = pvz_df_clustered)
  
  # Save models to cache
  save_model(rmodel_game, "rmodel_game")
  save_model(rmodel_enjoyment, "rmodel_enjoyment")
} else {
  # Load cached models
  rmodel_game <- load_model("rmodel_game")
  rmodel_enjoyment <- load_model("rmodel_enjoyment")
}

# Display models side-by-side using modelsummary
modelsummary(
  list("Well-being (SPANE)" = rmodel_game, "Enjoyment" = rmodel_enjoyment),
  estimate = "{estimate} ({std.error})",
  statistic = "t = {statistic}",
  gof_omit = "AIC|BIC|RMSE|Log",
  coef_rename = c(
    "(Intercept)" = "Intercept",
    "player_typeElite Allround Players" = "Elite Allround Players",
    "player_typeCasual PVP Players" = "Casual PVP Players",
    "hours" = "Hours Played",
    "genderMale" = "Gender (Male)",
    "genderOther" = "Gender (Other)",
    "genderPrefer not to say" = "Gender (Undisclosed)",
    "age" = "Age"
  ),
  title = "Robust Linear Mixed-Effects Models Predicting Well-being and Enjoyment",
  notes = "Reference category for play style: Casual PVE Players. Reference category for gender: Female. Country included as random effect."
)

# Extract coefficients for inline reporting
coefs_game <- coef(summary(rmodel_game))
coefs_enjoyment <- coef(summary(rmodel_enjoyment))
```

We fitted robust linear mixed-effects models predicting each outcome from player type, hours played, gender, and age, with random intercepts for country (Appendix 3). Casual PVE Players served as the reference category. Relative to these players, both Casual PVP Players (b = `r round(coefs_game["player_typeCasual PVP Players", "Estimate"], 2)`, SE = `r round(coefs_game["player_typeCasual PVP Players", "Std. Error"], 2)`, t = `r round(coefs_game["player_typeCasual PVP Players", "t value"], 2)`) and Elite Allround Players (b = `r round(coefs_game["player_typeElite Allround Players", "Estimate"], 2)`, SE = `r round(coefs_game["player_typeElite Allround Players", "Std. Error"], 2)`, t = `r round(coefs_game["player_typeElite Allround Players", "t value"], 2)`) reported significantly lower game-specific well-being. For enjoyment, Casual PVP Players again scored lower than Casual PVE Players (b = `r round(coefs_enjoyment["player_typeCasual PVP Players", "Estimate"], 2)`, SE = `r round(coefs_enjoyment["player_typeCasual PVP Players", "Std. Error"], 2)`, t = `r round(coefs_enjoyment["player_typeCasual PVP Players", "t value"], 2)`), whereas the difference between Elite Allround and Casual PVE Players was smaller and not statistically reliable (b = `r round(coefs_enjoyment["player_typeElite Allround Players", "Estimate"], 2)`, SE = `r round(coefs_enjoyment["player_typeElite Allround Players", "Std. Error"], 2)`, t = `r round(coefs_enjoyment["player_typeElite Allround Players", "t value"], 2)`). Hours played was positively associated with both outcomes (well-being: b = `r round(coefs_game["hours", "Estimate"], 2)`, SE = `r round(coefs_game["hours", "Std. Error"], 2)`, t = `r round(coefs_game["hours", "t value"], 2)`; enjoyment: b = `r round(coefs_enjoyment["hours", "Estimate"], 2)`, SE = `r round(coefs_enjoyment["hours", "Std. Error"], 2)`, t = `r round(coefs_enjoyment["hours", "t value"], 2)`).

## Does Playstyle Moderate the Playtime–Well-being Relationship?

```{r}
#| label: moderation-models
#| code-summary: "Fit moderation models (play style × hours interaction)"
#| message: false
#| warning: false
#| include: false

# Check if models should be refit or loaded from cache
if (eval_models || !model_exists("rmodel_moderation_game") || !model_exists("rmodel_moderation_enjoyment")) {
  # Create interaction models for our two key well-being outcomes
  rmodel_moderation_game <- rlmer(
    spane_game_balance ~ player_type * hours + (1|country),
    data = pvz_df_clustered
  )

  rmodel_moderation_enjoyment <- rlmer(
    enjoyment ~ player_type * hours + (1|country),
    data = pvz_df_clustered
  )
  
  # Save models to cache
  save_model(rmodel_moderation_game, "rmodel_moderation_game")
  save_model(rmodel_moderation_enjoyment, "rmodel_moderation_enjoyment")
} else {
  # Load cached models
  rmodel_moderation_game <- load_model("rmodel_moderation_game")
  rmodel_moderation_enjoyment <- load_model("rmodel_moderation_enjoyment")
}

# Display moderation models side-by-side using modelsummary
modelsummary(
  list("Well-being (SPANE)" = rmodel_moderation_game, "Enjoyment" = rmodel_moderation_enjoyment),
  estimate = "{estimate} ({std.error})",
  statistic = "t = {statistic}",
  gof_omit = "AIC|BIC|RMSE|Log",
  coef_rename = c(
    "(Intercept)" = "Intercept",
    "player_typeElite Allround Players" = "Elite Allround Players",
    "player_typeCasual PVP Players" = "Casual PVP Players",
    "hours" = "Hours Played",
    "player_typeElite Allround Players:hours" = "Elite Allround × Hours",
    "player_typeCasual PVP Players:hours" = "Casual PVP × Hours"
  ),
  title = "Moderation Models: Play Style × Hours Interaction",
  notes = "Reference category for play style: Casual PVE Players. Country included as random effect."
)

# Extract coefficients for inline reporting
coefs_mod_game <- coef(summary(rmodel_moderation_game))
coefs_mod_enjoyment <- coef(summary(rmodel_moderation_enjoyment))
```

To test whether the playtime–outcome association differed across playstyles, we added player type × hours interaction terms to the models (Appendix 4). Each interaction coefficient represents the difference in the playtime slope for a given playstyle relative to Casual PVE Players. None of the interaction terms reached significance: for well-being, neither Casual PVP Players (Δb = `r round(coefs_mod_game["player_typeCasual PVP Players:hours", "Estimate"], 2)`, SE = `r round(coefs_mod_game["player_typeCasual PVP Players:hours", "Std. Error"], 2)`, t = `r round(coefs_mod_game["player_typeCasual PVP Players:hours", "t value"], 2)`) nor Elite Allround Players (Δb = `r round(coefs_mod_game["player_typeElite Allround Players:hours", "Estimate"], 2)`, SE = `r round(coefs_mod_game["player_typeElite Allround Players:hours", "Std. Error"], 2)`, t = `r round(coefs_mod_game["player_typeElite Allround Players:hours", "t value"], 2)`) had a playtime slope that differed from the reference group. The same held for enjoyment (Casual PVP: Δb = `r round(coefs_mod_enjoyment["player_typeCasual PVP Players:hours", "Estimate"], 2)`, SE = `r round(coefs_mod_enjoyment["player_typeCasual PVP Players:hours", "Std. Error"], 2)`, t = `r round(coefs_mod_enjoyment["player_typeCasual PVP Players:hours", "t value"], 2)`; Elite Allround: Δb = `r round(coefs_mod_enjoyment["player_typeElite Allround Players:hours", "Estimate"], 2)`, SE = `r round(coefs_mod_enjoyment["player_typeElite Allround Players:hours", "Std. Error"], 2)`, t = `r round(coefs_mod_enjoyment["player_typeElite Allround Players:hours", "t value"], 2)`). Baseline playstyle differences in intercepts remained in the interaction models, particularly for Casual PVP Players (well-being: b = `r round(coefs_mod_game["player_typeCasual PVP Players", "Estimate"], 2)`, t = `r round(coefs_mod_game["player_typeCasual PVP Players", "t value"], 2)`; enjoyment: b = `r round(coefs_mod_enjoyment["player_typeCasual PVP Players", "Estimate"], 2)`, t = `r round(coefs_mod_enjoyment["player_typeCasual PVP Players", "t value"], 2)`).

```{r}
#| label: wellbeing-anova
#| message: false
#| warning: false
#| include: false

# Run ANOVAs to test for significant differences between player types
wellbeing_vars <- c("spane_game_balance", "enjoyment")

# Store ANOVA results and models
anova_results <- list()
anova_models <- list()

for (var in wellbeing_vars) {
  formula <- as.formula(paste(var, "~ player_type"))
  anova_model <- aov(formula, data = pvz_df_clustered)
  anova_models[[var]] <- anova_model
  anova_results[[var]] <- broom::tidy(anova_model)
  anova_results[[var]]$variable <- var
  
  # Add effect size
  anova_results[[var]]$eta_squared <- effectsize::eta_squared(anova_model)[["Eta2"]]
  
  # Add residual df
  anova_results[[var]]$df_residual <- summary(anova_model)[[1]]["Residuals", "Df"]
}

# Combine all results
all_anova_results <- bind_rows(anova_results)

# Display results in a nice table
all_anova_results |> 
  filter(term == "player_type") |>
  select(Variable = variable, df, SS = sumsq, F = statistic, p = p.value, `η²` = eta_squared) |>
  mutate(
    p = case_when(
      p < 0.001 ~ "< .001^†^",
      p < 0.05 ~ paste0(sprintf("%.3f", round(p, 3)), "^†^"),
      TRUE ~ sprintf("%.3f", round(p, 3))
    ),
    `η²` = round(`η²`, 3)
  )

# Post-hoc tests for variables with significant differences
posthoc_results <- list()

for (var in wellbeing_vars) {
  if (summary(anova_models[[var]])[[1]]["player_type", "Pr(>F)"] < 0.05) {
    posthoc <- TukeyHSD(anova_models[[var]])
    posthoc_results[[var]] <- broom::tidy(posthoc)
  }
}

# Extract key statistics for inline reporting
wellbeing_anova <- all_anova_results |> 
  filter(term == "player_type" & variable == "spane_game_balance")
enjoyment_anova <- all_anova_results |> 
  filter(term == "player_type" & variable == "enjoyment")

# Extract post-hoc p-values
wellbeing_posthoc <- posthoc_results[["spane_game_balance"]]
enjoyment_posthoc <- posthoc_results[["enjoyment"]]

# Get specific comparisons (TukeyHSD tidy output uses 'contrast' column)
wb_pvp_pve <- wellbeing_posthoc |> 
  filter(grepl("Casual PVP.*Casual PVE|Casual PVE.*Casual PVP", contrast))
enj_pvp_pve <- enjoyment_posthoc |> 
  filter(grepl("Casual PVP.*Casual PVE|Casual PVE.*Casual PVP", contrast))
enj_pvp_elite <- enjoyment_posthoc |> 
  filter(grepl("Casual PVP.*Elite|Elite.*Casual PVP", contrast))
```

```{r}
#| label: fig-anova-boxplots
#| fig-cap: "Distribution of well-being outcomes by play style. Panel A shows game-specific well-being (SPANE balance score); Panel B shows enjoyment. Boxes represent interquartile ranges with median lines; whiskers extend to 1.5 × IQR. Individual points show outliers."
#| fig-width: 10
#| fig-height: 5
#| message: false
#| warning: false

# Create boxplot for game-specific well-being
p_wellbeing <- ggplot(pvz_df_clustered, aes(x = player_type, y = spane_game_balance, fill = player_type)) +
  geom_boxplot(alpha = 0.7, outlier.shape = 21) +
  geom_jitter(width = 0.2, alpha = 0.2, size = 1) +
  scale_fill_brewer(palette = "Set2") +
  labs(
    title = "A. Game-Specific Well-being",
    x = NULL,
    y = "SPANE Balance Score"
  ) +
  theme_minimal() +
  theme(
    legend.position = "none",
    axis.text.x = element_text(angle = 15, hjust = 1),
    panel.grid.major.x = element_blank()
  )

# Create boxplot for enjoyment
p_enjoyment <- ggplot(pvz_df_clustered, aes(x = player_type, y = enjoyment, fill = player_type)) +
  geom_boxplot(alpha = 0.7, outlier.shape = 21) +
  geom_jitter(width = 0.2, alpha = 0.2, size = 1) +
  scale_fill_brewer(palette = "Set2") +
  labs(
    title = "B. Enjoyment",
    x = NULL,
    y = "Enjoyment Score"
  ) +
  theme_minimal() +
  theme(
    legend.position = "none",
    axis.text.x = element_text(angle = 15, hjust = 1),
    panel.grid.major.x = element_blank()
  )

# Combine plots
p_wellbeing + p_enjoyment
```

```{r}
#| label: dynamic-results-text
#| echo: false

# Function to format p-value for inline text
format_p_inline <- function(p) {
  if (length(p) == 0 || is.na(p[1])) return("N/A")
  p <- p[1]  # Take first value if vector
  if (p < .001) return("< .001")
  sprintf("= %.3f", p)
}
```

These playstyle differences were corroborated by one-way ANOVAs (well-being: *F*(`r wellbeing_anova$df`, `r wellbeing_anova$df_residual`) = `r sprintf("%.2f", wellbeing_anova$statistic)`, *p* `r format_p_inline(wellbeing_anova$p.value)`, η² = `r sprintf("%.3f", wellbeing_anova$eta_squared)`; enjoyment: *F*(`r enjoyment_anova$df`, `r enjoyment_anova$df_residual`) = `r sprintf("%.2f", enjoyment_anova$statistic)`, *p* `r format_p_inline(enjoyment_anova$p.value)`, η² = `r sprintf("%.3f", enjoyment_anova$eta_squared)`) and Tukey HSD post-hoc tests, which confirmed that Casual PVP Players scored lower than Casual PVE Players on both well-being (*p* `r format_p_inline(wb_pvp_pve$adj.p.value)`) and enjoyment (*p* `r format_p_inline(enj_pvp_pve$adj.p.value)`), and lower than Elite Allround Players on enjoyment (*p* `r format_p_inline(enj_pvp_elite$adj.p.value)`) (@fig-anova-boxplots).


```{r}
#| label: simple-slopes-functions
#| message: false
#| warning: false
#| include: false

# Function to calculate simple slopes for all player types
calculate_all_simple_slopes <- function(model, outcome_name) {
  # For robust mixed models, extract coefficients manually
  coefs <- fixef(model)
  vcov_matrix <- vcov(model)
  
  # Get player types
  player_types <- levels(pvz_df_clustered$player_type)
  
  # Base effect of hours (for reference level)
  hours_effect <- coefs["hours"]
  hours_se <- sqrt(vcov_matrix["hours", "hours"])
  
  # Create results dataframe
  results <- data.frame(
    player_type = player_types[1],  # reference level
    estimate = hours_effect,
    std.error = hours_se,
    statistic = hours_effect / hours_se,
    p.value = 2 * pt(abs(hours_effect / hours_se), 
                    df = nrow(pvz_df_clustered) - length(coefs), 
                    lower.tail = FALSE),
    stringsAsFactors = FALSE
  )
  
  # Calculate for other player types
  for (i in 2:length(player_types)) {
    pt <- player_types[i]
    interaction_term <- paste0("player_type", pt, ":hours")
    
    if (interaction_term %in% names(coefs)) {
      # Calculate slope
      slope <- hours_effect + coefs[interaction_term]
      
      # Calculate SE using delta method (simplified)
      var_hours <- vcov_matrix["hours", "hours"]
      var_inter <- vcov_matrix[interaction_term, interaction_term]
      cov_hours_inter <- vcov_matrix["hours", interaction_term]
      slope_se <- sqrt(var_hours + var_inter + 2*cov_hours_inter)
      
      # t-statistic and p-value
      t_stat <- slope / slope_se
      p_val <- 2 * pt(abs(t_stat), df = nrow(pvz_df_clustered) - length(coefs), lower.tail = FALSE)
      
      # Add to results
      results <- rbind(results, data.frame(
        player_type = pt,
        estimate = slope,
        std.error = slope_se,
        statistic = t_stat,
        p.value = p_val,
        stringsAsFactors = FALSE
      ))
    }
  }
  
  # Add confidence intervals and formatting
  results <- results %>%
    mutate(
      outcome = outcome_name,
      conf.low = estimate - 1.96 * std.error,
      conf.high = estimate + 1.96 * std.error,
      significant = p.value < 0.05,
      sig = case_when(
        p.value < 0.001 ~ "^†^",
        p.value < 0.01 ~ "^†^",
        p.value < 0.05 ~ "^†^",
        TRUE ~ ""
      )
    )
  
  return(results)
}

# Function to create interaction plots for all player types
create_interaction_plots <- function(model, outcome_name) {
  # Extract fixed effects
  fixed_effects <- fixef(model)
  
  # Get player types
  player_types <- levels(pvz_df_clustered$player_type)
  
  # Create a grid of values for hours
  hours_range <- range(pvz_df_clustered$hours, na.rm = TRUE)
  hours_grid <- seq(from = hours_range[1], to = hours_range[2], length.out = 100)
  
  # Get the coefficient for hours (main effect)
  hours_coef <- fixed_effects["hours"]
  
  # Create a dataframe to store predictions
  predictions <- data.frame()
  
  # For each player type, calculate predicted values across hours range
  for (i in 1:length(player_types)) {
    pt <- player_types[i]
    
    # Calculate intercept
    if (i == 1) {
      # Reference level
      intercept <- fixed_effects["(Intercept)"]
      slope <- hours_coef
    } else {
      # Other levels
      intercept <- fixed_effects["(Intercept)"] + fixed_effects[paste0("player_type", pt)]
      interaction_term <- paste0("player_type", pt, ":hours")
      slope <- hours_coef + fixed_effects[interaction_term]
    }
    
    # Calculate predictions
    pred <- intercept + slope * hours_grid
    
    # Add to predictions dataframe
    predictions <- rbind(predictions, data.frame(
      player_type = pt,
      hours = hours_grid,
      predicted = pred
    ))
  }
  
  # Create the plot
  p <- ggplot(predictions, aes(x = hours, y = predicted, color = player_type)) +
    geom_line(size = 1.2) +
    geom_ribbon(aes(ymin = predicted - 1.96 * 0.5, ymax = predicted + 1.96 * 0.5, fill = player_type), alpha = 0.2) +
    labs(
      title = paste("Interaction Plot for", outcome_name),
      x = "Hours Played",
      y = outcome_name,
      color = "Player Type",
      fill = "Player Type"
    ) +
    theme_minimal() +
    theme(
      plot.title = element_text(hjust = 0.5, face = "bold"),
      legend.position = "bottom",
      panel.grid.minor = element_blank()
    ) +
    scale_color_brewer(palette = "Set2") +
    scale_fill_brewer(palette = "Set2")
  
  return(p)
}

# Function to create Johnson-Neyman plot
create_jn_plot <- function(model, moderator = "player_type", predictor = "hours", outcome_name) {
  # For rlmerMod models, we'll use a custom approach
  # This function creates a simplified version of Johnson-Neyman plot
  
  # Extract fixed effects
  fixed_effects <- fixef(model)
  vcov_matrix <- vcov(model)
  
  # Get player types
  player_types <- levels(pvz_df_clustered$player_type)
  
  # Create a dataframe to store slopes and confidence intervals
  jn_data <- data.frame(
    player_type = player_types,
    slope = numeric(length(player_types)),
    se = numeric(length(player_types)),
    lower_ci = numeric(length(player_types)),
    upper_ci = numeric(length(player_types)),
    stringsAsFactors = FALSE
  )
  
  # Get the coefficient for hours (main effect)
  hours_coef <- fixed_effects["hours"]
  hours_se <- sqrt(vcov_matrix["hours", "hours"])
  
  # For each player type, calculate slope and CI
  for (i in 1:length(player_types)) {
    pt <- player_types[i]
    
    if (i == 1) {
      # Reference level
      jn_data$slope[i] <- hours_coef
      jn_data$se[i] <- hours_se
    } else {
      # Other levels
      interaction_term <- paste0("player_type", pt, ":hours")
      slope <- hours_coef + fixed_effects[interaction_term]
      
      # Calculate SE using delta method (simplified)
      var_hours <- vcov_matrix["hours", "hours"]
      var_inter <- vcov_matrix[interaction_term, interaction_term]
      cov_hours_inter <- vcov_matrix["hours", interaction_term]
      slope_se <- sqrt(var_hours + var_inter + 2*cov_hours_inter)
      
      jn_data$slope[i] <- slope
      jn_data$se[i] <- slope_se
    }
    
    # Calculate 95% confidence intervals
    jn_data$lower_ci[i] <- jn_data$slope[i] - 1.96 * jn_data$se[i]
    jn_data$upper_ci[i] <- jn_data$slope[i] + 1.96 * jn_data$se[i]
  }
  
  # Determine significance
  jn_data$significant <- (jn_data$lower_ci > 0) | (jn_data$upper_ci < 0)
  
  # Create the plot
  p <- ggplot(jn_data, aes(x = player_type, y = slope, color = significant)) +
    geom_point(size = 3) +
    geom_errorbar(aes(ymin = lower_ci, ymax = upper_ci), width = 0.2) +
    geom_hline(yintercept = 0, linetype = "dashed", color = "gray50") +
    labs(
      title = paste("Conditional Effect of", predictor, "on", outcome_name),
      subtitle = "by Player Type with 95% Confidence Intervals",
      x = "Player Type",
      y = paste("Effect of", predictor),
      color = "Significant"
    ) +
    theme_minimal() +
    theme(
      plot.title = element_text(hjust = 0.5, face = "bold"),
      plot.subtitle = element_text(hjust = 0.5),
      legend.position = "bottom",
      panel.grid.minor = element_blank()
    ) +
    scale_color_manual(values = c("gray50", "blue"))
  
  return(p)
}
```

```{r}
#| label: wellbeing-simple-slopes
#| message: false
#| warning: false
#| include: false

# Calculate simple slopes for game-specific well-being model
wellbeing_slopes <- calculate_all_simple_slopes(rmodel_moderation_game, "Game-Specific Well-being")

# Create a formatted table for game-specific well-being simple slopes
wellbeing_slopes_table <- wellbeing_slopes %>%
  mutate(
    estimate = round(estimate, 3),
    std.error = round(std.error, 3),
    statistic = round(statistic, 3),
    p.value = case_when(
      p.value < 0.001 ~ paste0("< .001", sig),
      p.value < 0.05 ~ paste0(sprintf("%.3f", round(p.value, 3)), sig),
      TRUE ~ sprintf("%.3f", round(p.value, 3))
    ),
    ci = paste0("[", round(conf.low, 3), ", ", round(conf.high, 3), "]")
  ) %>%
  select(`Player Type` = player_type, b = estimate, SE = std.error, t = statistic, p = p.value, `95% CI` = ci)
```

```{r}
#| label: enjoyment-simple-slopes
#| message: false
#| warning: false
#| include: false

# Calculate simple slopes for enjoyment model
enjoyment_slopes <- calculate_all_simple_slopes(rmodel_moderation_enjoyment, "Intrinsic Enjoyment")

# Create a formatted table for enjoyment simple slopes
enjoyment_slopes_table <- enjoyment_slopes %>%
  mutate(
    estimate = round(estimate, 3),
    std.error = round(std.error, 3),
    statistic = round(statistic, 3),
    p.value = case_when(
      p.value < 0.001 ~ paste0("< .001", sig),
      p.value < 0.05 ~ paste0(sprintf("%.3f", round(p.value, 3)), sig),
      TRUE ~ sprintf("%.3f", round(p.value, 3))
    ),
    ci = paste0("[", round(conf.low, 3), ", ", round(conf.high, 3), "]")
  ) %>%
  select(`Player Type` = player_type, b = estimate, SE = std.error, t = statistic, p = p.value, `95% CI` = ci)
```

```{r}
#| label: wellbeing-pairwise-comparisons
#| message: false
#| warning: false
#| include: false

# Create pairwise comparison of slopes for game-specific well-being
wellbeing_pairs <- data.frame(player_type1 = character(),
                        player_type2 = character(),
                        difference = numeric(),
                        std.error = numeric(),
                        statistic = numeric(),
                        p.value = numeric(),
                        stringsAsFactors = FALSE)

player_types <- levels(pvz_df_clustered$player_type)
for (i in 1:(length(player_types)-1)) {
  for (j in (i+1):length(player_types)) {
    pt1 <- player_types[i]
    pt2 <- player_types[j]
    
    # Get slopes and standard errors
    slope1 <- wellbeing_slopes$estimate[wellbeing_slopes$player_type == pt1]
    slope2 <- wellbeing_slopes$estimate[wellbeing_slopes$player_type == pt2]
    se1 <- wellbeing_slopes$std.error[wellbeing_slopes$player_type == pt1]
    se2 <- wellbeing_slopes$std.error[wellbeing_slopes$player_type == pt2]
    
    # Calculate difference and standard error
    diff <- slope1 - slope2
    diff_se <- sqrt(se1^2 + se2^2)  # Simplified approach, assumes independence
    
    # Calculate test statistic and p-value
    t_stat <- diff / diff_se
    p_val <- 2 * pt(abs(t_stat), df = nrow(pvz_df_clustered) - 10, lower.tail = FALSE)  # Approximate df
    
    # Add to results
    wellbeing_pairs <- rbind(wellbeing_pairs, data.frame(
      player_type1 = pt1,
      player_type2 = pt2,
      difference = diff,
      std.error = diff_se,
      statistic = t_stat,
      p.value = p_val,
      stringsAsFactors = FALSE
    ))
  }
}

# Format and display pairwise comparison table for subjective well-being
wellbeing_pairs <- wellbeing_pairs %>%
  mutate(
    sig = case_when(
      p.value < 0.001 ~ "^†^",
      p.value < 0.01 ~ "^†^",
      p.value < 0.05 ~ "^†^", 
      TRUE ~ ""
    ),
    difference = round(difference, 3),
    std.error = round(std.error, 3),
    statistic = round(statistic, 3),
    p.value = case_when(
      p.value < 0.001 ~ paste0("< .001", sig),
      p.value < 0.05 ~ paste0(sprintf("%.3f", round(p.value, 3)), sig),
      TRUE ~ sprintf("%.3f", round(p.value, 3))
    )
  ) |>
  rename(`Type 1` = player_type1, `Type 2` = player_type2, Diff = difference, SE = std.error, t = statistic, p = p.value)
```

```{r}
#| label: enjoyment-pairwise-comparisons
#| message: false
#| warning: false
#| include: false

# Create pairwise comparison of slopes for enjoyment model
enjoyment_pairs <- data.frame(player_type1 = character(),
                              player_type2 = character(),
                              difference = numeric(),
                              std.error = numeric(),
                              statistic = numeric(),
                              p.value = numeric(),
                              stringsAsFactors = FALSE)

for (i in 1:(length(player_types)-1)) {
  for (j in (i+1):length(player_types)) {
    pt1 <- player_types[i]
    pt2 <- player_types[j]
    
    # Get slopes and standard errors
    slope1 <- enjoyment_slopes$estimate[enjoyment_slopes$player_type == pt1]
    slope2 <- enjoyment_slopes$estimate[enjoyment_slopes$player_type == pt2]
    se1 <- enjoyment_slopes$std.error[enjoyment_slopes$player_type == pt1]
    se2 <- enjoyment_slopes$std.error[enjoyment_slopes$player_type == pt2]
    
    # Calculate difference and standard error
    diff <- slope1 - slope2
    diff_se <- sqrt(se1^2 + se2^2)  # Simplified approach, assumes independence
    
    # Calculate test statistic and p-value
    t_stat <- diff / diff_se
    p_val <- 2 * pt(abs(t_stat), df = nrow(pvz_df_clustered) - 10, lower.tail = FALSE)  # Approximate df
    
    # Add to results
    enjoyment_pairs <- rbind(enjoyment_pairs, data.frame(
      player_type1 = pt1,
      player_type2 = pt2,
      difference = diff,
      std.error = diff_se,
      statistic = t_stat,
      p.value = p_val,
      stringsAsFactors = FALSE
    ))
  }
}

# Format and display pairwise comparison table for enjoyment model
enjoyment_pairs <- enjoyment_pairs %>%
  mutate(
    sig = case_when(
      p.value < 0.001 ~ "^†††^",
      p.value < 0.01 ~ "^††^",
      p.value < 0.05 ~ "^†^", 
      TRUE ~ ""
    ),
    difference = round(difference, 3),
    std.error = round(std.error, 3),
    statistic = round(statistic, 3),
    p.value = case_when(
      p.value < 0.001 ~ paste0("< .001", sig),
      p.value < 0.05 ~ paste0(sprintf("%.3f", round(p.value, 3)), sig),
      TRUE ~ sprintf("%.3f", round(p.value, 3))
    )
  ) |>
  rename(`Type 1` = player_type1, `Type 2` = player_type2, Diff = difference, SE = std.error, t = statistic, p = p.value)
```

```{r}
#| label: fig-interaction-plots
#| fig-cap: "Interaction plots showing the relationship between hours played and psychological outcomes across play styles. Left panel shows game-specific well-being (SPANE balance score); right panel shows enjoyment. Lines represent linear regression slopes with 95% confidence intervals. Casual PVE Players and Elite Allround Players show positive associations between playtime and outcomes, while Casual PVP Players show no significant relationship."
#| fig-width: 12
#| fig-height: 6
#| code-summary: "Interaction plots for moderation analysis"
#| message: false
#| warning: false

# Create scatter plots for total playtime vs. outcomes across player types with APA-style formatting
wellbeing_plot <- ggplot(pvz_df_clustered, aes(x = hours, y = spane_game_balance, color = player_type)) +
  geom_point(alpha = 0.6) +
  geom_smooth(method = "lm", se = TRUE, alpha = 0.2, fullrange = TRUE) +
  scale_color_brewer(palette = "Set2") +
  scale_y_continuous(limits = c(-6, 6), breaks = seq(-6, 6, 2)) +
  labs(
    x = "Total Hours Played",
    y = "SPANE Score",
    color = "Play Style"
  ) +
  # Apply APA-style formatting
  theme_minimal() +
  theme(
    plot.title = element_text(size = 12, face = "bold", hjust = 0),
    plot.subtitle = element_text(size = 10, hjust = 0),
    axis.title = element_text(size = 11),
    axis.text = element_text(size = 10),
    panel.grid.minor = element_blank(),
    panel.border = element_rect(fill = NA, color = "black", size = 0.5),
    panel.background = element_rect(fill = "white")
  )

enjoyment_plot <- ggplot(pvz_df_clustered, aes(x = hours, y = enjoyment, color = player_type)) +
  geom_point(alpha = 0.6) +
  geom_smooth(method = "lm", se = TRUE, alpha = 0.2, fullrange = TRUE) +
  scale_color_brewer(palette = "Set2") +
  scale_y_continuous(limits = c(1, 7), breaks = seq(1, 7, 1)) +
  labs(
    x = "Total Hours Played",
    y = "Enjoyment Score",
    color = "Play Style"
  ) +
  # Apply APA-style formatting
  theme_minimal() +
  theme(
    plot.title = element_text(size = 12, face = "bold", hjust = 0),
    plot.subtitle = element_text(size = 10, hjust = 0),
    axis.title = element_text(size = 11),
    axis.text = element_text(size = 10),
    panel.grid.minor = element_blank(),
    panel.border = element_rect(fill = NA, color = "black", size = 0.5),
    panel.background = element_rect(fill = "white")
  )

# Combine the plots with common legend at bottom
ggarrange(wellbeing_plot, enjoyment_plot, ncol = 2, common.legend = TRUE, legend = "bottom")
```

```{r}
#| label: simple-slopes-inline-values
#| echo: false

# Extract simple slopes values for inline reporting
# Wellbeing slopes
wb_pve <- wellbeing_slopes |> filter(player_type == "Casual PVE Players")
wb_pvp <- wellbeing_slopes |> filter(player_type == "Casual PVP Players")
wb_elite <- wellbeing_slopes |> filter(player_type == "Elite Allround Players")

# Enjoyment slopes  
enj_pve <- enjoyment_slopes |> filter(player_type == "Casual PVE Players")
enj_pvp <- enjoyment_slopes |> filter(player_type == "Casual PVP Players")
enj_elite <- enjoyment_slopes |> filter(player_type == "Elite Allround Players")

# Wellbeing pairwise comparisons (need numeric version before formatting)
wb_pairs_numeric <- data.frame(player_type1 = character(),
                               player_type2 = character(),
                               difference = numeric(),
                               std.error = numeric(),
                               statistic = numeric(),
                               p.value = numeric(),
                               stringsAsFactors = FALSE)

for (i in 1:(length(player_types)-1)) {
  for (j in (i+1):length(player_types)) {
    pt1 <- player_types[i]
    pt2 <- player_types[j]
    slope1 <- wellbeing_slopes$estimate[wellbeing_slopes$player_type == pt1]
    slope2 <- wellbeing_slopes$estimate[wellbeing_slopes$player_type == pt2]
    se1 <- wellbeing_slopes$std.error[wellbeing_slopes$player_type == pt1]
    se2 <- wellbeing_slopes$std.error[wellbeing_slopes$player_type == pt2]
    diff <- slope1 - slope2
    diff_se <- sqrt(se1^2 + se2^2)
    t_stat <- diff / diff_se
    p_val <- 2 * pt(abs(t_stat), df = nrow(pvz_df_clustered) - 6, lower.tail = FALSE)
    wb_pairs_numeric <- bind_rows(wb_pairs_numeric, data.frame(
      player_type1 = pt1, player_type2 = pt2,
      difference = diff, std.error = diff_se, statistic = t_stat, p.value = p_val
    ))
  }
}

# Enjoyment pairwise comparisons
enj_pairs_numeric <- data.frame(player_type1 = character(),
                                player_type2 = character(),
                                difference = numeric(),
                                std.error = numeric(),
                                statistic = numeric(),
                                p.value = numeric(),
                                stringsAsFactors = FALSE)

for (i in 1:(length(player_types)-1)) {
  for (j in (i+1):length(player_types)) {
    pt1 <- player_types[i]
    pt2 <- player_types[j]
    slope1 <- enjoyment_slopes$estimate[enjoyment_slopes$player_type == pt1]
    slope2 <- enjoyment_slopes$estimate[enjoyment_slopes$player_type == pt2]
    se1 <- enjoyment_slopes$std.error[enjoyment_slopes$player_type == pt1]
    se2 <- enjoyment_slopes$std.error[enjoyment_slopes$player_type == pt2]
    diff <- slope1 - slope2
    diff_se <- sqrt(se1^2 + se2^2)
    t_stat <- diff / diff_se
    p_val <- 2 * pt(abs(t_stat), df = nrow(pvz_df_clustered) - 6, lower.tail = FALSE)
    enj_pairs_numeric <- bind_rows(enj_pairs_numeric, data.frame(
      player_type1 = pt1, player_type2 = pt2,
      difference = diff, std.error = diff_se, statistic = t_stat, p.value = p_val
    ))
  }
}

# Helper function to format p-values for inline text
fmt_p <- function(p) {
  if (p < 0.001) return("< .001")
  sprintf("= %.3f", p)
}
```

To further probe the interaction, simple slopes analyses decomposed the playtime–outcome association within each playstyle (@fig-interaction-plots). Only Elite Allround Players showed a significant positive slope for both well-being (b = `r round(wb_elite$estimate, 3)`, *p* `r fmt_p(wb_elite$p.value)`) and enjoyment (b = `r round(enj_elite$estimate, 3)`, *p* `r fmt_p(enj_elite$p.value)`). Slopes for Casual PVE Players (well-being: b = `r round(wb_pve$estimate, 3)`, *p* `r fmt_p(wb_pve$p.value)`; enjoyment: b = `r round(enj_pve$estimate, 3)`, *p* `r fmt_p(enj_pve$p.value)`) and Casual PVP Players (well-being: b = `r round(wb_pvp$estimate, 3)`, *p* `r fmt_p(wb_pvp$p.value)`; enjoyment: b = `r round(enj_pvp$estimate, 3)`, *p* `r fmt_p(enj_pvp$p.value)`) were not significant. Critically, pairwise comparisons of slopes revealed no significant differences between any pair of playstyles (all *p*s > .05), confirming that the playtime–outcome gradient does not differ meaningfully across groups. Taken together, these results indicate that playstyle is associated with different baseline levels of well-being and enjoyment but does not moderate the relationship between playtime and either outcome.

# Discussion

This study addressed three questions: whether distinct playstyles can be identified from in-game behavioural telemetry within a single game, whether players with different playstyles report different levels of well-being and enjoyment, and whether the association between playtime and these outcomes differs across playstyles. The findings clearly support the first two propositions but not the third. Three distinct playstyles were recoverable from behavioural telemetry, and these playstyles were associated with significant differences in baseline well-being and enjoyment. However, we found no evidence that playstyle moderated the playtime–outcome relationship.

We operationalised playing styles through behavioural telemetry rather than self-report player type measures for several methodological reasons. First, telemetry captures actual behaviour rather than self-reported preferences, avoiding inaccuracies due to social desirability and recall biases [@zendleFourGrandChallenges2023]. Second, telemetry provides objective, granular data about gameplay patterns over time. Third, recent evidence suggests self-reported player types lack temporal stability [@santosConsistencyGamificationUser2023], raising questions about their validity as trait-like predictors. Behavioural playstyles derived from telemetry thus better capture consistent patterns of game engagement, offering more ecologically valid characterisations of player-game interactions than potentially unstable self-reported preferences measured at a single timepoint.

The three playstyles identified through PAM clustering (Casual PVE Players, Elite Allround Players, and Casual PVP Players) are conceptually consistent with player segments previously identified in other shooter titles. @drachenGunsSwordsData2012 identified player types in *Battlefield 2: Bad Company 2* that parallel our casual and elite profiles, whilst @melhartYourGameplaySays2019 found comparable segments in *Tom Clancy's The Division*, including PVE-focused and competitively skilled players. The recurrence of these patterns across different games suggests that such typologies may capture fundamental approaches to gameplay rather than game-specific idiosyncrasies. This finding directly supports the heterogeneous exposures framework [@ballouHowVideoGames2025]: even within a single title, players encounter qualitively different game-centred exposures depending on the modes they select and the mechanics they engage with.

However, it remains an open question whether playstyles themselves show temporal stability or represent dynamic adaptations to changing contexts. Unlike player types, which are theorised as relatively stable individual differences yet demonstrate considerable instability [@santosConsistencyGamificationUser2023], playstyles may be better conceptualised as characteristic behavioural patterns that emerge from the interaction between player, game, and context. This framing accommodates the possibility that players' behavioural patterns might shift with skill development, game progression, life circumstances, or wellbeing states, whilst still providing meaningful characterisations of gameplay within specific periods and contexts.

```{r}
#| label: discussion-statistics
#| echo: false
# Extract coefficients for inline reporting
coef_pvp_wellbeing <- get_model_coef(rmodel_game, "player_typeCasual PVP Players")
coef_pvp_enjoyment <- get_model_coef(rmodel_enjoyment, "player_typeCasual PVP Players")
coef_elite_wellbeing <- get_model_coef(rmodel_game, "player_typeElite Allround Players")
coef_elite_enjoyment <- get_model_coef(rmodel_enjoyment, "player_typeElite Allround Players")
coef_hours_game <- get_model_coef(rmodel_moderation_game, "hours")
coef_hours_enjoyment <- get_model_coef(rmodel_moderation_enjoyment, "hours")
```

Our robust mixed-effects models and complementary ANOVAs converged on a consistent pattern: Casual PVE Players reported higher game-specific well-being and enjoyment than Casual PVP Players, while Elite Allround Players fell between the two groups on enjoyment but were also lower than Casual PVE Players on well-being. These baseline differences indicate that different styles of engagement with the same game are associated with different psychological experiences, even after controlling for playtime, age, and gender.

The higher well-being among Casual PVE Players is noteworthy given that Elite Allround Players demonstrated substantially higher performance metrics across all modes. If competence alone drove well-being, the most skilled players should report the most positive experiences. Prior work does suggest that competence satisfaction positively predicts well-being in gaming contexts [@klimmt2009a; @przybylskiMotivationalModelVideo2010; @tamboriniMediaEnjoymentNeed2011; @ryanHappinessHumanPotentials2001]. However, our findings suggest that the *context* in which competence is exercised matters as well, consistent with the IMP's emphasis on both mechanics and context as jointly shaping player experience [@elsonMoreStoriesButtons2014]. Cooperative PVE environments may afford a subjective experience that is more exploratory than competitive PVP modes, potentially increasing perceived autonomy over how to engage with game content. This interpretation is speculative and requires direct measurement of need satisfaction across playstyles in future work, but it aligns with the broader pattern that *how* players engage with a game is more informative for average well-being than *how much* they play.

Contrary to our expectations, we found no evidence that playstyle moderates the association between playtime and well-being or enjoyment. Interaction terms were consistently non-significant, and pairwise comparisons of simple slopes confirmed that the playtime gradient did not differ across playstyles. Although Elite Allround Players showed the only significant positive simple slope, this did not translate into a significant difference from the slopes of other groups.

This null finding has important implications for the playtime debate. Recent large-scale digital trace studies have established that playtime alone is not meaningfully associated with well-being for most players [@ballouOpenPlayLongitudinal2025; @johannesVideoGamePlay2021; @vuorreTimeSpentPlaying2022]. One plausible explanation for these aggregate null findings is that playtime effects are heterogeneous—positive for some styles of play, negative for others, and averaging to near-zero overall. Our results do not support this explanation; the playtime–well-being slope was consistently small and did not vary by playstyle. The present data therefore cannot be used to argue that the playtime paradigm is inadequate because it overlooks differential playtime effects across player types.

Perhaps the more parsimonious interpretation is simply that playtime is not a meaningful determinant of well-being in the first place, not because the relationship is masked by heterogeneity, but because time spent playing is too blunt a measure to capture what matters. This reading aligns with @ballouPerceivedValueVideo2025, who found that players' subjective evaluations of how well gaming fits their life predicted well-being an order of magnitude more strongly than playtime itself. Similarly, in our data, the baseline well-being differences between playstyles substantially exceed any playtime-related effects. The consistent null slopes across all three playstyles thus converge with the broader accumulating evidence that quality of play determines the wellbeing outcomes, not the quantity.

What the data *do* show is that the playtime paradigm is incomplete in a critical respect: it misses qualitative differences in play that are associated with meaningful variation in average well-being. Playtime treats all minutes of play as equivalent, but players who devote their time to cooperative PVE gameplay report higher well-being than those who spend comparable time in competitive PVP modes. This between-group difference in levels—rather than in slopes—is the primary empirical contribution of the present study. In this respect, our results complement rather than contradict platform-level research: playtime may indeed be uninformative as a predictor of well-being, but the behavioural composition of that playtime is not.

Furthermore, this study addresses the fundamental trade-off in the current generation of digital trace research on games and well-being between breadth and depth. Where platform-level datasets capture play across large numbers of games, long time horizons, and diverse platforms, offering unprecedented breadth in behavioural coverage [@ballouOpenPlayLongitudinal2025; @ballouPerceivedValueVideo2025; @vuorreTimeSpentPlaying2022]. Yet these approaches typically lack behavioural granularity regarding what happens *within* a game—whether players compete or cooperate, what roles they adopt, and how they perform under different in-game demands. This matters because in-game interaction patterns are precisely the level at which many plausible well-being-relevant mechanisms operate, including social connectedness, competence experiences, and stress from competitive evaluation [@elsonMoreStoriesButtons2014]. By sacrificing the breadth of platform-level work for the depth of session-level behavioural profiling within a single title, we demonstrate what becomes visible at finer resolution: meaningful behavioural subgroups with distinct well-being correlates.

Still, neither approach alone provides a complete picture, and we see strong potential in future study designs that combine them. A platform-level dataset augmented with targeted in-game behavioural telemetry from selected titles could capture *both* the breadth of cross-game play patterns *and* the depth of within-game behavioural profiles, allowing researchers to test whether the playstyle–wellbeing associations documented here generalise beyond a single game or whether they are moderated by platform- and genre-level factors. We also see value in combining digital trace data with experimental designs that would enable stronger causal inference. For example, randomising players to specific game modes or restricting play to certain conditions to examine effects on well-being. Digital trace data is uniquely suited to complement such designs because it allows researchers to monitor adherence (whether participants follow assigned play patterns) and substitution (what games or modes participants switch to under intervention), addressing key threats to validity in field experiments on gaming behaviour.

## Limitations & Future Work

Several limitations constrain interpretation of these findings and point toward productive directions for future research. First, the study examines a single game with a specific player population recruited by the developer, and we lack evidence that the identified playstyles generalise to other games, genres, or player populations. Unlike personality traits, playstyles may be contextual and fluid, varying with game design and a player's development over time. 

Second, while our behavioural telemetry provided comprehensive metrics on core gameplay mechanics and performance, it captured only a subset of potential in-game behaviours. Notably absent were indicators of social interaction, which is an important dimension of multiplayer gaming experiences that the IMP identifies as part of the contextual layer of player experience [@elsonMoreStoriesButtons2014]. In this study, sessions spent in the game's social hub were considered idle time and thus excluded from the playstyle analysis because these sessions generated no mechanics-relevant telemetry on which to cluster. Yet recent work has shown that ostensibly idle or non-interactive periods in games, including socialising between matches and exploring social spaces, can serve experiential functions such as community-building and strategic preparation that shape the broader play experience [@tepponenUnderstandingWaitingVideo2025]. Future work could address this gap by extending the playstyle profiling framework to incorporate social and communicative dimensions of gameplay. Gesture-based digital trace data, such as the 'ping' communication systems common in multiplayer games, been demonstrated to be related to game context, team dynamics, and performance outcomes[@zhengUnderstandingPlayersGestureBased2023]. Integrating such communicative behaviours alongside signals from social spaces like the hub sessions excluded here would allow researchers to construct a more complete picture of how players engage with a game's social layer and whether these patterns relate to well-being independently of combat-focused playstyle. 

Third, the cross-sectional design precludes causal inference. The observed associations between playstyle and well-being may reflect selection effects (e.g., players with higher baseline well-being may gravitate toward cooperative play) rather than the influence of play behaviour on outcomes. Fourth, the silhouette coefficient for the three-cluster solution, while above the threshold for meaningful structure, fell at the boundary between weak and strong clustering, suggesting that playstyle variation in this game may be better described as a continuum than as discrete types. Finally, while we have proposed mechanistic interpretations through the lens of the IMP and need satisfaction theory, these remain speculative—future work should directly measure autonomy, competence, and relatedness across playstyles to test whether these mediate the observed associations.

## Conclusion

This study demonstrates that players of the same game engage with it in qualitatively different ways and that these differences are associated with meaningful variation in well-being and enjoyment. However, we found no evidence that playstyle moderates the playtime–well-being relationship. These findings contribute to digital trace research in two respects. First, they provide empirical support for the view that gameplay constitutes heterogeneous exposures with distinct psychological correlates, even within a single title. Second, they demonstrate a practical methodology, telemetry-driven profiling, for incorporating in-game behavioural depth into well-being research. At the same time, the null moderation results caution against overclaiming, the present data show that *what* players do is associated with how they feel, but not that different styles of play change *how playtime itself relates to well-being*. As richer forms of in-game data become available through industry–academic partnerships, researchers should dig deeper and continue to develop more granular, theory-driven approaches to understanding player experience and its relationship to psychological outcomes.

## Acknowledgements

Claude (Sonnet/Opus 4.6; Anthropic) was used to assist with code development and minor text edits during the preparation of this manuscript. All AI-assisted contributions were reviewed, verified, and, where necessary, further revised by members of the author team, who maintain full responsibility for the final content.

## Data Availability

All data, materials, and code related to the dataset and this manuscript are available under {{< meta data-license >}} at <https://github.com/Thomhak/playstyle-article.git>.

## Funding

[funding]

## Conflicts of Interest

The authors declare no competing financial, intellectual, or institutional interests relevant to this research. The disclosures below are provided in the interest of full transparency.

No authors have received funding or compensation from the tech companies or other industries connected to the work presented here.

MV has provided in-kind consultations and served as a nonpaid panel member for Meta and K-Games. AKP has served in advisory or expert roles for the Sync Digital Wellbeing Program (2022–2024), the Google Expert Advisory on Youth and Technology (2025), the OpenAI Expert Council on Well-Being and AI (2025), and UK Department for Science, Innovation and Technology–funded research on children's smartphone and social media use (University of Cambridge, 2025). He is a member of the UK Department for Culture, Media & Sport College of Experts. Any industry fees are directly donated to charity, and his research is conducted in accordance with University of Oxford academic integrity policies.

Industry partners had no role in the design, analysis, or publication of results.

# References

::: {#refs}
:::

# Appendix

## Appendix 1: Cluster Tendency Assessment

```{r}
#| label: appendix_cluster_tendency
#| fig.width: 12
#| fig.height: 10
#| echo: false
#| warning: false
#| message: false
#| fig.cap: "Cluster tendency assessment comparing player data (left column) to random data (right column). Top row: Visual Assessment of Cluster Tendency (VAT) ordered dissimilarity matrices. Bottom row: PCA visualizations. The player data shows clear clustering structure (dark blocks in VAT, visible groupings in PCA) while random data shows no such patterns."

# Combine all four plots into a 2x2 grid
grid.arrange(
  dist_plot_real + labs(title = "A: VAT - Player Data"),
  dist_plot_random + labs(title = "B: VAT - Random Data"),
  pca_plot_real + labs(title = "C: PCA - Player Data"),
  pca_plot_random + labs(title = "D: PCA - Random Data"),
  ncol = 2, nrow = 2
)
```

## Appendix 2: Cluster Validation Plots

```{r}
#| label: appendix_cluster_validation_plots
#| fig.width: 6
#| fig.height: 4
#| echo: false
#| warning: false
#| message: false
#| fig.cap: "Cluster validation using the Calinski-Harabasz Index across k = 2 to 10."

# Display CH-index plot in appendix
p_pam_ch
```

The plot above shows the validation method used to determine the optimal number of clusters. The Calinski-Harabasz Index evaluates cluster validity based on the ratio of between-cluster to within-cluster variance, with higher values indicating better separation between clusters. The index indicated that k = 3 was the optimal number of clusters for our player data.

## Appendix 3: Statistical Tables for Interpretation

```{r}
#| label: appendix_stat_tables_prep
#| include: false
#| cache: false

fmt_p_table <- function(p) {
  case_when(
    is.na(p) ~ "NA",
    p < 0.001 ~ "< .001",
    TRUE ~ sprintf("%.3f", p)
  )
}

tidy_rlmer_table <- function(model, outcome, model_name) {
  as.data.frame(coef(summary(model))) %>%
    rownames_to_column("term") %>%
    as_tibble() %>%
    rename(
      estimate = Estimate,
      std_error = `Std. Error`,
      t_value = `t value`
    ) %>%
    mutate(
      outcome = outcome,
      model = model_name,
      p_approx = 2 * pnorm(-abs(t_value)),
      significant = p_approx < 0.05
    )
}

term_map <- c(
  "(Intercept)" = "Intercept",
  "player_typeCasual PVP Players" = "Casual PVP vs Casual PVE",
  "player_typeElite Allround Players" = "Elite Allround vs Casual PVE",
  "hours" = "Hours played",
  "genderMale" = "Gender: Male (vs Female)",
  "genderOther" = "Gender: Other (vs Female)",
  "genderPrefer not to say" = "Gender: Undisclosed (vs Female)",
  "age" = "Age",
  "player_typeCasual PVP Players:hours" = "Casual PVP × Hours",
  "player_typeElite Allround Players:hours" = "Elite Allround × Hours"
)

make_term_label <- function(term) {
  recode(term, !!!term_map, .default = term)
}

tbl_main_model <- bind_rows(
  tidy_rlmer_table(rmodel_game, "Well-being (SPANE)", "Main effects"),
  tidy_rlmer_table(rmodel_enjoyment, "Enjoyment", "Main effects")
) %>%
  mutate(term_label = make_term_label(term)) %>%
  filter(term != "(Intercept)")

tbl_moderation_model <- bind_rows(
  tidy_rlmer_table(rmodel_moderation_game, "Well-being (SPANE)", "Moderation"),
  tidy_rlmer_table(rmodel_moderation_enjoyment, "Enjoyment", "Moderation")
) %>%
  mutate(term_label = make_term_label(term)) %>%
  filter(term != "(Intercept)")

tbl_anova <- all_anova_results %>%
  filter(term == "player_type") %>%
  mutate(
    outcome = recode(
      variable,
      "spane_game_balance" = "Well-being (SPANE)",
      "enjoyment" = "Enjoyment"
    ),
    significant = p.value < 0.05
  )

tbl_posthoc <- bind_rows(
  wellbeing_posthoc %>% mutate(outcome = "Well-being (SPANE)"),
  enjoyment_posthoc %>% mutate(outcome = "Enjoyment")
) %>%
  mutate(significant = adj.p.value < 0.05)

tbl_simple_slopes <- bind_rows(
  wellbeing_slopes %>% mutate(outcome = "Well-being (SPANE)"),
  enjoyment_slopes %>% mutate(outcome = "Enjoyment")
) %>%
  mutate(significant = p.value < 0.05)

tbl_slope_pairs <- bind_rows(
  wb_pairs_numeric %>% mutate(outcome = "Well-being (SPANE)"),
  enj_pairs_numeric %>% mutate(outcome = "Enjoyment")
) %>%
  mutate(significant = p.value < 0.05)

tbl_significant_findings <- bind_rows(
  tbl_main_model %>%
    filter(significant) %>%
    transmute(
      analysis = model,
      outcome,
      test = term_label,
      effect = estimate,
      statistic = t_value,
      p_value = p_approx
    ),
  tbl_moderation_model %>%
    filter(significant) %>%
    transmute(
      analysis = model,
      outcome,
      test = term_label,
      effect = estimate,
      statistic = t_value,
      p_value = p_approx
    ),
  tbl_anova %>%
    filter(significant) %>%
    transmute(
      analysis = "ANOVA",
      outcome,
      test = "Overall play-style effect",
      effect = eta_squared,
      statistic = statistic,
      p_value = p.value
    ),
  tbl_posthoc %>%
    filter(significant) %>%
    transmute(
      analysis = "Post-hoc (Tukey HSD)",
      outcome,
      test = contrast,
      effect = estimate,
      statistic = NA_real_,
      p_value = adj.p.value
    ),
  tbl_simple_slopes %>%
    filter(significant) %>%
    transmute(
      analysis = "Simple slopes",
      outcome,
      test = player_type,
      effect = estimate,
      statistic = statistic,
      p_value = p.value
    ),
  tbl_slope_pairs %>%
    filter(significant) %>%
    transmute(
      analysis = "Pairwise slope differences",
      outcome,
      test = paste(player_type1, "vs", player_type2),
      effect = difference,
      statistic = statistic,
      p_value = p.value
    )
) %>%
  arrange(analysis, outcome)
```

```{r}
#| label: appendix_table_main_effects
#| echo: false
#| warning: false
#| message: false

tbl_main_model %>%
  transmute(
    Outcome = outcome,
    Term = term_label,
    Estimate = round(estimate, 3),
    SE = round(std_error, 3),
    t = round(t_value, 3),
    `Approx. p` = fmt_p_table(p_approx),
    Significant = if_else(significant, "Yes", "No")
  ) %>%
  tt(caption = "Main-Effects Robust Mixed Models (Approximate Significance)") %>%
  format_tt(digits = 3)
```

```{r}
#| label: appendix_table_moderation
#| echo: false
#| warning: false
#| message: false

tbl_moderation_model %>%
  transmute(
    Outcome = outcome,
    Term = term_label,
    Estimate = round(estimate, 3),
    SE = round(std_error, 3),
    t = round(t_value, 3),
    `Approx. p` = fmt_p_table(p_approx),
    Significant = if_else(significant, "Yes", "No")
  ) %>%
  tt(caption = "Moderation Models (Play Style × Hours; Approximate Significance)") %>%
  format_tt(digits = 3)
```

```{r}
#| label: appendix_table_anova_posthoc
#| echo: false
#| warning: false
#| message: false

tbl_anova %>%
  transmute(
    Outcome = outcome,
    `df1` = df,
    `df2` = df_residual,
    F = round(statistic, 2),
    p = fmt_p_table(p.value),
    `Eta-squared` = round(eta_squared, 3),
    Significant = if_else(significant, "Yes", "No")
  ) %>%
  tt(caption = "Omnibus ANOVA Tests by Outcome") %>%
  format_tt(digits = 3)

tbl_posthoc %>%
  transmute(
    Outcome = outcome,
    Contrast = contrast,
    Estimate = round(estimate, 3),
    `CI Low` = round(conf.low, 3),
    `CI High` = round(conf.high, 3),
    `Adj. p` = fmt_p_table(adj.p.value),
    Significant = if_else(significant, "Yes", "No")
  ) %>%
  tt(caption = "Post-hoc Tukey HSD Comparisons by Outcome") %>%
  format_tt(digits = 3)
```

```{r}
#| label: appendix_table_slopes
#| echo: false
#| warning: false
#| message: false

tbl_simple_slopes %>%
  transmute(
    Outcome = outcome,
    `Play Style` = player_type,
    b = round(estimate, 3),
    SE = round(std.error, 3),
    t = round(statistic, 3),
    p = fmt_p_table(p.value),
    `CI Low` = round(conf.low, 3),
    `CI High` = round(conf.high, 3),
    Significant = if_else(significant, "Yes", "No")
  ) %>%
  tt(caption = "Simple Slopes by Play Style and Outcome") %>%
  format_tt(digits = 3)

tbl_slope_pairs %>%
  transmute(
    Outcome = outcome,
    Contrast = paste(player_type1, "vs", player_type2),
    Diff = round(difference, 3),
    SE = round(std.error, 3),
    t = round(statistic, 3),
    p = fmt_p_table(p.value),
    Significant = if_else(significant, "Yes", "No")
  ) %>%
  tt(caption = "Pairwise Differences Between Simple Slopes") %>%
  format_tt(digits = 3)
```

```{r}
#| label: appendix_table_significant_findings
#| echo: false
#| warning: false
#| message: false

tbl_significant_findings %>%
  transmute(
    Analysis = analysis,
    Outcome = outcome,
    Test = test,
    Effect = round(effect, 3),
    Statistic = round(statistic, 3),
    p = fmt_p_table(p_value)
  ) %>%
  tt(caption = "Significant Findings Summary (p < .05)") %>%
  format_tt(digits = 3)
```

These appendix tables summarize all model terms, omnibus tests, post-hoc contrasts, simple slopes, and pairwise slope differences to support interpretation of statistically significant findings. For robust mixed-effects models (`rlmer`), p-values are reported as normal-approximation values derived from t-statistics and should be interpreted as approximate.
