---
title: "Player Playstyle Analysis"
author: "Thomas Hakman"
format: 
  html:
    toc: true
    toc-depth: 3
    code-fold: true
    theme: flatly
    fig-width: 8
    fig-height: 6
execute:
  echo: true
  message: false
  warning: false
  cache: true
editor: visual
bibliography: references.bib
---

```{r setup}
#| label: setup
#| include: false

# Set global options
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  cache = TRUE,
  fig.path = "figures/"
)
```

# Introduction

This document analyzes player telemetry and survey data from Plants vs. Zombies: Battle for Neighborville. The analysis aims to understand the relationship between player playstyles, game performance metrics, and well-being outcomes.

# Data Preparation

## Load Packages

```{r packages}
#| label: load-packages

library(pacman)
p_load(
  readxl,         # Reading Excel files
  knitr,          # Document generation
  here,           # File path management
  lubridate,      # Date handling
  scales,         # Visualization scales
  janitor,        # Data cleaning
  tidyverse,      # Data wrangling and visualization
  oii,            # Oxford Internet Institute functions
  psych,          # Psychological analysis tools
  cluster,        # Clustering algorithms
  factoextra,     # Cluster visualization and validation
  lme4,           # Mixed models
  sjPlot          # Tables and plots for models
)
```

## Load Raw Data

The data comes from a study conducted by EA and the Oxford Internet Institute. Players completed a survey on well-being and motivations, and EA provided telemetry data on their gameplay.

```{r raw-data-paths}
#| label: raw-data-paths

# Store all file paths in a central location for easy updates
data_paths <- list(
  survey = here("data/raw/ea/Oxford PvZ - Pilot Wave + Wave 1 Download.xlsx"),
  telemetry_files = list.files(
    here("data/raw/ea"), 
    pattern = "OII_PVZ_", 
    full.names = TRUE
  )
)

# Function to read both waves of telemetry data into a single table
read_both_waves <- function(stub = "OII_PVZ_Authentications", ...) {
  # Filenames of wave 1 and 2 authentications files
  list.files(
    here("data/raw/ea"), pattern = stub, full.names = TRUE
  ) %>% 
    # Put wave 1 first
    rev() %>% 
    # Read into a list of data frames
    map(read_csv, ...) %>% 
    # Bind into one data frame with wave ID
    bind_rows(.id = "wave")
}
```

## Load Survey Data

```{r survey-load}
#| label: survey-load

survey <- read_excel(data_paths$survey)

survey <- survey %>% 
  mutate(
    # Ensure date data type
    date = mdy_hm(date),
    # Additional date variable for filtering telemetry
    start_date = date
  ) %>% 
  # Rename ID column to match telemetry files
  rename(player_id = UID)
```

## Clean Survey Data

```{r survey-clean}
#| label: survey-clean

# Rename variables with meaningful names
survey <- 
  survey %>% 
  select(
    # meta-info
    player_id,
    start_date, date,
    survey_duration = qtime,
    of_age = Q2,
    consent = Q3,
    
    # demographics
    played = Q1,
    country = hCountry,
    survey_language = PRELANG,
    age = D1,
    gender = D2,
    gender_other = D2r3oe,
    experience = D3,
    
    # well-being
    spane_ = starts_with("A1"), # general well-being in the two weeks
    straightliner_spane = hA1Straight,
    
    # motivations
    played_with_others = NSat1, # filter question for relatedness items
    autonomy_ = NSat2r1:NSat2r3,
    competence_ = NSat2r4:NSat2r6,
    relatedness_ = NSat2r7:NSat2r9,
    enjoyment_ = NSat2r10:NSat2r13,
    extrinsic_ = NSat2r14:NSat2r17,
    straightliner_motivations = hNSat2Straight,
    
    # self-reported play
    active_play_hours = GT1c1,
    active_play_minutes = GT1c2,
    within_estimate = GT2,
    between_estimate = GT3,
    spane_game_ = starts_with("GT4")
  )

# Assign proper variable types and factor levels
survey <- survey %>%   
  mutate(
    across(
      c(
        player_id,
        of_age:survey_language,
        gender,
        played_with_others
      ),
      as.factor
    ),
    
    across(
      c(of_age:played, played_with_others),
      ~ fct_recode(
        .x,
        "Yes" = "1",
        "No" = "2"
      )
    ),
    
    gender = fct_recode(
      gender,
      "Male" = "1",
      "Female" = "2",
      "Other" = "3",
      "Prefer not to say" = "4"
    ),
    
    country = fct_recode(
      country,
      "US" = "1",
      "UK" = "2",
      "Canada" = "5"
    ),
    
    survey_language = fct_recode(
      survey_language,
      "English" = "1",
      "French" = "2"
    )
  )

# Apply reverse scoring for appropriate items
survey <- survey %>% 
  mutate(
    across(
      c(
        starts_with("spane"), # 1 was never
        between_estimate, # those estimates also had 1: much more than typically
        within_estimate,
        relatedness_3,
        enjoyment_2,
        enjoyment_3
      ),
      ~ 8 - .x
    )
  )

# Create indicators for straightliners
survey <- survey %>% 
  mutate(across(contains("straightliner"), ~.x==1))
```

## Create Survey Scales

```{r survey-scales}
#| label: survey-scales

# General SPANE scale (well-being)
survey <- 
  survey %>% 
  mutate(
    spane_positive = rowMeans(
      select(
        .,
        spane_1,
        spane_3,
        spane_5,
        spane_7,
        spane_10,
        spane_12
      ),
      na.rm = TRUE
    ),
    spane_negative = rowMeans(
      select(
        .,
        spane_2,
        spane_4,
        spane_6,
        spane_8,
        spane_9,
        spane_11
      ),
      na.rm = TRUE
    ),
    spane_balance = spane_positive - spane_negative
  )

# Motivation scales
survey <- survey %>% 
  mutate(
    autonomy = rowMeans(
      select(., starts_with("autonomy")), na.rm = TRUE
    ),
    competence = rowMeans(
      select(., starts_with("competence")), na.rm = TRUE
    ),
    relatedness = rowMeans(
      select(., starts_with("relatedness")), na.rm = TRUE
    ),
    enjoyment = rowMeans(
      select(., starts_with("enjoyment")), na.rm = TRUE
    ),
    extrinsic = rowMeans(
      select(., starts_with("extrinsic")), na.rm = TRUE
    )
  )

# SPANE because of playing game
survey <- survey %>% 
  mutate(
    spane_game_positive = rowMeans(
      select(
        .,
        spane_game_1,
        spane_game_3,
        spane_game_5,
        spane_game_7,
        spane_game_10,
        spane_game_12
      ),
      na.rm = TRUE
    ),
    spane_game_negative = rowMeans(
      select(
        .,
        spane_game_2,
        spane_game_4,
        spane_game_6,
        spane_game_8,
        spane_game_9,
        spane_game_11
      ),
      na.rm = TRUE
    ),
    spane_game_balance = spane_game_positive - spane_game_negative
  )

# Hours of estimated play
survey <- survey %>%
  mutate(active_play_minutes = active_play_minutes / 60) %>% 
  mutate(active_play = rowSums(select(., starts_with("active_play")), na.rm = TRUE))

# Save processed survey data
write_rds(survey, here("data/processed/survey_processed.rds"))
```

# Telemetry Data Processing

## Load Game Time Data

```{r game-time-load}
#| label: game-time-load

game_time_raw <- read_both_waves("OII_PVZ_Game_Time") %>%
  # Only include survey participants 
  filter(player_id %in% survey$player_id)
```

## Process Game Time Data

```{r game-time-process}
#| label: game-time-process

# Handle missing start times
game_time <- game_time_raw %>% 
  mutate(
    # If start time is missing, set to end time (will result in 0 duration)
    game_start_time = coalesce(game_start_time, game_end_time)
  ) %>%
  # Collapse rows by session
  group_by(player_id, game_session) %>% 
  mutate(
    game_start_time = min(game_start_time),
    game_end_time = max(game_end_time),
  ) %>% 
  # Take the first of each (identical) person-session combo
  slice(1) %>% 
  ungroup()

# Filter to sessions within 14 days before survey
game_time_2week <- game_time %>% 
  left_join(select(survey, player_id, start_date)) %>% 
  filter(game_start_time >= (start_date - days(14)))

# Save processed game time data
write_rds(game_time_2week, here("data/processed/game_time_2week.rds"))
```

## Load Authentication Data

```{r auth-load}
#| label: auth-load

authentications_raw <- read_both_waves("OII_PVZ_Authentications") %>%
  # Only include survey participants
  filter(player_id %in% survey$player_id)

# Filter to 14 days before survey
authentications_2week <- authentications_raw %>% 
  left_join(select(survey, player_id, start_date)) %>% 
  filter(date >= (as_date(start_date) - days(14)))

# Count authentications per player
authentications_players <- authentications_2week %>% 
  count(player_id, name = "authentications")

# Save processed authentication data
write_rds(authentications_players, here("data/processed/authentications_players.rds"))
```

## Load Character Performance Data

```{r character-load}
#| label: character-load

characters_raw <- read_both_waves("OII_PVZ_Character") %>%
  # Only include survey participants
  filter(player_id %in% survey$player_id)

# Add session timing info and filter to 14 days before survey
characters_2week <- characters_raw %>%
  left_join(
    game_time_2week %>% 
      select(player_id, game_session, game_start_time, game_end_time),
    by = c("player_id", "game_session")
  ) %>%
  left_join(select(survey, player_id, start_date)) %>% 
  filter(game_start_time >= (start_date - days(14)))

# Save for further processing
write_rds(characters_2week, here("data/processed/characters_2week.rds"))
```

# Game Mode Classification

```{r game-mode-classification}
#| label: game-mode-classification

# Create game mode classification variables
characters_2week <- characters_2week %>%
  mutate(
    # Social hub sessions (no combat)
    social_hub = case_when(
      game_level_key == "dsub_social_terrain_summer" & is.na(score) ~ 1,
      TRUE ~ 0
    ),
    
    # PvE sessions
    pve_session = case_when(
      game_level_key %in% c(
        "dsub_region_cheesemines", 
        "dsub_region_dreadwood", 
        "dsub_region_towncenter"
      ) ~ 1,
      TRUE ~ 0
    ),
    
    # PvP sessions
    pvp_session = case_when(
      game_level_key %in% c(
        "dsub_rush_dreadwood", "dsub_rush_towncenter", 
        "dsub_tvm_cheesemines", "dsub_tvm_zrefinery", 
        "dsub_tvm_suburbs", "dsub_tvm_dreadwood", 
        "dsub_rush_farm", "dsub_rush_cheesemines", 
        "dsub_rush_shoreline", "dsub_rush_resort", 
        "dsub_4v4_battlearena_v2", "dsub_4v4_superarena"
      ) ~ 1,
      # Giddy Park with combat data is also PvP
      game_level_key == "dsub_social_terrain_summer" & !is.na(score) ~ 1,
      TRUE ~ 0
    )
  )

# Remove sessions without performance data (mainly social hub)
characters_2week_nona <- characters_2week %>% 
  drop_na(total_kill_count:shots_hit)

# Save processed data
write_rds(characters_2week, here("data/processed/characters_with_modes.rds"))
write_rds(characters_2week_nona, here("data/processed/characters_with_modes_nona.rds"))
```

# Game Time Calculations

```{r game-time-calculations}
#| label: game-time-calculations

# Calculate total game time per player
game_time_players_total <- characters_2week %>% 
  mutate(duration = game_end_time - game_start_time) %>% 
  mutate(seconds = as.numeric(duration)) %>% 
  group_by(player_id) %>% 
  summarise(seconds = sum(seconds)) %>%
  mutate(hours = seconds / 60 / 60) %>% 
  ungroup()

# Calculate PVP game time per player
game_time_players_pvp <- characters_2week %>% 
  filter(pvp_session == 1) %>%
  mutate(duration = game_end_time - game_start_time) %>% 
  mutate(pvp_seconds = as.numeric(duration)) %>% 
  group_by(player_id) %>% 
  summarise(pvp_seconds = sum(pvp_seconds)) %>% 
  mutate(pvp_hours = pvp_seconds / 60 / 60) %>% 
  ungroup()

# Calculate PVE game time per player  
game_time_players_pve <- characters_2week %>% 
  filter(pve_session == 1) %>%
  mutate(duration = game_end_time - game_start_time) %>% 
  mutate(pve_seconds = as.numeric(duration)) %>% 
  group_by(player_id) %>% 
  summarise(pve_seconds = sum(pve_seconds)) %>% 
  mutate(pve_hours = pve_seconds / 60 / 60) %>% 
  ungroup()

# Calculate social hub time per player
game_time_players_social <- characters_2week %>% 
  filter(social_hub == 1) %>%
  mutate(duration = game_end_time - game_start_time) %>% 
  mutate(social_seconds = as.numeric(duration)) %>% 
  group_by(player_id) %>% 
  summarise(social_seconds = sum(social_seconds)) %>% 
  mutate(social_hours = social_seconds / 60 / 60) %>% 
  ungroup()

# Combine game time data
game_time_players <- purrr::reduce(
  list(
    game_time_players_total, 
    game_time_players_social, 
    game_time_players_pve, 
    game_time_players_pvp
  ),
  left_join, by = "player_id"
)

# Replace NA values with 0 (no time spent in that mode)
game_time_players[is.na(game_time_players)] <- 0

# Save processed game time data
write_rds(game_time_players, here("data/processed/game_time_players.rds"))
```

# Performance Metrics Calculation

```{r performance-calculations}
#| label: performance-calculations

# Calculate overall performance metrics per player
characters_players <- characters_2week %>%
  group_by(player_id) %>% 
  summarise(across(total_kill_count:shots_hit, sum, na.rm = TRUE)) %>% 
  ungroup()

# Calculate game sessions per player
game_sessions_total <- characters_2week %>% 
  count(player_id) %>%
  rename(n_total_game_sessions = n)

# Calculate sessions by game mode
game_sessions <- characters_2week %>% 
  group_by(player_id) %>% 
  summarise(across(social_hub:pvp_session, sum, na.rm = TRUE)) %>% 
  ungroup()

# PvP-specific performance metrics
pvp_telemetry <- characters_2week %>%
  filter(pvp_session == 1) %>%
  group_by(player_id) %>% 
  summarise(across(total_kill_count:shots_hit, sum, na.rm = TRUE)) %>%
  rename_with(~ paste0("pvp_", .), -player_id)

# PvE-specific performance metrics
pve_telemetry <- characters_2week %>%
  filter(pve_session == 1) %>%
  group_by(player_id) %>% 
  summarise(across(total_kill_count:shots_hit, sum, na.rm = TRUE)) %>%
  rename_with(~ paste0("pve_", .), -player_id)

# Combine all telemetry data
telemetry_players <- purrr::reduce(
  list(
    characters_players,
    game_time_players,
    game_sessions,
    game_sessions_total,
    pvp_telemetry,
    pve_telemetry
  ),
  left_join, by = "player_id"
)

# Replace NA values with 0
telemetry_players[is.na(telemetry_players)] <- 0

# Save processed telemetry data
write_rds(telemetry_players, here("data/processed/telemetry_players.rds"))
```

# Derived Performance Metrics

```{r derived-metrics}
#| label: derived-metrics

# Calculate derived performance metrics
telemetry_players <- telemetry_players %>%
  # Handle divisions by zero for K/D ratio
  mutate(
    across(
      c(total_death_count, pvp_total_death_count, pve_total_death_count),
      ~ ifelse(.x == 0, 1, .x)
    )
  ) %>%
  # K/D ratios
  mutate(
    total_kd = total_kill_count / total_death_count,
    pve_kd = pve_total_kill_count / pve_total_death_count,
    pvp_kd = pvp_total_kill_count / pvp_total_death_count
  ) %>%
  # Hit accuracy ratios
  mutate(
    total_hit_accuracy = shots_hit / shots_fired,
    pve_hit_accuracy = pve_shots_hit / pve_shots_fired,
    pvp_hit_accuracy = pvp_shots_hit / pvp_shots_fired
  ) %>%
  # Critical hit ratios
  mutate(
    total_crit_ratio = critical_hit_count / shots_hit,
    pve_crit_ratio = pve_critical_hit_count / pve_shots_hit,
    pvp_crit_ratio = pvp_critical_hit_count / pvp_shots_hit
  ) %>%
  # Damage per game
  mutate(
    total_average_damage = damage_dealt / n_total_game_sessions,
    pve_average_damage = pve_damage_dealt / pve_session,
    pvp_average_damage = pvp_damage_dealt / pvp_session
  ) %>%
  # Damage per second
  mutate(
    total_dps = damage_dealt / seconds,
    pve_dps = pve_damage_dealt / pve_seconds,
    pvp_dps = pvp_damage_dealt / pvp_seconds
  ) %>%
  # Score per game
  mutate(
    total_average_score = score / n_total_game_sessions,
    pve_average_score = pve_score / pve_session,
    pvp_average_score = pvp_score / pvp_session
  ) %>%
  # Score per second
  mutate(
    total_sps = score / seconds,
    pve_sps = pve_score / pve_seconds,
    pvp_sps = pvp_score / pvp_seconds
  )

# Replace NaN values with 0
telemetry_players <- telemetry_players %>%
  mutate(across(everything(), ~ifelse(is.nan(.x), 0, .x)))

# Save processed metrics
write_rds(telemetry_players, here("data/processed/telemetry_with_metrics.rds"))
```

# Combine Survey and Telemetry Data

```{r merge-data}
#| label: merge-data

# Merge survey with telemetry data
pvz_df <- survey %>% 
  left_join(telemetry_players, by = "player_id")

# Calculate ratio metrics
pvz_df <- pvz_df %>%
  mutate(
    ratio_social_hours = social_hours / hours,
    ratio_pve_hours = pve_hours / hours,
    ratio_pvp_hours = pvp_hours / hours
  )

# Remove participants with very low playtime (< 2 minutes)
pvz_df_clean <- pvz_df %>%
  filter(hours > 0.03)

# Save final dataset
write_rds(pvz_df_clean, here("data/processed/pvz_complete.rds"))
```

# Exploratory Data Analysis

## Demographic Statistics

```{r demographic-stats}
#| label: demographic-stats

# Demographic summary
demographic_summary <- pvz_df_clean %>%
  summarise(
    n_participants = n(),
    mean_age = mean(age, na.rm = TRUE),
    sd_age = sd(age, na.rm = TRUE),
    median_age = median(age, na.rm = TRUE),
    n_male = sum(gender == "Male", na.rm = TRUE),
    pct_male = mean(gender == "Male", na.rm = TRUE) * 100,
    n_female = sum(gender == "Female", na.rm = TRUE),
    pct_female = mean(gender == "Female", na.rm = TRUE) * 100
  )

knitr::kable(demographic_summary, caption = "Demographic Summary")
```

## Game Time Distribution

```{r gametime-distribution}
#| label: gametime-distribution

# Plot distribution of total game time
ggplot(pvz_df_clean, aes(x = hours)) +
  geom_histogram(binwidth = 2, fill = "steelblue", color = "white") +
  labs(
    title = "Distribution of Total Play Time",
    x = "Hours (2-week period)",
    y = "Count"
  ) +
  theme_minimal()

# Plot distribution of game time by mode
pvz_df_clean %>%
  select(player_id, pvp_hours, pve_hours, social_hours) %>%
  pivot_longer(
    cols = c(pvp_hours, pve_hours, social_hours),
    names_to = "mode",
    values_to = "hours"
  ) %>%
  mutate(mode = factor(
    mode,
    levels = c("pvp_hours", "pve_hours", "social_hours"),
    labels = c("PvP", "PvE", "Social Hub")
  )) %>%
  ggplot(aes(x = hours, fill = mode)) +
  geom_histogram(binwidth = 2, color = "white", position = "dodge") +
  labs(
    title = "Distribution of Play Time by Game Mode",
    x = "Hours (2-week period)",
    y = "Count",
    fill = "Game Mode"
  ) +
  scale_fill_brewer(palette = "Set2") +
  theme_minimal()
```

## Performance Metrics Distribution

```{r performance-metrics}
#| label: performance-metrics

# Plot distribution of key performance metrics
pvz_df_clean %>%
  select(player_id, total_kd, total_hit_accuracy, total_dps) %>%
  pivot_longer(
    cols = c(total_kd, total_hit_accuracy, total_dps),
    names_to = "metric",
    values_to = "value"
  ) %>%
  mutate(metric = factor(
    metric,
    levels = c("total_kd", "total_hit_accuracy", "total_dps"),
    labels = c("Kill/Death Ratio", "Hit Accuracy", "Damage per Second")
  )) %>%
  ggplot(aes(x = value)) +
  geom_histogram(fill = "steelblue", color = "white") +
  facet_wrap(~ metric, scales = "free") +
  labs(
    title = "Distribution of Performance Metrics",
    x = "Value",
    y = "Count"
  ) +
  theme_minimal()
```

# Player Clustering Analysis

## Preparing Clustering Data

```{r cluster-prep}
#| label: cluster-prep

# Select variables for clustering based on playstyle
cluster_data <- pvz_df_clean %>% 
  select(
    # Game time distribution
    ratio_pvp_hours, ratio_pve_hours,
    # Performance metrics
    pvp_kd, pvp_hit_accuracy, pvp_dps,
    pve_kd, pve_hit_accuracy, pve_dps
  )

# Standardize data for clustering
cluster_data_scaled <- scale(cluster_data)
```

## Determining Optimal Number of Clusters

```{r optimal-clusters}
#| label: optimal-clusters

# Elbow method to determine optimal number of clusters
fviz_nbclust(
  cluster_data_scaled, 
  kmeans, 
  method = "wss"
) +
  labs(title = "Elbow Method for Optimal k")

# Silhouette method to validate cluster number
fviz_nbclust(
  cluster_data_scaled, 
  kmeans, 
  method = "silhouette"
) +
  labs(title = "Silhouette Method for Optimal k")
```

## K-means Clustering

```{r kmeans-clustering}
#| label: kmeans-clustering

# Set seed for reproducibility
set.seed(123)

# Perform k-means clustering with optimal k=5
km_res <- kmeans(cluster_data_scaled, centers = 5, nstart = 25)

# Add cluster assignment to dataset
pvz_df_clean <- pvz_df_clean %>%
  mutate(
    cluster = km_res$cluster,
    player_type = case_when(
      cluster == 1 ~ "Elite Players",
      cluster == 2 ~ "PvE Adventurers",
      cluster == 3 ~ "PvP Novices",
      cluster == 4 ~ "PvP Specialists",
      cluster == 5 ~ "Casual Players"
    )
  )

# Visualize the clusters
fviz_cluster(
  list(data = cluster_data_scaled, cluster = km_res$cluster),
  palette = "Set2",
  ellipse.type = "convex",
  ggtheme = theme_minimal()
) +
  labs(title = "Player Clusters")

# Create a table summarizing cluster characteristics
cluster_summary <- pvz_df_clean %>%
  group_by(player_type) %>%
  summarise(
    n = n(),
    hours_mean = mean(hours),
    pvp_ratio = mean(ratio_pvp_hours),
    pve_ratio = mean(ratio_pve_hours),
    pvp_kd = mean(pvp_kd),
    pve_kd = mean(pve_kd),
    pvp_accuracy = mean(pvp_hit_accuracy),
    pve_accuracy = mean(pve_hit_accuracy)
  )

knitr::kable(cluster_summary, 
             caption = "Summary of Player Type Clusters", 
             digits = 2)

# Save clustering results
write_rds(pvz_df_clean, here("data/processed/pvz_with_clusters.rds"))
```

# Relationship Between Playtime and Well-Being

## Basic Linear Regression

```{r basic-regression}
#| label: basic-regression

# Mean-center hours for interpretability
pvz_df_clean <- pvz_df_clean %>%
  mutate(hours_c = hours - mean(hours))

# Basic linear model of wellbeing based on game time
wellbeing_model_0 <- lm(spane_game_balance ~ hours_c, data = pvz_df_clean)

# Summarize model
summary(wellbeing_model_0)

# Visualize relationship
ggplot(pvz_df_clean, aes(x = hours, y = spane_game_balance)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "lm", color = "blue") +
  labs(
    title = "Relationship Between Play Time and Game-Related Well-Being",
    x = "Hours Played (2-week period)",
    y = "Game-Related Well-Being (SPANE Balance)"
  ) +
  theme_minimal()
```

## Moderation by Player Type

```{r player-type-moderation}
#| label: player-type-moderation

# Model with player type moderation
wellbeing_model_mod <- lm(spane_game_balance ~ hours_c * player_type, data = pvz_df_clean)

# Summarize model
summary(wellbeing_model_mod)

# Visualize interaction
library(interactions)
interact_plot(
  wellbeing_model_mod,
  pred = "hours_c",
  modx = "player_type",
  plot.points = TRUE,
  point.alpha = 0.3,
  colors = "Set2",
  x.label = "Hours Played (centered)",
  y.label = "Game-Related Well-Being"
) +
  theme_minimal() +
  labs(title = "Relationship Between Play Time and Well-Being by Player Type")
```

## Mixed Effects Model

```{r mixed-model}
#| label: mixed-model

# Random intercept model
library(lme4)
mixed_model_1 <- lmer(
  spane_game_balance ~ hours_c + (1 | player_type), 
  data = pvz_df_clean
)

# Random slope and intercept model
mixed_model_2 <- lmer(
  spane_game_balance ~ hours_c + (1 + hours_c | player_type), 
  data = pvz_df_clean
)

# Compare models
anova(mixed_model_1, mixed_model_2)

# Summary of best model
summary(mixed_model_2)

# Visualize random effects
library(sjPlot)
plot_model(mixed_model_2, type = "re")
```

# Discussion and Conclusion

This analysis has revealed several distinct player types based on their gameplay patterns and performance metrics. The relationship between play time and well-being varies across these player types, suggesting that the impact of gaming is not uniform for all players.

Key findings:

1. We identified five player types: Elite Players, PvE Adventurers, PvP Novices, PvP Specialists, and Casual Players.

2. Overall, there is a weak positive relationship between play time and game-related well-being.

3. This relationship is moderated by player type, with certain groups showing stronger positive associations than others.

4. Performance metrics like K/D ratio and hit accuracy vary substantially across player types and may influence the play time-wellbeing relationship.

These findings contribute to our understanding of how different playstyles might affect player experiences and well-being outcomes in video games.

# References

::: {#refs}
::: 