---
title: "Player Playstyle Analysis"
author: "Thomas Hakman"
format: 
  html:
    toc: true
    toc-depth: 3
    code-fold: true
    theme: flatly
    fig-width: 8
    fig-height: 6
execute:
  echo: true
  message: false
  warning: false
  cache: true
editor: visual
bibliography: references.bib
---

```{r setup}
#| label: setup
#| include: false

# Set global options
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  cache = TRUE,
  fig.path = "figures/"
)
```

# Introduction

This document analyzes player telemetry and survey data from Plants vs. Zombies: Battle for Neighborville. The analysis aims to understand the relationship between player playstyles, game performance metrics, and well-being outcomes.

# Data Preparation

## Load Packages

```{r packages}
#| label: load-packages
# Load required packages
library(tidyverse)
library(here)
library(lubridate)
library(scales)
library(janitor)
library(magrittr)
library(oii)
library(psych)
library(cluster)
library(factoextra)
library(lme4)
library(sjPlot)
library(interactions)
library(readxl)
library(knitr)
library(clustertend)
library(gridExtra)
## Load Raw Data

The data comes from a study conducted by EA and the Oxford Internet Institute. Players completed a survey on well-being and motivations, and EA provided telemetry data on their gameplay.

```{r raw-data-paths}
#| label: raw-data-paths

# Store all file paths in a central location for easy updates
data_paths <- list(
  survey = here("data/raw/ea/Oxford PvZ - Pilot Wave + Wave 1 Download.xlsx"),
  telemetry_files = list.files(
    here("data/raw/ea"), 
    pattern = "OII_PVZ_", 
    full.names = TRUE
  )
)

# Function to read both waves of telemetry data into a single table
read_both_waves <- function(stub = "OII_PVZ_Authentications", ...) {
  # Find files matching the pattern
  here("data/raw/ea") %>%
    list.files(pattern = stub, full.names = TRUE) %>%
    # Put wave 1 first
    rev() %>%
    # Read into a list of data frames with purrr's map
    map(read_csv, ...) %>%
    # Bind into one data frame with wave ID
    bind_rows(.id = "wave")
}
```

## Load Survey Data

```{r survey-load}
#| label: survey-load

survey <- read_excel(data_paths$survey)

survey <- survey %>% 
  mutate(
    # Ensure date data type
    date = mdy_hm(date),
    # Additional date variable for filtering telemetry
    start_date = date
  ) %>% 
  # Rename ID column to match telemetry files
  rename(player_id = UID)
```

## Clean Survey Data

```{r survey-clean}
#| label: survey-clean

# Rename variables with meaningful names
survey <- 
  survey %>% 
  select(
    # meta-info
    player_id,
    start_date, date,
    survey_duration = qtime,
    of_age = Q2,
    consent = Q3,
    
    # demographics
    played = Q1,
    country = hCountry,
    survey_language = PRELANG,
    age = D1,
    gender = D2,
    gender_other = D2r3oe,
    experience = D3,
    
    # well-being
    spane_ = starts_with("A1"), # general well-being in the two weeks
    straightliner_spane = hA1Straight,
    
    # motivations
    played_with_others = NSat1, # filter question for relatedness items
    autonomy_ = NSat2r1:NSat2r3,
    competence_ = NSat2r4:NSat2r6,
    relatedness_ = NSat2r7:NSat2r9,
    enjoyment_ = NSat2r10:NSat2r13,
    extrinsic_ = NSat2r14:NSat2r17,
    straightliner_motivations = hNSat2Straight,
    
    # self-reported play
    active_play_hours = GT1c1,
    active_play_minutes = GT1c2,
    within_estimate = GT2,
    between_estimate = GT3,
    spane_game_ = starts_with("GT4")
  )

# Assign proper variable types and factor levels
survey <- survey %>%   
  mutate(
    across(
      c(
        player_id,
        of_age:survey_language,
        gender,
        played_with_others
      ),
      as.factor
    ),
    
    across(
      c(of_age:played, played_with_others),
      ~ fct_recode(
        .x,
        "Yes" = "1",
        "No" = "2"
      )
    ),
    
    gender = fct_recode(
      gender,
      "Male" = "1",
      "Female" = "2",
      "Other" = "3",
      "Prefer not to say" = "4"
    ),
    
    country = fct_recode(
      country,
      "US" = "1",
      "UK" = "2",
      "Canada" = "5"
    ),
    
    survey_language = fct_recode(
      survey_language,
      "English" = "1",
      "French" = "2"
    )
  )

# Apply reverse scoring for appropriate items
survey <- survey %>% 
  mutate(
    across(
      c(
        starts_with("spane"), # 1 was never
        between_estimate, # those estimates also had 1: much more than typically
        within_estimate,
        relatedness_3,
        enjoyment_2,
        enjoyment_3
      ),
      ~ 8 - .x
    )
  )

# Create indicators for straightliners
survey <- survey %>% 
  mutate(across(contains("straightliner"), ~.x==1))
```

## Create Survey Scales

```{r survey-scales}
#| label: survey-scales

# General SPANE scale (well-being)
survey <- survey %>% 
  mutate(
    # Use purrr to create positive and negative scale means
    spane_positive = map_dbl(1:nrow(.), ~mean(c(
      survey$spane_1[.x], 
      survey$spane_3[.x], 
      survey$spane_5[.x], 
      survey$spane_7[.x], 
      survey$spane_10[.x], 
      survey$spane_12[.x]
    ), na.rm = TRUE)),
    
    spane_negative = map_dbl(1:nrow(.), ~mean(c(
      survey$spane_2[.x], 
      survey$spane_4[.x], 
      survey$spane_6[.x], 
      survey$spane_8[.x], 
      survey$spane_9[.x], 
      survey$spane_11[.x]
    ), na.rm = TRUE)),
    
    spane_balance = spane_positive - spane_negative
  )

# Motivation scales
survey <- survey %>% 
  mutate(
    # Use purrr's map functions to compute means for different motivation scales
    autonomy = pmap_dbl(select(., starts_with("autonomy")), 
                       ~mean(c(...), na.rm = TRUE)),
    competence = pmap_dbl(select(., starts_with("competence")), 
                         ~mean(c(...), na.rm = TRUE)),
    relatedness = pmap_dbl(select(., starts_with("relatedness")), 
                          ~mean(c(...), na.rm = TRUE)),
    enjoyment = pmap_dbl(select(., starts_with("enjoyment")), 
                        ~mean(c(...), na.rm = TRUE)),
    extrinsic = pmap_dbl(select(., starts_with("extrinsic")), 
                        ~mean(c(...), na.rm = TRUE))
  )

# SPANE because of playing game
survey <- survey %>% 
  mutate(
    # Use purrr for game-related SPANE scales
    spane_game_positive = map_dbl(1:nrow(.), ~mean(c(
      survey$spane_game_1[.x], 
      survey$spane_game_3[.x], 
      survey$spane_game_5[.x], 
      survey$spane_game_7[.x], 
      survey$spane_game_10[.x], 
      survey$spane_game_12[.x]
    ), na.rm = TRUE)),
    
    spane_game_negative = map_dbl(1:nrow(.), ~mean(c(
      survey$spane_game_2[.x], 
      survey$spane_game_4[.x], 
      survey$spane_game_6[.x], 
      survey$spane_game_8[.x], 
      survey$spane_game_9[.x], 
      survey$spane_game_11[.x]
    ), na.rm = TRUE)),
    
    spane_game_balance = spane_game_positive - spane_game_negative
  )

# Hours of estimated play
survey <- survey %>%
  mutate(active_play_minutes = active_play_minutes / 60) %>% 
  mutate(active_play = rowSums(select(., starts_with("active_play")), na.rm = TRUE))

# Save processed survey data
write_csv(survey, here("data/processed/survey_processed.csv"))
```

# Telemetry Data Processing

## Load Game Time Data

```{r game-time-load}
#| label: game-time-load

game_time_raw <- read_both_waves("OII_PVZ_Game_Time") %>%
  # Only include survey participants 
  filter(player_id %in% survey$player_id)
```

## Process Game Time Data

```{r game-time-process}
#| label: game-time-process

# Handle missing start times
game_time <- game_time_raw %>% 
  mutate(
    # If start time is missing, set to end time (will result in 0 duration)
    game_start_time = coalesce(game_start_time, game_end_time)
  ) %>%
  # Collapse rows by session
  group_by(player_id, game_session) %>% 
  mutate(
    game_start_time = min(game_start_time),
    game_end_time = max(game_end_time),
  ) %>% 
  # Take the first of each (identical) person-session combo
  slice(1) %>% 
  ungroup()

# Filter to sessions within 14 days before survey
game_time_2week <- game_time %>% 
  left_join(select(survey, player_id, start_date)) %>% 
  filter(game_start_time >= (start_date - days(14)))

# Save processed game time data
write_csv(game_time_2week, here("data/processed/game_time_2week.csv"))
```

## Load Authentication Data

```{r auth-load}
#| label: auth-load

authentications_raw <- read_both_waves("OII_PVZ_Authentications") %>%
  # Only include survey participants
  filter(player_id %in% survey$player_id)

# Filter to 14 days before survey
authentications_2week <- authentications_raw %>% 
  left_join(select(survey, player_id, start_date)) %>% 
  filter(date >= (as_date(start_date) - days(14)))

# Count authentications per player
authentications_players <- authentications_2week %>% 
  count(player_id, name = "authentications")

# Save processed authentication data
write_csv(authentications_players, here("data/processed/authentications_players.csv"))
```

## Load Character Performance Data

```{r character-load}
#| label: character-load

characters_raw <- read_both_waves("OII_PVZ_Character") %>%
  # Only include survey participants
  filter(player_id %in% survey$player_id)

# Add session timing info and filter to 14 days before survey
characters_2week <- characters_raw %>%
  left_join(
    game_time_2week %>% 
      select(player_id, game_session, game_start_time, game_end_time),
    by = c("player_id", "game_session")
  ) %>%
  left_join(select(survey, player_id, start_date)) %>% 
  filter(game_start_time >= (start_date - days(14)))

# Save for further processing
write_csv(characters_2week, here("data/processed/characters_2week.csv"))
```

# Game Mode Classification

```{r game-mode-classification}
#| label: game-mode-classification

# Create game mode classification variables
characters_2week <- characters_2week %>%
  mutate(
    # Social hub sessions (no combat)
    social_hub = case_when(
      game_level_key == "dsub_social_terrain_summer" & is.na(score) ~ 1,
      TRUE ~ 0
    ),
    
    # PvE sessions
    pve_session = case_when(
      game_level_key %in% c(
        "dsub_region_cheesemines", 
        "dsub_region_dreadwood", 
        "dsub_region_towncenter"
      ) ~ 1,
      TRUE ~ 0
    ),
    
    # PvP sessions
    pvp_session = case_when(
      game_level_key %in% c(
        "dsub_rush_dreadwood", "dsub_rush_towncenter", 
        "dsub_tvm_cheesemines", "dsub_tvm_zrefinery", 
        "dsub_tvm_suburbs", "dsub_tvm_dreadwood", 
        "dsub_rush_farm", "dsub_rush_cheesemines", 
        "dsub_rush_shoreline", "dsub_rush_resort", 
        "dsub_4v4_battlearena_v2", "dsub_4v4_superarena"
      ) ~ 1,
      # Giddy Park with combat data is also PvP
      game_level_key == "dsub_social_terrain_summer" & !is.na(score) ~ 1,
      TRUE ~ 0
    )
  )

# Remove sessions without performance data (mainly social hub)
characters_2week_nona <- characters_2week %>% 
  drop_na(total_kill_count:shots_hit)

# Save processed data
write_csv(characters_2week, here("data/processed/characters_with_modes.csv"))
write_csv(characters_2week_nona, here("data/processed/characters_with_modes_nona.csv"))
```

# Game Time Calculations

```{r game-time-calculations}
#| label: game-time-calculations

# Define a function to calculate time per player for a specific game mode
calculate_game_time <- function(data, mode_filter, time_prefix) {
  data %>%
    filter({{ mode_filter }} == 1) %>%
    mutate(duration = game_end_time - game_start_time) %>%
    mutate(seconds = as.numeric(duration)) %>%
    group_by(player_id) %>%
    summarise(seconds = sum(seconds)) %>%
    mutate(hours = seconds / 60 / 60) %>%
    rename_with(~ paste0(time_prefix, "_", .), .cols = c("seconds", "hours")) %>%
    ungroup()
}

# Calculate total game time per player
game_time_players_total <- characters_2week %>% 
  mutate(duration = game_end_time - game_start_time) %>% 
  mutate(seconds = as.numeric(duration)) %>% 
  group_by(player_id) %>% 
  summarise(seconds = sum(seconds)) %>%
  mutate(hours = seconds / 60 / 60) %>% 
  ungroup()

# Calculate game time for different modes using our function
game_time_players_pvp <- calculate_game_time(characters_2week, pvp_session, "pvp")
game_time_players_pve <- calculate_game_time(characters_2week, pve_session, "pve")
game_time_players_social <- calculate_game_time(characters_2week, social_hub, "social")

# Combine game time data with purrr's reduce
game_time_players <- list(
  game_time_players_total, 
  game_time_players_social, 
  game_time_players_pve, 
  game_time_players_pvp
) %>%
  reduce(left_join, by = "player_id")

# Replace NA values with 0 (no time spent in that mode)
game_time_players[is.na(game_time_players)] <- 0

# Save processed game time data
write_csv(game_time_players, here("data/processed/game_time_players.csv"))
```

# Performance Metrics Calculation

```{r performance-setup}
#| label: performance-setup
#| cache: true
#| dependson: c("game-mode-classification", "game-time-calculations")

# Create an empty list to store intermediate results
metrics_list <- list()
```

```{r overall-performance}
#| label: overall-performance
#| cache: true
#| dependson: c("performance-setup")

# Calculate overall performance metrics per player
characters_players <- characters_2week %>%
  group_by(player_id) %>% 
  summarise(across(total_kill_count:shots_hit, sum, na.rm = TRUE)) %>% 
  ungroup()

# Store in our list
metrics_list$characters_players <- characters_players

# Force garbage collection
gc()
```

```{r game-sessions}
#| label: game-sessions
#| cache: true
#| dependson: c("performance-setup")

# Calculate game sessions per player
game_sessions_total <- characters_2week %>% 
  count(player_id) %>%
  rename(n_total_game_sessions = n)

# Store in our list
metrics_list$game_sessions_total <- game_sessions_total

# Force garbage collection
gc()
```

```{r sessions-by-mode}
#| label: sessions-by-mode
#| cache: true
#| dependson: c("performance-setup")

# Calculate sessions by game mode
game_sessions <- characters_2week %>% 
  group_by(player_id) %>% 
  summarise(across(social_hub:pvp_session, sum, na.rm = TRUE)) %>% 
  ungroup()

# Store in our list
metrics_list$game_sessions <- game_sessions

# Force garbage collection
gc()
```

```{r pvp-performance}
#| label: pvp-performance
#| cache: true
#| dependson: c("performance-setup")

# PvP-specific performance metrics
pvp_telemetry <- characters_2week %>%
  filter(pvp_session == 1) %>%
  group_by(player_id) %>% 
  summarise(across(total_kill_count:shots_hit, sum, na.rm = TRUE)) %>%
  rename_with(~ paste0("pvp_", .), -player_id)

# Store in our list
metrics_list$pvp_telemetry <- pvp_telemetry

# Force garbage collection
gc()
```

```{r pve-performance}
#| label: pve-performance
#| cache: true
#| dependson: c("performance-setup")

# PvE-specific performance metrics
pve_telemetry <- characters_2week %>%
  filter(pve_session == 1) %>%
  group_by(player_id) %>% 
  summarise(across(total_kill_count:shots_hit, sum, na.rm = TRUE)) %>%
  rename_with(~ paste0("pve_", .), -player_id)

# Store in our list
metrics_list$pve_telemetry <- pve_telemetry

# Force garbage collection
gc()
```

```{r combine-metrics}
#| label: combine-metrics
#| cache: true
#| dependson: c("overall-performance", "game-sessions", "sessions-by-mode", "pvp-performance", "pve-performance")

# Load additional data we need from previous processing
metrics_list$game_time_players <- game_time_players

# Combine telemetry data using purrr's reduce function
telemetry_players <- list(
  metrics_list$characters_players,
  metrics_list$game_time_players,
  metrics_list$game_sessions,
  metrics_list$game_sessions_total,
  metrics_list$pvp_telemetry,
  metrics_list$pve_telemetry
) %>%
  reduce(left_join, by = "player_id")

# Replace NA values with 0
telemetry_players <- telemetry_players %>%
  mutate(across(everything(), ~replace_na(.x, 0)))

# Clear intermediate objects and list to free memory
rm(metrics_list)
gc()

# Save processed telemetry data
write_csv(telemetry_players, here("data/processed/telemetry_players.csv"))
```

# Derived Performance Metrics

```{r derived-metrics}
#| label: derived-metrics

# Define a function to calculate performance ratios
calculate_ratio <- function(numerator, denominator) {
  numerator / denominator
}

# Calculate derived performance metrics
telemetry_players <- telemetry_players %>%
  # Handle divisions by zero for K/D ratio
  mutate(
    across(
      c(total_death_count, pvp_total_death_count, pve_total_death_count),
      ~ ifelse(.x == 0, 1, .x)
    )
  ) %>%
  # Calculate all ratio metrics with purrr
  mutate(
    # K/D ratios
    total_kd = calculate_ratio(total_kill_count, total_death_count),
    pve_kd = calculate_ratio(pve_total_kill_count, pve_total_death_count),
    pvp_kd = calculate_ratio(pvp_total_kill_count, pvp_total_death_count),
    
    # Hit accuracy ratios
    total_hit_accuracy = calculate_ratio(shots_hit, shots_fired),
    pve_hit_accuracy = calculate_ratio(pve_shots_hit, pve_shots_fired),
    pvp_hit_accuracy = calculate_ratio(pvp_shots_hit, pvp_shots_fired),
    
    # Critical hit ratios
    total_crit_ratio = calculate_ratio(critical_hit_count, shots_hit),
    pve_crit_ratio = calculate_ratio(pve_critical_hit_count, pve_shots_hit),
    pvp_crit_ratio = calculate_ratio(pvp_critical_hit_count, pvp_shots_hit),
    
    # Damage per game
    total_average_damage = calculate_ratio(damage_dealt, n_total_game_sessions),
    pve_average_damage = calculate_ratio(pve_damage_dealt, pve_session),
    pvp_average_damage = calculate_ratio(pvp_damage_dealt, pvp_session),
    
    # Damage per second
    total_dps = calculate_ratio(damage_dealt, seconds),
    pve_dps = calculate_ratio(pve_damage_dealt, pve_seconds),
    pvp_dps = calculate_ratio(pvp_damage_dealt, pvp_seconds),
    
    # Score per game
    total_average_score = calculate_ratio(score, n_total_game_sessions),
    pve_average_score = calculate_ratio(pve_score, pve_session),
    pvp_average_score = calculate_ratio(pvp_score, pvp_session),
    
    # Score per second
    total_sps = calculate_ratio(score, seconds),
    pve_sps = calculate_ratio(pve_score, pve_seconds),
    pvp_sps = calculate_ratio(pvp_score, pvp_seconds)
  )

# Replace NaN values with 0
telemetry_players <- telemetry_players %>%
  mutate(across(everything(), ~ifelse(is.nan(.x), 0, .x)))

# Save processed metrics
write_csv(telemetry_players, here("data/processed/telemetry_with_metrics.csv"))
```

# Combine Survey and Telemetry Data

```{r merge-data}
#| label: merge-data

# Merge survey with telemetry data
pvz_df <- survey %>% 
  left_join(telemetry_players, by = "player_id")

# Calculate ratio metrics
pvz_df <- pvz_df %>%
  mutate(
    ratio_social_hours = social_hours / hours,
    ratio_pve_hours = pve_hours / hours,
    ratio_pvp_hours = pvp_hours / hours
  )

# Remove participants with very low playtime (< 2 minutes)
pvz_df_clean <- pvz_df %>%
  filter(hours > 0.03)

# Save final dataset
write_csv(pvz_df_clean, here("data/processed/pvz_complete.csv"))
```

# Handling Extreme Playtime Outliers

```{r outlier-handling}
#| label: outlier-handling
#| fig.width: 10
#| fig.height: 6

# Examine the distribution of hours played
ggplot(pvz_df_clean, aes(x = hours)) +
  geom_histogram(binwidth = 5, fill = "steelblue", color = "white") +
  geom_vline(xintercept = mean(pvz_df_clean$hours), color = "red", linetype = "dashed") +
  geom_vline(xintercept = mean(pvz_df_clean$hours) + 5*sd(pvz_df_clean$hours), 
             color = "darkred", linetype = "dashed") +
  labs(
    title = "Distribution of Hours Played with Outlier Threshold",
    subtitle = "Red dashed line = mean, Dark red dashed line = mean + 5 SD",
    x = "Hours (2-week period)",
    y = "Count"
  ) +
  theme_minimal()

# Calculate statistical outlier threshold using mean + 5 SD
outlier_threshold <- mean(pvz_df_clean$hours) + 5*sd(pvz_df_clean$hours)
print(paste("Outlier threshold (mean + 5 SD):", round(outlier_threshold, 1), "hours"))

# Identify extreme outliers
extreme_outliers <- pvz_df_clean %>%
  filter(hours > outlier_threshold) %>%
  select(player_id, hours, pvp_hours, pve_hours, social_hours, pvp_kd, player_type)

# Display the outliers
print("Players with extreme hours played (outliers):")
knitr::kable(extreme_outliers, digits = 1)

# Create dataset without extreme playtime outliers
pvz_df_no_outliers <- pvz_df_clean %>%
  filter(hours <= outlier_threshold)

# Compare summary statistics before and after outlier removal
comparison <- bind_rows(
  pvz_df_clean %>% 
    summarise(
      dataset = "With outliers",
      n = n(),
      mean_hours = mean(hours),
      median_hours = median(hours),
      sd_hours = sd(hours),
      max_hours = max(hours)
    ),
  pvz_df_no_outliers %>% 
    summarise(
      dataset = "Without outliers",
      n = n(),
      mean_hours = mean(hours),
      median_hours = median(hours),
      sd_hours = sd(hours),
      max_hours = max(hours)
    )
)

# Display comparison
knitr::kable(comparison, digits = 1,
             caption = "Comparison of hours played statistics before and after outlier removal")

# Save dataset without extreme outliers for further analysis
write_csv(pvz_df_no_outliers, here("data/processed/pvz_no_outliers.csv"))

# Use the filtered dataset for subsequent analyses
# This preserves the original dataset while allowing for more robust statistical analysis
pvz_df_analysis <- pvz_df_no_outliers
```

# Exploratory Data Analysis

## Demographic Statistics

```{r demographic-stats}
#| label: demographic-stats

# Demographic summary
demographic_summary <- pvz_df_analysis %>%
  summarise(
    n_participants = n(),
    mean_age = mean(age, na.rm = TRUE),
    sd_age = sd(age, na.rm = TRUE),
    median_age = median(age, na.rm = TRUE),
    n_male = sum(gender == "Male", na.rm = TRUE),
    pct_male = mean(gender == "Male", na.rm = TRUE) * 100,
    n_female = sum(gender == "Female", na.rm = TRUE),
    pct_female = mean(gender == "Female", na.rm = TRUE) * 100
  )

knitr::kable(demographic_summary, caption = "Demographic Summary")
```

## Game Time Distribution

```{r gametime-distribution}
#| label: gametime-distribution

# Plot distribution of total game time
ggplot(pvz_df_analysis, aes(x = hours)) +
  geom_histogram(binwidth = 2, fill = "steelblue", color = "white") +
  labs(
    title = "Distribution of Total Play Time",
    x = "Hours (2-week period)",
    y = "Count"
  ) +
  theme_minimal()

# Plot distribution of game time by mode
pvz_df_analysis %>%
  select(player_id, pvp_hours, pve_hours, social_hours) %>%
  pivot_longer(
    cols = c(pvp_hours, pve_hours, social_hours),
    names_to = "mode",
    values_to = "hours"
  ) %>%
  mutate(mode = factor(
    mode,
    levels = c("pvp_hours", "pve_hours", "social_hours"),
    labels = c("PvP", "PvE", "Social Hub")
  )) %>%
  ggplot(aes(x = hours, fill = mode)) +
  geom_histogram(binwidth = 2, color = "white", position = "dodge") +
  labs(
    title = "Distribution of Play Time by Game Mode",
    x = "Hours (2-week period)",
    y = "Count",
    fill = "Game Mode"
  ) +
  scale_fill_brewer(palette = "Set2") +
  theme_minimal()
```

## Performance Metrics Distribution

```{r performance-metrics}
#| label: performance-metrics

# Plot distribution of key performance metrics
pvz_df_analysis %>%
  select(player_id, total_kd, total_hit_accuracy, total_dps) %>%
  pivot_longer(
    cols = c(total_kd, total_hit_accuracy, total_dps),
    names_to = "metric",
    values_to = "value"
  ) %>%
  mutate(metric = factor(
    metric,
    levels = c("total_kd", "total_hit_accuracy", "total_dps"),
    labels = c("Kill/Death Ratio", "Hit Accuracy", "Damage per Second")
  )) %>%
  ggplot(aes(x = value)) +
  geom_histogram(fill = "steelblue", color = "white") +
  facet_wrap(~ metric, scales = "free") +
  labs(
    title = "Distribution of Performance Metrics",
    x = "Value",
    y = "Count"
  ) +
  theme_minimal()
```

# Player Clustering Analysis


## Preparing Clustering Data

```{r cluster-prep}
#| label: cluster-prep

# Select variables for clustering based on playstyle
cluster_data <- pvz_df_analysis %>% 
  select(
    # Game time distribution
    ratio_pvp_hours, ratio_pve_hours, ratio_social_hours,
    
    # Performance metrics - PvP
    pvp_kd, pvp_hit_accuracy, pvp_dps, pvp_crit_ratio, 
    pvp_average_damage, pvp_average_score,
    
    # Performance metrics - PvE
    pve_kd, pve_hit_accuracy, pve_dps, pve_crit_ratio,
    pve_average_damage, pve_average_score
    
  )

# Standardize data for clustering
cluster_data_scaled <- scale(cluster_data)
```


## Assessing cluster tendency

Before performing the cluster analysis, we need to assess whether the data has a natural tendency to form clusters. This helps determine if clustering is appropriate for our dataset.

```{r cluster-tendency}
#| label: cluster-tendency

# Load the clustertend package for assessing clustering tendency
if (!requireNamespace("clustertend", quietly = TRUE)) {
  install.packages("clustertend")
}
library(clustertend)

# Prepare the data for cluster tendency assessment


# Create a random dataset with the same dimensions for comparison
set.seed(123)
random_df <- apply(cluster_data_scaled, 2,
                   function(x){runif(length(x), min(x), max(x))})
random_df <- as.data.frame(random_df)
```

### Visual Assessment of Clustering Tendency

We can visually assess whether our dataset has a natural cluster structure by comparing it with a randomly generated dataset using PCA visualization.

```{r visual-tendency}
#| label: visual-tendency
#| fig.width: 12
#| fig.height: 5

# PCA visualization of our actual dataset
p1 <- fviz_pca_ind(prcomp(cluster_data_scaled), 
                   title = "PCA - Player Playstyle Data",
                   geom = "point", 
                   ggtheme = theme_minimal(),
                   legend = "none")

# PCA visualization of random dataset for comparison
p2 <- fviz_pca_ind(prcomp(random_df), 
                   title = "PCA - Random Data",
                   geom = "point", 
                   ggtheme = theme_minimal(),
                   legend = "none")

# Display both visualizations side by side
grid.arrange(p1, p2, ncol = 2)
```

The PCA visualization shows that our playstyle data (left) appears to have more structure than the random data (right). In the playstyle data, we can observe potential groupings or clusters, whereas the random data shows a more uniform distribution without clear patterns. This visual assessment suggests natural clusters may exist in our dataset.

### Statistical Assessment with Hopkins Statistic

The Hopkins statistic quantitatively assesses clustering tendency. It measures whether a dataset is significantly more clusterable than a random dataset:

- H close to 0: Data has a high clustering tendency
- H around 0.5: Data is randomly distributed with no meaningful clusters
- H close to 1: Data is uniformly distributed

```{r hopkins-test}
#| label: hopkins-test

# Compute Hopkins statistic for our dataset
set.seed(123)
h_stat <- hopkins(cluster_data_scaled, n = nrow(cluster_data_scaled)-1)
cat("Hopkins statistic:", h_stat, "\n")
cat("Clustering tendency (1-H):", 1-h_stat, "\n")
```

The Hopkins statistic is well below the 0.5 threshold, indicating that our data has a strong clustering tendency. This value suggests that cluster analysis is an appropriate technique for identifying player types in our dataset.

For reference, the Hopkins statistic for the well-known Iris dataset (a benchmark for clustering algorithms) is around 0.18, so our value indicates an even stronger clustering tendency.

### Dissimilarity Matrix Visualization

The dissimilarity matrix visualization shows the pairwise distances between observations. Areas with similar colors indicate potential clusters:

```{r distance-matrix}
#| label: distance-matrix
#| fig.width: 8
#| fig.height: 8

# Compute dissimilarity matrix
dist_matrix <- get_dist(cluster_data_scaled, method = "pearson")
fviz_dist(dist_matrix, 
          show_labels = FALSE, 
          gradient = list(low = "#00AFBB", mid = "white", high = "#FC4E07"),
          title = "Distance Matrix Visualization - Player Data")
```

```{r random-distance-matrix}
#| label: random-distance-matrix
#| fig.width: 8
#| fig.height: 8

# Create a random dataset with the same dimensions for comparison
set.seed(123)
random_df <- apply(cluster_data_scaled, 2,
                   function(x){runif(length(x), min(x), max(x))})
random_df <- as.data.frame(random_df)

# Compute dissimilarity matrix for random data
dist_matrix_random <- get_dist(random_df, method = "pearson")
fviz_dist(dist_matrix_random, 
          show_labels = FALSE, 
          gradient = list(low = "#00AFBB", mid = "white", high = "#FC4E07"),
          title = "Distance Matrix Visualization - Random Data")
```
The distance matrix visualization reveals blocks of similar colors (blue), which indicate groups of observations that are similar to each other. This pattern is characteristic of data with natural clusters and further supports our conclusion that the player data contains distinguishable playstyle groups.

Based on both visual and statistical assessments, we can confidently proceed with cluster analysis to identify distinct player types in our dataset.


## Determining Optimal Number of Clusters

```{r optimal-clusters}
#| label: optimal-clusters

# Elbow method to determine optimal number of clusters
fviz_nbclust(
  cluster_data_scaled, 
  kmeans, 
  method = "wss"
) +
  labs(title = "Elbow Method for Optimal k")

# Silhouette method to validate cluster number
fviz_nbclust(
  cluster_data_scaled, 
  kmeans, 
  method = "silhouette"
) +
  labs(title = "Silhouette Method for Optimal k")

# Gap statistic method for determining optimal number of clusters
# This compares the within-cluster dispersion to a null reference distribution
set.seed(125)  # For reproducibility
fviz_nbclust(
  cluster_data_scaled, 
  kmeans, 
  method = "gap_stat", 
  nboot = 500,  # Number of bootstrap samples
  verbose = FALSE  # Hide computing progression
) +
  labs(title = "Gap Statistic Method for Optimal k")
```

## K-means Clustering

```{r kmeans-clustering}
#| label: kmeans-clustering

# Set seed for reproducibility
set.seed(125)

# Perform k-means clustering with optimal k=6
km_res <- kmeans(cluster_data_scaled, centers = 8, nstart = 25)

# Visualize the clusters
fviz_cluster(
  list(data = cluster_data_scaled, cluster = km_res$cluster),
  palette = "Set2",
  ellipse.type = "convex",
  ggtheme = theme_minimal()
) +
  labs(title = "Player Clusters")

# Examine cluster centers to understand what each cluster represents
cluster_centers <- as.data.frame(km_res$centers)
print("Cluster centers (standardized values):")
knitr::kable(cluster_centers, digits = 2)

# Examine actual metric values by cluster
cluster_characteristics <- pvz_df_analysis %>%
  mutate(cluster = km_res$cluster) %>%
  group_by(cluster) %>%
  summarise(
    n_players = n(),
    hours_mean = mean(hours),
    pvp_ratio = mean(ratio_pvp_hours),
    pve_ratio = mean(ratio_pve_hours),
    social_ratio = mean(ratio_social_hours),
    pvp_kd = mean(pvp_kd),
    pve_kd = mean(pve_kd),
    pvp_accuracy = mean(pvp_hit_accuracy),
    pve_accuracy = mean(pve_hit_accuracy),
    pvp_dps = mean(pvp_dps),
    pve_dps = mean(pve_dps)
  )

print("Characteristics of each cluster:")
knitr::kable(cluster_characteristics, digits = 2)

# Based on the analysis of cluster centers and characteristics,
# assign meaningful names to each cluster
pvz_df_analysis <- pvz_df_analysis %>%
    mutate(
      cluster = km_res$cluster,
      player_type = case_when(
        cluster == 1 ~ "Elite Players",
        cluster == 2 ~ "PvE Adventurers",
        cluster == 3 ~ "PvP Novices",
        cluster == 4 ~ "PvP Specialists",
        cluster == 5 ~ "Casual Players",
        cluster == 6 ~ "Social Hubbers"
      )
    )

# Now create a table summarizing cluster characteristics with the assigned names
cluster_summary <- pvz_df_analysis %>%
  group_by(player_type) %>%
  summarise(
    n = n(),
    hours_mean = mean(hours),
    pvp_ratio = mean(ratio_pvp_hours),
    pve_ratio = mean(ratio_pve_hours),
    pvp_kd = mean(pvp_kd),
    pve_kd = mean(pve_kd),
    pvp_accuracy = mean(pvp_hit_accuracy),
    pve_accuracy = mean(pve_hit_accuracy)
  )

knitr::kable(cluster_summary, 
             caption = "Summary of Player Type Clusters", 
             digits = 2)

# Save clustering results
write_csv(pvz_df_analysis, here("data/processed/pvz_with_clusters.csv"))
```


## cluster validation



# Relationship Between Playtime and Well-Being

## Basic Linear Regression

```{r basic-regression}
#| label: basic-regression

# Mean-center hours for interpretability
pvz_df_analysis <- pvz_df_analysis %>%
  mutate(hours_c = hours - mean(hours))

# Basic linear model of wellbeing based on game time
wellbeing_model_0 <- lm(spane_game_balance ~ hours_c, data = pvz_df_analysis)

# Summarize model
summary(wellbeing_model_0)

# Visualize relationship
ggplot(pvz_df_analysis, aes(x = hours, y = spane_game_balance)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "lm", color = "blue") +
  labs(
    title = "Relationship Between Play Time and Game-Related Well-Being",
    x = "Hours Played (2-week period)",
    y = "Game-Related Well-Being (SPANE Balance)"
  ) +
  theme_minimal()
```

## Moderation by Player Type

```{r player-type-moderation}
#| label: player-type-moderation

# Make sure the interactions package is available
library(interactions)

# Model with player type moderation
wellbeing_model_mod <- lm(spane_game_balance ~ hours_c * player_type, data = pvz_df_analysis)

# Summarize model
summary(wellbeing_model_mod)

# Visualize interaction
interact_plot(
  wellbeing_model_mod,
  pred = "hours_c",
  modx = "player_type",
  plot.points = TRUE,
  point.alpha = 0.3,
  colors = "Set2",
  x.label = "Hours Played (centered)",
  y.label = "Game-Related Well-Being"
) +
  theme_minimal() +
  labs(title = "Relationship Between Play Time and Well-Being by Player Type")
```

## Mixed Effects Model

```{r mixed-model}
#| label: mixed-model

# Random intercept model
library(lme4)
mixed_model_1 <- lmer(
  spane_game_balance ~ hours_c + (1 | player_type), 
  data = pvz_df_analysis
)

# Random slope and intercept model
mixed_model_2 <- lmer(
  spane_game_balance ~ hours_c + (1 + hours_c | player_type), 
  data = pvz_df_analysis
)

# Compare models
anova(mixed_model_1, mixed_model_2)

# Summary of best model
summary(mixed_model_2)

# Visualize random effects
library(sjPlot)
plot_model(mixed_model_2, type = "re")
```

# Discussion and Conclusion

This analysis has revealed several distinct player types based on their gameplay patterns and performance metrics. The relationship between play time and well-being varies across these player types, suggesting that the impact of gaming is not uniform for all players.

Key findings:

1. We identified five player types: Elite Players, PvE Adventurers, PvP Novices, PvP Specialists, and Casual Players.

2. Overall, there is a weak positive relationship between play time and game-related well-being.

3. This relationship is moderated by player type, with certain groups showing stronger positive associations than others.

4. Performance metrics like K/D ratio and hit accuracy vary substantially across player types and may influence the play time-wellbeing relationship.

These findings contribute to our understanding of how different playstyles might affect player experiences and well-being outcomes in video games.

# References

::: {#refs}
::: 