---
title: "Player Playstyle Data Preprocessing"
author: "Thomas Hakman"
format: 
  html:
    toc: true
    toc-depth: 3
    code-fold: true
    theme: flatly
    fig-width: 8
    fig-height: 6
execute:
  echo: true
  message: false
  warning: false
  cache: false
editor: visual
---

```{r setup}
#| label: setup
#| include: false

# Set global options
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  cache = FALSE,
  fig.path = "figures/"
)
```

# Introduction

This document processes player telemetry and survey data from Plants vs. Zombies: Battle for Neighborville. The preprocessing steps prepare the data for analysis of the relationship between playtime, playstyle, and well-being outcomes.

# Data Preparation

## Load Packages

```{r packages}
#| label: load-packages
# Install pacman if not already installed
if (!require("pacman")) install.packages("pacman")

# Load required packages using pacman
pacman::p_load(
  here,       # For project-relative file paths
  lubridate,  # For date/time handling
  scales,     # For visualization scaling
  janitor,    # For data cleaning
  magrittr,   # For additional pipe operators
  psych,      # For psychological statistics
  readxl,     # For reading Excel files
  knitr,      # For dynamic report generation
  tidyverse   # For data manipulation and visualization (loaded last to avoid masking)
)
```

## Load Helper Functions

```{r helper-functions}
#| label: helper-functions

# Helper functions for PvZ data analysis
# These functions support the main preprocessing steps

# Function to read both waves of telemetry data into a single table
read_both_waves <- function(stub = "OII_PVZ_Authentications", data_dir = "data/raw/ea", ...) {
  # Find files matching the pattern
  files <- list.files(
    here::here(data_dir), pattern = stub, full.names = TRUE
  )
  
  # Validate that files were found
  if (length(files) == 0) {
    stop(
      "No files found matching pattern: ", stub, " in ", data_dir, "\n",
      "Please download data from: https://osf.io/cjd6z/files/4gp3r"
    )
  }
  
  # Filenames of wave 1 and 2 files
  files %>% 
    # Put wave 1 first (they are in reverse order)
    rev() %>% 
    # Read into a list of data frames
    map(read_csv, ...) %>% 
    # Bind into one data frame with wave ID
    bind_rows(.id = "wave")
}

# Function to create performance metrics from player data
calculate_performance_metrics <- function(df, prefix = "total") {
  # Handle cases where denominators might be zero
  df <- df %>%
    mutate(
      # Set death count to 1 if 0 to avoid division by zero
      across(
        contains("death_count"), 
        ~ ifelse(.x == 0, 1, .x)
      )
    )
  
  # Create column names based on prefix
  kill_count_col <- paste0(prefix, ifelse(prefix == "", "", "_"), "kill_count")
  death_count_col <- paste0(prefix, ifelse(prefix == "", "", "_"), "death_count")
  shots_hit_col <- paste0(prefix, ifelse(prefix == "", "", "_"), "shots_hit")
  shots_fired_col <- paste0(prefix, ifelse(prefix == "", "", "_"), "shots_fired")
  crit_hit_col <- paste0(prefix, ifelse(prefix == "", "", "_"), "critical_hit_count")
  damage_col <- paste0(prefix, ifelse(prefix == "", "", "_"), "damage_dealt")
  sessions_col <- paste0(prefix, ifelse(prefix == "", "", "_"), "sessions")
  seconds_col <- paste0(prefix, ifelse(prefix == "", "", "_"), "seconds")
  score_col <- paste0(prefix, ifelse(prefix == "", "", "_"), "score")
  
  # Calculate and return metrics
  df %>%
    mutate(
      # KD ratio
      !!paste0(prefix, ifelse(prefix == "", "", "_"), "kd") := 
        .data[[kill_count_col]] / .data[[death_count_col]],
      
      # Hit accuracy
      !!paste0(prefix, ifelse(prefix == "", "", "_"), "hit_accuracy") := 
        .data[[shots_hit_col]] / .data[[shots_fired_col]],
      
      # Critical hit ratio
      !!paste0(prefix, ifelse(prefix == "", "", "_"), "crit_ratio") := 
        .data[[crit_hit_col]] / .data[[shots_hit_col]],
      
      # Damage per second
      !!paste0(prefix, ifelse(prefix == "", "", "_"), "dps") := 
        .data[[damage_col]] / .data[[seconds_col]],
      
      # Damage per game
      !!paste0(prefix, ifelse(prefix == "", "", "_"), "average_damage") := 
        .data[[damage_col]] / .data[[sessions_col]],
      
      # Score per second
      !!paste0(prefix, ifelse(prefix == "", "", "_"), "sps") := 
        .data[[score_col]] / .data[[seconds_col]],
      
      # Score per game
      !!paste0(prefix, ifelse(prefix == "", "", "_"), "average_score") := 
        .data[[score_col]] / .data[[sessions_col]]
    ) %>%
    # Replace NaN with 0
    mutate(
      across(
        everything(), 
        ~ ifelse(is.nan(.x), 0, .x)
      )
    )
}

# Function to summarize mode-specific playing time
summarize_mode_hours <- function(df, mode_col, start_col = "game_start_time", end_col = "game_end_time") {
  df %>%
    filter(.data[[mode_col]] == 1) %>%
    mutate(duration = .data[[end_col]] - .data[[start_col]]) %>%
    mutate(seconds = as.numeric(duration)) %>%
    group_by(player_id) %>%
    summarise(
      seconds = sum(seconds, na.rm = TRUE),
      .groups = "drop"
    ) %>%
    mutate(hours = seconds / 60 / 60) %>%
    rename_with(
      ~ paste0(gsub("_session", "", mode_col), "_", .),
      c(seconds, hours)
    )
}

# Function to calculate SPANE scales
calculate_spane <- function(df, prefix = "spane") {
  positive_items <- c(1, 3, 5, 7, 10, 12)
  negative_items <- c(2, 4, 6, 8, 9, 11)
  
  df %>%
    mutate(
      # Positive affect
      !!paste0(prefix, "_positive") := rowMeans(
        select(
          .,
          paste0(prefix, "_", positive_items)
        ),
        na.rm = TRUE
      ),
      # Negative affect
      !!paste0(prefix, "_negative") := rowMeans(
        select(
          .,
          paste0(prefix, "_", negative_items)
        ),
        na.rm = TRUE
      ),
      # Affect balance
      !!paste0(prefix, "_balance") := .data[[paste0(prefix, "_positive")]] - 
        .data[[paste0(prefix, "_negative")]]
    )
}

# Function to create motivation scales
calculate_motivation_scales <- function(df) {
  df %>%
    mutate(
      autonomy = rowMeans(
        select(., starts_with("autonomy_")), 
        na.rm = TRUE
      ),
      competence = rowMeans(
        select(., starts_with("competence_")), 
        na.rm = TRUE
      ),
      relatedness = rowMeans(
        select(., starts_with("relatedness_")), 
        na.rm = TRUE
      ),
      enjoyment = rowMeans(
        select(., starts_with("enjoyment_")), 
        na.rm = TRUE
      ),
      extrinsic = rowMeans(
        select(., starts_with("extrinsic_")), 
        na.rm = TRUE
      )
    )
}
```

## Data Source and Download Instructions

The data comes from a study conducted by EA and the Oxford Internet Institute investigating the relationship between video game play and well-being.

**Data Repository**: <https://osf.io/cjd6z/>

**Direct Download**: <https://osf.io/cjd6z/files/4gp3r>

### To Download the Data:

1.  Visit the OSF project page: <https://osf.io/cjd6z/>
2.  Download the complete dataset from: <https://osf.io/cjd6z/files/4gp3r>
3.  Extract the downloaded files to `data/raw/ea/` in the project directory
4.  The data directory should contain:
    -   `Oxford PvZ - Pilot Wave + Wave 1 Download.xlsx` (survey data)
    -   `OII_PVZ_Authentications_*.csv` (authentication logs)
    -   `OII_PVZ_Character_*.csv` (character performance data)
    -   `OII_PVZ_Game_Time_*.csv` (session timing data)

### Citation

If you use this data, please cite the original study:

> Johannes, N., Vuorre, M., Magnusson, K., & Przybylski, A. K. (2021). Video game play is positively correlated with well-being. *Royal Society Open Science*, 8(2), 202049. <https://doi.org/10.1098/rsos.202049>

## Load Raw Data

```{r raw-data-paths}
#| label: raw-data-paths

# Store all file paths in a central location for easy updates
data_paths <- list(
  survey = here("data/raw/ea/Oxford PvZ - Pilot Wave + Wave 1 Download.xlsx"),
  telemetry_dir = here("data/raw/ea")
)

# Check if data directory exists
if (!dir.exists(data_paths$telemetry_dir)) {
  stop(
    "Data directory not found: ", data_paths$telemetry_dir, "\n",
    "Please download the data from: https://osf.io/cjd6z/files/4gp3r\n",
    "Extract to: data/raw/ea/"
  )
}

# Check if survey file exists
if (!file.exists(data_paths$survey)) {
  stop(
    "Survey file not found: ", data_paths$survey, "\n",
    "Please download the data from: https://osf.io/cjd6z/files/4gp3r\n",
    "Extract to: data/raw/ea/"
  )
}

# List available telemetry files
telemetry_files <- list.files(
  data_paths$telemetry_dir, 
  pattern = "OII_PVZ_", 
  full.names = TRUE
)

# Validate that we have the required telemetry files
required_patterns <- c("Authentications", "Character", "Game_Time")
for (pattern in required_patterns) {
  matching_files <- grep(pattern, telemetry_files, value = TRUE)
  if (length(matching_files) == 0) {
    stop(
      "Missing required telemetry file matching pattern: ", pattern, "\n",
      "Please download the complete dataset from: https://osf.io/cjd6z/files/4gp3r"
    )
  }
}

cat("✓ All required data files found\n")
cat("Survey file:", data_paths$survey, "\n")
cat("Telemetry files:", length(telemetry_files), "files\n")
```

## Load Survey Data

```{r survey-load}
#| label: survey-load

survey <- read_excel(data_paths$survey)

# Validate survey data structure
expected_cols <- c("UID", "date", "Q1", "Q2", "Q3", "hCountry")
missing_cols <- setdiff(expected_cols, names(survey))
if (length(missing_cols) > 0) {
  stop(
    "Survey data is missing expected columns: ", 
    paste(missing_cols, collapse = ", "), "\n",
    "Please verify you have the correct survey file."
  )
}

survey <- survey %>% 
  mutate(
    # Ensure date data type
    date = mdy_hm(date),
    # Additional date variable for filtering telemetry
    start_date = date
  ) %>% 
  # Rename ID column to match telemetry files
  rename(player_id = UID)

cat("✓ Survey data loaded:", nrow(survey), "participants\n")
```

## Clean Survey Data

```{r survey-clean}
#| label: survey-clean

# Rename variables with meaningful names
survey <- 
  survey |> 
  select(
    # meta-info
    player_id,
    start_date, date,
    survey_duration = qtime,
    of_age = Q2,
    consent = Q3,
    
    # demographics
    played = Q1,
    country = hCountry,
    survey_language = PRELANG,
    age = D1,
    gender = D2,
    gender_other = D2r3oe,
    experience = D3,
    
    # well-being
    starts_with("A1"), # general well-being in the two weeks
    straightliner_spane = hA1Straight,
    
    # motivations
    played_with_others = NSat1, # filter question for relatedness items
    NSat2r1:NSat2r3,
    NSat2r4:NSat2r6,
    NSat2r7:NSat2r9,
    NSat2r10:NSat2r13,
    NSat2r14:NSat2r17,
    straightliner_motivations = hNSat2Straight,
    
    # self-reported play
    active_play_hours = GT1c1,
    active_play_minutes = GT1c2,
    within_estimate = GT2,
    between_estimate = GT3,
    starts_with("GT4")
  ) |>
  rename_with(~str_replace(.x, "A1r", "spane_"), starts_with("A1")) |>
  rename_with(~str_replace(.x, "NSat2r", "autonomy_"), NSat2r1:NSat2r3) |>
  rename_with(~str_replace(.x, "NSat2r", "competence_"), NSat2r4:NSat2r6) |>
  rename_with(~str_replace(.x, "NSat2r", "relatedness_"), NSat2r7:NSat2r9) |>
  rename_with(~str_replace(.x, "NSat2r", "enjoyment_"), NSat2r10:NSat2r13) |>
  rename_with(~str_replace(.x, "NSat2r", "extrinsic_"), NSat2r14:NSat2r17) |>
  rename_with(~str_replace(.x, "GT4r", "spane_game_"), starts_with("GT4"))

# Assign proper variable types and factor levels
survey <- survey |>   
  mutate(
    across(
      c(
        player_id,
        of_age:survey_language,
        gender,
        played_with_others
      ),
      as.factor
    ),
    
    across(
      c(of_age:played, played_with_others),
      ~ fct_recode(
        .x,
        "Yes" = "1",
        "No" = "2"
      )
    ),
    
    gender = fct_recode(
      gender,
      "Male" = "1",
      "Female" = "2",
      "Other" = "3",
      "Prefer not to say" = "4"
    ),
    
    country = fct_recode(
      country,
      "US" = "1",
      "UK" = "2",
      "Canada" = "5"
    ),
    
    survey_language = fct_recode(
      survey_language,
      "English" = "1",
      "French" = "2"
    )
  )

# Apply reverse scoring (8 - x) for items where original coding was inverted
# After reversal, higher values = more frequent / more positive for all items
survey <- survey |> 
  mutate(
    across(
      c(
        starts_with("spane"),   # All SPANE items: original 1 = very often, 7 = never
        between_estimate,       # Original 1 = much more than typically
        within_estimate,        # Original 1 = much more than typically
        relatedness_9,          # 3rd relatedness item (NSat2r9) — negatively worded
        enjoyment_11,           # 2nd enjoyment item (NSat2r11) — negatively worded
        enjoyment_12            # 3rd enjoyment item (NSat2r12) — negatively worded
      ),
      ~ 8 - .x
    )
  )

# Create indicators for straightliners
survey <- survey |> 
  mutate(across(contains("straightliner"), ~.x==1))
```

## Create Survey Scales

```{r survey-scales}
#| label: survey-scales

# General SPANE scale (well-being)
survey <- survey |> 
  mutate(
    spane_positive = rowMeans(pick(spane_1, spane_3, spane_5, spane_7, spane_10, spane_12), na.rm = TRUE),
    spane_negative = rowMeans(pick(spane_2, spane_4, spane_6, spane_8, spane_9, spane_11), na.rm = TRUE),
    spane_balance = spane_positive - spane_negative
  )

# Motivation scales
survey <- survey |> 
  mutate(
    autonomy = rowMeans(pick(starts_with("autonomy")), na.rm = TRUE),
    competence = rowMeans(pick(starts_with("competence")), na.rm = TRUE),
    relatedness = rowMeans(pick(starts_with("relatedness")), na.rm = TRUE),
    enjoyment = rowMeans(pick(starts_with("enjoyment")), na.rm = TRUE),
    extrinsic = rowMeans(pick(starts_with("extrinsic")), na.rm = TRUE)
  )

# SPANE because of playing game
survey <- survey |> 
  mutate(
    spane_game_positive = rowMeans(pick(spane_game_1, spane_game_3, spane_game_5, spane_game_7, spane_game_10, spane_game_12), na.rm = TRUE),
    spane_game_negative = rowMeans(pick(spane_game_2, spane_game_4, spane_game_6, spane_game_8, spane_game_9, spane_game_11), na.rm = TRUE),
    spane_game_balance = spane_game_positive - spane_game_negative
  )

# Hours of estimated play
survey <- survey |>
  mutate(
    active_play_minutes = active_play_minutes / 60,
    active_play = active_play_hours + active_play_minutes
  )

# Save processed survey data
write_csv(survey, here("data/processed/survey_processed.csv"))
```

# Telemetry Data Processing

## Load Game Time Data

```{r game-time-load}
#| label: game-time-load

game_time_raw <- read_both_waves("OII_PVZ_Game_Time") %>%
  # Only include survey participants 
  filter(player_id %in% survey$player_id)
```

## Process Game Time Data

```{r game-time-process}
#| label: game-time-process

# Handle missing start times
game_time <- game_time_raw %>% 
  mutate(
    # If start time is missing, set to end time (will result in 0 duration)
    game_start_time = coalesce(game_start_time, game_end_time)
  ) %>%
  # Collapse rows by session
  group_by(player_id, game_session) %>% 
  mutate(
    game_start_time = min(game_start_time),
    game_end_time = max(game_end_time),
  ) %>% 
  # Take the first of each (identical) person-session combo
  slice(1) %>% 
  ungroup()

# Filter to sessions within 14 days before survey
game_time_2week <- game_time %>% 
  left_join(select(survey, player_id, start_date)) %>% 
  filter(game_start_time >= (start_date - days(14)))

# NOTE: In later processing, we handle the fact that multiple characters can be
# used in a single game session by ensuring each session is counted only once in our
# game time calculations. We also cap unrealistically long sessions to 12 hours.

# Save processed game time data
write_csv(game_time_2week, here("data/processed/game_time_2week.csv"))
```

## Load Authentication Data

```{r auth-load}
#| label: auth-load

authentications_raw <- read_both_waves("OII_PVZ_Authentications") %>%
  # Only include survey participants
  filter(player_id %in% survey$player_id)

# Filter to 14 days before survey
authentications_2week <- authentications_raw %>% 
  left_join(select(survey, player_id, start_date)) %>% 
  filter(date >= (as_date(start_date) - days(14)))

# Count authentications per player
authentications_players <- authentications_2week %>% 
  count(player_id, name = "authentications")

# Save processed authentication data
write_csv(authentications_players, here("data/processed/authentications_players.csv"))
```

## Load Character Performance Data

```{r character-load}
#| label: character-load

characters_raw <- read_both_waves("OII_PVZ_Character") %>%
  # Only include survey participants
  filter(player_id %in% survey$player_id)

# Add session timing info and filter to 14 days before survey
characters_2week <- characters_raw %>%
  left_join(
    game_time_2week %>% 
      select(player_id, game_session, game_start_time, game_end_time),
    by = c("player_id", "game_session")
  ) %>%
  left_join(select(survey, player_id, start_date)) %>% 
  filter(game_start_time >= (start_date - days(14)))

# Save for further processing
write_csv(characters_2week, here("data/processed/characters_2week.csv"))
```

# Game Mode Classification

```{r game-mode-classification}
#| label: game-mode-classification

# Create game mode classification variables
characters_2week <- characters_2week %>%
  mutate(
    # Social hub sessions (no combat)
    social_hub = case_when(
      game_level_key == "dsub_social_terrain_summer" & is.na(score) ~ 1,
      TRUE ~ 0
    ),
    
    # PvE sessions
    pve_session = case_when(
      game_level_key %in% c(
        "dsub_region_cheesemines", 
        "dsub_region_dreadwood", 
        "dsub_region_towncenter"
      ) ~ 1,
      TRUE ~ 0
    ),
    
    # PvP sessions
    pvp_session = case_when(
      game_level_key %in% c(
        "dsub_rush_dreadwood", "dsub_rush_towncenter", 
        "dsub_tvm_cheesemines", "dsub_tvm_zrefinery", 
        "dsub_tvm_suburbs", "dsub_tvm_dreadwood", 
        "dsub_rush_farm", "dsub_rush_cheesemines", 
        "dsub_rush_shoreline", "dsub_rush_resort", 
        "dsub_4v4_battlearena_v2", "dsub_4v4_superarena"
      ) ~ 1,
      # Giddy Park with combat data is also PvP
      game_level_key == "dsub_social_terrain_summer" & !is.na(score) ~ 1,
      TRUE ~ 0
    )
  )

# Remove sessions without performance data (mainly social hub)
characters_2week_nona <- characters_2week %>% 
  drop_na(total_kill_count:shots_hit)

# Save processed data
write_csv(characters_2week, here("data/processed/characters_with_modes.csv"))
write_csv(characters_2week_nona, here("data/processed/characters_with_modes_nona.csv"))
```

# Game Time Calculations

```{r game-time-calculations}
#| label: game-time-calculations

# Define a function to calculate time per player for a specific game mode
calculate_game_time <- function(data, mode_column, time_prefix) {
  # Apply mode filter directly using the column name as a string
  filtered_data <- data %>%
    filter(!!rlang::sym(mode_column) == 1)
  
  # Calculate time
  filtered_data %>%
    mutate(duration = game_end_time - game_start_time) %>%
    mutate(seconds = as.numeric(duration)) %>%
    group_by(player_id) %>%
    summarise(
      seconds = sum(seconds, na.rm = TRUE)
    ) %>%
    mutate(hours = seconds / 60 / 60) %>%
    rename_with(~ paste0(time_prefix, "_", .), .cols = c("seconds", "hours")) %>%
    ungroup()
}

# Get unique game sessions with their mode information
game_sessions_modes <- characters_2week %>%
  select(player_id, game_session, social_hub, pve_session, pvp_session, game_start_time, game_end_time) %>%
  # Group by player and session to get unique sessions with mode flags
  group_by(player_id, game_session) %>%
  summarise(
    # If any character in a session was used in a specific mode, count the session as that mode
    social_hub = max(social_hub, na.rm = TRUE),
    pve_session = max(pve_session, na.rm = TRUE),
    pvp_session = max(pvp_session, na.rm = TRUE),
    # Preserve the start and end times
    game_start_time = first(game_start_time),
    game_end_time = first(game_end_time),
    .groups = "drop"
  )

# Calculate total game time per player (only count each session once)
game_time_players_total <- game_sessions_modes %>% 
  mutate(duration = game_end_time - game_start_time) %>% 
  mutate(seconds = as.numeric(duration)) %>% 
  group_by(player_id) %>% 
  summarise(seconds = sum(seconds, na.rm = TRUE)) %>%
  mutate(hours = seconds / 60 / 60) %>% 
  ungroup()

# Calculate game time for different modes using our function
game_time_players_pvp <- calculate_game_time(game_sessions_modes, "pvp_session", "pvp")
game_time_players_pve <- calculate_game_time(game_sessions_modes, "pve_session", "pve")
game_time_players_social <- calculate_game_time(game_sessions_modes, "social_hub", "social")

# Combine game time data with purrr's reduce
game_time_players <- list(
  game_time_players_total, 
  game_time_players_social, 
  game_time_players_pve, 
  game_time_players_pvp
) %>%
  reduce(left_join, by = "player_id")

# Replace NA values with 0 (no time spent in that mode)
game_time_players[is.na(game_time_players)] <- 0

# Save processed game time data
write_csv(game_time_players, here("data/processed/game_time_players.csv"))
```

# Performance Metrics Calculation

```{r performance-setup}
#| label: performance-setup
#| cache: false
#| dependson: c("game-mode-classification", "game-time-calculations")

# Create an empty list to store intermediate results
metrics_list <- list()
```

```{r overall-performance}
#| label: overall-performance
#| cache: false
#| dependson: c("performance-setup")

# Calculate overall performance metrics per player
characters_players <- characters_2week %>%
  group_by(player_id) %>% 
  summarise(across(total_kill_count:shots_hit, sum, na.rm = TRUE)) %>% 
  ungroup()

# Store in our list
metrics_list$characters_players <- characters_players

# Force garbage collection
gc()
```

```{r game-sessions}
#| label: game-sessions
#| cache: false
#| dependson: c("performance-setup")

# Calculate game sessions per player
game_sessions_total <- game_sessions_modes %>% 
  count(player_id) %>%
  rename(n_total_game_sessions = n)

# Store in our list
metrics_list$game_sessions_total <- game_sessions_total

# Force garbage collection
gc()
```

```{r sessions-by-mode}
#| label: sessions-by-mode
#| cache: false
#| dependson: c("performance-setup")

# Calculate sessions by game mode
game_sessions <- game_sessions_modes %>% 
  group_by(player_id) %>% 
  summarise(across(social_hub:pvp_session, sum, na.rm = TRUE)) %>% 
  ungroup()

# Store in our list
metrics_list$game_sessions <- game_sessions

# Force garbage collection
gc()
```

```{r pvp-performance}
#| label: pvp-performance
#| cache: false
#| dependson: c("performance-setup")

# PvP-specific performance metrics
pvp_telemetry <- characters_2week %>%
  filter(pvp_session == 1) %>%
  group_by(player_id) %>% 
  summarise(across(total_kill_count:shots_hit, sum, na.rm = TRUE)) %>%
  rename_with(~ paste0("pvp_", .), -player_id)

# Store in our list
metrics_list$pvp_telemetry <- pvp_telemetry

# Force garbage collection
gc()
```

```{r pve-performance}
#| label: pve-performance
#| cache: false
#| dependson: c("performance-setup")

# PvE-specific performance metrics
pve_telemetry <- characters_2week %>%
  filter(pve_session == 1) %>%
  group_by(player_id) %>% 
  summarise(across(total_kill_count:shots_hit, sum, na.rm = TRUE)) %>%
  rename_with(~ paste0("pve_", .), -player_id)

# Store in our list
metrics_list$pve_telemetry <- pve_telemetry

# Force garbage collection
gc()
```

```{r combine-metrics}
#| label: combine-metrics
#| cache: false
#| dependson: c("overall-performance", "game-sessions", "sessions-by-mode", "pvp-performance", "pve-performance")

# Load additional data we need from previous processing
metrics_list$game_time_players <- game_time_players

# Combine telemetry data using purrr's reduce function
telemetry_players <- list(
  metrics_list$characters_players,
  metrics_list$game_time_players,
  metrics_list$game_sessions,
  metrics_list$game_sessions_total,
  metrics_list$pvp_telemetry,
  metrics_list$pve_telemetry
) %>%
  reduce(left_join, by = "player_id")

# Replace NA values with 0
telemetry_players <- telemetry_players %>%
  mutate(across(everything(), ~replace_na(.x, 0)))

# Print summary of capped sessions
cat("Summary of capped sessions (>12 hours):\n")
cat("Total capped sessions:", sum(telemetry_players$capped_sessions), "\n")
cat("Players with capped sessions:", sum(telemetry_players$capped_sessions > 0), "\n")
cat("Maximum capped sessions per player:", max(telemetry_players$capped_sessions), "\n")

# Clear intermediate objects and list to free memory
rm(metrics_list)
gc()

# Save processed telemetry data
write_csv(telemetry_players, here("data/processed/telemetry_players.csv"))
```

# Derived Performance Metrics

```{r derived-metrics}
#| label: derived-metrics

# Calculate derived performance metrics
telemetry_players <- telemetry_players %>%
  # Handle divisions by zero for K/D ratio
  mutate(
    across(
      c(total_death_count, pvp_total_death_count, pve_total_death_count),
      ~ ifelse(.x == 0, 1, .x)
    )
  ) %>%
  # Calculate all ratio metrics
  mutate(
    # K/D ratios
    total_kd = total_kill_count / total_death_count,
    pve_kd = pve_total_kill_count / pve_total_death_count,
    pvp_kd = pvp_total_kill_count / pvp_total_death_count,
    
    # Hit accuracy ratios
    total_hit_accuracy = shots_hit / shots_fired,
    pve_hit_accuracy = pve_shots_hit / pve_shots_fired,
    pvp_hit_accuracy = pvp_shots_hit / pvp_shots_fired,
    
    # Critical hit ratios
    total_crit_ratio = critical_hit_count / shots_hit,
    pve_crit_ratio = pve_critical_hit_count / pve_shots_hit,
    pvp_crit_ratio = pvp_critical_hit_count / pvp_shots_hit,
    
    # Damage per game
    total_average_damage = damage_dealt / n_total_game_sessions,
    pve_average_damage = pve_damage_dealt / pve_session,
    pvp_average_damage = pvp_damage_dealt / pvp_session,
    
    # Damage per second
    total_dps = damage_dealt / seconds,
    pve_dps = pve_damage_dealt / pve_seconds,
    pvp_dps = pvp_damage_dealt / pvp_seconds,
    
    # Score per game
    total_average_score = score / n_total_game_sessions,
    pve_average_score = pve_score / pve_session,
    pvp_average_score = pvp_score / pvp_session,
    
    # Score per second
    total_sps = score / seconds,
    pve_sps = pve_score / pve_seconds,
    pvp_sps = pvp_score / pvp_seconds
  )

# Replace NaN values with 0
telemetry_players <- telemetry_players %>%
  mutate(across(everything(), ~ifelse(is.nan(.x), 0, .x)))

# Save processed metrics
write_csv(telemetry_players, here("data/processed/telemetry_with_metrics.csv"))
```

# Participants and Procedure

# Combine Survey and Telemetry Data

```{r merge-data}
#| label: merge-data

# Merge survey with telemetry data
pvz_df <- survey %>% 
  right_join(telemetry_players, by = "player_id")

# Check for participants without valid performance metrics
# These would be participants with zeros or NAs for all key performance indicators
pvz_df <- pvz_df %>%
  # Create a flag for participants with no meaningful telemetry data
  mutate(
    has_telemetry = !(
      # Check if key performance metrics are all zeros or NAs
      (total_kill_count == 0 | is.na(total_kill_count)) &
      (total_death_count == 0 | is.na(total_death_count)) &
      (damage_dealt == 0 | is.na(damage_dealt)) &
      (shots_fired == 0 | is.na(shots_fired)) &
      (n_total_game_sessions == 0 | is.na(n_total_game_sessions))
    )
  )

# Print summary of participants with and without telemetry
cat("Participants without valid telemetry:", sum(!pvz_df$has_telemetry, na.rm = TRUE), "\n")
cat("Participants with valid telemetry:", sum(pvz_df$has_telemetry, na.rm = TRUE), "\n")

# Remove participants without valid telemetry data
pvz_df_clean_final <- pvz_df %>%
  filter(has_telemetry == TRUE)

# Calculate ratio metrics
pvz_df_clean_final <- pvz_df_clean_final %>%
  mutate(
    ratio_social_hours = social_hours / hours,
    ratio_pve_hours = pve_hours / hours,
    ratio_pvp_hours = pvp_hours / hours
  )

# Keep participants with ANY measurable playtime (>0 hours)

pvz_df_clean_final <- pvz_df_clean_final %>%
  filter(hours > 0)

# Print final dataset dimensions
cat("Final dataset dimensions:", dim(pvz_df_clean_final)[1], "participants x", dim(pvz_df_clean_final)[2], "variables\n")

# Save final dataset
write_csv(pvz_df_clean_final, here("data/processed/pvz_complete.csv"))

# Update pvz_df_clean to match the final version for consistency
pvz_df_clean <- pvz_df_clean_final
```

## Sample Construction Summary

# Outlier Handling

```{r outlier-removal}
#| label: outlier-removal

# Get list of numeric variables to check for outliers
numeric_vars <- pvz_df_clean_final %>%
  select(where(is.numeric)) %>%
  colnames()

# Calculate z-scores for each numeric variable
pvz_with_z_final <- pvz_df_clean_final %>%
  mutate(across(all_of(numeric_vars), 
                ~ scale(.x)[,1],
                .names = "{.col}_z"))

# Create a list of variables with outliers (>6 SD)
z_cols_final <- colnames(pvz_with_z_final)[grep("_z$", colnames(pvz_with_z_final))]
outlier_vars_final <- c()
outlier_counts <- c()

# Count outliers per variable (using 6 SD threshold)
for(z_col in z_cols_final) {
  orig_col <- sub("_z$", "", z_col)
  outlier_count <- sum(abs(pvz_with_z_final[[z_col]]) > 6, na.rm = TRUE)
  
  if(outlier_count > 0) {
    outlier_vars_final <- c(outlier_vars_final, orig_col)
    outlier_counts <- c(outlier_counts, outlier_count)
  }
}

# Create a summary table of outliers
outlier_summary <- tibble(
  name = outlier_vars_final,
  Extremes = outlier_counts,
  Extremes_p = scales::percent(outlier_counts / nrow(pvz_df_clean_final), accuracy = .01)
)

# Display the summary
print(outlier_summary)

# Create a cleaned version with outliers replaced by NA
pvz_clean_final <- pvz_df_clean_final

# Replace outliers with NA
for(i in seq_along(outlier_vars_final)) {
  var <- outlier_vars_final[i]
  z_var <- paste0(var, "_z")
  pvz_clean_final[[var]] <- ifelse(abs(pvz_with_z_final[[z_var]]) > 6, NA, pvz_df_clean_final[[var]])
}

# Save the cleaned dataset
write_csv(pvz_clean_final, here("data/pvz_clean_final.csv"))

# Print summary statistics
cat("\nOutlier handling summary:\n")
cat("Threshold: 6 SD (aligned with original paper)\n")
cat("Number of variables with extreme outliers:", length(outlier_vars_final), "\n")
cat("Total number of data points identified as outliers:", sum(outlier_counts), "\n")

# Count participants with at least one outlier
participants_with_outliers <- c()
for(i in seq_along(outlier_vars_final)) {
  var <- outlier_vars_final[i]
  z_var <- paste0(var, "_z")
  participants_with_outliers <- c(
    participants_with_outliers, 
    pvz_with_z_final$player_id[abs(pvz_with_z_final[[z_var]]) > 6]
  )
}

cat("Number of participants with at least one extreme value:", 
    length(unique(participants_with_outliers)), "\n")
cat("\nFinal output saved: data/pvz_clean_final.csv (n =", nrow(pvz_clean_final), ")\n")
```

# Next Steps

The preprocessing is complete. The processed data has been saved to `data/pvz_clean_final.csv` with n = 470 participants.

**Key preprocessing decisions:** - **Outlier threshold**: 6 SD (aligned with original paper) - **Playtime filter**: hours \> 0 (any measurable playtime) - **Telemetry requirement**: Valid combat/gameplay metrics

The main analysis script (`manuscript.qmd`) will apply additional filters: 1. Combat engagement/performance filter (removes 49 participants with no usable non-zero combat/mode indicators) 2. Mode-specific completeness filter (removes 8 participants missing required clustering metrics) 3. Final analysis sample: n = 413