---
title: "Player Playstyle Data Preprocessing"
author: "Thomas Hakman"
format: 
  html:
    toc: true
    toc-depth: 3
    code-fold: true
    theme: flatly
    fig-width: 8
    fig-height: 6
execute:
  echo: true
  message: false
  warning: false
  cache: false
editor: visual
---

```{r setup}
#| label: setup
#| include: false

# Set global options
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  cache = FALSE,
  fig.path = "figures/"
)
```

# Introduction

This document processes player telemetry and survey data from Plants vs. Zombies: Battle for Neighborville. The preprocessing steps prepare the data for analysis of the relationship between playtime, playstyle, and well-being outcomes.

# Data Preparation

## Load Packages

```{r packages}
#| label: load-packages
# Install pacman if not already installed
if (!require("pacman")) install.packages("pacman")

# Load required packages using pacman
pacman::p_load(
  tidyverse,  # For data manipulation and visualization
  here,       # For project-relative file paths
  lubridate,  # For date/time handling
  scales,     # For visualization scaling
  janitor,    # For data cleaning
  magrittr,   # For additional pipe operators
  psych,      # For psychological statistics
  readxl,     # For reading Excel files
  knitr       # For dynamic report generation
)
```

## Load Helper Functions

```{r helper-functions}
#| label: helper-functions

# Helper functions for PvZ data analysis
# These functions support the main preprocessing steps

# Function to read both waves of telemetry data into a single table
read_both_waves <- function(stub = "OII_PVZ_Authentications", data_dir = "data/raw/ea", ...) {
  # Find files matching the pattern
  files <- list.files(
    here::here(data_dir), pattern = stub, full.names = TRUE
  )
  
  # Validate that files were found
  if (length(files) == 0) {
    stop(
      "No files found matching pattern: ", stub, " in ", data_dir, "\n",
      "Please download data from: https://osf.io/cjd6z/files/4gp3r"
    )
  }
  
  # Filenames of wave 1 and 2 files
  files %>% 
    # Put wave 1 first (they are in reverse order)
    rev() %>% 
    # Read into a list of data frames
    purrr::map(readr::read_csv, ...) %>% 
    # Bind into one data frame with wave ID
    dplyr::bind_rows(.id = "wave")
}

# Function to create performance metrics from player data
calculate_performance_metrics <- function(df, prefix = "total") {
  # Handle cases where denominators might be zero
  df <- df %>%
    dplyr::mutate(
      # Set death count to 1 if 0 to avoid division by zero
      dplyr::across(
        dplyr::contains("death_count"), 
        ~ ifelse(.x == 0, 1, .x)
      )
    )
  
  # Create column names based on prefix
  kill_count_col <- paste0(prefix, ifelse(prefix == "", "", "_"), "kill_count")
  death_count_col <- paste0(prefix, ifelse(prefix == "", "", "_"), "death_count")
  shots_hit_col <- paste0(prefix, ifelse(prefix == "", "", "_"), "shots_hit")
  shots_fired_col <- paste0(prefix, ifelse(prefix == "", "", "_"), "shots_fired")
  crit_hit_col <- paste0(prefix, ifelse(prefix == "", "", "_"), "critical_hit_count")
  damage_col <- paste0(prefix, ifelse(prefix == "", "", "_"), "damage_dealt")
  sessions_col <- paste0(prefix, ifelse(prefix == "", "", "_"), "sessions")
  seconds_col <- paste0(prefix, ifelse(prefix == "", "", "_"), "seconds")
  score_col <- paste0(prefix, ifelse(prefix == "", "", "_"), "score")
  
  # Calculate and return metrics
  df %>%
    dplyr::mutate(
      # KD ratio
      !!paste0(prefix, ifelse(prefix == "", "", "_"), "kd") := 
        .data[[kill_count_col]] / .data[[death_count_col]],
      
      # Hit accuracy
      !!paste0(prefix, ifelse(prefix == "", "", "_"), "hit_accuracy") := 
        .data[[shots_hit_col]] / .data[[shots_fired_col]],
      
      # Critical hit ratio
      !!paste0(prefix, ifelse(prefix == "", "", "_"), "crit_ratio") := 
        .data[[crit_hit_col]] / .data[[shots_hit_col]],
      
      # Damage per second
      !!paste0(prefix, ifelse(prefix == "", "", "_"), "dps") := 
        .data[[damage_col]] / .data[[seconds_col]],
      
      # Damage per game
      !!paste0(prefix, ifelse(prefix == "", "", "_"), "average_damage") := 
        .data[[damage_col]] / .data[[sessions_col]],
      
      # Score per second
      !!paste0(prefix, ifelse(prefix == "", "", "_"), "sps") := 
        .data[[score_col]] / .data[[seconds_col]],
      
      # Score per game
      !!paste0(prefix, ifelse(prefix == "", "", "_"), "average_score") := 
        .data[[score_col]] / .data[[sessions_col]]
    ) %>%
    # Replace NaN with 0
    dplyr::mutate(
      dplyr::across(
        tidyselect::everything(), 
        ~ ifelse(is.nan(.x), 0, .x)
      )
    )
}

# Function to summarize mode-specific playing time
summarize_mode_hours <- function(df, mode_col, start_col = "game_start_time", end_col = "game_end_time") {
  df %>%
    dplyr::filter(.data[[mode_col]] == 1) %>%
    dplyr::mutate(duration = .data[[end_col]] - .data[[start_col]]) %>%
    dplyr::mutate(seconds = as.numeric(duration)) %>%
    dplyr::group_by(player_id) %>%
    dplyr::summarise(
      seconds = sum(seconds, na.rm = TRUE),
      .groups = "drop"
    ) %>%
    dplyr::mutate(hours = seconds / 60 / 60) %>%
    dplyr::rename_with(
      ~ paste0(gsub("_session", "", mode_col), "_", .),
      c(seconds, hours)
    )
}

# Function to calculate SPANE scales
calculate_spane <- function(df, prefix = "spane") {
  positive_items <- c(1, 3, 5, 7, 10, 12)
  negative_items <- c(2, 4, 6, 8, 9, 11)
  
  df %>%
    dplyr::mutate(
      # Positive affect
      !!paste0(prefix, "_positive") := rowMeans(
        dplyr::select(
          .,
          paste0(prefix, "_", positive_items)
        ),
        na.rm = TRUE
      ),
      # Negative affect
      !!paste0(prefix, "_negative") := rowMeans(
        dplyr::select(
          .,
          paste0(prefix, "_", negative_items)
        ),
        na.rm = TRUE
      ),
      # Affect balance
      !!paste0(prefix, "_balance") := .data[[paste0(prefix, "_positive")]] - 
        .data[[paste0(prefix, "_negative")]]
    )
}

# Function to create motivation scales
calculate_motivation_scales <- function(df) {
  df %>%
    dplyr::mutate(
      autonomy = rowMeans(
        dplyr::select(., dplyr::starts_with("autonomy_")), 
        na.rm = TRUE
      ),
      competence = rowMeans(
        dplyr::select(., dplyr::starts_with("competence_")), 
        na.rm = TRUE
      ),
      relatedness = rowMeans(
        dplyr::select(., dplyr::starts_with("relatedness_")), 
        na.rm = TRUE
      ),
      enjoyment = rowMeans(
        dplyr::select(., dplyr::starts_with("enjoyment_")), 
        na.rm = TRUE
      ),
      extrinsic = rowMeans(
        dplyr::select(., dplyr::starts_with("extrinsic_")), 
        na.rm = TRUE
      )
    )
}
```

## Data Source and Download Instructions

The data comes from a study conducted by EA and the Oxford Internet Institute investigating the relationship between video game play and well-being.

**Data Repository**: <https://osf.io/cjd6z/>

**Direct Download**: <https://osf.io/cjd6z/files/4gp3r>

### To Download the Data:

1.  Visit the OSF project page: <https://osf.io/cjd6z/>
2.  Download the complete dataset from: <https://osf.io/cjd6z/files/4gp3r>
3.  Extract the downloaded files to `data/raw/ea/` in the project directory
4.  The data directory should contain:
    -   `Oxford PvZ - Pilot Wave + Wave 1 Download.xlsx` (survey data)
    -   `OII_PVZ_Authentications_*.csv` (authentication logs)
    -   `OII_PVZ_Character_*.csv` (character performance data)
    -   `OII_PVZ_Game_Time_*.csv` (session timing data)

### Citation

If you use this data, please cite the original study:

> Johannes, N., Vuorre, M., Magnusson, K., & Przybylski, A. K. (2021). Video game play is positively correlated with well-being. *Royal Society Open Science*, 8(2), 202049. <https://doi.org/10.1098/rsos.202049>

## Load Raw Data

```{r raw-data-paths}
#| label: raw-data-paths

# Store all file paths in a central location for easy updates
data_paths <- list(
  survey = here("data/raw/ea/Oxford PvZ - Pilot Wave + Wave 1 Download.xlsx"),
  telemetry_dir = here("data/raw/ea")
)

# Check if data directory exists
if (!dir.exists(data_paths$telemetry_dir)) {
  stop(
    "Data directory not found: ", data_paths$telemetry_dir, "\n",
    "Please download the data from: https://osf.io/cjd6z/files/4gp3r\n",
    "Extract to: data/raw/ea/"
  )
}

# Check if survey file exists
if (!file.exists(data_paths$survey)) {
  stop(
    "Survey file not found: ", data_paths$survey, "\n",
    "Please download the data from: https://osf.io/cjd6z/files/4gp3r\n",
    "Extract to: data/raw/ea/"
  )
}

# List available telemetry files
telemetry_files <- list.files(
  data_paths$telemetry_dir, 
  pattern = "OII_PVZ_", 
  full.names = TRUE
)

# Validate that we have the required telemetry files
required_patterns <- c("Authentications", "Character", "Game_Time")
for (pattern in required_patterns) {
  matching_files <- grep(pattern, telemetry_files, value = TRUE)
  if (length(matching_files) == 0) {
    stop(
      "Missing required telemetry file matching pattern: ", pattern, "\n",
      "Please download the complete dataset from: https://osf.io/cjd6z/files/4gp3r"
    )
  }
}

cat("✓ All required data files found\n")
cat("Survey file:", data_paths$survey, "\n")
cat("Telemetry files:", length(telemetry_files), "files\n")
```

## Load Survey Data

```{r survey-load}
#| label: survey-load

survey <- read_excel(data_paths$survey)

# Validate survey data structure
expected_cols <- c("UID", "date", "Q1", "Q2", "Q3", "hCountry")
missing_cols <- setdiff(expected_cols, names(survey))
if (length(missing_cols) > 0) {
  stop(
    "Survey data is missing expected columns: ", 
    paste(missing_cols, collapse = ", "), "\n",
    "Please verify you have the correct survey file."
  )
}

survey <- survey %>% 
  mutate(
    # Ensure date data type
    date = mdy_hm(date),
    # Additional date variable for filtering telemetry
    start_date = date
  ) %>% 
  # Rename ID column to match telemetry files
  rename(player_id = UID)

cat("✓ Survey data loaded:", nrow(survey), "participants\n")
```

## Clean Survey Data

```{r survey-clean}
#| label: survey-clean

# Rename variables with meaningful names
survey <- 
  survey |> 
  dplyr::select(
    # meta-info
    player_id,
    start_date, date,
    survey_duration = qtime,
    of_age = Q2,
    consent = Q3,
    
    # demographics
    played = Q1,
    country = hCountry,
    survey_language = PRELANG,
    age = D1,
    gender = D2,
    gender_other = D2r3oe,
    experience = D3,
    
    # well-being
    starts_with("A1"), # general well-being in the two weeks
    straightliner_spane = hA1Straight,
    
    # motivations
    played_with_others = NSat1, # filter question for relatedness items
    NSat2r1:NSat2r3,
    NSat2r4:NSat2r6,
    NSat2r7:NSat2r9,
    NSat2r10:NSat2r13,
    NSat2r14:NSat2r17,
    straightliner_motivations = hNSat2Straight,
    
    # self-reported play
    active_play_hours = GT1c1,
    active_play_minutes = GT1c2,
    within_estimate = GT2,
    between_estimate = GT3,
    starts_with("GT4")
  ) |>
  rename_with(~str_replace(.x, "A1r", "spane_"), starts_with("A1")) |>
  rename_with(~str_replace(.x, "NSat2r", "autonomy_"), NSat2r1:NSat2r3) |>
  rename_with(~str_replace(.x, "NSat2r", "competence_"), NSat2r4:NSat2r6) |>
  rename_with(~str_replace(.x, "NSat2r", "relatedness_"), NSat2r7:NSat2r9) |>
  rename_with(~str_replace(.x, "NSat2r", "enjoyment_"), NSat2r10:NSat2r13) |>
  rename_with(~str_replace(.x, "NSat2r", "extrinsic_"), NSat2r14:NSat2r17) |>
  rename_with(~str_replace(.x, "GT4r", "spane_game_"), starts_with("GT4"))

# Assign proper variable types and factor levels
survey <- survey |>   
  mutate(
    across(
      c(
        player_id,
        of_age:survey_language,
        gender,
        played_with_others
      ),
      as.factor
    ),
    
    across(
      c(of_age:played, played_with_others),
      ~ fct_recode(
        .x,
        "Yes" = "1",
        "No" = "2"
      )
    ),
    
    gender = fct_recode(
      gender,
      "Male" = "1",
      "Female" = "2",
      "Other" = "3",
      "Prefer not to say" = "4"
    ),
    
    country = fct_recode(
      country,
      "US" = "1",
      "UK" = "2",
      "Canada" = "5"
    ),
    
    survey_language = fct_recode(
      survey_language,
      "English" = "1",
      "French" = "2"
    )
  )

# Apply reverse scoring for appropriate items
survey <- survey |> 
  mutate(
    across(
      c(
        starts_with("spane"), # 1 was never
        between_estimate, # those estimates also had 1: much more than typically
        within_estimate,
        relatedness_9,  # This is the 3rd relatedness item (NSat2r9)
        enjoyment_11,   # This is the 2nd enjoyment item (NSat2r11)
        enjoyment_12    # This is the 3rd enjoyment item (NSat2r12)
      ),
      ~ 8 - .x
    )
  )

# Create indicators for straightliners
survey <- survey |> 
  mutate(across(contains("straightliner"), ~.x==1))
```

## Create Survey Scales

```{r survey-scales}
#| label: survey-scales

# General SPANE scale (well-being)
survey <- survey |> 
  mutate(
    spane_positive = rowMeans(pick(spane_1, spane_3, spane_5, spane_7, spane_10, spane_12), na.rm = TRUE),
    spane_negative = rowMeans(pick(spane_2, spane_4, spane_6, spane_8, spane_9, spane_11), na.rm = TRUE),
    spane_balance = spane_positive - spane_negative
  )

# Motivation scales
survey <- survey |> 
  mutate(
    autonomy = rowMeans(pick(starts_with("autonomy")), na.rm = TRUE),
    competence = rowMeans(pick(starts_with("competence")), na.rm = TRUE),
    relatedness = rowMeans(pick(starts_with("relatedness")), na.rm = TRUE),
    enjoyment = rowMeans(pick(starts_with("enjoyment")), na.rm = TRUE),
    extrinsic = rowMeans(pick(starts_with("extrinsic")), na.rm = TRUE)
  )

# SPANE because of playing game
survey <- survey |> 
  mutate(
    spane_game_positive = rowMeans(pick(spane_game_1, spane_game_3, spane_game_5, spane_game_7, spane_game_10, spane_game_12), na.rm = TRUE),
    spane_game_negative = rowMeans(pick(spane_game_2, spane_game_4, spane_game_6, spane_game_8, spane_game_9, spane_game_11), na.rm = TRUE),
    spane_game_balance = spane_game_positive - spane_game_negative
  )

# Hours of estimated play
survey <- survey |>
  mutate(
    active_play_minutes = active_play_minutes / 60,
    active_play = active_play_hours + active_play_minutes
  )

# Save processed survey data
write_csv(survey, here("data/processed/survey_processed.csv"))
```

# Telemetry Data Processing

## Load Game Time Data

```{r game-time-load}
#| label: game-time-load

game_time_raw <- read_both_waves("OII_PVZ_Game_Time") %>%
  # Only include survey participants 
  filter(player_id %in% survey$player_id)
```

## Process Game Time Data

```{r game-time-process}
#| label: game-time-process

# Handle missing start times
game_time <- game_time_raw %>% 
  mutate(
    # If start time is missing, set to end time (will result in 0 duration)
    game_start_time = coalesce(game_start_time, game_end_time)
  ) %>%
  # Collapse rows by session
  group_by(player_id, game_session) %>% 
  mutate(
    game_start_time = min(game_start_time),
    game_end_time = max(game_end_time),
  ) %>% 
  # Take the first of each (identical) person-session combo
  slice(1) %>% 
  ungroup()

# Filter to sessions within 14 days before survey
game_time_2week <- game_time %>% 
  left_join(dplyr::select(survey, player_id, start_date)) %>% 
  filter(game_start_time >= (start_date - days(14)))

# NOTE: In later processing, we handle the fact that multiple characters can be
# used in a single game session by ensuring each session is counted only once in our
# game time calculations. We also cap unrealistically long sessions to 12 hours.

# Save processed game time data
write_csv(game_time_2week, here("data/processed/game_time_2week.csv"))
```

## Load Authentication Data

```{r auth-load}
#| label: auth-load

authentications_raw <- read_both_waves("OII_PVZ_Authentications") %>%
  # Only include survey participants
  filter(player_id %in% survey$player_id)

# Filter to 14 days before survey
authentications_2week <- authentications_raw %>% 
  left_join(dplyr::select(survey, player_id, start_date)) %>% 
  filter(date >= (as_date(start_date) - days(14)))

# Count authentications per player
authentications_players <- authentications_2week %>% 
  count(player_id, name = "authentications")

# Save processed authentication data
write_csv(authentications_players, here("data/processed/authentications_players.csv"))
```

## Load Character Performance Data

```{r character-load}
#| label: character-load

characters_raw <- read_both_waves("OII_PVZ_Character") %>%
  # Only include survey participants
  filter(player_id %in% survey$player_id)

# Add session timing info and filter to 14 days before survey
characters_2week <- characters_raw %>%
  left_join(
    game_time_2week %>% 
      dplyr::select(player_id, game_session, game_start_time, game_end_time),
    by = c("player_id", "game_session")
  ) %>%
  left_join(dplyr::select(survey, player_id, start_date)) %>% 
  filter(game_start_time >= (start_date - days(14)))

# Save for further processing
write_csv(characters_2week, here("data/processed/characters_2week.csv"))
```

# Game Mode Classification

```{r game-mode-classification}
#| label: game-mode-classification

# Create game mode classification variables
characters_2week <- characters_2week %>%
  mutate(
    # Social hub sessions (no combat)
    social_hub = case_when(
      game_level_key == "dsub_social_terrain_summer" & is.na(score) ~ 1,
      TRUE ~ 0
    ),
    
    # PvE sessions
    pve_session = case_when(
      game_level_key %in% c(
        "dsub_region_cheesemines", 
        "dsub_region_dreadwood", 
        "dsub_region_towncenter"
      ) ~ 1,
      TRUE ~ 0
    ),
    
    # PvP sessions
    pvp_session = case_when(
      game_level_key %in% c(
        "dsub_rush_dreadwood", "dsub_rush_towncenter", 
        "dsub_tvm_cheesemines", "dsub_tvm_zrefinery", 
        "dsub_tvm_suburbs", "dsub_tvm_dreadwood", 
        "dsub_rush_farm", "dsub_rush_cheesemines", 
        "dsub_rush_shoreline", "dsub_rush_resort", 
        "dsub_4v4_battlearena_v2", "dsub_4v4_superarena"
      ) ~ 1,
      # Giddy Park with combat data is also PvP
      game_level_key == "dsub_social_terrain_summer" & !is.na(score) ~ 1,
      TRUE ~ 0
    )
  )

# Remove sessions without performance data (mainly social hub)
characters_2week_nona <- characters_2week %>% 
  drop_na(total_kill_count:shots_hit)

# Save processed data
write_csv(characters_2week, here("data/processed/characters_with_modes.csv"))
write_csv(characters_2week_nona, here("data/processed/characters_with_modes_nona.csv"))
```

# Game Time Calculations

```{r game-time-calculations}
#| label: game-time-calculations

# Define a function to calculate time per player for a specific game mode
calculate_game_time <- function(data, mode_column, time_prefix) {
  # Apply mode filter directly using the column name as a string
  filtered_data <- data %>%
    filter(!!rlang::sym(mode_column) == 1)
  
  # Calculate time
  filtered_data %>%
    mutate(duration = game_end_time - game_start_time) %>%
    mutate(seconds = as.numeric(duration)) %>%
    group_by(player_id) %>%
    summarise(
      seconds = sum(seconds, na.rm = TRUE)
    ) %>%
    mutate(hours = seconds / 60 / 60) %>%
    rename_with(~ paste0(time_prefix, "_", .), .cols = c("seconds", "hours")) %>%
    ungroup()
}

# Get unique game sessions with their mode information
game_sessions_modes <- characters_2week %>%
  dplyr::select(player_id, game_session, social_hub, pve_session, pvp_session, game_start_time, game_end_time) %>%
  # Group by player and session to get unique sessions with mode flags
  group_by(player_id, game_session) %>%
  summarise(
    # If any character in a session was used in a specific mode, count the session as that mode
    social_hub = max(social_hub, na.rm = TRUE),
    pve_session = max(pve_session, na.rm = TRUE),
    pvp_session = max(pvp_session, na.rm = TRUE),
    # Preserve the start and end times
    game_start_time = first(game_start_time),
    game_end_time = first(game_end_time),
    .groups = "drop"
  )

# Calculate total game time per player (only count each session once)
game_time_players_total <- game_sessions_modes %>% 
  mutate(duration = game_end_time - game_start_time) %>% 
  mutate(seconds = as.numeric(duration)) %>% 
  group_by(player_id) %>% 
  summarise(seconds = sum(seconds, na.rm = TRUE)) %>%
  mutate(hours = seconds / 60 / 60) %>% 
  ungroup()

# Calculate game time for different modes using our function
game_time_players_pvp <- calculate_game_time(game_sessions_modes, "pvp_session", "pvp")
game_time_players_pve <- calculate_game_time(game_sessions_modes, "pve_session", "pve")
game_time_players_social <- calculate_game_time(game_sessions_modes, "social_hub", "social")

# Combine game time data with purrr's reduce
game_time_players <- list(
  game_time_players_total, 
  game_time_players_social, 
  game_time_players_pve, 
  game_time_players_pvp
) %>%
  reduce(left_join, by = "player_id")

# Replace NA values with 0 (no time spent in that mode)
game_time_players[is.na(game_time_players)] <- 0

# Save processed game time data
write_csv(game_time_players, here("data/processed/game_time_players.csv"))
```

# Performance Metrics Calculation

```{r performance-setup}
#| label: performance-setup
#| cache: false
#| dependson: c("game-mode-classification", "game-time-calculations")

# Create an empty list to store intermediate results
metrics_list <- list()
```

```{r overall-performance}
#| label: overall-performance
#| cache: false
#| dependson: c("performance-setup")

# Calculate overall performance metrics per player
characters_players <- characters_2week %>%
  group_by(player_id) %>% 
  summarise(across(total_kill_count:shots_hit, sum, na.rm = TRUE)) %>% 
  ungroup()

# Store in our list
metrics_list$characters_players <- characters_players

# Force garbage collection
gc()
```

```{r game-sessions}
#| label: game-sessions
#| cache: false
#| dependson: c("performance-setup")

# Calculate game sessions per player
game_sessions_total <- game_sessions_modes %>% 
  count(player_id) %>%
  rename(n_total_game_sessions = n)

# Store in our list
metrics_list$game_sessions_total <- game_sessions_total

# Force garbage collection
gc()
```

```{r sessions-by-mode}
#| label: sessions-by-mode
#| cache: false
#| dependson: c("performance-setup")

# Calculate sessions by game mode
game_sessions <- game_sessions_modes %>% 
  group_by(player_id) %>% 
  summarise(across(social_hub:pvp_session, sum, na.rm = TRUE)) %>% 
  ungroup()

# Store in our list
metrics_list$game_sessions <- game_sessions

# Force garbage collection
gc()
```

```{r pvp-performance}
#| label: pvp-performance
#| cache: false
#| dependson: c("performance-setup")

# PvP-specific performance metrics
pvp_telemetry <- characters_2week %>%
  filter(pvp_session == 1) %>%
  group_by(player_id) %>% 
  summarise(across(total_kill_count:shots_hit, sum, na.rm = TRUE)) %>%
  rename_with(~ paste0("pvp_", .), -player_id)

# Store in our list
metrics_list$pvp_telemetry <- pvp_telemetry

# Force garbage collection
gc()
```

```{r pve-performance}
#| label: pve-performance
#| cache: false
#| dependson: c("performance-setup")

# PvE-specific performance metrics
pve_telemetry <- characters_2week %>%
  filter(pve_session == 1) %>%
  group_by(player_id) %>% 
  summarise(across(total_kill_count:shots_hit, sum, na.rm = TRUE)) %>%
  rename_with(~ paste0("pve_", .), -player_id)

# Store in our list
metrics_list$pve_telemetry <- pve_telemetry

# Force garbage collection
gc()
```

```{r combine-metrics}
#| label: combine-metrics
#| cache: false
#| dependson: c("overall-performance", "game-sessions", "sessions-by-mode", "pvp-performance", "pve-performance")

# Load additional data we need from previous processing
metrics_list$game_time_players <- game_time_players

# Combine telemetry data using purrr's reduce function
telemetry_players <- list(
  metrics_list$characters_players,
  metrics_list$game_time_players,
  metrics_list$game_sessions,
  metrics_list$game_sessions_total,
  metrics_list$pvp_telemetry,
  metrics_list$pve_telemetry
) %>%
  reduce(left_join, by = "player_id")

# Replace NA values with 0
telemetry_players <- telemetry_players %>%
  mutate(across(everything(), ~replace_na(.x, 0)))

# Print summary of capped sessions
cat("Summary of capped sessions (>12 hours):\n")
cat("Total capped sessions:", sum(telemetry_players$capped_sessions), "\n")
cat("Players with capped sessions:", sum(telemetry_players$capped_sessions > 0), "\n")
cat("Maximum capped sessions per player:", max(telemetry_players$capped_sessions), "\n")

# Clear intermediate objects and list to free memory
rm(metrics_list)
gc()

# Save processed telemetry data
write_csv(telemetry_players, here("data/processed/telemetry_players.csv"))
```

# Derived Performance Metrics

```{r derived-metrics}
#| label: derived-metrics

# Define a function to calculate performance ratios
calculate_ratio <- function(numerator, denominator) {
  numerator / denominator
}

# Calculate derived performance metrics
telemetry_players <- telemetry_players %>%
  # Handle divisions by zero for K/D ratio
  mutate(
    across(
      c(total_death_count, pvp_total_death_count, pve_total_death_count),
      ~ ifelse(.x == 0, 1, .x)
    )
  ) %>%
  # Calculate all ratio metrics with purrr
  mutate(
    # K/D ratios
    total_kd = calculate_ratio(total_kill_count, total_death_count),
    pve_kd = calculate_ratio(pve_total_kill_count, pve_total_death_count),
    pvp_kd = calculate_ratio(pvp_total_kill_count, pvp_total_death_count),
    
    # Hit accuracy ratios
    total_hit_accuracy = calculate_ratio(shots_hit, shots_fired),
    pve_hit_accuracy = calculate_ratio(pve_shots_hit, pve_shots_fired),
    pvp_hit_accuracy = calculate_ratio(pvp_shots_hit, pvp_shots_fired),
    
    # Critical hit ratios
    total_crit_ratio = calculate_ratio(critical_hit_count, shots_hit),
    pve_crit_ratio = calculate_ratio(pve_critical_hit_count, pve_shots_hit),
    pvp_crit_ratio = calculate_ratio(pvp_critical_hit_count, pvp_shots_hit),
    
    # Damage per game
    total_average_damage = calculate_ratio(damage_dealt, n_total_game_sessions),
    pve_average_damage = calculate_ratio(pve_damage_dealt, pve_session),
    pvp_average_damage = calculate_ratio(pvp_damage_dealt, pvp_session),
    
    # Damage per second
    total_dps = calculate_ratio(damage_dealt, seconds),
    pve_dps = calculate_ratio(pve_damage_dealt, pve_seconds),
    pvp_dps = calculate_ratio(pvp_damage_dealt, pvp_seconds),
    
    # Score per game
    total_average_score = calculate_ratio(score, n_total_game_sessions),
    pve_average_score = calculate_ratio(pve_score, pve_session),
    pvp_average_score = calculate_ratio(pvp_score, pvp_session),
    
    # Score per second
    total_sps = calculate_ratio(score, seconds),
    pve_sps = calculate_ratio(pve_score, pve_seconds),
    pvp_sps = calculate_ratio(pvp_score, pvp_seconds)
  )

# Replace NaN values with 0
telemetry_players <- telemetry_players %>%
  mutate(across(everything(), ~ifelse(is.nan(.x), 0, .x)))

# Save processed metrics
write_csv(telemetry_players, here("data/processed/telemetry_with_metrics.csv"))
```

# Participants and Procedure

# Combine Survey and Telemetry Data

```{r merge-data}
#| label: merge-data

# Merge survey with telemetry data
pvz_df <- survey %>% 
  right_join(telemetry_players, by = "player_id")

# Check for participants without valid performance metrics
# These would be participants with zeros or NAs for all key performance indicators
pvz_df <- pvz_df %>%
  # Create a flag for participants with no meaningful telemetry data
  mutate(
    has_telemetry = !(
      # Check if key performance metrics are all zeros or NAs
      (total_kill_count == 0 | is.na(total_kill_count)) &
      (total_death_count == 0 | is.na(total_death_count)) &
      (damage_dealt == 0 | is.na(damage_dealt)) &
      (shots_fired == 0 | is.na(shots_fired)) &
      (n_total_game_sessions == 0 | is.na(n_total_game_sessions))
    )
  )

# Print summary of participants with and without telemetry
cat("Participants without valid telemetry:", sum(!pvz_df$has_telemetry, na.rm = TRUE), "\n")
cat("Participants with valid telemetry:", sum(pvz_df$has_telemetry, na.rm = TRUE), "\n")

# Remove participants without valid telemetry data
pvz_df_clean_final <- pvz_df %>%
  filter(has_telemetry == TRUE)

# Calculate ratio metrics
pvz_df_clean_final <- pvz_df_clean_final %>%
  mutate(
    ratio_social_hours = social_hours / hours,
    ratio_pve_hours = pve_hours / hours,
    ratio_pvp_hours = pvp_hours / hours
  )

# Keep participants with ANY measurable playtime (>0 hours)

pvz_df_clean_final <- pvz_df_clean_final %>%
  filter(hours > 0)

# Print final dataset dimensions
cat("Final dataset dimensions:", dim(pvz_df_clean_final)[1], "participants x", dim(pvz_df_clean_final)[2], "variables\n")

# Save final dataset
write_csv(pvz_df_clean_final, here("data/processed/pvz_complete.csv"))

# Update pvz_df_clean to match the final version for consistency
pvz_df_clean <- pvz_df_clean_final
```

## Sample Construction Summary

As detailed in @johannesVideoGamePlay2021, participants were recruited by Electronic Arts directly from their active player base via email invitations sent in two waves: the first in early August 2020 to 50,000 players, and the second in late September 2020 to 200,000 players. In total, 518 PvZ players completed the well-being survey (approximately 0.21% response rate). For details on the full dataset, which also includes *Animal Crossing: New Horizons* players, see @johannesVideoGamePlay2021.

Because our research questions concern mechanics-based playstyles, we applied several inclusion criteria to ensure the analytic sample consisted of players who actively engaged with the game's core mechanics. We first excluded participants without matching telemetry data (n = 47). We then excluded one participant with zero recorded playtime in the telemetry window. Next, we excluded 49 participants whose telemetry contained no non-zero performance or mode-engagement indicators, followed by 8 additional participants who did not have the required non-missing mode-specific performance values for clustering. These participants did not provide sufficient behavioural information to derive mechanics-based playstyles. The final analytic sample consisted of 413 combat-active players (M age = 35.0, SD = 11.8; 319 male, 81 female, 1 other, 12 undisclosed), representing 79.7% of the original 518 respondents.

# Outlier Handling

We identify and handle extreme outliers in the dataset using a 6 SD threshold. Following the methodology from the original paper, values more than 6 standard deviations from a variable mean were replaced with NA on a case-by-case basis. In this dataset, this procedure replaced 87 values (across 48 variables; 26 participants affected) and did not reduce the final analytic sample used in the clustering pipeline.

```{r outlier-removal}
#| label: outlier-removal

# Get list of numeric variables to check for outliers
numeric_vars <- pvz_df_clean_final %>%
  dplyr::select(where(is.numeric)) %>%
  colnames()

# Calculate z-scores for each numeric variable
pvz_with_z_final <- pvz_df_clean_final %>%
  mutate(across(all_of(numeric_vars), 
                ~ scale(.x)[,1],
                .names = "{.col}_z"))

# Create a list of variables with outliers (>6 SD)
z_cols_final <- colnames(pvz_with_z_final)[grep("_z$", colnames(pvz_with_z_final))]
outlier_vars_final <- c()
outlier_counts <- c()

# Count outliers per variable (using 6 SD threshold)
for(z_col in z_cols_final) {
  orig_col <- sub("_z$", "", z_col)
  outlier_count <- sum(abs(pvz_with_z_final[[z_col]]) > 6, na.rm = TRUE)
  
  if(outlier_count > 0) {
    outlier_vars_final <- c(outlier_vars_final, orig_col)
    outlier_counts <- c(outlier_counts, outlier_count)
  }
}

# Create a summary table of outliers
outlier_summary <- tibble(
  name = outlier_vars_final,
  Extremes = outlier_counts,
  Extremes_p = scales::percent(outlier_counts / nrow(pvz_df_clean_final), accuracy = .01)
)

# Display the summary
print(outlier_summary)

# Create a cleaned version with outliers replaced by NA
pvz_clean_final <- pvz_df_clean_final

# Replace outliers with NA
for(i in seq_along(outlier_vars_final)) {
  var <- outlier_vars_final[i]
  z_var <- paste0(var, "_z")
  pvz_clean_final[[var]] <- ifelse(abs(pvz_with_z_final[[z_var]]) > 6, NA, pvz_df_clean_final[[var]])
}

# Save the cleaned dataset
write_csv(pvz_clean_final, here("data/pvz_clean_final.csv"))

# Print summary statistics
cat("\nOutlier handling summary:\n")
cat("Threshold: 6 SD (aligned with original paper)\n")
cat("Number of variables with extreme outliers:", length(outlier_vars_final), "\n")
cat("Total number of data points identified as outliers:", sum(outlier_counts), "\n")

# Count participants with at least one outlier
participants_with_outliers <- c()
for(i in seq_along(outlier_vars_final)) {
  var <- outlier_vars_final[i]
  z_var <- paste0(var, "_z")
  participants_with_outliers <- c(
    participants_with_outliers, 
    pvz_with_z_final$player_id[abs(pvz_with_z_final[[z_var]]) > 6]
  )
}

cat("Number of participants with at least one extreme value:", 
    length(unique(participants_with_outliers)), "\n")
cat("\nFinal output saved: data/pvz_clean_final.csv (n =", nrow(pvz_clean_final), ")\n")
```

# Next Steps

The preprocessing is complete. The processed data has been saved to `data/pvz_clean_final.csv` with n = 470 participants. 

**Key preprocessing decisions:**
- **Outlier threshold**: 6 SD (aligned with original paper)
- **Playtime filter**: hours > 0 (any measurable playtime)
- **Telemetry requirement**: Valid combat/gameplay metrics

The main analysis script (`manuscript.qmd`) will apply additional filters:
1. Combat engagement/performance filter (removes 49 participants with no usable non-zero combat/mode indicators)
2. Mode-specific completeness filter (removes 8 participants missing required clustering metrics)
3. Final analysis sample: n = 413
