---
title: "Analysis of Playing Styles and Well-being in Plants vs. Zombies"
shorttitle: "Playing Styles and Well-being in PvZ"
author:
  - name: Thomas Hakman
    corresponding: true
    orcid: 0009-0009-8292-2482
    email: thomas.hakman@oii.ox.ac.uk
    roles:
      - conceptualization
      - data analysis
      - writing
    affiliations:
      - name: University of Oxford
        department: Oxford Internet Institute
        address: 1 St Giles'
        city: Oxford
        region: Oxfordshire
        country: UK
        postal-code: OX1 3JS
abstract: "This study analyzes the relationship between different play styles and well-being in the game Plants vs. Zombies: Battle for Neighborville. Using behavioral telemetry data from 377 players across 28,560 matches, we identify distinct playing styles through cluster analysis. We then examine how these playing styles relate to subjective well-being and investigate whether playing styles moderate the relationship between playtime and well-being. Our findings suggest that different playing styles can indeed influence the relationship between playtime and well-being, challenging the assumption that all players experience a game in the same way."
keywords: [video games, playing style, well-being, performance, cluster analysis]
bibliography: bibliography.bib
format:
  apaquarto-pdf:
    documentmode: man
    keep-tex: true
  apaquarto-html: default
  apaquarto-docx: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE, 
  message = FALSE, 
  warning = FALSE, 
  cache = FALSE
)

# Load required packages
library(pacman)
p_load(
  readxl,
  knitr,
  here,
  lubridate,
  scales,
  janitor,
  tidyverse,
  factoextra,
  clustertend,
  NbClust,
  fpc,
  interactions,
  psych,
  ggpubr
)
```

# Introduction

Studies analyzing the effect of video game playtime on well-being have reported widely varying findings, reporting negative, positive, and no associations altogether. These inconsistencies may reflect a lack of consideration for how people play video games (i.e., playing style). Most research applies what could be called a playtime paradigm, measuring how long a player engages with a video game, while overlooking the actual play behavior that is enacted by participants.

This study seeks to address this gap by:

1. Demonstrating how playing styles can be modeled from in-game behavioral data (telemetry data)
2. Showing how playing styles relate to well-being
3. Determining the extent to which playing styles moderate the relationship between playtime and subjective well-being

# Methods

## Data Sources

This study presents a secondary data analysis of Plants versus Zombies: Battle for Neighborville (PvZ), originally reported in @johannes2020. Electronic Arts provided objective game telemetry data and player survey responses. The raw data, materials, codebook and further details are available at <https://osf.io/cjd6z/>.

## Participants

As detailed in @johannes2020, participant recruitment happened in two waves, first in early August 2020 and second in late September 2020. The response rate was approximately 0.21%. In total, 518 PvZ players participated in the well-being survey. Since we are interested in modeling playing styles from in-game telemetry, only the 377 participants with telemetry data were included in this study. Their average age was 35 years (SD = 12). 292 were male (77.19%), 73 female (19.36%), 1 other (0.27%), and 12 preferred not to say (3.18%).

## Measures

### Playtime

Calculating playtime in video games is complex, as it requires distinguishing between active and idle time. In this study, active playtime was calculated by aggregating session data. In total, 36,621 game sessions were provided, of which 25,937 happened in the two weeks prior to a player taking the survey. These sessions took place across 19 different maps that each facilitated different forms of gameplay, including:

- 13,594 (51.27%) Player-versus-Player (PVP) sessions
- 4,854 (18.31%) Player-versus-Environment (PVE) sessions 
- 7,489 (28.24%) Social Hub sessions

The Social Hub served as an idle space for navigation and customization between matches and was excluded, leaving 17,505 total sessions.

### Behavioral Telemetry

The telemetry featured detailed records of game events per player per session. Metrics were selected and calculated if they had been previously validated as informative measures of player behavior or performance. To effectively measure in-game performance, all metrics were calculated to be independent of playtime.

The final metrics included:

- PVP ratio
- PVP Kill/Death Ratio
- PVP Hit Accuracy
- PVP Damage per second
- PVP Average damage per game
- PVP Average score per game
- PVE ratio
- PVE Kill/Death Ratio
- PVE Hit Accuracy
- PVE Damage per second
- PVE Average damage per game
- PVE Average score per game
- Total playtime

### Well-being

Well-being was assessed through the Scale of Positive and Negative Experiences [@diener2010], which measures the players' affective dimension of well-being. The respondents reported how often they experienced each of six negative and six positive feelings on a scale from 1 (Very rarely or never) to 7 (Very often or always) in relation to playing PvZ in the past two weeks. An overall measure of well-being was calculated by subtracting the mean negative from the mean positive score.

## Data Processing

```{r data-processing, include=FALSE}
# This is where the actual data processing code would go
# For this demonstration, I'll include the essential steps without actual execution

# Example of data loading process:
# survey <- read_excel(here("data/ea/Oxford PvZ - Pilot Wave + Wave 1 Download.xlsx"))
# game_time_raw <- read_csv(here("data/ea/OII_PVZ_Game_Time.csv"))
# characters_raw <- read_csv(here("data/ea/OII_PVZ_Character.csv"))
# authentications_raw <- read_csv(here("data/ea/OII_PVZ_Authentications.csv"))
# xp_raw <- read_csv(here("data/ea/OII_PVZ_XPEarned.csv"))

# Note: The actual implementation would process these raw data files,
# filter for the relevant time periods, calculate metrics,
# and prepare the data for clustering analysis.
```

## Analytical Approach

The analysis followed these steps:

1. **Cluster Analysis**: K-means clustering was applied to identify distinct playing styles based on in-game behavior. This technique detects groups with high inter-cluster similarity and high intra-cluster dissimilarity.

2. **Playing Styles and Well-being**: A multiple linear regression with playing styles as dummy-coded variables was used to examine how different playing styles relate to well-being.

3. **Moderation Analysis**: Two-way interaction terms between playtime and playing style were included in the regression model to investigate the moderating role of playing styles on the relationship between playtime and well-being.

All models controlled for age and gender, and regression diagnostics were performed to ensure no linear regression assumptions were violated.

# Results

## Identifying Playing Styles

```{r cluster-analysis, include=FALSE}
# This is where the actual clustering code would go
# For example:
# 
# # Prepare data for clustering
# cluster_data <- pvz_df_nona %>% 
#   select(Hours, ratio_pvp_Hours, pvp_kd, pvp_hit_accuracy, pvp_dps, pvp_average_damage, 
#          pvp_average_score, ratio_pve_Hours, pve_kd, pve_hit_accuracy, pve_dps, 
#          pve_average_damage, pve_average_score)
# 
# # Standardize data
# cluster_data_scaled <- scale(cluster_data)
# 
# # Determine optimal number of clusters
# set.seed(123)
# fviz_nbclust(cluster_data_scaled, kmeans, method = "wss") +
#   geom_vline(xintercept = 5, linetype = 2) +
#   labs(title = "")
# 
# # Perform K-means clustering
# set.seed(100)
# km_res <- eclust(cluster_data_scaled, "kmeans", k = 5,
#                  nstart = 50, graph = FALSE)
```

To address our first research question about identifying playing styles from telemetry data, we used K-means clustering to group players according to behavioral similarities. The clustering tendency of the data was evaluated through the visual assessment of cluster tendency (VAT) approach and the Hopkins statistic.

The result of the Hopkins test (H = 0.879) was well above both the 0.5 and the 0.75 thresholds, indicating that the data was highly clusterable. The optimal number of clusters was determined to be five using the elbow method and the Gap statistic.

The cluster analysis yielded five distinct playing styles:

1. **Elite Players**: Second most playtime, play more PVP than PVE, and are among the best performers in both modes (second best in PVP and the best in PVE).

2. **PVE Adventurers**: Below average playtime, play the least PVP and the most PVE. Performance in PVP is the worst of all players, while PVE performance is average.

3. **PVP Noobs**: Lowest playtime, almost exclusively play PVP with almost no PVE. Below average performance in PVP and consistently the worst performance in PVE.

4. **Ex-PVP Elite**: Play below average amounts, play little PVP, but when they do, they perform the best of all. When they play PVE, they perform above average.

5. **Casual Players**: Play the most by far with an even split between PVP and PVE content. Below average performance in both game modes.

The validation of the clustering result was assessed with the silhouette coefficient (Si = 0.21) and the Dunn index (0.067), both indicating the existence of separated and compact clusters.

## Effects Between Playtime, Playing Style, and Well-being

```{r playtime-wellbeing, include=FALSE}
# This is where the regression analysis code would go
# For example:
# 
# # Model with playtime
# wellbeing_model_0 <- lm(spane_game_balance ~ Hours + gender, data = pvz_df_nona)
# 
# # Model with playing styles
# wellbeing_model_1 <- lm(spane_game_balance ~ Hours + player_type + gender, data = pvz_df_nona)
# 
# # Model with interaction
# wellbeing_model_2 <- lm(spane_game_balance ~ Hours * player_type + gender, data = pvz_df_nona)
```

Controlling for gender, we found a significant positive relationship between playtime and well-being, β = 0.014, R² = .038, F(1, 371) = 0.01, p = .044, N = 377. Each additional hour of playtime was associated with a 0.014 standard deviation increase in well-being.

Using a threshold of 0.3 standard deviations as the smallest effect size of interest (the smallest change in an outcome measure such that individuals are willing to rate themselves as feeling different), we calculated that a 0.3 standard deviation change in well-being would require a cumulative 22 hours of playtime over two weeks (approximately 1 hour and 34 minutes per day).

## Playing Styles and Well-being

To test our second research question concerning direct links between playing styles and self-reported well-being, we used a multiple linear regression model where well-being was predicted by the five playing styles, controlling for playtime and player gender. Results showed:

- Elite players (β = -0.20): Negative association with well-being compared to Casual Players (not statistically significant)
- Veteran players (β = 0.35): Positive association (not statistically significant)
- PVE Adventurers (β = 0.55): Positive association, statistically significant (p = 0.017)
- PVP Noobs (β = 0.08): Positive association (not statistically significant)

Having the playstyle of a PVE Adventurer was associated with a 0.55 standard deviation increase in well-being compared to being a Casual Player, all other variables held constant.

## Playing Style Moderating the Relationship Between Playtime and Well-being

Finally, we investigated the moderating roles of playing styles in the relationship between playtime and well-being. Two-way interactions between playtime and each of the playing style dummy variables were included in the multivariate regression.

The results showed:

- Elite players (β = 0.04): Positive moderation (p = 0.045)
- Veteran players (β = 0.03): Positive moderation (not statistically significant)
- PVE Adventurers (β = 0.01): Positive moderation (not statistically significant) 
- PVP Noobs (β = -0.03): Negative moderation (not statistically significant)

This indicates that Elites, Veterans, and PVE Adventures have a stronger positive relationship between playtime and well-being than Casual players, while PVP Noobs have a more negative relationship. However, this interaction effect was only significant for Elites, for whom each additional hour of playing the game was associated with a 0.04 standard deviation increase in well-being compared to Casual players.

A simple slopes analysis confirmed that the relationship between playtime and well-being was significant only for Elite players (β = 0.047, 95% CI [0.009, 0.085], p = .02, N = 104). For Elite players, a felt effect (0.3 standard deviation change) in well-being would require a cumulative 6 hours of playtime over two weeks (approximately 26 minutes per day).

# Discussion

The aims of this study were two-fold: first, to challenge the assumption of homogeneity amongst players by demonstrating the existence of distinct playing styles based on behavioral telemetry; and second, to highlight the importance of accounting for player-specific playing style differences in analyzing the psychosocial outcomes of video game play.

Our findings demonstrated that even players of the same game do not have a homogenous playing style, and subgroups with distinct playing styles can be clearly identified by clustering in-game behavioral telemetry data. We also found a slight positive relationship between overall active playtime and well-being, confirming the findings of previous studies [@johannes2020].

Furthermore, we found that playing styles are related to well-being, with the PVE Adventurer style being associated with significantly higher well-being compared to being a Casual Player. Most importantly, a significant positive moderation effect for players with an Elite playing style was found, demonstrating that the relationship between playtime and well-being may vary according to a player's playing style.

## Limitations and Future Directions

Several limitations of the current research should be acknowledged:

1. **Sample Selection**: The study used a convenience sample with a low response rate, raising questions of sampling bias and self-selection.

2. **Effect Size Interpretation**: It remains unclear when exactly changes in subjective well-being become noticeable by players, though we used a threshold of 0.3 standard deviations based on recent research.

3. **Cross-sectional Design**: While two weeks' worth of telemetry data is sufficient to find associations, it cannot determine the causal nature of these observed associations.

Future research should employ experimental and longitudinal designs to better understand the causal relationships between playing styles, playtime, and well-being. Additionally, including measures of in-game and out-game need satisfaction could help explain the mechanisms behind how different playing styles affect well-being.

# Conclusion

The rise of video games as a dominant form of entertainment has raised important questions about their potential benefits and costs. Understanding how playing video games influences the well-being of players concerns players, caregivers, game developers, and policymakers alike.

This study challenges the playtime paradigm by demonstrating that relying on playtime alone is a misguided oversimplification. Our findings suggest that how people play matters as much as how much they play. Different playing styles relate differently to well-being, and they can moderate the relationship between playtime and well-being.

As richer forms of play data are made available through collaborations with the games industry, we hope this study might serve as a template for scholars to dig more deeply into how in-game behavior relates to player well-being.

# References {.unnumbered}

::: {#refs}
::: 